---
title: "alpha/beta diversity and the significantly different taxa of anti CD3 samples"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

```{r}
library(tidyverse, quietly = T)
library(ggpubr)
library(vdbR)
connect_database()
get_table_from_database('asv_alpha_diversity_ag')
```

# the metaphlan results

```{r}
fns <- list.files('../data/anticd3/', pattern = '.txt', full.names = T)

taxa <- fns %>% 
  set_names(fns) %>% 
  map(~ read_tsv(., skip = 4, col_types = 'ccddd') %>% 
          select(species  = `#clade_name`, counts = estimated_number_of_reads_from_the_clade) %>% 
          filter(str_detect(species, 's__')) ) %>% 
  bind_rows(.id = 'sampleid') %>% 
  filter(!str_detect(species, 'k__Viruses')) %>% 
  mutate(sampleid = str_replace(sampleid, '../data/anticd3//Sample_',''),
         sampleid = str_replace(sampleid, '_IGO_.+$',''))

# make a table of the pheno of the samples
pheno <- tibble(
  sampleid = taxa %>% distinct(sampleid) %>% pull(sampleid),
  grp = c(rep('control', 3), rep('grp1', 3),rep('grp2', 3))
)
```

# alpha and beta diversity

```{r}
# to look at the alpha diversity
library(vegan)
cts_relab <- fns %>% 
  set_names(fns) %>% 
  map(~ read_tsv(., skip = 4, col_types = 'ccddd') %>% 
          select(species  = `#clade_name`, relative_abundance) %>% 
          filter(str_detect(species, 's__')) ) %>% 
  bind_rows(.id = 'sampleid') %>% 
  filter(!str_detect(species, 'k__Viruses')) %>% 
  mutate(sampleid = str_replace(sampleid, '../data/anticd3//Sample_',''),
         sampleid = str_replace(sampleid, '_IGO_.+$',''))

cts_ <- cts_relab %>% 
  spread(key = 'species', value = 'relative_abundance', fill = 0) %>% 
  column_to_rownames('sampleid')



meta <- diversity(cts_, index = 'inv') %>% 
  enframe(name = 'sampleid', value = 'inv') %>% 
  inner_join(pheno)

all_alpha <- meta %>% 
  ggboxplot(x = 'grp', y = 'inv', add = 'dotplot', ylab = 'inverse simpson',
            title = 'Alpha diversity (grp1: lower grp2: higher)',
            add.params = list(color = 'forestgreen', size = 2, alpha = 1, shape = 20)) +
  stat_compare_means(comparisons= list(c('control', 'grp1'),
                                       c('control', 'grp2'),
                                       c('grp1', 'grp2')),
										label= "p.format",
										method= 'wilcox.test',
										correct=FALSE) +
  theme_bw()


all_alpha
```
```{r}
# beta diversity
# calculate the Bray curtis beta diversity 

dist_ <- vegdist(cts_, method = 'bray')
eigen <- cmdscale(dist_, eig = T)$eig
percent_var <- signif(eigen/sum(eigen), 3)*100
bc <- cmdscale(dist_, k = 2)
beta_df <- bc %>%
  as.data.frame() %>%
  rownames_to_column('sampleid')  %>% 
  full_join(meta)

beta_df %>% 
  ggscatter(x = 'V1', y = 'V2', color =  'grp', palette = 'lancet',  alpha = 1) +
  labs(title = 'Bray curtis distance') +
  xlab(paste0("PC 1 [",percent_var[1],"%]")) +
  ylab(paste0("PC 2 [",percent_var[2],"%]")) +
  theme(aspect.ratio=1, legend.position = 'right')

```

# Sig different taxa at species level (with FDR < 0.25)

## Corncob

```{r}
library(corncob)
library(phyloseq)
cts <- taxa %>% 
  spread('sampleid','counts', fill = 0) %>% 
  column_to_rownames('species')

# I think I need to build a phyloseq object for this
all.equal(pheno$sampleid, colnames(cts))

# the 6 samples that are control and grp1
cts1 <- taxa %>% 
  filter(!str_detect(sampleid, 'D7_3')) %>% 
  spread('sampleid','counts', fill = 0) %>% 
  column_to_rownames('species')
samples1 = sample_data(pheno %>% filter(grp != 'grp2') %>% column_to_rownames('sampleid'))

OTU1 = otu_table(cts1, taxa_are_rows = T)
physeq1 = phyloseq(OTU1,  samples1)

set.seed(1)
dv_analysis1 <- differentialTest(
  formula = ~ grp,
  phi.formula = ~ grp,
  formula_null = ~ 1,
  phi.formula_null = ~ grp,
  test = "LRT", boot = FALSE,
  data = physeq1,
  fdr_cutoff = 0.25) 


# the 6 samples that are control and grp2
cts2 <- taxa %>% 
  filter(!str_detect(sampleid, 'D7_2')) %>% 
  spread('sampleid','counts', fill = 0) %>% 
  column_to_rownames('species')
samples2 = sample_data(pheno %>% filter(grp != 'grp1') %>% column_to_rownames('sampleid'))

OTU2 = otu_table(cts2, taxa_are_rows = T)
physeq2 = phyloseq(OTU2,  samples2)


dv_analysis2 <- differentialTest(
  formula = ~ grp,
  phi.formula = ~ grp,
  formula_null = ~ 1,
  phi.formula_null = ~ grp,
  test = "LRT", boot = FALSE,
  data = physeq2,
  fdr_cutoff = 0.25)
```

The sig different species between control and lower from Corncob

```{r}
# the sig different species between control and lower
Corncob1 <- dv_analysis1$significant_taxa
Corncob1
```

The sig different species between control and higher from Corncob

```{r}
# the sig different species between control and higher
Corncob2 <- dv_analysis2$significant_taxa
Corncob2
```

## Maaslin2

```{r}
library(Maaslin2)

cts_1 <- cts_relab %>% 
  filter(!str_detect(sampleid, 'D7_3')) %>% 
  spread('species','relative_abundance', fill = 0) %>% 
  column_to_rownames('sampleid')

samples_1 <- pheno %>% 
  filter(!str_detect(sampleid, 'D7_3')) %>% 
  column_to_rownames('sampleid')

all.equal(rownames(samples_1), rownames(cts_1))
  
Maaslin1 <- Maaslin2(
    input_data = cts_1, 
    input_metadata = samples_1, 
    normalization = "TSS",
    min_abundance = 0.0,
    min_prevalence = 0.0,
    max_significance = 0.25,
    output = "../data/output_shotgun-Maaslin1", 
    fixed_effects = c("grp"),
    reference = c("grp,control"))


sig1 <- read_tsv('../data/output_shotgun-Maaslin1/significant_results.tsv') 
```
No sig species between control and lower from Maaslin2

```{r}
cts_2 <- cts_relab %>% 
  filter(!str_detect(sampleid, 'D7_2')) %>% 
  spread('species','relative_abundance', fill = 0) %>% 
  column_to_rownames('sampleid')

samples_2 <- pheno %>% 
  filter(!str_detect(sampleid, 'D7_2')) %>% 
  column_to_rownames('sampleid')

all.equal(rownames(samples_2), rownames(cts_2))
  
Maaslin2 <- Maaslin2(
    input_data = cts_2, 
    input_metadata = samples_2, 
    normalization = "TSS",
    min_abundance = 0.0,
    min_prevalence = 0.0,
    max_significance = 0.25,
    output = "../data/output_shotgun-Maaslin2", 
    fixed_effects = c("grp"),
    reference = c("grp,control"))

sig2 <- read_tsv('../data/output_shotgun-Maaslin2/significant_results.tsv') 
```
The species sig different between control and higher from Maaslin2

```{r}
sig2$feature
```

## Metagenomeseq

```{r}
library(metagenomeSeq)

pheno_data1 <- pheno %>% 
  filter(!str_detect(sampleid, 'D7_3')) %>% 
  column_to_rownames('sampleid') %>% 
  mutate(grp = factor(grp, levels = c('control','grp1')))

counts1 <- cts %>% 
  select(!starts_with('D7_3'))
  
all.equal(colnames(counts1) , rownames(pheno_data1))


# assemble the object for metagenomeSeq

phenotypeData1 <- AnnotatedDataFrame(pheno_data1)

d1 <- newMRexperiment(counts1,phenoData=phenotypeData1)

p1 = cumNormStatFast(d1)


d1 = filterData(d1, present = 1, depth = 1) 
d1 <- cumNorm(d1, p = p1) 
pd1 <- pData(d1) 
mod1 <- model.matrix(~1 + grp, data = pd1) 
d1_res = fitFeatureModel(d1, mod1) 


whole1 <- tibble(species = d1_res@taxa,
                pval = d1_res@pvalues) %>% 
  mutate(padj = p.adjust(pval, method = 'BH')) %>% 
  filter(padj < 0.25)

```
No sig species between control and lower from Metagenomeseq

```{r}

pheno_data2 <- pheno %>% 
  filter(!str_detect(sampleid, 'D7_2')) %>% 
  column_to_rownames('sampleid') %>% 
  mutate(grp = factor(grp, levels = c('control','grp2')))

counts2 <- cts %>% 
  select(!starts_with('D7_2'))
  
all.equal(colnames(counts2) , rownames(pheno_data2))


# assemble the object for metagenomeSeq

phenotypeData2 <- AnnotatedDataFrame(pheno_data2)

d2 <- newMRexperiment(counts2,phenoData=phenotypeData2)

p2 = cumNormStatFast(d2)


d2 = filterData(d2, present = 1, depth = 1) 
d2 <- cumNorm(d2, p = p2) 
pd2 <- pData(d2) 
mod2 <- model.matrix(~1 + grp, data = pd2) 
d2_res = fitFeatureModel(d2, mod2) 


whole2 <- tibble(species = d2_res@taxa,
                pval = d2_res@pvalues) %>% 
  mutate(padj = p.adjust(pval, method = 'BH')) %>% 
  filter(padj < 0.25)
```
The species sig different between control and higher from Metagenomeseq

```{r}
whole2$species
```

## Conclusion

All three tools identified s__Oscillibacter_sp_1_3 as the sig different species between control and higher.