---
title: "Create and filter count matrix"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(phyloseq)
library(phangorn)
library(dada2)
library(DECIPHER)
library(tidyverse)
library(kableExtra)
```

```{r}
# load the big table from the db
source('~/db.R')
CTS <- get_data_from_query_OTU(0,'asv_counts_ag')
SEQ <- get_data_from_query_OTU(0,'asv_sequences_ag')
ANNOT <- get_data_from_query_OTU(0, 'asv_annotation_blast_ag')
```


## Make sure the sampleids in your table can be found in our database (need to clean when you can't)

```{r}
# load the pheno table (contains at least the sampleid column and a column with phenotype information that you are interested in, such as Toxicity Y and N, Responding status: responder and non_responder)
pheno <- read_csv('../data/final_pheno.csv') %>% 
  # some samples have "..pool........" attached at the end of the sampleid, remove them
  mutate(postgres_ID = str_replace(postgres_ID, '\\.\\.pool.+$',''))


# lets look at the key information in this pheno table
pheno %>% 
  select(Patient_ID, Sample_ID, postgres_ID, Timepoint, Response_Status )
```

- There are several unconventional sampleids in the `Sample_ID` column, such as DP.1. That's because they are from external insitutions. Luckily John helped to link those to the sampleids in our own database (only need a little bit cleaning like talked above). 

```{r}
# look at some details of the samples

# how many sample total?
nrow(pheno)
```
```{r}
# how many patients total?
# this can be found by the patient ID 
pheno %>% 
  dplyr::count(Patient_ID)
```
So only 2 of them have only one time point of sample, but all the others have both day 0 and day 28 data.

```{r}
# how many responders and non_responders?
# every sample has a responding status variable, but it really depends on the day 28 response. In other words, for the sample patient's samples, day 0 response will be the same as day 28's.

pheno %>% 
  filter(Timepoint == 'd28') %>% 
  dplyr::count(Response_Status)

```



```{r}
# Do we have all of these samples counts in the database
# how many samples counts we have
have_counts <-CTS %>% 
  filter(sampleid %in% pheno$postgres_ID) %>% 
  dplyr::count(sampleid) 

nrow(have_counts)  # that one missing sample is not there because it failed sequencing due to some amplification problem.
```

n is the number of ASV identified in each sample. 

```{r}
# now recreate the pheno table but with only the samples that we have 
pheno <- pheno %>% 
  filter(postgres_ID %in% have_counts$sampleid)
```



```{r}
# get the unique asv_seq to build a tree.
seqtab <- CTS %>% 
  filter(sampleid %in% pheno$postgres_ID) %>% 
  select(asv_key, sampleid, count) %>% 
  left_join(SEQ %>% 
              dplyr::rename(asv_key = key), by  = 'asv_key')

seqs <- seqtab %>% 
  distinct(asv_sequence) %>% 
  pull(asv_sequence)

names(seqs) <- seqs

# tree building and the last step takes a lot of time!
alignment <- AlignSeqs(DNAStringSet(seqs), anchor=NA)
phang.align <- phyDat(as(alignment, "matrix"), type="DNA")
dm <- dist.ml(phang.align)
treeNJ <- NJ(dm) # Note, tip order != sequence order
fit = pml(treeNJ, data=phang.align)
fitGTR <- update(fit, k=4, inv=0.2)
fitGTR <- optim.pml(fitGTR, model="GTR", optInv=TRUE, optGamma=TRUE,
 rearrangement = "stochastic", control = pml.control(trace = 0))
detach("package:phangorn", unload=TRUE)
```


