---
title: "Look at the patients to be included in the cohort and also alpha diversity of the samples"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

```{r}
library(tidyverse)
library(ggpubr)
```

```{r}
# the ethnicity data on the MSK set that was used in the nejm paper (“frozen set”) here: smb://skimcs.mskcc.org/vandenbrinklab/deep_sequencing/Clinical Annotation/Ethnicity/ethnicity_nejm_frozen_set.csv
pheno <- read_csv('/Volumes/vandenBrinkLab/deep_sequencing/Clinical Annotation/Ethnicity/ethnicity_nejm_frozen_set.csv') %>% 
  select(-X1, -X1_1)
```

Question: the ethnicity_simple should be used ??

```{r}
pheno %>% 
  head()
```
```{r}
pheno %>% 
  count(ethnicity_simple) %>% 
  mutate(grp = if_else(ethnicity_simple == 'White', 'white', if_else(ethnicity_simple == 'other', 'other', 'nonwhite'))) %>% 
  group_by(grp) %>% 
  summarise(grp_sum = sum(n)) %>% 
  # removing the other cuz don't know their race
  filter(grp != 'other') %>% 
  mutate(grp_perc = round(grp_sum/sum(grp_sum)*100, 2))
```

```{r}
# the patients mrn that gonna be used in this analysis: white and nonwhite (excluding the other)
target <- pheno %>% 
  filter(ethnicity_simple != 'other') %>% 
  select(mrn=patient_id, ethnicity=ethnicity_simple)

nrow(target)
```

## Find all the samples for the patients and get the day relative to transplant

```{r}
source('~/db.R')
castori <- get_data_from_query_OTU(0, 'samples_castori_ag')
allo <- get_data_from_query_OTU(0, 'patient_allo_ag')
# get the alpha diversity from the db
aldiv <- get_data_from_query_OTU(0, 'asv_alpha_diversity_ag')
```

```{r}
DF <- target %>% 
  left_join(allo %>% 
              filter(str_detect(indication, fixed('initial', ignore_case = T))) %>%  
              select(mrn, hct), by  = 'mrn') %>% 
  left_join(castori %>% 
              filter(sampletype == 'Stool' | sampletype == 'stool') %>% 
              select(mrn, sampleid, datecollection), by = 'mrn') %>% 
  mutate(drt = datecollection - hct) %>% 
  # filter on the drt < 1000 & drt > -100
  filter(drt < 1000 & drt > -100)
```

```{r}
# look at the dist of drt
DF %>% 
  gghistogram('drt', bins = 30,
              title = 'Distribution of day relative to transplant')
```

```{r}
# join to get the alpha diversity and also the count total to see if need to filter
all <- DF %>% 
  left_join(aldiv %>% 
              select(sampleid, simpson_reciprocal, shannon, count_total), by = 'sampleid') %>% 
  # there are duplicated sequenced samples remove the lower count ones
  arrange(sampleid, desc(count_total)) %>% 
  distinct(sampleid, .keep_all = T)
```

```{r}
# look at this all table
all %>% 
  summary 
```
```{r}
# We may not have the alpha div and the total counts in the db for a given sample in castori
# the reason may be: 
# the samples are collected but not sequenced
# or couldn't be sequenced because failed to amplify
# or finished sequencing but haven't gone thru the pipeline thus not in the db (new samples)
all_temp <- all %>% 
  filter(!is.na(count_total)) %>% 
  filter(count_total >= 500) %>% 
  # simplify the ethnicity
  mutate(eth = if_else(ethnicity == 'White', 'white', 'nonwhite'))
```


## alpha diversity

- are there ethnicity differences in alpha diversity? 
- Can look at first baseline sample (earliest sample per patient collected day -30 to day -6). 
- Can look at the median diversity per patient in the window day 7-21 as the "periengraftment period".

### baseline

```{r}
# in the baseline window: collected day -30 to day -6
df_baseline <- all_temp %>% 
  filter(drt >= -30 & drt <= -6) 
```

```{r}
# earliest baseline sample
df_baseline_first <- df_baseline %>% 
  group_by(mrn) %>% 
  arrange(drt) %>% 
  slice(1)

nrow(df_baseline_first)
```


```{r}
df_baseline_first %>% 
  ggboxplot(x = 'eth', y = 'simpson_reciprocal', add = 'jitter', xlab = 'ethnicity',
            title = 'Alpha div simposon reciprocal comparison') +
  stat_compare_means(comparisons = list(c('white', 'nonwhite')),
                    label = "p.signif",
                    method = 'wilcox.test',
                    correct=FALSE)


df_baseline_first %>% 
  ggboxplot(x = 'eth', y = 'shannon', add = 'jitter', xlab = 'ethnicity',
            title = 'Alpha div shannon comparison') +
  stat_compare_means(comparisons = list(c('white', 'nonwhite')),
                    label = "p.signif",
                    method = 'wilcox.test',
                    correct=FALSE)
```

### peri-engraftment

```{r}
# peri 

df_peri <- all_temp %>% 
  filter(drt >= 7 & drt <= 21) 

# median div in the above window
df_peri_med_div <- df_peri %>% 
  group_by(mrn, eth) %>% 
  summarise(med_sr = median(simpson_reciprocal),
            med_shan = median(shannon))

nrow(df_peri_med_div)
```

```{r}
df_peri_med_div %>% 
  ggboxplot(x = 'eth', y = 'med_sr', add = 'jitter', xlab = 'ethnicity',
            title = 'Median alpha div simposon reciprocal peri comparison') +
  stat_compare_means(comparisons = list(c('white', 'nonwhite')),
                    label = "p.signif",
                    method = 'wilcox.test',
                    correct=FALSE)

 
df_peri_med_div %>% 
  ggboxplot(x = 'eth', y = 'med_shan', add = 'jitter', xlab = 'ethnicity',
            title = 'Median alpha div Shannon peri comparison') +
  stat_compare_means(comparisons = list(c('white', 'nonwhite')),
                    label = "p.signif",
                    method = 'wilcox.test',
                    correct=FALSE)
```


```{r}
# save the two time period tables for use in other scripts
df_baseline %>% 
  write_csv('../data/df_baseline.csv')

df_peri %>% 
  write_csv('../data/df_peri.csv')
```


```{r}
# aldiv %>% 
#   filter(sampleid == 'FMT.0061M')
# # find the samples that don't have alpha div value in the db 
# nodiv <- all %>% 
#   filter(is.na(count_total)) %>% 
#   select(sampleid)
# 
# nodiv %>% 
#   write_csv('../data/no_alpha_div_sampleid.csv')
```

