---
title: "Aggregate counts at different taxa level"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

```{r}
library(tidyverse)
```

## Input: a counts matrix, with asv_keys as the row names, and sampleids as the columns

```{r}
samps <- read_csv('../../ppp_Nilu/data/msk duke sdtable 03 02 20.csv') %>% 
  rename(sampleid = sampleid_clean)
```

## Connect to the db and load the tables

```{r}
source('~/db.R')
source('~/db_connect_simple.R')
get_table_from_database('counts_ag')
#q2_str = get_data_from_query_OTU("2_str"); #get the taxonomy.  This has full taxonomy
if(!exists("q2_str")){
  q2_str = get_data_from_query_OTU("2_str")
}
```
   

```{r}
# get the counts 
# remove the pool number
# get the total number of counts for all the sample
CTS_total <- counts_ag %>% 
  group_by(sampleid) %>% 
  summarise(total_count = sum(count))

CTS <- counts_ag %>% 
  left_join(CTS_total) %>% 
  mutate(sampleid = str_replace(sampleid, '\\.\\.pool.+$','')) %>% 
  # remove the lower coverage one
  arrange(sampleid, otu_key, desc(total_count)) %>% 
  distinct(sampleid, otu_key, .keep_all = T)


cts <- CTS %>% 
  filter(sampleid %in% samps$sampleid)

cts %>% count(sampleid) %>% nrow
```

```{r}
cts <- cts %>% 
  select(otu_key, sampleid, count) %>% 
  spread(key = 'sampleid', value = 'count', fill = 0) %>% 
  arrange(otu_key) %>% 
  column_to_rownames('otu_key') %>% 
  as.matrix()

cts %>% 
  write.csv('../data/nilu_cts.csv', row.names = T, quote = F)


cts <- read.csv('../data/nilu_cts.csv', row.names = 1, check.names = F)
```


## Summarize count matrix at different taxa levels

```{r}
# for the specific samples above 
annot <- q2_str %>% 
  mutate(key = as.character(key)) %>% 
  filter(key %in% rownames(cts))

colnames(annot)
```


```{r}
# the species level summarization
cts_species <- cts %>% 
  as.data.frame() %>% 
  rownames_to_column('key') %>% 
  full_join(annot %>%  select(key, s_full), by  = 'key') %>% 
  select(-key) %>% 
  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
  group_by(sampleid, s_full) %>% 
  summarise(cnt = sum(count)) %>% 
  # get the total count from the db to calculate the relab
  left_join(CTS_total, by = 'sampleid') %>% 
  mutate(relab = cnt/total_count) %>% 
  select(sampleid, s_full, relab) %>% 
  spread(key = 'sampleid', value = 'relab') %>% 
  column_to_rownames('s_full') 

# the genus level summarization
cts_genus <- cts %>% 
  as.data.frame() %>% 
  rownames_to_column('key') %>% 
  full_join(annot %>%  select(key, g_full), by  = 'key') %>% 
  select(-key) %>% 
  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
  group_by(sampleid, g_full) %>% 
  summarise(cnt = sum(count)) %>% 
  # get the total count from the db to calculate the relab
  left_join(CTS_total, by = 'sampleid') %>% 
  mutate(relab = cnt/total_count) %>% 
  select(sampleid, g_full, relab) %>% 
  spread(key = 'sampleid', value = 'relab') %>% 
  column_to_rownames('g_full') 

# the family level summarization
cts_family <- cts %>% 
  as.data.frame() %>% 
  rownames_to_column('key') %>% 
  full_join(annot %>%  select(key, f_full), by  = 'key') %>% 
  select(-key) %>% 
  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
  group_by(sampleid, f_full) %>% 
  summarise(cnt = sum(count)) %>% 
  # get the total count from the db to calculate the relab
  left_join(CTS_total, by = 'sampleid') %>% 
  mutate(relab = cnt/total_count) %>% 
  select(sampleid, f_full, relab) %>% 
  spread(key = 'sampleid', value = 'relab') %>% 
  column_to_rownames('f_full') 

# the order level summarization
cts_order <- cts %>% 
  as.data.frame() %>% 
  rownames_to_column('key') %>% 
  full_join(annot %>%  select(key, o_full), by  = 'key') %>% 
  select(-key) %>% 
  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
  group_by(sampleid, o_full) %>% 
  summarise(cnt = sum(count)) %>% 
  # get the total count from the db to calculate the relab
  left_join(CTS_total, by = 'sampleid') %>% 
  mutate(relab = cnt/total_count) %>% 
  select(sampleid, o_full, relab) %>% 
  spread(key = 'sampleid', value = 'relab') %>% 
  column_to_rownames('o_full') 

# the class level summarization
cts_class <- cts %>% 
  as.data.frame() %>% 
  rownames_to_column('key') %>% 
  full_join(annot %>%  select(key, c_full), by  = 'key') %>% 
  select(-key) %>% 
  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
  group_by(sampleid, c_full) %>% 
  summarise(cnt = sum(count)) %>% 
  # get the total count from the db to calculate the relab
  left_join(CTS_total, by = 'sampleid') %>% 
  mutate(relab = cnt/total_count) %>% 
  select(sampleid, c_full, relab) %>% 
  spread(key = 'sampleid', value = 'relab') %>% 
  column_to_rownames('c_full') 


# the phylum level summarization
cts_phylum <- cts %>% 
  as.data.frame() %>% 
  rownames_to_column('key') %>% 
  full_join(annot %>%  select(key, p_full), by  = 'key') %>% 
  select(-key) %>% 
  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
  group_by(sampleid, p_full) %>% 
  summarise(cnt = sum(count)) %>% 
  # get the total count from the db to calculate the relab
  left_join(CTS_total, by = 'sampleid') %>% 
  mutate(relab = cnt/total_count) %>% 
  select(sampleid, p_full, relab) %>% 
  spread(key = 'sampleid', value = 'relab') %>% 
  column_to_rownames('p_full')

# convert the count in otu cts table to relab
cts_relab <- cts %>% 
  as.data.frame() %>% 
  rownames_to_column('key') %>% 
  full_join(annot %>%  select(key, otu_full), by  = 'key') %>% 
  select(-key) %>% 
  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.)-1])  %>% 
  left_join(CTS_total, by = 'sampleid') %>% 
  mutate(relab = count/total_count) %>% 
  select(sampleid, otu_full, relab) %>% 
  spread(key = 'sampleid', value = 'relab')  %>% 
  column_to_rownames('otu_full')
```


```{r}
# bind all those matrix together including the otu one
ALL <- rbind(cts_relab,
             cts_species,
             cts_genus,
             cts_family,
             cts_order,
             cts_class,
             cts_phylum)
```

*The output table(ALL) will be a matrix with samples in the columns, and features in the rows. The features include ASV_key, full taxa up till species level, full taxa up till genus level... full taxa up till phylum level.*


## See the number of features in each taxa level in the combined big table (also a way to subset different level)


```{r}
# look at the number of features in each taxa group
tally_feature <- data_frame(feature = rownames(ALL)) %>% 
  mutate(grp = if_else(str_detect(feature, 'otu__OTU'), 'OTU',
                       if_else(str_detect(feature, 's__'),'species',
                               if_else(str_detect(feature, 'g__'),'genus',
                                       if_else(str_detect(feature, 'f__'),'family',
                                               if_else(str_detect(feature, 'o__'),'order',
                                                       if_else(str_detect(feature, 'c__'),'class', 'phylum'))))))) %>% 
  mutate(grp = factor(grp, levels = c('OTU',
                                      'species',
                                      'genus',
                                      'family',
                                      'order',
                                      'class',
                                      'phylum')))

  
tally_feature %>% 
  count(grp)
```

