---
title: "Aggregate counts at different taxa level"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

```{r}
library(tidyverse)
```


## Input: a counts matrix, with asv_keys as the row names, and sampleids as the columns

```{r}
samps <- read_csv('../../ppp_Nilu/data/msk duke sdtable 03 02 20.csv') %>% 
  rename(sampleid = sampleid_clean)
```

## Connect to the db and load the tables

```{r}
source('~/db.R')
source('~/db_connect_simple.R')
CTS <- get_data_from_query_OTU(0, 'counts_ag')
#q2_str = get_data_from_query_OTU("2_str"); #get the taxonomy.  This has full taxonomy
if(!exists("q2_str")){
  q2_str = get_data_from_query_OTU("2_str")
}
```
   

```{r}
# get the counts 
# no duplicates
# remove the pool number
CTSR <- counts_ag %>% 
  mutate(sampleid = str_replace(sampleid, '\\.\\.pool.+$',''))


cts <- CTSR %>% 
  filter(sampleid %in% samps$sampleid)
cts %>% count(sampleid)



```

```{r}
cts_ma <- cts %>% 
  select(otu_key, sampleid, count) %>% 
  spread(key = 'sampleid', value = 'count', fill = 0) %>% 
  arrange(otu_key) %>% 
  column_to_rownames('otu_key') %>% 
  as.matrix()

```


## Summarize count matrix at different taxa levels

```{r}
# for the specific samples above 
annot <- asv_annotation_blast_ag %>% 
  filter(asv_key %in% rownames(cts)) %>% 
  mutate(ordr =  if_else(ordr == '', str_glue('of_class_{class}'), ordr),
         family =  if_else(family == '', str_glue('of_order_{ordr}'), family),
         genus =  if_else(genus == '', str_glue('of_family_{family}'), genus),
         species =  if_else(species == '', str_glue('of_genus_{genus}'), species)) %>% 
  mutate(taxa_species = str_glue('k__{kingdom}|p__{phylum}|c__{class}|o__{ordr}|f__{family}|g__{genus}|s__{species}'),
         taxa_genus = str_glue('k__{kingdom}|p__{phylum}|c__{class}|o__{ordr}|f__{family}|g__{genus}'),
         taxa_family = str_glue('k__{kingdom}|p__{phylum}|c__{class}|o__{ordr}|f__{family}'),
         taxa_order = str_glue('k__{kingdom}|p__{phylum}|c__{class}|o__{ordr}'),
         taxa_class = str_glue('k__{kingdom}|p__{phylum}|c__{class}'),
         taxa_phylum = str_glue('k__{kingdom}|p__{phylum}'))


# the species level summarization
cts_species <- cts %>% 
  as.data.frame() %>% 
  rownames_to_column('asv_key') %>% 
  full_join(annot %>%  select(asv_key, taxa_species), by  = 'asv_key') %>% 
  select(-asv_key) %>% 
  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
  group_by(sampleid, taxa_species) %>% 
  summarise(cnt = sum(count)) %>% 
  # get the total count from the db to calculate the relab
  left_join(asv_counts_ag %>% distinct(sampleid,count_total ), by = 'sampleid') %>% 
  mutate(relab = cnt/count_total) %>% 
  select(sampleid, taxa_species, relab) %>% 
  spread(key = 'sampleid', value = 'relab') %>% 
  column_to_rownames('taxa_species') 

# the genus level summarization
cts_genus <- cts %>% 
  as.data.frame() %>% 
  rownames_to_column('asv_key') %>% 
  full_join(annot %>%  select(asv_key, taxa_genus), by  = 'asv_key') %>% 
  select(-asv_key) %>% 
  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
  group_by(sampleid, taxa_genus) %>% 
  summarise(cnt = sum(count)) %>% 
  # get the total count from the db to calculate the relab
  left_join(asv_counts_ag %>% distinct(sampleid,count_total ), by = 'sampleid') %>% 
  mutate(relab = cnt/count_total) %>% 
  select(sampleid, taxa_genus, relab) %>% 
  spread(key = 'sampleid', value = 'relab') %>% 
  column_to_rownames('taxa_genus') 

# the family level summarization
cts_family <- cts %>% 
  as.data.frame() %>% 
  rownames_to_column('asv_key') %>% 
  full_join(annot %>%  select(asv_key, taxa_family), by  = 'asv_key') %>% 
  select(-asv_key) %>% 
  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
  group_by(sampleid, taxa_family) %>% 
  summarise(cnt = sum(count)) %>% 
  # get the total count from the db to calculate the relab
  left_join(asv_counts_ag %>% distinct(sampleid,count_total ), by = 'sampleid') %>% 
  mutate(relab = cnt/count_total) %>% 
  select(sampleid, taxa_family, relab) %>% 
  spread(key = 'sampleid', value = 'relab') %>% 
  column_to_rownames('taxa_family') 

# the order level summarization
cts_order <- cts %>% 
  as.data.frame() %>% 
  rownames_to_column('asv_key') %>% 
  full_join(annot %>%  select(asv_key, taxa_order), by  = 'asv_key') %>% 
  select(-asv_key) %>% 
  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
  group_by(sampleid, taxa_order) %>% 
  summarise(cnt = sum(count)) %>% 
  # get the total count from the db to calculate the relab
  left_join(asv_counts_ag %>% distinct(sampleid,count_total ), by = 'sampleid') %>% 
  mutate(relab = cnt/count_total) %>% 
  select(sampleid, taxa_order, relab) %>% 
  spread(key = 'sampleid', value = 'relab') %>% 
  column_to_rownames('taxa_order') 

# the class level summarization
cts_class <- cts %>% 
  as.data.frame() %>% 
  rownames_to_column('asv_key') %>% 
  full_join(annot %>%  select(asv_key, taxa_class), by  = 'asv_key') %>% 
  select(-asv_key) %>% 
  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
  group_by(sampleid, taxa_class) %>% 
  summarise(cnt = sum(count)) %>% 
  # get the total count from the db to calculate the relab
  left_join(asv_counts_ag %>% distinct(sampleid,count_total ), by = 'sampleid') %>% 
  mutate(relab = cnt/count_total) %>% 
  select(sampleid, taxa_class, relab) %>% 
  spread(key = 'sampleid', value = 'relab') %>% 
  column_to_rownames('taxa_class') 


# the phylum level summarization
cts_phylum <- cts %>% 
  as.data.frame() %>% 
  rownames_to_column('asv_key') %>% 
  full_join(annot %>%  select(asv_key, taxa_phylum), by  = 'asv_key') %>% 
  select(-asv_key) %>% 
  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
  group_by(sampleid, taxa_phylum) %>% 
  summarise(cnt = sum(count)) %>% 
  # get the total count from the db to calculate the relab
  left_join(asv_counts_ag %>% distinct(sampleid,count_total ), by = 'sampleid') %>% 
  mutate(relab = cnt/count_total) %>% 
  select(sampleid, taxa_phylum, relab) %>% 
  spread(key = 'sampleid', value = 'relab') %>% 
  column_to_rownames('taxa_phylum') 


# convert the count in asv cts table to relab
cts_relab <- cts %>% 
  as.data.frame() %>% 
  rownames_to_column('asv_key') %>% 
  gather(key = 'sampleid', value = 'count', names(.)[2]:names(.)[ncol(.)])  %>% 
  left_join(asv_counts_ag %>% distinct(sampleid,count_total ), by = 'sampleid') %>% 
  mutate(relab = count/count_total) %>% 
  select(sampleid, asv_key, relab) %>% 
  spread(key = 'sampleid', value = 'relab')  %>% 
  column_to_rownames('asv_key') 

# bind all those matrix together including the asv one
ALL <- rbind(cts_relab,
             cts_species,
             cts_genus,
             cts_family,
             cts_order,
             cts_class,
             cts_phylum)
```

*The output table(ALL) will be a matrix with samples in the columns, and features in the rows. The features include ASV_key, full taxa up till species level, full taxa up till genus level... full taxa up till phylum level.*


## See the number of features in each taxa level in the combined big table (also a way to subset different level)


```{r}
# look at the number of features in each taxa group
tally_feature <- data_frame(feature = rownames(ALL)) %>% 
  mutate(grp = if_else(str_detect(feature, '^asv'), 'ASV',
                       if_else(str_detect(feature, 's__'),'species',
                               if_else(str_detect(feature, 'g__'),'genus',
                                       if_else(str_detect(feature, 'f__'),'family',
                                               if_else(str_detect(feature, 'o__'),'order',
                                                       if_else(str_detect(feature, 'c__'),'class', 'phylum'))))))) %>% 
  mutate(grp = factor(grp, levels = c('ASV',
                                      'species',
                                      'genus',
                                      'family',
                                      'order',
                                      'class',
                                      'phylum')))


tally_feature %>% 
  count(grp)
```

