---
title: "Match the species name"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

```{r}
library(tidyverse)
library(ggpubr)
library(kableExtra)
```

## Load the normalized absolute quantification from BioMe

```{r}
normcts <- read_csv('../data/frombiome/Merged Client Report_only_abundance.csv', 
                    col_types = cols(.default = col_number(), `Sample ID` = col_character())) %>% 
  rename_all(~ gsub(" ", "_", .)) %>% 
  gather('species', 'normcts', names(.)[2]: names(.)[ncol(.)])
```


```{r}
# count the number of NA in each sample
normcts %>% 
  filter(is.na(normcts)) %>% 
  count(Sample_ID) %>% arrange(n)

all_NA_rm <- normcts %>% 
  filter(is.na(normcts)) %>% 
  count(Sample_ID) %>% 
  filter(n == 106) %>% 
  pull(Sample_ID)

# 20 samples to be reomved completely
normcts__ <- normcts %>% 
  filter(! Sample_ID %in% all_NA_rm) %>% 
  rename(sampleid = Sample_ID)

# how many effective samples went into comparison 
normcts__ %>% 
  distinct(sampleid) %>% 
  nrow
  
```

## Load our relab of the species of the same samples

```{r}
# our spp
source('~/db_connect_simple.R')
connect_database(config_file = '~/dbConfig.txt')

get_table_from_database('asv_annotation_blast_ag')
get_table_from_database_predefined_filter('asv_counts_ag')


ourcts <- asv_counts_ag %>% 
  filter(sampleid %in% normcts__$sampleid) %>% 
  select(asv_key, sampleid, count) %>% 
  spread(key = 'sampleid', value = 'count', fill = 0) %>% 
  arrange(asv_key)  


annot <- asv_annotation_blast_ag %>% 
  filter(asv_key %in% ourcts$asv_key) %>% 
  mutate(ordr =  if_else(ordr == '', str_glue('of_class_{class}'), ordr),
         family =  if_else(family == '', str_glue('of_order_{ordr}'), family),
         genus =  if_else(genus == '', str_glue('of_family_{family}'), genus),
         species =  if_else(species == '', str_glue('of_genus_{genus}'), species)) %>% 
  mutate(taxa_species = str_glue('k__{kingdom}|p__{phylum}|c__{class}|o__{ordr}|f__{family}|g__{genus}|s__{species}'))


# replace the asv_key with the genus taxa level
# summarize from asv level to genus level
# get the relative abundance for the genera
ourcts_all <- ourcts %>% 
  full_join(annot %>%  select(asv_key, taxa_species), by  = 'asv_key') %>% 
  select(-asv_key) %>% 
  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
  group_by(sampleid, taxa_species) %>% 
  summarise(cnt = sum(count)) %>% 
  # get the total count from the db to calculate the relab
  left_join(asv_counts_ag %>% distinct(sampleid,count_total ), by = 'sampleid') %>% 
  mutate(relab = cnt/count_total) %>% 
  select(sampleid, taxa_species, relab) %>% 
  mutate(species = str_extract(taxa_species, 's__.+$'),
         species = str_replace(species, '^s__','')) %>% 
  select(sampleid, species, relab) %>% 
  ungroup() %>% 
  # get rid of the [] in the genus name
  mutate(species = str_replace_all(species, '\\[|\\]',''))
```

## How many are overlapping and how many are missing

```{r}
# the overlap of the species name
biome_spp <- normcts__ %>% 
  distinct(species) %>% 
  pull(species)

our_spp <- ourcts_all %>% 
  ungroup() %>% 
  distinct(species)%>%  
  pull(species)

# a venn diagram
library(ggvenn)
two_list <- list(
  our_spp = our_spp,
  biome_spp = biome_spp)

ggvenn(two_list)
```


```{r}
# overlap

only_biome <- tibble(species = setdiff(biome_spp, our_spp)) %>% 
  mutate(genus = str_extract( species,"[^_]+"))
  

our_Bacteroides <- asv_annotation_blast_ag %>% 
  filter(genus == 'Bacteroides') %>% 
  distinct(species)


# check the mean norm counts in the biome table for the missing spp in our counts
mean_missing <- normcts__ %>% 
  filter(species %in% only_biome$species) %>% 
  group_by(species) %>% 
  summarise(mean_normcts = mean(normcts, na.rm = T),
            median_normcts = median(normcts, na.rm = T)) %>% 
  mutate(in_blast_db = species %in% asv_annotation_blast_ag$species)


mean_missing %>% 
  gghistogram(x = 'mean_normcts', title = 'Mean norm cts in the biome table of the spp that is missing in our cts table') 

mean_missing %>% 
  gghistogram(x = 'median_normcts', title = 'Median norm cts in the biome table of the spp that is missing in our cts table') 

color_false <- which(mean_missing$in_blast_db == FALSE)

mean_missing %>% 
  kable(caption = 'species missing in our quantification of the same samples') %>% 
  kable_styling(full_width = F) %>%
  row_spec(color_false,  color = "red")

```

## Look at some missing species

```{r}
# take a look at the ones that have mean_normcts > 2000
big <- mean_missing %>% 
  filter(mean_normcts > 2000)
```


```{r}
# the biggest difference 
our_Streptococcus <- asv_annotation_blast_ag %>% 
  filter(genus == 'Streptococcus') %>% 
  distinct(species)

# Streptococcus_thermophilus seems like a very common bacteria
# but in our blast db
asv_annotation_blast_ag %>% 
  filter(str_detect(species, '_thermophilus'))
```


- others
- Bifidobacterium_longum also seems quite common
- also Lactobacillus_paracasei
- what about Enterococcus_hirae 
- what to do with Bifidobacterium_longum_subsp._longum	 this kind of format


