---
title: "look at the kraken output"
author: "Angel"
date: "2022-08-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
library(vdbR)
connect_database()
get_table_from_database('shotgun_lookup_ad')
```

Percentage of fragments covered by the clade rooted at this taxon
Number of fragments covered by the clade rooted at this taxon
Number of fragments assigned directly to this taxon
A rank code, indicating (U)nclassified, (R)oot, (D)omain, (K)ingdom, (P)hylum, (C)lass, (O)rder, (F)amily, (G)enus, or (S)pecies. Taxa that are not at any of these 10 ranks have a rank code that is formed by using the rank code of the closest ancestor rank with a number indicating the distance from that rank. E.g., "G2" is a rank code indicating a taxon is between genus and species and the grandparent taxon is at the genus rank.
NCBI taxonomic ID number
Indented scientific name

```{bash}

for file in ../data/archaea/*_kraken2.report
  do echo $file
  ./../data/KrakenTools/kreport2mpa.py  --report-file $file --output $file.out  --percentages  --intermediate-ranks
  done
```

```{r}
fns <- list.files('../data/archaea/', full.names = T, pattern = 'out$') 
all <- tibble(fn = fns)
test <- read_delim(fns[1])
kraken <- fns %>% 
  set_names(fns) %>% 
  map(~ read_delim(., col_names = F, col_types = 'cd')  %>% 
        select(taxa = X1, perc = X2)) %>% 
  bind_rows(.id = 'sampleid') %>% 
  mutate(
         sampleid = str_extract(sampleid, '//.+_'),
         sampleid = str_replace(sampleid, '//',''),
         sampleid = str_replace(sampleid, '_*+$',''))

ids <- kraken %>% 
  distinct(sampleid) %>% 
  mutate(fid = str_replace(sampleid, 'Sample_', ''),
         fid = str_replace(fid, '_IGO.+$', '')) %>% 
  inner_join(shotgun_lookup_ad %>% select(fid, cleaned_sampleid = sampleid))

archaea_k <- kraken %>% 
  filter(str_detect(taxa, 'k__Archaea$'))

archaea <- kraken %>% 
  filter(str_detect(taxa, 'k__Archaea')) %>% 
  left_join(ids) %>% 
  select(-fid)
archaea %>% write_csv('../data/kraken_shotgun_archaea_perc.csv')

```







