---
title: "the new 30 cleaning"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
library(readxl)
library(lubridate)
```

```{r}
# the computrition output of the new 30
new <- read_excel('../data/source/new30_all.xlsx') %>% 
  distinct(Food_NSC)

# the ones that we already have food code
uni <- read_excel('../data/source/eaten_FNDDS.xlsx', sheet = 1)

uniq_food <- uni %>% 
  distinct(item1_value_raw_raw,  .keep_all = T)

length(intersect(uniq_food$item1_value_raw_raw, new$Food_NSC))

# the ones that don't have exact match with the food nsc thus don't have a code
unknown <- new %>% 
  filter(! Food_NSC %in% uniq_food$item1_value_raw_raw) %>% 
  arrange(Food_NSC) %>% 
  # remove the title lines (not food)
  filter(!str_detect(Food_NSC, '^\\*\\*\\*\\*\\*\\*\\*'))

unknown %>% 
  write_csv('../data/source/new30_unknown_food_nsc.csv')

```
```{r}
# 
final <- read_csv('../data/finalized/all_patients_record_w_food_code.csv') %>% 
  filter(Food_NSC %in% c('Cheese and Crackers Snack Pack','Cheese and Crackers, SF','2017 Panini, Mozzarella Cheese & Tomato (3.5)','2017 Sandwich, Tomato & Mozzarella Pesto Panini','Pie, Apple a La Mode','2017 Salad, Shrimp Caesar - Entree (1.5)','Lettuce and Tomato Garnish'))
```
```{r}
# the food in the first 67 patients that have 2 items food
multiple <- read_csv('../data/finalized/all_patients_record_w_food_code.csv') %>% 
  filter(Food_NSC %in% c('Cheese and Crackers Snack Pack','Cheese and Crackers, SF','2017 Panini, Mozzarella Cheese & Tomato (3.5)','2017 Sandwich, Tomato & Mozzarella Pesto Panini','Pie, Apple a La Mode','2017 Salad, Shrimp Caesar - Entree (1.5)','Lettuce and Tomato Garnish')) %>% 
  select(Food_NSC: Sodium_mg) %>% 
  distinct()

multiple %>% 
  write_csv('../data/source/foods_to_be_split.csv')

# the current cleaned table
final <- read_csv('../data/finalized/all_patients_record_w_food_code.csv')
```

# new 30

## in the list of unique foods with new food codes, ignore the NA ones and ignore the one with 2 items, and combine with the codes I previously have

```{r}
new_unique <- read_excel('../data/new30_FNDDS_Peter.xlsx') %>% 
  filter(Food.Code != 'NA') %>% 
  filter(is.na(...6)) %>% 
  select(Food_NSC:FNDDS.Main.Food.Description)


# the previous unique food and food code I have 
uniq_food_code <- uniq_food %>% 
  distinct(item1_value_raw_raw, Food.Code, FNDDS.Main.Food.Description) %>% 
  rename(Food_NSC = item1_value_raw_raw) %>% 
  mutate(Food.Code = as.character(Food.Code))

# ALL of the unique food nsc and its code!
all_code <- bind_rows(
  uniq_food_code,
  new_unique
)

all_code %>% 
  write_csv('../data/cleaned_diet_data/all_food_nsc_with_code.csv')
```

## get the food code for all items in the new 30 pts

```{r}
all_code <- read_csv('../data/cleaned_diet_data/all_food_nsc_with_code_FINAL.csv')

new_all <- read_excel('../data/source/new30_all.xlsx')  %>% 
  select(- ...1, - Food) %>% 
  left_join(all_code, by = 'Food_NSC') %>% 
  filter(Por_eaten != 'na') %>% 
  filter(!str_detect(Food_NSC, '^\\*\\*\\*\\*\\*\\*\\*')) %>% 
  mutate(Por_eaten = if_else(str_detect(Por_eaten, '^.'), str_glue('0{Por_eaten}'), Por_eaten)) %>% 
  mutate(Por_eaten = as.numeric(Por_eaten)) %>% 
  filter(Por_eaten > 0) %>% 
  select(-Por_ordered)  %>% 
  # remove the + symbol
   mutate(Calories_kcal = str_replace_all(Calories_kcal, '\\+$',''),
           Proteing_g = str_replace_all(Proteing_g, '\\+$',''),
           Fat_g = str_replace_all(Fat_g, '\\+$',''),
           Carbohydrates_g = str_replace_all(Carbohydrates_g, '\\+$',''),
           Fibers_g = str_replace_all(Fibers_g, '\\+$',''),
           Sugars_g = str_replace_all(Sugars_g, '\\+$',''),
           Sodium_mg = str_replace_all(Sodium_mg, '\\+$','')) %>% 
  mutate_at(vars(ends_with('g')), funs(as.numeric)) %>% 
  mutate(Calories_kcal = as.numeric(Calories_kcal))  %>% 
  select(-Food.Code, -FNDDS.Main.Food.Description) 
  

new_all %>% 
  write_csv('../data/cleaned_diet_data/new_all_cleaned_with_NA.csv')
```


```{r}
# find the distinct foods (also the portion eaten) that have NA or blanks
new_all_blanks <- new_all %>% 
  select(Food_NSC: Sodium_mg) %>% 
  distinct()  %>% 
  filter(Por_eaten != 'na') %>% 
  #mutate(Por_eaten = if_else(str_detect(Por_eaten, '^.'), str_glue('0{Por_eaten}'), Por_eaten)) %>% 
  mutate(Por_eaten = as.numeric(Por_eaten)) %>% 
  filter(Por_eaten > 0) %>% 
  filter(!str_detect(Food_NSC, '^\\*\\*\\*\\*\\*\\*\\*')) %>% 
  filter(is.na(Calories_kcal) | 
           is.na(Proteing_g) |
           is.na(Fat_g) |
           is.na(Carbohydrates_g) |
           is.na(Fibers_g) |
           is.na(Sugars_g) |
           is.na(Sodium_mg) ) %>% 
  arrange(Food_NSC)

 
new_all_blanks %>% 
  write_csv('../data/source/new_30_records_with_NA.csv')
```

 
## THE NA corrected version from peter

```{r}
na_cor <- read_excel('../data/new_30_records_with_NA_Peter.xlsx') %>% 
  filter(Calories_kcal != 'NA') %>% 
  mutate_at(vars(ends_with('g')), funs(as.numeric)) %>% 
  mutate(Calories_kcal = as.numeric(Calories_kcal)) %>% 
  distinct()
  


# merge into the original new 30 df
new_all_split_NA <- new_all %>% 
  split(.$Food_NSC %in% na_cor$Food_NSC)

cor_all_NA <- new_all_split_NA %>% 
  pluck('TRUE') %>% 
  select(MRN: Por_eaten) %>% 
  left_join(na_cor, by = c('Food_NSC','Por_eaten'))

NA_cor_all_df <- bind_rows(
  new_all_split_NA %>% 
    pluck('FALSE'),
  cor_all_NA
)  
colnames(NA_cor_all_df)

new <- NA_cor_all_df %>% 
  select(-Name) %>% 
  mutate(Date = mdy(Date))

new %>% 
  summary()
```


```{r}
# something that doesn't sound right ....
new %>% 
  filter(Calories_kcal %in% c(702,949,1822)) %>% 
  write_csv('../data/!question/suspicious_ones_from_new_30_patients.csv')

high_ques <- new %>% 
  filter(Calories_kcal %in% c(702,949,1822)) 

```

```{r}
new %>% 
  filter(str_detect(Food_NSC, '2017 Quesadilla, Chicken')) %>% 
  distinct() %>% 
  write_csv('../data/!question/Quesadilla_chicken.csv')
```

```{r}
# in the *new* table the NA ones , the only thing worth lookng at is the toast BR ones
new %>% 
  #filter(is.na(Fibers_g)) %>% 
  filter(str_detect(Food_NSC, 'Toast')) %>% 
  arrange(Food_NSC) %>% 
  write_csv('../data/!question/toast_I_have_question.csv')
```

**Reply from Peter:**
Quesadilla is okay as is. It was a lot of food and a lot of kcal...

Also, I have attached the toast sheet back here. You can remove all of those as they are instructions to "toast the bread... but they still come up." So, they can be entirely removed. 

```{r}
# replace the two extreme ones: Brocollini and Philly Cheese Steak with the corrected values
suspicious2 <- read_csv('../data/!question/suspicious_ones_from_new_30_patients_peter.csv') %>% 
  filter(!str_detect(Food_NSC, 'Quesadilla'))

new_cor1_split <- new %>% 
  split(.$Calories_kcal %in% c(702,949))

corrected_new_cor1_split <- new_cor1_split %>% 
  pluck('TRUE') %>% 
  select(MRN: Por_eaten) %>% 
  left_join(suspicious2 %>% select(Food_NSC:Sodium_mg),  by = c("Food_NSC", "Unit", "Por_eaten")) 

new_cor1 <- bind_rows(
  corrected_new_cor1_split,
  new_cor1_split %>% pluck('FALSE')
)

new_cor1 %>% 
  summary 
```
```{r}
# the nas in the current new_cor1 they can all be removed.

new_cor2 <- new_cor1 %>%     
  filter(!is.na(Sugars_g))
```


# combine the above like essential diet columns from old 67 and the new 30 to create a ALL table

```{r}
old <- read_csv('../data/!question/all_cor6_old_67.csv')
 
combined <- bind_rows(old, new_cor2)
combined %>% 
  write_csv('../data/!question/all_97_cor.csv')
```

**Now jump to another script to examine the values more thoroughly**

# now find the unique item names with calories == 0

```{r}
cal_0 <- combined %>% 
  filter(Calories_kcal == 0) %>% 
  distinct(Food_NSC,Por_eaten ) %>% 
  arrange(Food_NSC) 

cal_0 %>% 
  write_csv('../data/cal_0_food_nsc_all.csv')

combined %>% 
  filter(Food_NSC == 'Specials, Apple Cinnamon Crepes (2)')
```


```{r}
# get the unit for these items 
new_w_unit <- read_excel('../data/source/new30_all.xlsx')  %>% 
  select(- ...1, - Food) %>% 
  filter(Por_eaten != 'na') %>% 
  mutate(Por_eaten = as.numeric(Por_eaten)) %>% 
  filter(Por_eaten > 0) %>% 
  distinct(Food_NSC, Unit, Por_eaten)

# old pts with unit 
old <- read_excel('../data/source/new_list.xls') %>% 
  filter(!str_detect(Food_NSC, '^\\*\\*\\*\\*\\*\\*\\*')) %>% 
  filter(Por_eaten != 'na') %>% 
  mutate(Por_eaten = as.numeric(Por_eaten)) %>% 
  filter(Por_eaten > 0) %>% 
  # I have a code for them
  filter(Food_NSC %in% all_code$Food_NSC) %>%  
  mutate(Date = mdy(Date)) %>% 
  # get rid of the + in the values
  mutate(Calories_kcal = str_replace_all(Calories_kcal, '\\+$',''),
           Proteing_g = str_replace_all(Proteing_g, '\\+$',''),
           Fat_g = str_replace_all(Fat_g, '\\+$',''),
           Carbohydrates_g = str_replace_all(Carbohydrates_g, '\\+$',''),
           Fibers_g = str_replace_all(Fibers_g, '\\+$',''),
           Sugars_g = str_replace_all(Sugars_g, '\\+$',''),
           Sodium_mg = str_replace_all(Sodium_mg, '\\+$','')) %>% 
    mutate(Calories_kcal = as.numeric(Calories_kcal),
           Proteing_g = as.numeric(Proteing_g),
           Fat_g = as.numeric(Fat_g),
           Carbohydrates_g = as.numeric(Carbohydrates_g),
           Fibers_g = as.numeric(Fibers_g),
           Sugars_g = as.numeric(Sugars_g),
           Sodium_mg = as.numeric(Sodium_mg))  %>% 
  distinct(Food_NSC, Unit, Por_eaten)

all_w_unit <- bind_rows(
  old, new_w_unit
)

cal_0_w_unit <- all_w_unit %>% 
  inner_join(cal_0) %>% 
  arrange(Food_NSC)

cal_0_w_unit %>% 
  write_csv('../data/cal_0_food_nsc_all_with_unit.csv')
```


# what's next:

peter send me the updated gram weight for the watery stuff

```{r}
cal0_peter <- read_excel('../data/cal_0_food_nsc_all_with_unit_Peter.xls')

# the things that should be removed from combined
na_rm <- cal0_peter %>% 
  filter(`gram weight` == 'NA') %>% 
  distinct(Food_NSC) %>% 
  pull(Food_NSC)

combined2 <- combined %>% 
  filter(! Food_NSC %in% na_rm)

cal0_peter_items <- cal0_peter %>% 
  filter(`gram weight` != 'NA')  %>% 
  rename(total_weight = `gram weight`) %>% 
  mutate(total_weight = as.numeric(total_weight)) %>% 
  distinct(Food_NSC) %>% 
  pull(Food_NSC)

cal0_peter_clean <- cal0_peter %>% 
  filter(`gram weight` != 'NA')  %>% 
  rename(total_weight = `gram weight`) %>% 
  mutate(total_weight = as.numeric(total_weight)) %>% 
  distinct()

cal0_split <- combined2 %>% 
  split(.$Food_NSC %in% cal0_peter_items)

cal0_join_with_other <- cal0_split %>% 
  pluck('TRUE') %>%  
  left_join(cal0_peter_clean) %>% 
  filter(is.na(total_weight))  %>% 
  select(-Unit, -total_weight)


# for the stuff in the above result that is NA in the gram weight goddamn messed up calories for the food! do a calculation instead
cal0_weight <- cal0_split %>% 
  pluck('TRUE') %>%  
  left_join(cal0_peter_clean) %>% 
  filter(!is.na(total_weight))  %>% 
  select(-Unit) %>% 
  # join by the food code
  left_join(all_code, by = 'Food_NSC')

# most of the cal!=0 stuff that needs the fndds conversion
other_need <- bind_rows(
  cal0_split %>% pluck('FALSE'),
  cal0_join_with_other
) %>% 
  # join by the food code
  left_join(all_code, by = 'Food_NSC')

```

```{r}
cal0_weight %>% 
  write_csv('../data/cleaned_diet_data/cal0_weight_with_code.csv')

other_need %>% 
  write_csv('../data/cleaned_diet_data/other_need_with_code.csv')
```


## correct the Specials, Apple Cinnamon Crepes (2)

```{r}
# one thing I need to correct is this   Specials, Apple Cinnamon Crepes in other_need
other_need %>% 
  filter(str_detect(Food_NSC, 'Apple Cinnamon'))
```
 ok I will manually correct in the csv file
