---
title: "the patient info meta data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script is for organizing the meta data of the patient info, including transplant and nutrition info. intensity and source from the patient allo ag table. age, sex and admit_bmi from the nutrition_demographics_ag table.

```{r}
library(tidyverse)
library(readxl)
library(lubridate)
source('~/db_connect_simple.R')
connect_database(config_file = '~/dbConfig.txt')
#dbListTables(con) %>% data.frame() %>% filter(grepl('nutri',.))
```

```{r}
dtb <- read_csv('../data/cleaned_diet_data/FINAL_97_with_code_all_foods_drt_with_EN_finalized.csv')
```


## the demographic info

```{r}
# the current nutritional info that we have

#nutri <- get_table_from_database('nutrition_ag')
get_table_from_database('nutrition_demographics_ag')

# the info from this table that I wanna integrate into my meta
demo <- nutrition_demographics_ag %>% 
  select(mrn, sex, age_at_hsct, admit_bmi) %>% 
  filter(mrn %in% dtb$mrn)

nrow(demo)

# THERE ARE A DOZEN NA FOR THE PATIENTS the sex age and bmi column
demo %>% 
  filter(is.na(sex))

# ignore this table for now


# patients that missing the weight
wt <- nutrition_demographics_ag %>% 
  filter(mrn %in% dtb$mrn)  %>% 
  filter(is.na(admit_wt)) %>% 
  pull(mrn)
```

## the transplant info, only the source and intensity for now

```{r}
# getting the conmorbitity index from the table
list_table_from_database(pattern = 'comorbidity')
get_table_from_database('comorbidity_index_ag')
get_table_from_database('patients_msk_comorbidity_index_ag')

ci_pt <- patients_msk_comorbidity_index_ag %>% 
  filter(mrn %in% dtb$mrn)

have_ci_pts <- ci_pt %>% 
  filter(!is.na(ci))

need_ci_pts <- setdiff(dtb$mrn, have_ci_pts %>% pull(mrn))
# the above is the list that I need to have the CI

length(intersect(wt, need_ci_pts))

need_info <- tibble(
  mrn = need_ci_pts
) %>% 
  left_join(nutrition_demographics_ag %>% filter(mrn %in% dtb$mrn) %>% select(mrn, admit_wt))

need_info %>% 
  write_csv('../data/!question/patients_missing_CI_and_wt_diet.csv')


ci_needed <- ptb %>% 
  filter(mrn %in% need_ci_pts) %>% 
  dplyr::select(mrn, hct)

ci_needed %>% 
  write_csv('../data/!question/CI_needed.csv')
```

```{r}
# all the weights of the patients
filled <- read_csv('../data/!question/patients_missing_HCTCI_diet_peter.csv') %>% 
  select(mrn, admit_wt)

have_had <- nutrition_demographics_ag %>% 
  filter(mrn %in% dtb$mrn) %>% 
  select(mrn, admit_wt) %>% 
  filter(!is.na(admit_wt))

full_wt <- filled %>% 
  full_join(have_had)
```


also getting the age and sex from this table

```{r}
get_table_from_database('patient_allo_ag')

# Collapse MSKCC messy diseases into broad, simplified categories
patient_allo_ag$disease.simple[patient_allo_ag$disease == "Leukemia" &
                                         patient_allo_ag$sub_hist == "Acute" &
                                         patient_allo_ag$sub_sub_hist == "Myelogenous"] <- "AML"
patient_allo_ag$disease.simple[patient_allo_ag$disease == "Leukemia" &
                                         patient_allo_ag$sub_hist == "Acute" &
                                         patient_allo_ag$sub_sub_hist == "Lymphoblastic"] <- "ALL"
patient_allo_ag$disease.simple[patient_allo_ag$disease == "Leukemia" &
                                         patient_allo_ag$sub_hist == "Chronic" &
                                         patient_allo_ag$sub_sub_hist == "Myelogenous"] <- "CML"
patient_allo_ag$disease.simple[patient_allo_ag$disease == "Leukemia" &
                                         patient_allo_ag$sub_hist == "Chronic" &
                                         patient_allo_ag$sub_sub_hist == "Myelomonocytic"] <- "MDS/MPN"
patient_allo_ag$disease.simple[patient_allo_ag$disease == "Leukemia (Second Primary)" &
                                         patient_allo_ag$sub_hist == "Acute" &
                                         patient_allo_ag$sub_sub_hist == "Myelogenous"] <- "AML"
patient_allo_ag$disease.simple[patient_allo_ag$disease == "Leukemia" &
                                         patient_allo_ag$sub_hist == "Acute" &
                                         (patient_allo_ag$sub_sub_hist == "Monoblastic" | 
                                            patient_allo_ag$sub_sub_hist == "Monocytic" |
                                            patient_allo_ag$sub_sub_hist == "Myelomonocytic" |
                                            patient_allo_ag$sub_sub_hist =="Promyelocytic")   ] <- "AML"
patient_allo_ag$disease.simple[grepl("Aplastic", patient_allo_ag$disease)] <- "AA"
patient_allo_ag$disease.simple[grepl("Aplastic", patient_allo_ag$sub_hist)] <- "AA"
patient_allo_ag$disease.simple[grepl("Myelodysplastic|Myeloproliferative", patient_allo_ag$disease)] <- "MDS/MPN"
patient_allo_ag$disease.simple[grepl("Non-Hodgkin", patient_allo_ag$disease)] <- "NHL"
patient_allo_ag$disease.simple[grepl("Myeloma", patient_allo_ag$disease)] <- "Myeloma"
patient_allo_ag$disease.simple[grepl("Hodgkin's Disease", patient_allo_ag$disease)] <- "Hodgkins"
patient_allo_ag$disease.simple[grepl("CLL", patient_allo_ag$sub_sub_hist)] <- "CLL"
patient_allo_ag$disease.simple %>% table(useNA ="ifany") %>% data.frame() %>% arrange(-Freq)
patient_allo_ag$disease.simple[grepl("Prolymphocytic", patient_allo_ag$sub_sub_hist)] <- "NHL"
patient_allo_ag$disease.simple[grepl("Biphenotypic|Plasmacytoid Dendritic|Ambiguous|NK-LGL", patient_allo_ag$sub_sub_hist)] <- "other"
patient_allo_ag$disease.simple[grepl("Non-Malignant Hemat Disorders", patient_allo_ag$disease)] <- "other"
patient_allo_ag$disease.simple %>% table(useNA  = "ifany")
```


```{r}
other <- patient_allo_ag %>% 
  filter(str_detect(indication, fixed('initial', ignore_case = T))) %>% 
  select(mrn, source , intensity, hct, age, sex, disease.simple) %>%  
  filter(mrn %in% dtb$mrn) %>% 
  mutate(source = if_else(str_detect(source, 'Cord' ), 
                             'cord',
                             if_else(str_detect(source, 'SBA|CD34'), 
                                     'TCD',
                                     'unmodified'))) %>% 
  mutate(intensity = if_else(intensity %in% c('Ablative','ABLATIVE'), 'ablative',
                                         if_else(intensity %in% c('NONABL','Nonablative'), 'nonablative','reduced'))) %>% 
  mutate(disease.simple = if_else(is.na(disease.simple), 'AML', disease.simple))
```


```{r}

ptb <- bind_rows(other, new1) %>% 
  full_join(full_wt)
```

```{r}
# the break down of the disease type
ptb %>% 
  count(disease.simple) %>% 
  mutate(disea_perc = round(n/97*100,1)) %>% 
  arrange(-disea_perc)


ptb %>% 
  count(source) %>% 
  mutate(source_perc = round(n/97*100,1)) %>% 
  arrange(-source_perc)
```


```{r}
# output this table this is the patient meta table
ptb %>% 
  write_csv('../data/cleaned_patients/diet_patients_97.csv')
```


## a more comprehensive summary of the data

```{r}

ptb <- read_csv('../data/cleaned_patients/diet_patients_97.csv')
range(ptb$age)
mean(ptb$age)

ci <- read_csv('../data/cleaned_patients/angel_hctci.csv')  %>% 
  select(mrn, hct_ci) %>% 
  rename(ci = hct_ci)

old_ci <- ci_pt %>% 
  select(mrn, ci) %>% 
  filter(!is.na(ci))

CI <- bind_rows(
  ci, old_ci
)

PTB <- ptb %>% 
  full_join(CI) 

PTB %>% 
  write_csv('../data/cleaned_patients/diet_patients_97.csv')
```


```{r}
library(gtsummary) 
PTB <- read_csv('../data/cleaned_patients/diet_patients_97.csv')

final <- PTB %>% 
  select(age, sex, Disease = disease.simple, source,  intensity) %>% 
  mutate(Disease = case_when(
    Disease == 'NHL' ~ "Non-Hodgkin's lymphoma",
    # Myelodysplastic/myeloproliferative neoplasms
    Disease == 'MDS/MPN' ~ 'MDS/MPN',
    Disease == 'AML' ~ 'Acute myeloid leukemia',
    Disease == 'ALL' ~ 'Acute lymphoid leukemia',
    Disease == 'CLL' ~ 'Chronic lymphocytic leukemia',
    Disease == 'CML' ~ 'Other',
    Disease == 'Hodgkins' ~ "Other",
    Disease == 'Myeloma' ~ 'Myeloma', 
    Disease == 'AA' ~ 'Other'
  )) %>% 
  mutate(source = case_when(
    source == 'unmodified' ~ "Unmodified bone marrow or PBSC",
    source == 'cord' ~ "Cord blood",
    source == 'TCD' ~ "T-cell depleted PBSC"
  )) %>% 
  mutate(intensity = case_when(
    intensity == 'nonablative' ~ "Nonmyeloablative",
    intensity == 'ablative' ~ "Ablative",
    intensity == 'reduced' ~ "Reduced intensity"
  )) %>% 
  mutate(intensity = fct_reorder(intensity, intensity, .fun=length, .desc = T),
         source = fct_reorder(source, source, .fun=length, .desc = T),
         Disease = fct_reorder(Disease, Disease, .fun=length, .desc = T),
         sex = fct_reorder(sex, sex, .fun=length, .desc = T)) %>% 
  rename(`Graft type` = source, 
         #`HCT-CI` = ci,
         `Intensity of conditioning regimen` = intensity ) 

final %>% 
  tbl_summary(
    type = list(all_integer() ~ "continuous"), 
    statistic = list(all_continuous() ~ "{mean} ({sd})")
  ) %>%
  bold_labels() %>% 
  as_gt() %>%
  #gt::tab_footnote(footnote = "MDS/MPN = Myelodysplastic/myeloproliferative neoplasms") %>% 
  gt::gtsave(filename = "../figs/paper/003_table_1_patient_summary.pdf")

```



## for Tsoni's survival analysis I need to send the data

```{r}
faith <- read_tsv('../data/cleaned_diet_data/FINAL_97_faith_pd/alpha-diversity.tsv') %>% 
  separate(X1, into = c('mrn', 'fdrt'), sep = 'd', convert = T)  %>% 
  mutate(mrn = as.numeric(str_replace(mrn, 'P',''))) 

# a summary of cal per day per patient 
day_cal <- dtb %>% 
  group_by(mrn, fdrt) %>% 
  summarise(cal = sum(Calories_kcal))

send <- faith %>% 
  inner_join(day_cal) %>% 
  left_join(ptb)

send %>% 
  write_csv('../data/diet_pt_daily_cal_and_faith.csv')

colnames(send)
```

## add the donor source 

```{r}
get_table_from_database('patient_allo_ag')
donor<- patient_allo_ag %>% 
  filter(mrn %in% dtb$mrn) %>% 
  distinct(mrn, .keep_all = T) %>% 
  select(mrn, hla) %>% 
  count(hla)
  mutate(donor_clean = if_else(str_detect(donor, 'Unrelated'), 'Unrelated', 'Related')) %>% 
  count(donor_clean) %>% 
  mutate(perc = round(n/sum(n)*100))
```

