---
title: "integrating enteral nutrition info"
output: html_document   
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)
```

This is going to include all patients enteral info, but currently only have 67's.

# fill in the diet matrix  

```{r}
dtb <- read_csv('../data/cleaned_diet_data/FINAL_97_with_code_all_foods_drt.csv') %>% 
  select(mrn,fdrt, Meal, Food_NSC, Por_eaten:Sodium_g, Food_code:description, dehydrated_weight )

```


## create *0* counts for the entero nutrition info (TPN)

EN needs to have the specific nutrition information!

```{r}
entero <- read_csv('../data/cleaned_diet_data/TPN_EN patients.csv') %>% 
  filter(mrn %in% dtb$mrn) %>% 
  rename(fdrt = Days.Post.HSCT) %>% 
  mutate(nut_type = if_else(TPN.1..EN.2 == 1, "TPN", "EN")) %>% 
  select(mrn, fdrt, nut_type) %>% 
  filter(nut_type == 'TPN')

```
 
```{r}
# to see if there is overlap between the normal diet and the above one
overlap <- entero %>% 
  inner_join(dtb) %>% 
  distinct(mrn, fdrt, nut_type)

overlap %>% 
  distinct(nut_type)

# since all of the overlap are TPN which doesn't go thru colon, so ignore them in the entero table
```
  
```{r}
entero_clean <- entero %>% 
  left_join(overlap, by = c("mrn", "fdrt")) %>% 
  filter(is.na(nut_type.y))  %>% 
  rename(nut_type = nut_type.x) %>% 
  select(-nut_type.y) %>% 
  select(- nut_type)
```
 
```{r}
ftb_colnames <- colnames(dtb)
```
 
 
```{r}
# use all of the foods this patients has ever eaten as the food_nsc
create_fake_food_cts_for_enteral_patient <- function(mrn_, fdrt_){
  this_pt = dtb %>% 
    filter(mrn == mrn_) %>% 
    distinct(mrn, Food_NSC, Food_code, description) %>% 
    mutate(fdrt = fdrt_, 
           Meal = "fake meal",
           Por_eaten = 0,
           Calories_kcal = 0, 
           Proteing_g = 0, 
           Fat_g = 0,
           Carbohydrates_g = 0,
           Fibers_g = 0,
           Sugars_g =0,
           Sodium_g =0,
           dehydrated_weight = 0) %>% 
    select(all_of(ftb_colnames))
}
         
fake_enteral <- pmap_dfr(entero_clean, function(mrn, fdrt){
  create_fake_food_cts_for_enteral_patient(mrn, fdrt)
} )

```
 
## create *NA* counts for the missing days 

No need to create cuz we are gonna ignore it.


## the combined table(include fake data)

```{r}
filled <- bind_rows(dtb, fake_enteral) 
```

```{r}
# output the above table
filled %>% 
  write_csv('../data/cleaned_diet_data/FINAL_97_with_code_all_foods_drt_with_enteralTPN.csv')
```


## combine the EN data (that has a dedicated food code for them)

```{r}
# the ptb that has the transplant date
ptb <- read_csv('../data/cleaned_patients/diet_patients_97.csv')
dtb <- read_csv('../data/cleaned_diet_data/FINAL_97_with_code_all_foods_drt_with_enteralTPN.csv')

en <- read_csv('../data/cleaned_diet_data/NutritionSupport_DietPatients.csv') %>% 
  filter(!is.na(Food.Code)) %>% 
  mutate(meal = 'EN',
         Por_eaten = 1) %>% 
  select(mrn = MRN, 
         Date = `Date Nutrition Support`,
         Meal = meal,
         Food_NSC = `EN formula`,
         Por_eaten,
         Calories_kcal = `EN calories`,
         Proteing_g = `EN PRO`,
         Fat_g = `EN fat`,
         Carbohydrates_g = `EN CHO`,
         Fibers_g = `EN Fiber`,
         Sugars_g = `EN Sugar`,
         Sodium_g = `EN Sodium`,
         Food_code = Food.Code,
         description = FNDDS.Main.Food.Description
         ) %>% 
  mutate(Sodium_g = Sodium_g/1000) %>% 
  mutate(dehydrated_weight = Proteing_g + Fat_g + Carbohydrates_g + Fibers_g + Sugars_g + Sodium_g) %>% 
  mutate(Date = mdy(Date)) %>% 
  left_join(ptb %>% select(mrn, hct), by = 'mrn') %>% 
  mutate(fdrt = Date - hct) %>% 
  select(colnames(dtb))


en %>% 
  filter(`Nutrition Support` == 'TPN') %>% 
  distinct(MRN)
```

```{r}
# add the above table to the all table and have the final all!!!!
FINAL <- bind_rows(dtb, en)

FINAL %>% 
  write_csv('../data/cleaned_diet_data/FINAL_97_with_code_all_foods_drt_with_enteralTPN.csv')


# and update the no fake diet counts table 
real <- read_csv('../data/cleaned_diet_data/FINAL_97_with_code_all_foods_drt.csv')
 
no_fake <- bind_rows(
  en %>% 
    mutate(fdrt = as.numeric(fdrt)),
  real %>% 
    select(colnames(en))
)

no_fake %>% 
  write_csv('../data/cleaned_diet_data/FINAL_97_with_code_all_foods_drt_W_EN.csv')
```


```{r}
# update the food code and desc database

current <- read_csv('../data/cleaned_diet_data/all_food_nsc_with_code_FINAL.csv') 

en_code <- en %>% 
  distinct(Food_NSC, Food_code, description) %>% 
  rename(Food.Code = Food_code,
         FNDDS.Main.Food.Description = description)

final_code <- bind_rows(current, en_code)
final_code %>% 
  write_csv('../data/cleaned_diet_data/all_food_nsc_with_code_FINAL.csv')
```

  