---
title: "UMAP overlay"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
source('~/db_connect_simple.R')
connect_database(config_file = '~/dbConfig.txt')

get_table_from_database_predefined_filter('asv_alpha_diversity_ag')
get_table_from_database('asv_annotation_blast_ag')
```

# load the data

```{r}
# the umap coords
dcoords <- read_csv('../softwares/phylo-umap/taxumap/results/embedding_for_angel.csv')

link <- read_csv('../data/cleaned_diet_data/deidentify_dsample_map.csv')

coord <- dcoords %>% 
  inner_join(link) %>% 
  select(-fid, -pt) %>% 
  mutate(p1sdrt = fdrt + 1,
         p2sdrt = fdrt + 2,
         p3sdrt = fdrt + 3,
         p4sdrt = fdrt + 4,
         p5sdrt = fdrt + 5) 

coord_f <- coord %>% 
  select(mrn, p1sdrt, p2sdrt, p3sdrt, p4sdrt, p5sdrt)
```


# overlay the mcrobiome data

## find the pairing stool data (closest after the diet sample)

### **Using all the stool samples (including the empirical affected ones)** 

```{r}
stb <- read_csv('../data/cleaned_stool/selected_stool_samples_type_abx.csv')  
```

```{r}
closest_s <- apply(coord_f, 1, function(Row){
  df = stb %>% 
    filter(mrn == Row[['mrn']]) %>% 
    filter(sdrt %in% c(Row[['p1sdrt']], Row[['p2sdrt']], Row[['p3sdrt']], Row[['p4sdrt']], Row[['p5sdrt']]))
  if(nrow(df) == 0) {
    ret = 'NA'
  }else {
    ret = df %>% 
       arrange(sdrt) %>% 
      slice(1) %>% 
      pull(sampleid)
  }
  return(ret)
})  %>% 
  set_names(coord %>% pull(index_column)) %>% 
  bind_rows() %>% 
  gather('index_column', 'sampleid') %>% 
  full_join(coord)

# some diet data just don't have stool in the following five days
# so remove them for now
closest_all <-  closest_s %>% 
  filter(sampleid != "NA")
```

### **porphylactic only stool samples**

```{r}
closest_s_porphylactic_only <- apply(coord_f, 1, function(Row){
  df = stb %>% 
    filter(abx == 'prophylactic') %>% 
    filter(mrn == Row[['mrn']]) %>% 
    filter(sdrt %in% c(Row[['p1sdrt']], Row[['p2sdrt']], Row[['p3sdrt']], Row[['p4sdrt']], Row[['p5sdrt']]))
  if(nrow(df) == 0) {
    ret = 'NA'
  }else {
    ret = df %>% 
       arrange(sdrt) %>% 
      slice(1) %>% 
      pull(sampleid)
  }
  return(ret)
})  %>% 
  set_names(coord %>% pull(index_column)) %>% 
  bind_rows() %>% 
  gather('index_column', 'sampleid') %>% 
  full_join(coord) %>% 
  filter(sampleid != "NA") %>% 
  inner_join(asv_alpha_diversity_ag %>% select(sampleid, simpson_reciprocal))
```


## get the stool alpha diversity from db

```{r}

closest_all_stool <- closest_all %>% 
  inner_join(asv_alpha_diversity_ag %>% select(sampleid, simpson_reciprocal))
```
 
## overlay the stool alpha diversity

```{r}
library(viridis)
closest_all_stool %>% 
  ggplot(aes(x = `taxumap-LLLL-1`, y = `taxumap-LLLL-2`, color = simpson_reciprocal)) +
  geom_point(alpha = 0.7, size = 3) +
  scale_color_viridis() 

```

```{r}
closest_s_porphylactic_only_plot <- closest_s_porphylactic_only %>% 
  mutate(sqrtSR = sqrt(simpson_reciprocal)) 

brks <- seq(0, 1, 0.1)

quants_sqrtSR <- brks %>% 
  set_names(brks) %>% 
  map_dfr(~ quantile(closest_s_porphylactic_only_plot$sqrtSR, .)) %>% 
  gather('brks', 'value')
 
closest_s_porphylactic_only_plot_df <- closest_s_porphylactic_only_plot %>% 
  arrange(sqrtSR) %>% 
  mutate(brks = cut(closest_s_porphylactic_only_plot$sqrtSR, 
                    breaks = format(quants_sqrtSR %>% pull(value), scientific=F),
                    include.lowest = T)) %>% 
  filter(!is.na(brks))


closest_s_porphylactic_only_plot_df %>% 
  ggplot(aes(x = `taxumap-LLLL-1`, y = `taxumap-LLLL-2`, color = brks)) +
  geom_point(alpha = 0.7, size = 3) +
  scale_colour_manual(values = rev(brewer.pal(11,"RdYlBu"))) +
  labs(title = 'food umap but colored by closest stool samples sqrt simposon reciprocal')


```




## make this version of the food maps with the dominant food groups and the second dominant food groups 

Only show the diet samples paired with a stool sample

```{r}
fcts_fg <- read_csv('../data/cleaned_diet_data/summarize_food_groups_pt_daily.csv') 
link <- read_csv('../data/cleaned_diet_data/deidentify_dsample_map.csv')

fcts_fg_dom <- fcts_fg %>%
  group_by(mrn, fdrt) %>% 
  arrange(-grp_frac, .by_group = T) %>% 
  slice(1) %>% 
  ungroup() %>% 
  transmute(fid = str_glue('P{mrn}d{fdrt}'),
            fg1_dom = fg1_name) %>% 
  inner_join(link %>% select(fid, index_column))


closest_s_porphylactic_only_fg <- closest_s_porphylactic_only %>% 
  inner_join(fcts_fg_dom)

closest_s_porphylactic_only_fg %>% 
  ggplot(aes(x = `taxumap-LLLL-1`, y = `taxumap-LLLL-2`, color = fg1_dom))+
  geom_point(alpha = 0.6, size = 3) +
  scale_color_brewer(palette = "Set1")
```


## overlay the dominant taxa (phyla)

```{r}
# summarize at phylum level for each sample 
asv_counts_ag <- read_csv('../data/asv_counts_ag_Feb17.csv')

stb_porphy <- stb %>% 
  filter(abx == 'prophylactic')

cts <- asv_counts_ag %>% 
  filter(sampleid %in% stb_porphy$sampleid) %>% 
  select(asv_key, sampleid, count) %>% 
  spread(key = 'sampleid', value = 'count', fill = 0) %>% 
  arrange(asv_key)  


annot <- asv_annotation_blast_ag %>% 
  filter(asv_key %in% cts$asv_key) %>% 
  mutate(ordr =  if_else(ordr == '', str_glue('of_class_{class}'), ordr),
         family =  if_else(family == '', str_glue('of_order_{ordr}'), family),
         genus =  if_else(genus == '', str_glue('of_family_{family}'), genus),
         species =  if_else(species == '', str_glue('of_genus_{genus}'), species))

# replace the asv_key with the genus taxa level
# summarize from asv level to genus level
# get the relative abundance for the genera
cts_phy <- cts %>% 
  full_join(annot %>%  select(asv_key, phylum), by  = 'asv_key') %>% 
  select(-asv_key) %>% 
  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
  group_by(sampleid, phylum) %>% 
  summarise(cnt = sum(count)) %>% 
  # get the total count from the db to calculate the relab
  left_join(asv_counts_ag %>% distinct(sampleid,count_total ), by = 'sampleid') %>% 
  mutate(relab = cnt/count_total) 

cts_phy %>% 
  write_csv('../data/cleaned_stool/prophylactic_only_stool_samples_phylum_counts.csv')
```

```{r}
# summarize at order level
cts_order <- cts %>% 
  full_join(annot %>%  select(asv_key, phylum), by  = 'asv_key') %>% 
  select(-asv_key) %>% 
  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
  group_by(sampleid, phylum) %>% 
  summarise(cnt = sum(count)) %>% 
  # get the total count from the db to calculate the relab
  left_join(asv_counts_ag %>% distinct(sampleid,count_total ), by = 'sampleid') %>% 
  mutate(relab = cnt/count_total) 
```


```{r}
# find the phyla with biggest abundance in each sample 
cts_phy_dominant <- cts_phy %>% 
  group_by(sampleid) %>%  
  arrange(-relab, .by_group = T) %>% 
  slice(1)

cts_phy_dominant %>% 
  ungroup() %>% 
  distinct(phylum)

# the sampleid and the corresponding demoniant phylum
closest_s_porphylactic_domi <- closest_s_porphylactic_only %>% 
  inner_join(cts_phy_dominant %>% select(sampleid, phylum))

```

```{r}
library(RColorBrewer)
closest_s_porphylactic_domi %>% 
  ggplot(aes(x = `taxumap-LLLL-1`, y = `taxumap-LLLL-2`, color = phylum)) +
  geom_point(alpha = 0.6, size = 3) +
  scale_color_brewer(palette = "Dark2")
  
```

