---
title: "Procrustes including all prophylactic stool samples"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
library(vegan)
```

# get all the prophylactic stool samples that have 5 days previous diet data

## load the data and the database access

```{r}
dtb <- read_csv('../data/cleaned_diet_data/FINAL_97_with_code_all_foods_drt_with_enteralTPN.csv')

## !!!! here we look at the stool samples affected by prophylactic only
stb <- read_csv('../data/cleaned_stool/selected_stool_samples_type_abx.csv') %>% 
  filter(abx == 'prophylactic')
```

## sort fdrt for each patient

```{r}
pt_fdrt <- dtb %>% 
  distinct(mrn, fdrt)

stb_loop <- stb %>%  
  select(mrn, sdrt) %>% 
  transmute(mrn = mrn,
            p1d = sdrt-1,
            p2d = sdrt-2, 
            p3d = sdrt-3,
            p4d = sdrt-4,
            p5d = sdrt-5) 
 
check_p5d_diet <-  function(mrn_, p1d_, p2d_, p3d_, p4d_, p5d_){
  num = pt_fdrt %>% 
    filter(mrn == mrn_) %>% 
    filter(fdrt %in% c(p1d_, p2d_ ,p3d_, p4d_, p5d_ )) %>% 
    nrow()
  return(num)
}

check_p5d_diet_df <- pmap(stb_loop, function(mrn, p1d, p2d, p3d, p4d, p5d){
    check_p5d_diet(mrn, p1d, p2d, p3d, p4d, p5d)
  }) %>% 
  set_names(stb %>% pull(sampleid)) %>% 
  bind_rows() %>% 
  gather('sampleid', 'pNd_num') 

p5d_samples <- check_p5d_diet_df %>% 
  filter(pNd_num == 5) %>% 
  pull(sampleid)  
 
stb_loop_sum <- stb %>%  
  filter(sampleid %in% p5d_samples) %>% 
  mutate(p1d = sdrt-1,
            p2d = sdrt-2, 
            p3d = sdrt-3,
            p4d = sdrt-4,
            p5d = sdrt-5) 
```

# sum of previous 1,2,3,4,5 days of diet --- pairing with each stool sample

```{r}
p1d <- apply(stb_loop_sum, 1, function(Row){
  df = dtb %>%   
        filter(mrn == Row[['mrn']]) %>% 
        filter(fdrt %in% c(Row[['p1d']])) %>% 
        group_by(Food_code) %>% 
        summarise(sum_dewt = sum(dehydrated_weight))
}) %>% 
  set_names(stb_loop_sum %>% pull(sampleid)) %>% 
  bind_rows(.id = 'sampleid') %>% 
  spread(key = 'sampleid', value = 'sum_dewt', fill = 0)

p1d <- p1d[, colSums(p1d != 0) > 0]
 
p2d <- apply(stb_loop_sum, 1, function(Row){
  df = dtb %>%  
        filter(mrn == Row[['mrn']]) %>% 
        filter(fdrt %in% c(Row[['p1d']], Row[['p2d']])) %>% 
        group_by(Food_code) %>% 
        summarise(sum_dewt = sum(dehydrated_weight))
}) %>% 
  set_names(stb_loop_sum %>% pull(sampleid)) %>% 
  bind_rows(.id = 'sampleid') %>% 
  spread(key = 'sampleid', value = 'sum_dewt', fill = 0)

p2d <- p2d[, colSums(p2d != 0) > 0]

p3d <- apply(stb_loop_sum, 1, function(Row){
  df = dtb %>%  
        filter(mrn == Row[['mrn']]) %>% 
        filter(fdrt %in% c(Row[['p1d']], Row[['p2d']], Row[['p3d']])) %>% 
        group_by(Food_code) %>% 
        summarise(sum_dewt = sum(dehydrated_weight))
}) %>% 
  set_names(stb_loop_sum %>% pull(sampleid)) %>% 
  bind_rows(.id = 'sampleid') %>% 
  spread(key = 'sampleid', value = 'sum_dewt', fill = 0)

p3d <- p3d[, colSums(p3d != 0) > 0]


p4d <- apply(stb_loop_sum, 1, function(Row){
  df = dtb %>%  
        filter(mrn == Row[['mrn']]) %>% 
        filter(fdrt %in% c(Row[['p1d']], Row[['p2d']], Row[['p3d']], Row[['p4d']])) %>% 
        group_by(Food_code) %>% 
        summarise(sum_dewt = sum(dehydrated_weight))
}) %>% 
  set_names(stb_loop_sum %>% pull(sampleid)) %>% 
  bind_rows(.id = 'sampleid') %>% 
  spread(key = 'sampleid', value = 'sum_dewt', fill = 0) 

p4d <- p4d[, colSums(p4d != 0) > 0]

p5d <- apply(stb_loop_sum, 1, function(Row){
  df = dtb %>%  
        filter(mrn == Row[['mrn']]) %>% 
        filter(fdrt %in% c(Row[['p1d']], Row[['p2d']], Row[['p3d']], Row[['p4d']], Row[['p5d']])) %>% 
        group_by(Food_code) %>% 
        summarise(sum_dewt = sum(dehydrated_weight))
}) %>% 
  set_names(stb_loop_sum %>% pull(sampleid)) %>% 
  bind_rows(.id = 'sampleid') %>% 
  spread(key = 'sampleid', value = 'sum_dewt', fill = 0)  

p5d <- p5d[, colSums(p5d != 0) > 0]
```


```{r}
# the species counts data would be :
cts_long <- read_csv('../data/cleaned_stool/prophylactic_only_stool_samples_species_counts.csv') %>% 
  filter(sampleid %in% p5d_samples) %>% 
  select(sampleid, taxa_species, relab) 
 
# since the samples in the diet counts are different so need to subset according to each
p1d_scts <- cts_long %>% 
  filter(sampleid %in% colnames(p1d)[2:ncol(p1d)]) %>% 
  spread('sampleid', 'relab', fill = 0)

p2d_scts <- cts_long %>% 
  filter(sampleid %in% colnames(p2d)[2:ncol(p2d)]) %>% 
  spread('sampleid', 'relab', fill = 0)

p3d_scts <- cts_long %>% 
  filter(sampleid %in% colnames(p3d)[2:ncol(p3d)]) %>% 
  spread('sampleid', 'relab', fill = 0)

p4d_scts <- cts_long %>% 
  filter(sampleid %in% colnames(p4d)[2:ncol(p4d)]) %>% 
  spread('sampleid', 'relab', fill = 0)

p5d_scts <- cts_long %>% 
  filter(sampleid %in% colnames(p5d)[2:ncol(p5d)]) %>% 
  spread('sampleid', 'relab', fill = 0)
```

# output for procrustes in qiime

```{r}
# output the food id cts table (need to be in a similar format to fit my previously written snakemake rules)
dcts <- list(p1d = p1d,
             p2d = p2d,
             p3d = p3d,
             p4d = p4d,
             p5d = p5d)

for (i in 1:length(dcts)){
  dcts[[i]] %>% 
    write_tsv(str_glue('../data/finalized/paired/pNday/prophylactic_{names(dcts)[i]}_diet_foodID_dehydrated_weight_per_pt.tsv'))
}


scts <- list(
  p1d_scts = p1d_scts,
  p2d_scts = p2d_scts,
  p3d_scts = p3d_scts,
  p4d_scts = p4d_scts,
  p5d_scts = p5d_scts
)


for (i in 1:length(scts)){
  scts[[i]] %>% 
    write_tsv(str_glue('../data/finalized/paired/pNday/prophylactic_{names(scts)[i]}_stool_relab_species.tsv'))
}
```

# import qiime results and do protest
 
```{r}
fns_both <- tibble(
  sfolder = list.files('../data/finalized/paired/pNday/', pattern = 'prophylactic_p.d_scts_stool_relab_species_braycurtis_pcoa$', full.names = T),
  dfolder = list.files('../data/finalized/paired/pNday/', pattern = 'prophylactic_p.d_diet_foodID_dehydrated_weight_per_pt_unweighted_unifrac_pcoa$', full.names = T)
) %>% 
  mutate(ID = str_replace(sfolder, '_scts_stool_relab_species_braycurtis_pcoa', ''),
         ID = str_replace(ID, '../data/finalized/paired/pNday//prophylactic_','')) %>% 
  mutate(sfn = str_glue('{sfolder}/ordination.txt'),
         dfn = str_glue('{dfolder}/ordination.txt'))


# a function to do the procrustes as a pair
procrustes_Nday <- map2_dfc(fns_both$dfn, fns_both$sfn, function(dfn, sfn){
  dpcoa = read_tsv(dfn, skip = 9, col_names = F)  %>% 
    filter(! X1 %in% c('Biplot','Site constraints')) %>% 
    column_to_rownames('X1') %>% 
    as.matrix()
  
  spcoa = read_tsv(sfn, skip = 9, col_names = F)  %>% 
    filter(! X1 %in% c('Biplot','Site constraints')) %>% 
    column_to_rownames('X1') %>% 
    as.matrix()
  
  protest_res = protest(X = dpcoa, Y = spcoa,  permutations = 999)
  return( protest_res$t0)
}) %>% 
  gather(key = 'pNd', value = 'corre') %>% 
  mutate(pNd = fns_both$ID) %>% 
  arrange(corre)
```

# An experiment 

the diet sample 10 days and 20 days after the stool sample collection

## the stool samples that have d sample 10 days and 20 days

```{r}

stb_loop <- stb %>%  
  select(mrn, sdrt) %>% 
  transmute(mrn = mrn,
            a10d = sdrt+10,
            a20d = sdrt+20
            ) 
 
check_if_diet <-  function(mrn_, a10d_, a20d_){
  num = pt_fdrt %>% 
    filter(mrn == mrn_) %>% 
    filter(fdrt %in% c(a10d_, a20d_  )) %>% 
    nrow()
  return(num)
}

check_if_diet_df <- pmap(stb_loop, function(mrn, a10d, a20d){
    check_if_diet(mrn, a10d, a20d)
  }) %>% 
  set_names(stb %>% pull(sampleid)) %>% 
  bind_rows() %>% 
  gather('sampleid', 'pNd_num') 

a2d_samples <- check_if_diet_df %>% 
  filter(pNd_num == 2) %>% 
  pull(sampleid)  
 
stb_loop_sum <- stb %>%  
  filter(sampleid %in% a2d_samples) %>% 
  mutate(a10d = sdrt+10,
        a20d = sdrt+20) 
```

```{r}
a10d <- apply(stb_loop_sum, 1, function(Row){
  df = dtb %>%   
        filter(mrn == Row[['mrn']]) %>% 
        filter(fdrt %in% c(Row[['a10d']])) %>% 
        group_by(Food_code) %>% 
        summarise(sum_dewt = sum(dehydrated_weight))
}) %>% 
  set_names(stb_loop_sum %>% pull(sampleid)) %>% 
  bind_rows(.id = 'sampleid') %>% 
  spread(key = 'sampleid', value = 'sum_dewt', fill = 0)

a10d <- a10d[, colSums(a10d != 0) > 0]

a20d <- apply(stb_loop_sum, 1, function(Row){
  df = dtb %>%   
        filter(mrn == Row[['mrn']]) %>% 
        filter(fdrt %in% c(Row[['a20d']])) %>% 
        group_by(Food_code) %>% 
        summarise(sum_dewt = sum(dehydrated_weight))
}) %>% 
  set_names(stb_loop_sum %>% pull(sampleid)) %>% 
  bind_rows(.id = 'sampleid') %>% 
  spread(key = 'sampleid', value = 'sum_dewt', fill = 0)

a20d <- a20d[, colSums(a20d != 0) > 0]


a10d_scts <- cts_long %>% 
  filter(sampleid %in% colnames(a10d)[2:ncol(a10d)]) %>% 
  spread('sampleid', 'relab', fill = 0)

a20d_scts <- cts_long %>% 
  filter(sampleid %in% colnames(a20d)[2:ncol(a20d)]) %>% 
  spread('sampleid', 'relab', fill = 0)


a10d %>% 
  write_tsv('../data/finalized/paired/pNday/experiment_a10d_diet_foodID_dehydrated_weight_per_pt.tsv')

a20d %>% 
  write_tsv('../data/finalized/paired/pNday/experiment_a20d_diet_foodID_dehydrated_weight_per_pt.tsv')

a10d_scts %>% 
  write_tsv(str_glue('../data/finalized/paired/pNday/experiment_a10d_scts_stool_relab_species.tsv'))

a20d_scts %>% 
  write_tsv(str_glue('../data/finalized/paired/pNday/experiment_a20d_scts_stool_relab_species.tsv'))
```
 
```{r}
fns_both <- tibble(
  sfolder = list.files('../data/finalized/paired/pNday/', pattern = 'experiment_a..d_scts_stool_relab_species_braycurtis_pcoa$', full.names = T),
  dfolder = list.files('../data/finalized/paired/pNday/', pattern = 'experiment_a..d_diet_foodID_dehydrated_weight_per_pt_unweighted_unifrac_pcoa$', full.names = T)
) %>% 
  mutate(ID = str_replace(sfolder, '_scts_stool_relab_species_braycurtis_pcoa', ''),
         ID = str_replace(ID, '../data/finalized/paired/pNday//experiment_','')) %>% 
  mutate(sfn = str_glue('{sfolder}/ordination.txt'),
         dfn = str_glue('{dfolder}/ordination.txt'))
# a function to do the procrustes as a pair
procrustes_Nday <- map2_dfc(fns_both$dfn, fns_both$sfn, function(dfn, sfn){
  dpcoa = read_tsv(dfn, skip = 9, col_names = F)  %>% 
    filter(! X1 %in% c('Biplot','Site constraints')) %>% 
    column_to_rownames('X1') %>% 
    as.matrix()
  
  spcoa = read_tsv(sfn, skip = 9, col_names = F)  %>% 
    filter(! X1 %in% c('Biplot','Site constraints')) %>% 
    column_to_rownames('X1') %>% 
    as.matrix()
  
  protest_res = protest(X = dpcoa, Y = spcoa,  permutations = 999)
  return( list(pval  = protest_res$signif,
               corr = protest_res$t0))
}) %>% 
  gather(key = 'pNd', value = 'corre') %>% 
  mutate(pNd = fns_both$ID) %>% 
  arrange(corre)
```
 
 