---
title: "Linear mixed effects model stool alpha diversity"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

```{r}
library(lmerTest)
library(tidyverse) 
```

# the data table

```{r}
ntb <- read_csv('../data/cleaned_diet_data/paired_mean_p3d_diet_meta.csv')
stb <- read_csv('../data/cleaned_stool/selected_stool_samples_type_abx.csv')

# pulling the alpha diversity 
source('~/db_connect_simple.R')
connect_database(config_file = '~/dbConfig.txt')
get_table_from_database_predefined_filter('asv_alpha_diversity_ag')
 
asv_counts_ag <- read_csv('../data/asv_counts_ag_Feb17.csv')
get_table_from_database('asv_annotation_blast_ag')

dat <- stb %>% 
  inner_join(asv_alpha_diversity_ag %>% select(sampleid, simpson_reciprocal)) %>% 
  full_join(ntb) 
  # making factors factors

fac_col <- c('mrn','source', 'intensity', 'sex', 'abx')

dat[fac_col] <- lapply(dat[fac_col], factor)  

colnames(dat)

dat %>% 
  write_csv('../data/cleaned_stool/all_samples_meta.csv')
```

# the model testing using the food groups

```{r}
mod1 <- lmer(simpson_reciprocal ~ fg_egg +
              fg_fruit+
               fg_grain+
               fg_meat+
               fg_milk+
               fg_oils+
               fg_sweets+
               fg_veggie+ 
               abx + 
               source + 
               (1 | mrn), REML = FALSE, data = dat)
#summary(mod1) 
```

```{r}
mod2 <- lmer(simpson_reciprocal ~ fg_egg +
              fg_fruit+
               fg_grain+
               fg_meat+
               fg_milk+
               fg_oils+
               fg_sweets+
               fg_veggie+ 
               abx + 
               (1 | mrn), REML = FALSE, data = dat)
#summary(mod2)

anova(mod1, mod2)
```

**Need to include source **

```{r}
mod3 <- lmer(simpson_reciprocal ~ fg_egg +
              fg_fruit+
               fg_grain+
               fg_meat+
               fg_milk+
               fg_oils+
               fg_sweets+
               fg_veggie+ 
               abx + 
               source + 
               sex +
               (1 | mrn), REML = FALSE, data = dat)
#summary(mod3)

anova(mod1, mod3)

```
**No need to include sex**

```{r}
mod4 <- lmer(simpson_reciprocal ~ fg_egg +
              fg_fruit+
               fg_grain+
               fg_meat+
               fg_milk+
               fg_oils+
               fg_sweets+
               fg_veggie+ 
               abx + 
               source + 
               age +
               (1 | mrn), REML = FALSE, data = dat)
#summary(mod4)
 
anova(mod1, mod4)
```
**Need to include age**   

```{r}
mod5 <- lmer(simpson_reciprocal ~ fg_egg +
              fg_fruit+
               fg_grain+
               fg_meat+
               fg_milk+
               fg_oils+
               fg_sweets+
               fg_veggie+ 
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)
#summary(mod5)
 
anova(mod5, mod4)
```
**abx and mrn interaction term**

```{r}
mod6 <-  lmer(simpson_reciprocal ~ fg_egg +
              fg_fruit+
               fg_grain+
               fg_meat+
               fg_milk+
               #fg_oils+
               fg_sweets+
               fg_veggie+ 
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)
#summary(mod6)
 
anova(mod5, mod6)
```
**yes fg_oils**

```{r}
mod7 <-  lmer(simpson_reciprocal ~ fg_egg +
              fg_fruit+
               fg_grain+
               fg_meat+
               #fg_milk+
               fg_oils+
               fg_sweets+
               fg_veggie+ 
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)
#summary(mod7)
 
anova(mod5, mod7)
```

**yes fg_milk**

```{r}
mod8 <-  lmer(simpson_reciprocal ~ fg_egg +
              fg_fruit+
               fg_grain+
               #fg_meat+
               fg_milk+
               fg_oils+
               fg_sweets+
               fg_veggie+ 
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)
#summary(mod8)
 
anova(mod5, mod8)
```

**yes meat**

```{r}
mod9 <-  lmer(simpson_reciprocal ~ fg_egg +
              fg_fruit+
               #fg_grain+
               fg_meat+
               fg_milk+
               fg_oils+
               fg_sweets+
               fg_veggie+ 
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)
#summary(mod9)
 
anova(mod5, mod9)
```
**no grain**

```{r}
mod10 <-  lmer(simpson_reciprocal ~ fg_egg +
              #fg_fruit+
               #fg_grain+
               fg_meat+
               fg_milk+
               fg_oils+
               fg_sweets+
               fg_veggie+ 
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)
#summary(mod10)
 
anova(mod10, mod9)
```

**yes fruit**

```{r}
mod11 <-  lmer(simpson_reciprocal ~ fg_egg +
              fg_fruit+
               #fg_grain+
               fg_meat+
               fg_milk+
               fg_oils+
               #fg_sweets+
               fg_veggie+ 
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)
#summary(mod11)
 
anova(mod11, mod9)
```

**no fg_sweets**

```{r}
mod12 <-  lmer(simpson_reciprocal ~ fg_egg +
              fg_fruit+
               #fg_grain+
               fg_meat+
               fg_milk+
               fg_oils+
               #fg_sweets+
               #fg_veggie+ 
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)
#summary(mod12)
 
anova(mod11, mod12)
```

**no fg_veggie**

```{r}
mod13 <-  lmer(simpson_reciprocal ~ 
              fg_fruit+
               #fg_grain+
               fg_meat+
               fg_milk+
               fg_oils+
               #fg_sweets+
               #fg_veggie+ 
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)
#summary(mod13)
 
anova(mod13, mod12)
```

**no fg_egg**

```{r}
mod14 <-  lmer(simpson_reciprocal ~ 
              fg_fruit+
               #fg_grain+
               fg_meat+
               fg_milk+
               fg_oils+
               #fg_sweets+
               #fg_veggie+ 
                sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)
#summary(mod14)
 
anova(mod13, mod14)
```

**yes sdrt**

```{r}
mod15 <-  lmer(simpson_reciprocal ~ 
              fg_fruit+
               #fg_grain+
               fg_meat+
               fg_milk+
               fg_oils+
               #fg_sweets+
               #fg_veggie+ 
                sdrt+
               #abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)
#summary(mod15)
 
anova(mod15, mod14)
```

**yes abx!**

```{r}
mod16 <-  lmer(simpson_reciprocal ~ 
              fg_fruit+
               #fg_grain+
               fg_meat+
               fg_milk+
               fg_oils+
               #fg_sweets+
               #fg_veggie+ 
                sdrt+
               #abx + 
               source + 
                intensity +
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)
#summary(mod16)
 
anova(mod16, mod14)
```

**no intensity!**



# model including nutrients terms

```{r}
mod17 <-  lmer(simpson_reciprocal ~ 
              fg_fruit+
               fg_meat+
               fg_milk+
               fg_oils+
                Nut_Fibers +
               sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)
#summary(mod17)
 
anova(mod17, mod14)
```

**no Nut_Fibers**

```{r}
mod18 <-  lmer(simpson_reciprocal ~ 
              fg_fruit+
               fg_meat+
               fg_milk+
               fg_oils+
                #Nut_Fibers +
                Nut_Carbohydrates+
               sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)
#summary(mod18)
 
anova(mod18, mod14)
```

**no Nut_Carbohydrates**

```{r}
mod19 <-  lmer(simpson_reciprocal ~ 
              fg_fruit+
               fg_meat+
               fg_milk+
               fg_oils+
                #Nut_Fibers +
                #Nut_Carbohydrates+
                Nut_Fat +
               sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)
#summary(mod19)
 
anova(mod19, mod14)
```

**no Nut_Fat**

```{r}
mod20 <-  lmer(simpson_reciprocal ~ 
              fg_fruit+
               fg_meat+
               fg_milk+
               fg_oils+
                #Nut_Fibers +
                #Nut_Carbohydrates+
                #Nut_Fat +
                Nut_Proteing +
               sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)
#summary(mod20)
 
anova(mod20, mod14)
```


**no Nut_Proteing cuz it's not significant **

```{r}
mod21 <-  lmer(simpson_reciprocal ~ 
              fg_fruit+
               fg_meat+
               fg_milk+
               fg_oils+
                #Nut_Fibers +
                #Nut_Carbohydrates+
                #Nut_Fat +
                #Nut_Proteing +
                Nut_Sodium+
               sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)
#summary(mod21)
 
anova(mod21, mod14)
```

**no Nut_Sodium**

```{r}
mod22 <-  lmer(simpson_reciprocal ~ 
              fg_fruit+
               fg_meat+
               fg_milk+
               fg_oils+
                #Nut_Fibers +
                #Nut_Carbohydrates+
                #Nut_Fat +
                #Nut_Proteing +
                #Nut_Sodium+
                Nut_Sugars+
               sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)
#summary(mod22)
 
anova(mod22, mod14)
```

**no sugars**

no need to include any nutrients terms 

# the final model with stool alpha diversity as the response var

```{r}
summary(mod14)
```

- the older the person, the slightly lower the diversity
- prophylactically treated patients have higher diversity than empirical
- the later time points into transplant the slightly lower the diversity
- oil and meat help boost diversity 
- fruit decreases diversity 


This is in agreement with my spearman correlation plot except with veggie 

# using enterococcus relab as the response

```{r}
# getting entercoccus genus relab for all the samples 
cts <- asv_counts_ag %>% 
    filter(sampleid %in% dat$sampleid) %>% 
    select(asv_key, sampleid, count) %>% 
    spread(key = 'sampleid', value = 'count', fill = 0) %>% 
    arrange(asv_key)  
  
annot <- asv_annotation_blast_ag %>% 
    filter(asv_key %in% cts$asv_key) %>% 
    mutate(ordr =  if_else(ordr == '', str_glue('of_class_{class}'), ordr),
           family =  if_else(family == '', str_glue('of_order_{ordr}'), family),
           genus =  if_else(genus == '', str_glue('of_family_{family}'), genus),
           species =  if_else(species == '', str_glue('of_genus_{genus}'), species)) %>% 
    mutate(taxa_genus = str_glue('k__{kingdom}|p__{phylum}|c__{class}|o__{ordr}|f__{family}|g__{genus}'))
  
  # replace the asv_key with the genus taxa level
  # summarize from asv level to genus level
  # get the relative abundance for the species
cts_relab <- cts %>% 
    full_join(annot %>%  select(asv_key, taxa_genus), by  = 'asv_key') %>% 
    select(-asv_key) %>% 
    gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
    group_by(sampleid, taxa_genus) %>% 
    summarise(cnt = sum(count)) %>% 
    # get the total count from the db to calculate the relab
    left_join(asv_counts_ag %>% distinct(sampleid,count_total ), by = 'sampleid') %>% 
    mutate(relab = cnt/count_total)

cts_relab %>% 
  write_csv('../data/cleaned_stool/ALL_stool_samples_genus_counts.csv')
```

```{r}
entercoccus <- cts_relab %>% 
  filter(str_detect(taxa_genus, 'g__Enterococcus')) %>% 
  select(sampleid, relab)

dat_e <- dat %>% 
  full_join(entercoccus) %>% 
  rename(entercoccus_relab = relab)
```

```{r}
mod_1 <- lmer(entercoccus_relab ~ 
              fg_fruit+
               fg_meat+ 
               fg_milk+
               fg_oils+
                sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat_e)

summary(mod_1)
```

