---
title: "the other way around linear mixed effects"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:  
    toc: true
    toc_float: true 
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

```{r}
library(tidyverse)
library(lmerTest)
library(ggpubr)
```

```{r}
dat <- read_csv('../data/cleaned_stool/all_samples_meta_p2d.csv') %>% 
  mutate(timebin = cut_width(sdrt, 7, boundary=0, closed = 'left')) %>% 
  mutate(intensity = factor(intensity, levels = c('nonablative','reduced','ablative'))) %>% 
  mutate(abx = factor(abx, levels = c('non_empirical','empirical')))

fac_col <- c('mrn','source',  'sex', 'abx')

dat[fac_col] <- lapply(dat[fac_col], factor)  
colnames(dat)

summary(dat)

dat2 <- dat %>% 
  #mutate(abx = as.character(abx)) %>% 
  filter(abx == 'non_empirical')
dat %>% 
  distinct(mrn, .keep_all = T)  %>% 
  dplyr::count(intensity)
```

```{r}
# visualize how the alpha div change according to time
dat %>% 
  ggscatter(x = 'sdrt', y = 'simpson_reciprocal',alpha = 0.5, 
   add = "loess", conf.int = TRUE) 
```


# important food groups one by one

```{r}
mod23 <-  lmer(simpson_reciprocal ~ 
              #fg_fruit+
               #fg_meat+
               #fg_milk+
               #fg_oils+
                sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)

mod23_2 <-  lmer(simpson_reciprocal ~ 
              #fg_fruit+
               #fg_meat+
               #fg_milk+
               #fg_oils+
                sdrt^2 +
                #sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)
anova(mod23, mod23_2)
# It doesn't seem I need to use the time square term
```


```{r}
mod24 <-  lmer(simpson_reciprocal ~ 
              #fg_fruit+
               fg_meat+
               #fg_milk+
               #fg_oils+
                sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)

anova(mod23, mod24)

# yes meat!
```


```{r}
mod25 <-  lmer(simpson_reciprocal ~ 
              fg_fruit+
               #fg_meat+
               #fg_milk+
               #fg_oils+
                sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)

anova(mod23, mod25)
# no fruit 
```
```{r}
mod26 <-  lmer(simpson_reciprocal ~ 
              #fg_fruit+
               #fg_meat+
               fg_milk+
               #fg_oils+
                sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)

anova(mod23, mod26)
# yes milk!
```
```{r}
mod27 <-  lmer(simpson_reciprocal ~ 
              #fg_fruit+
               #fg_meat+
               #fg_milk+
               fg_oils+
                sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)

anova(mod23, mod27)
# yes oil!
```
```{r}
mod34 <-  lmer(simpson_reciprocal ~ 
              #fg_fruit+
               #fg_meat+
               #fg_milk+
               #fg_oils+
                fg_egg+
                sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)

anova(mod23, mod34)
# yes egg!
```

```{r}
mod35 <-  lmer(simpson_reciprocal ~ 
              #fg_fruit+
               #fg_meat+
               #fg_milk+
               #fg_oils+
                #fg_egg+
                fg_grain+
                sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)

anova(mod23, mod35)
# yes grain!
```
```{r}
mod36 <-  lmer(simpson_reciprocal ~ 
              #fg_fruit+
               #fg_meat+
               #fg_milk+
               #fg_oils+
                #fg_egg+
                #fg_grain+
                fg_sweets+
                sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)

anova(mod23, mod36)
#no sweets!
```
```{r}
mod37 <-  lmer(simpson_reciprocal ~ 
              #fg_fruit+
               #fg_meat+
               #fg_milk+
               #fg_oils+
                #fg_egg+
                #fg_grain+
                #fg_sweets+
                fg_legume+
                sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)

anova(mod23, mod37)
# no legume
```
```{r}
mod38 <-  lmer(simpson_reciprocal ~ 
              #fg_fruit+
               #fg_meat+
               #fg_milk+
               #fg_oils+
                #fg_egg+
                #fg_grain+
                #fg_sweets+
                #fg_legume+
                fg_veggie+
                sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)

anova(mod23, mod38)
# yes fg_veggie!
```

 
# if only looking at nutrients

```{r}
mod28 <-  lmer(simpson_reciprocal ~ 
                Nut_Carbohydrates +
                sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)

anova(mod23, mod28)

# yes Nut_Carbohydrates!
```

```{r}
mod29 <-  lmer(simpson_reciprocal ~ 
                Nut_Fat +
                sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)

anova(mod23, mod29)

# yes Nut_Fat! 
```

```{r}
mod30 <-  lmer(simpson_reciprocal ~ 
                Nut_Proteing +
                sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)

anova(mod23, mod30)

# yes protein!
```
```{r}
mod31 <-  lmer(simpson_reciprocal ~ 
                Nut_Fibers +
                sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)

anova(mod23, mod31)

# no fiber! 
```
```{r}
mod32 <-  lmer(simpson_reciprocal ~ 
                Nut_Sugars +
                sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)

anova(mod23, mod32)
# no Nut_Sugars
```
```{r}
mod33 <-  lmer(simpson_reciprocal ~ 
                Nut_Sodium +
                sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = FALSE, data = dat)

anova(mod23, mod33)
# yes Nut_Sodium
```

**sodium, fat, protein is important**

# use some functions to do backward model selection

```{r}
# compute correlation matrix for the fg and nut
fmetrics <- dat %>% 
  select(fg_egg:Nut_Sugars)

f_cor <- cor(fmetrics, method = "pearson")

library(caret)
f_cor_res <- findCorrelation(f_cor, names = T)
f_cor_res
```
```{r}
mod99 <-  lmer(simpson_reciprocal ~ 
              fg_fruit+
               fg_meat+
               fg_milk+
               fg_oils+
                fg_egg+
                fg_grain+
                fg_sweets+
                fg_legume+
                fg_veggie+
                Nut_Carbohydrates+
                Nut_Fat+
                Nut_Fibers+
                Nut_Proteing+
                Nut_Sodium+
                Nut_Sugars+
                sdrt+
               abx + 
               source + 
               age +
               (1 + abx | mrn), REML = F, data = dat)
step(mod99)

head(dat$timebin)
```

# finalizing the model: what I care about and wanna model?



```{r}
mod102 <-  lmer(simpson_reciprocal ~ 
              fg_fruit+
               fg_meat+
               fg_milk+
               fg_oils+
                fg_egg+ 
                fg_grain+
                fg_sweets+
                fg_legume+
                fg_veggie+
               # sdrt +
               abx + 
               intensity + 
               (1 | mrn) +
                (1 | timebin), REML = F, data = dat )
summary(mod102)
```

```{r}
mod103 <-  lmer(simpson_reciprocal ~ 
              fg_fruit+
               fg_meat+
               fg_milk+
               fg_oils+
                fg_egg+ 
                fg_grain+
                fg_sweets+
                fg_legume+
                fg_veggie+
               abx + 
               intensity + 
                abx:fg_fruit +
               (1 | mrn) +
                (1 | timebin), REML = F, data = dat )
summary(mod103)
```

```{r}
dat_ <- dat %>% 
  mutate(tot_food = rowSums(.[11:19]))

mod104 <-  lmer(simpson_reciprocal ~ 
             tot_food +
               abx + 
               intensity + 
               (1 | mrn) +
                (1 | timebin), REML = F, data = dat_ )
summary(mod104)
```


# why is fruit decreasing the diversity 

## pariwise correlation between the food group columns

```{r}
library(arrangements)

fgs <- dat %>% 
  select(starts_with('fg_'))

pw_cor <- cor(fgs, method = 'spearman')

library(pheatmap)

pheatmap(
  pw_cor,
  #color=bwrPalette, 
  #annotation_col = annot,
  show_rownames = TRUE,
  show_colnames = TRUE,
  #filename = '../data/spearman.pairwise.cor.noTrans.jpeg',
  height = 8,
  width = 8,
  angle_col = '45',
  display_numbers = T,
  cluster_rows =T,
  cluster_cols=T
)


# 
fgs_ <- fgs %>% 
  mutate(rowsum = rowSums(.))
pw_cor_ <- cor(fgs_, method = 'spearman')

pheatmap(
  pw_cor_,
  #color=bwrPalette, 
  #annotation_col = annot,
  show_rownames = TRUE,
  show_colnames = TRUE,
  #filename = '../data/spearman.pairwise.cor.noTrans.jpeg',
  height = 8,
  width = 8,
  angle_col = '45',
  display_numbers = T,
  cluster_rows =T,
  cluster_cols=T
)
```


## when do patients eat fruits?

```{r}
dat %>% 
  mutate(fg_fruit = scale(fg_fruit)) %>% 
  ggplot(aes(x = timebin, y = fg_fruit)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = 'jitter', alpha = 0.5, size = 2, col = 'forestgreen') +
  theme_bw()
```

```{r}
# calculate the average fruit consumption by each timebin
dat %>% 
  group_by(timebin) %>% 
  summarise(timebin_ave = mean(fg_fruit))
```

## Plot individual patients over time with fruit and diversity shown

```{r}
dat %>% 
  select(mrn, sdrt, simpson_reciprocal, fg_fruit) %>% 
  ggscatter(x = 'fg_fruit', y = 'simpson_reciprocal', 
            add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE, # Add confidence interval
   cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
   cor.coeff.args = list(method = "pearson", label.x = 3, label.sep = "\n")) +
  facet_wrap(~mrn)  +
  theme_bw()+
  theme(
  strip.text.x = element_blank()
) +
  ggsave('../figs/97_fruit_diversity_no_time.pdf', width = 20, height = 17)
```

```{r}
dat %>% 
  select(mrn, sdrt, simpson_reciprocal, fg_fruit) %>% 
  #mutate(fg_fruit = sqrt(fg_fruit)) %>% 
  gather('type', 'value', simpson_reciprocal:fg_fruit) %>% 
  ggscatter(x = 'sdrt', y = 'value', color = 'type', palette = 'nejm') +
  geom_line(aes(group = type, col = type)) +
  facet_wrap(~mrn)  +
  theme_bw()+
  theme(
  strip.text.x = element_blank()
) +
  ggsave('../figs/97_fruit_diversity.pdf', width = 20, height = 17)
```

