---
title: "the other way around linear mixed effects"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:  
    toc: true
    toc_float: true 
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

```{r}
library(tidyverse)
library(lmerTest)
library(ggpubr)
```


```{r}
dat <- read_csv('../data/cleaned_stool/all_samples_meta_p2d.csv') %>% 
  mutate(timebin = cut_width(sdrt, 7, boundary=0, closed = 'left')) %>% 
  mutate(intensity = factor(intensity, levels = c('nonablative','reduced','ablative'))) %>% 
  mutate(abx = factor(abx, levels = c('non_empirical','empirical')),
         mrn = factor(mrn))

fac_col <- c('mrn','source',  'sex', 'abx')

dat[fac_col] <- lapply(dat[fac_col], factor)  
colnames(dat)

summary(dat)

dat2 <- dat %>% 
  #mutate(abx = as.character(abx)) %>% 
  filter(abx == 'non_empirical')
dat %>% 
  distinct(mrn, .keep_all = T)  %>% 
  dplyr::count(intensity)
```

```{r}
# visualize how the alpha div change according to time
dat %>% 
  ggscatter(x = 'sdrt', y = 'simpson_reciprocal',alpha = 0.5, 
   add = "loess", conf.int = TRUE) 
```



# finalizing the model: what I care about and wanna model?

```{r}
mod102 <-  lmer(log(simpson_reciprocal) ~ 
              fg_fruit+
               fg_meat+
               fg_milk+
               fg_oils+
                fg_egg+ 
                fg_grain+
                fg_sweets+
                fg_legume+
                fg_veggie+
               abx + 
               intensity + 
               (1 | mrn) +
                (1 | timebin), REML = F, data = dat )
summary(mod102)
```

# why is fruit decreasing the diversity 

## pariwise correlation between the food group columns

```{r}
library(arrangements)

fgs <- dat %>% 
  select(starts_with('fg_'))

pw_cor <- cor(fgs, method = 'spearman')

library(pheatmap)

pheatmap(
  pw_cor,
  #color=bwrPalette, 
  #annotation_col = annot,
  show_rownames = TRUE,
  show_colnames = TRUE,
  #filename = '../data/spearman.pairwise.cor.noTrans.jpeg',
  height = 8,
  width = 8,
  angle_col = '45',
  display_numbers = T,
  cluster_rows =T,
  cluster_cols=T
)


# 
fgs_ <- fgs %>% 
  mutate(rowsum = rowSums(.))
pw_cor_ <- cor(fgs_, method = 'spearman')

pheatmap(
  pw_cor_,
  #color=bwrPalette, 
  #annotation_col = annot,
  show_rownames = TRUE,
  show_colnames = TRUE,
  #filename = '../data/spearman.pairwise.cor.noTrans.jpeg',
  height = 8,
  width = 8,
  angle_col = '45',
  display_numbers = T,
  cluster_rows =T,
  cluster_cols=T
)
```


## when do patients eat fruits?

```{r}
dat %>% 
  mutate(fg_fruit = scale(fg_fruit)) %>% 
  ggplot(aes(x = timebin, y = fg_fruit)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = 'jitter', alpha = 0.5, size = 2, col = 'forestgreen') +
  theme_bw()
```

```{r}
# calculate the average fruit consumption by each timebin
dat %>% 
  group_by(timebin) %>% 
  summarise(timebin_ave = mean(fg_fruit))
```

## Plot individual patients over time with fruit and diversity shown



```{r}
dat %>% 
  select(mrn, sdrt, simpson_reciprocal, fg_fruit) %>% 
  ggplot() +
  geom_line(mapping = aes(x = sdrt, y = fg_fruit), color = '#1B9E77') +
  geom_point(mapping = aes(x = sdrt, y = fg_fruit), color = '#1B9E77') +
  geom_line(mapping = aes(x = sdrt, y = simpson_reciprocal), color = '#A6761D') +
  geom_point(mapping = aes(x = sdrt, y = simpson_reciprocal), color = '#A6761D') +
  #scale_color_manual(values = c('#1B9E77','#A6761D')) +
  facet_wrap(~mrn, scales = 'free_y')  +
  #scale_y_continuous("Fruit consumption (g)", sec.axis = sec_axis(~ . / 4, name = "simpson_reciprocal")) +
  theme_bw()+
  theme(
  strip.text.x = element_blank(),
  legend.position = 'bottom'
) +
  ggsave('../figs/97_fruit_diversity.pdf', width = 20, height = 17)

range(dat$simpson_reciprocal)
range(dat$fg_fruit)
 
```









 
# use sugar as the predictor directly 

```{r}
mod_sugar <-  lmer(simpson_reciprocal ~ 
              Nut_Sugars +
               abx + 
               intensity + 
               (1 | mrn) +
                (1 | timebin), REML = F, data = dat )
summary(mod_sugar)
```

Interesting, the same direction but not significant

```{r}
mod_nut_all <-  lmer(simpson_reciprocal ~ 
              Nut_Sugars + 
                Nut_Carbohydrates +
                Nut_Fat +
                Nut_Fibers +
                Nut_Proteing +
                Nut_Sodium +
               abx + 
               intensity + 
               (1 | mrn) +
                (1 | timebin), REML = F, data = dat )
summary(mod_nut_all)
```

# Does total calorie intake still hold up as a predictor of alpha diversity?

```{r}
# get the caloric intake 
nutable <- ''

source('~/db_connect_simple.R')
connect_database(config_file = '~/dbConfig.txt')

dbListTables(con) %>% data.frame() %>% filter(grepl("nutrition",.))

get_table_from_database('nutrition_ag')
get_table_from_database('nutrition_demographics_ag')
```

