---
title: "My whole new take on vegan"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Reference: http://ubio.bioinfo.cnio.es/Cursos/CEU_MDA07_practicals/Further%20reading/Oksanen%202005/R-%20Vegan%20tutorial%20-%20Multivariate%20analysis%20of%20ecological%20communities%20by%20Oksanen.pdf

```{r}
library(vegan)
library(tidyverse)
```

```{r}
data(varespec)
```

## the genus level count data (raw count) and filter

```{r}
meta <- read_csv('../data/cleaned_stool/all_samples_meta_p2d.csv') %>% 
  mutate(intensity = factor(intensity, levels = c('nonablative','reduced','ablative'))) %>% 
  mutate(abx = factor(abx, levels = c('non_empirical','empirical')),
         mrn = factor(mrn)) %>% 
  arrange(sampleid)
```


```{r}
gcts <- read_csv('../data/cleaned_stool/ALL_stool_samples_genus_counts.csv') %>% 
  select(sampleid, taxa_genus, cnt) %>% 
  filter(sampleid %in% meta$sampleid) %>% 
  arrange(sampleid)

gctsm <- gcts %>% 
  spread(key = 'taxa_genus', value = 'cnt') %>% 
  column_to_rownames('sampleid') %>% 
  as.matrix()

num_thre <- floor(nrow(gctsm) * 0.1)

gctsmf <- gctsm[ ,colSums(gctsm >= 5) >= num_thre ]
  
```

## the mds of the count table 

```{r}
set.seed(1)
gctsmf.mds <- metaMDS(gctsmf, trace = T,
                      distance = "bray", k = 4, maxit = 1000,try = 40, trymax = 100, parallel = 10 ) 
gctsmf.mds
 
stressplot(gctsmf.mds)

```

The fit between the community distances (data dissimilarity) and ordination distances is indicated by the monotonic step line (red).
Two stress values are provided:
Non-metric fit - this is the modified Kruskal's stress value and is sqrt(1-r2)
Linear fit - this is the R2 value of the correlation between the ordination values and the ordination values predicted from the (monotonic) regression line.

## if it's really linear

If the response really is linear and vectors are appropriate, the fitted surface is a plane whose gradient is parallel to the arrow, and the fitted contours are equally spaced parallel lines perpendicular to
the arrow. 

```{r}
ef <- envfit(gctsmf.mds ~ fg_grain, meta, 
                   strata = meta$mrn,
                   permutations = 999)
plot(gctsmf.mds)
plot(ef)
with(meta, ordisurf(gctsmf.mds, fg_grain, add = TRUE))
```

```{r}
ef <- envfit(gctsmf.mds ~ fg_egg, meta, 
                   strata = meta$mrn,
                   permutations = 999)
plot(gctsmf.mds) 
plot(ef)
with(meta, ordisurf(gctsmf.mds, fg_egg, add = TRUE))
```
```{r}
ef <- envfit(gctsmf.mds ~ fg_fruit, meta, 
                   strata = meta$mrn,
                   permutations = 999)
plot(gctsmf.mds) 
plot(ef)
with(meta, ordisurf(gctsmf.mds, fg_fruit, add = TRUE))
```

```{r}
ef <- envfit(gctsmf.mds ~ fg_legume, meta, 
                   strata = meta$mrn,
                   permutations = 999)
plot(gctsmf.mds) 
plot(ef)
with(meta, ordisurf(gctsmf.mds, fg_legume, add = TRUE))
```

```{r}
ef <- envfit(gctsmf.mds ~ fg_meat, meta, 
                   strata = meta$mrn,
                   permutations = 999)
plot(gctsmf.mds) 
plot(ef)
with(meta, ordisurf(gctsmf.mds, fg_meat, add = TRUE))
```
```{r}
ef <- envfit(gctsmf.mds ~ fg_milk, meta, 
                   strata = meta$mrn,
                   permutations = 999)
plot(gctsmf.mds)      
plot(ef)
with(meta, ordisurf(gctsmf.mds, fg_milk, add = TRUE))
```

```{r}
ef <- envfit(gctsmf.mds ~ fg_oils, meta, 
                   strata = meta$mrn,
                   permutations = 999)
plot(gctsmf.mds)      
plot(ef)
with(meta, ordisurf(gctsmf.mds, fg_oils, add = TRUE))
```
```{r}
ef <- envfit(gctsmf.mds ~ fg_sweets, meta, 
                   strata = meta$mrn,
                   permutations = 999)
plot(gctsmf.mds)      
plot(ef)
with(meta, ordisurf(gctsmf.mds, fg_sweets, add = TRUE))
```

```{r}
ef <- envfit(gctsmf.mds ~ fg_veggie, meta, 
                   strata = meta$mrn,
                   permutations = 999)
plot(gctsmf.mds)      
plot(ef)
tmp <- with(meta, ordisurf(gctsmf.mds, fg_veggie, add = TRUE))
```


```{r}
mgcv::vis.gam(tmp)
```
```{r}
# put all of the linear ones together
ef <- envfit(gctsmf.mds ~ fg_veggie + fg_grain + fg_egg + fg_legume + fg_milk, meta, 
                   strata = meta$mrn,
                   permutations = 999)
plot(gctsmf.mds)      
plot(ef)
with(meta, ordisurf(gctsmf.mds, fg_veggie, add = TRUE, col = 'green'))
with(meta, ordisurf(gctsmf.mds, fg_grain, add = TRUE, col = 'brown'))
with(meta, ordisurf(gctsmf.mds, fg_egg, add = TRUE, col = 'blue'))
with(meta, ordisurf(gctsmf.mds, fg_legume, add = TRUE, col = 'orange'))
with(meta, ordisurf(gctsmf.mds, fg_milk, add = TRUE, col = 'gray'))
# legume doesn't look strictly perpendicular so remove 
```



```{r}
ef <- envfit(gctsmf.mds ~ sdrt, meta, 
                   strata = meta$mrn,
                   permutations = 999)
plot(gctsmf.mds)       
plot(ef)
with(meta, ordisurf(gctsmf.mds, sdrt, add = TRUE))
```

## what about the factors

```{r}
ef <- envfit(gctsmf.mds ~ abx + intensity, meta, 
                   strata = meta$mrn,
                   permutations = 999)
ef
```

## no need to use constrained ordination

In principle, constrained ordination should be used with designed a priori
constraints only.(so in our case the diet project, we shouldn't use the constrained version then.)

## the full model with both vectors and factors

```{r}
full <- envfit(gctsmf.mds ~ fg_veggie + fg_grain + fg_egg + fg_milk +
                 abx + intensity , meta, 
                   strata = meta$mrn,
                   permutations = 10000)
full
```

```{r}
# FDR correction
full_res <- enframe(full$vectors$r, value = 'r2') %>% 
  full_join(enframe(full$vectors$pvals, value = 'pval'))
full_res 
```


## analysis of similarities 