---
title: "Tease apart nutritional drink"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

```{r}
library(tidyverse) 
```

```{r}
dtb <- read_csv('../data/cleaned_diet_data/FINAL_97_with_code_all_foods_drt_with_EN_and_TPN_Oct8.csv') %>% 
  mutate(Food_code = as.character(Food_code))
```

## regrouping fruit juice (citrus and non-citrus ) into sweet beverage group

```{r}
all_level_label <- read_tsv('../data/source/NodeLabelsMCT.txt', col_types = 'cc')

fruit2 <- read_tsv('../data/source/NodeLabelsMCT.txt', col_types = 'cc')  %>% 
  filter(str_detect(Level.code, '^6')) %>% 
  filter(str_length(Level.code) == 2)

sweets2 <- read_tsv('../data/source/NodeLabelsMCT.txt', col_types = 'cc')  %>% 
  filter(str_detect(Level.code, '^9')) %>% 
  filter(str_length(Level.code) == 2)

all_fruit <- dtb %>% 
  filter(str_detect(Food_code, '^6')) %>% 
  distinct(Food_NSC, Food_code) %>% 
  mutate(L2_grp = str_sub(Food_code, start = 1, end = 2))  %>% 
  left_join(fruit2 %>% rename(L2_grp = Level.code), by  = 'L2_grp') %>% 
  arrange(L2_grp) %>% 
  # to revise the food code for the fruit juice citrus and non-citrus
  mutate(is_juice = if_else(str_detect(Food_NSC, fixed('juice', ignore_case=TRUE)), T, F)) %>% 
  # change the food code for the citrus juice 
  mutate(Food_code = if_else(L2_grp == '61' & is_juice == 'TRUE', '92530610', Food_code)) %>% 
  # change the food code for the non-citrus juice
  mutate(Food_code = if_else(L2_grp == '64', '92510610', Food_code))  %>% 
  # also updating the description 
  mutate(L2_grp = str_sub(Food_code, start = 1, end = 2)) %>% 
  mutate(Main.food.description = if_else(Food_code == '92510610', 'Fruit juice drink',
                                                    if_else(Food_code == '92530610', 'Fruit juice drink, with high vitamin C', Main.food.description)))

juice_changed <- all_fruit %>% 
  filter(str_detect(Food_code, '^9')) %>% 
  select(-is_juice, -L2_grp) %>% 
  rename(description = Main.food.description)

# no need to change the food code for the raw orange, it is in its right form, the citrus fruit is separated from "other fruit". 

# so what really happened is I split the fruit juice (citrus and non-citrus ) into sweet beverage group
```


```{r}
# updating the above info in the dtb 
dtbj_split  <- dtb %>% 
  split(.$Food_NSC %in% juice_changed$Food_NSC)
  
# replace the food code and description for the things that changed 
dtbj_split_change <-  dtbj_split %>% 
  pluck('TRUE') %>% 
  select(-Food_code, -description) %>% 
  left_join(juice_changed, by  = 'Food_NSC') %>% 
  select(colnames(dtb))

dtbj <- bind_rows( 
  dtbj_split_change,
  dtbj_split %>% pluck('FALSE')
)
```

Find some 

## updating the food code to create a new food group

```{r}
# updating the food code of the nutritional drinks so that I can tease them apart
ndrink <- read_csv('../data/cleaned_diet_data/foods_sweets_group_95_tsoni_annotated.csv') %>% 
  filter(subcategory.tsoni == 'nutritional_supplement_often_urged_by_doctors_when_not_eating') %>% 
  mutate(Food_code = as.character(Food_code)) %>% 
  mutate(grp = if_else(str_detect(Food_code, '^951070'), 'ENfood', 'ndrink')) %>% 
  split(.$grp)

dtbn <- dtb %>% 
  mutate(Food_code = as.character(Food_code)) %>% 
  mutate(Food_code = if_else(Food_NSC %in% ndrink$Food_NSC, 
                             str_replace(Food_code, '^9','N'),
                             Food_code))

dtbn %>% 
  write_csv('../data/cleaned_diet_data/FINAL_97_with_code_all_foods_drt_with_EN_and_TPN_Oct8_UPDATED10.csv')

ndrink %>% 
  arrange(Food_NSC) %>%  
  write_csv('../data/ndrink.csv')
```
 
### split between EN fluid food and the nutritional drink

```{r}
en_food_nsc <- ndrink %>% pluck('ENfood') %>% pull(Food_NSC)
  
ndrink_food_nsc <- ndrink %>% pluck('ndrink') %>% pull(Food_NSC)

dtbnv <- dtb %>% 
  mutate(Food_code = as.character(Food_code)) %>% 
  mutate(Food_code = if_else(Food_NSC %in% ndrink_food_nsc, 
                             str_replace(Food_code, '^9','N'),
                             if_else(Food_NSC %in% en_food_nsc, 
                                     str_replace(Food_code, '^9','E'), Food_code)))
```
 

## summarise on the dtbn to per day level

```{r}
nodes <- read_tsv('../data/source/NodeLabels.txt', col_types = 'cc') %>% 
  filter(nchar(Level.code) == 1) %>% 
  rename(fgrp1 = Level.code, 
         fdesc = Main.food.description)

fgrps_df <- dtbnv %>% 
  select(mrn, fdrt, dehydrated_weight, Food_code) %>% 
  mutate(fgrp1 = str_sub(Food_code, 1, 1))


# the total dehydated weight for a patient for one day for a single food group
total_per_group <- fgrps_df %>% 
  group_by(mrn, fdrt, fgrp1) %>% 
  summarise(grp_tol = sum(dehydrated_weight)) %>% 
  mutate(fg1_name = case_when(
    fgrp1 == '1' ~ 'fg_milk',
    fgrp1 == '2' ~ 'fg_meat',
    fgrp1 == '3' ~ 'fg_egg',
    fgrp1 == '4' ~ 'fg_legume',
    fgrp1 == '5' ~ 'fg_grain',
    fgrp1 == '6' ~ 'fg_fruit',
    fgrp1 == '7' ~ 'fg_veggie',
    fgrp1 == '8' ~ 'fg_oils', 
    fgrp1 == '9' ~ 'fg_sweets',
    fgrp1 == 'N' ~ 'fg_ndrink',
    fgrp1 == 'E' ~ 'fg_EN'
  ))
```

## average on the previous two days diet food groups

```{r}
stb <- read_csv('../data/cleaned_stool/selected_stool_samples_type_abx.csv') 


stb_pair <- stb %>%  
  select(mrn, sdrt) %>% 
  transmute(mrn = mrn,
            p1d = sdrt-1,
            p2d = sdrt-2) 

mean_p2d_diet <-  function(mrn_, p1d_, p2d_){
  df = total_per_group %>% 
    filter(mrn == mrn_) %>% 
    filter(fdrt %in% c(p1d_, p2d_  )) %>% 
    group_by(fg1_name) %>% 
    summarise(ave_fg = sum(grp_tol)/2)
  return(df)
}

mean_p2d_df <- pmap(stb_pair, function(mrn, p1d, p2d){
    mean_p2d_diet(mrn, p1d, p2d)
  }) %>% 
  set_names(stb %>% pull(sampleid)) %>% 
  bind_rows(.id = 'sampleid') %>% 
  spread(key = 'fg1_name', value = 'ave_fg', fill = 0) %>% 
  inner_join(stb, by = "sampleid") %>% 
  select(-age, -sex, -source) 

mean_p2d_df %>% 
  write_csv('../data/cleaned_stool/all_samples_meta_p2d_fg11.csv')
```

## linear mixed model again

```{r}
library(lmerTest)     

dat <- mean_p2d_df %>% 
  mutate(timebin = cut_width(sdrt, 7, boundary=0, closed = 'left')) %>% 
  mutate(intensity = factor(intensity, levels = c('nonablative','reduced','ablative'))) %>% 
  mutate(mrn = as.factor(mrn))

dat %>% 
  select(-mrn) %>%  
  summary
```


```{r}
mod_fg11 <-  lmer(log(simpson_reciprocal) ~ 
              fg_fruit+
               fg_meat+
               fg_milk+
               fg_oils+
                fg_egg+ 
                fg_grain+
                fg_sweets+  
                fg_legume+
                fg_veggie+
                fg_ndrink+
                fg_EN+
               empirical +  
               intensity + 
               (1 | mrn) +
                (1 | timebin), REML = F, data = dat )
summary(mod_fg11) 
```
 
## brms model

```{r} 
library(brms)  
library(ggmcmc)
model_brm <- brm( log(simpson_reciprocal)~ 
                fg_fruit+
                fg_meat+
                fg_milk+
                fg_oils+
                fg_egg+ 
                fg_grain+
                fg_sweets+
                fg_legume+
                fg_veggie+
                fg_ndrink+
                fg_EN+
               empirical + 
               intensity + 
               (1 | mrn) +
                (1 | timebin),  
              data = dat, 
              warmup = 1000, iter = 3000, 
              cores = 8, chains = 2, 
              seed = 123) 
plot(model_brm)
summary(model_brm)
```

```{r}
# trying to make the centered at 0 posterior distribution plot 
posterior_model <- posterior_samples(model_brm) %>% 
  select(starts_with('b')) %>% 
  select(-b_Intercept) %>% 
  gather('item', 'value') %>% 
  group_by(item) %>% 
  mutate(meanperitem = mean(value))%>%
  ungroup()

library(ggthemes)
library(ggridges)
ggplot()+
  ggridges::geom_density_ridges(data  = posterior_model, 
                                aes(x      = value,
                                    y      = reorder(as.factor(item), meanperitem),
                                    height = ..density..),
                                scale = 0.8, size = 0.25,
                                #fill = 'forestgreen',
                                alpha = 0.6) +
  geom_vline(xintercept = 0, col = "red") +
  theme_ridges() +
  scale_x_continuous(limits = c(-1.38,0.25)) +
  ggsave('../figs/posteiror_brms.pdf', width = 10, height = 10)

# make a version that only has food groups
posterior_model_only_food_group <- posterior_model %>% 
  filter(str_detect(item, '_fg_'))

ggplot()+
  ggridges::geom_density_ridges(data  = posterior_model_only_food_group, 
                                aes(x      = value,
                                    y      = reorder(as.factor(item), meanperitem),
                                    height = ..density..),
                                scale = 1.5, size = 0.25,
                                alpha = 0.6) +
  geom_vline(xintercept = 0, col = "red") +
  #scale_x_continuous(limits = c(-0.02,0.03)) +
  labs(y = 'Food groups',
       title = 'Posterior distribution of food group level error of regression coefficient per food group') +
  theme(axis.text.y = element_text(size=20)) +
  theme_tufte() +
  ggsave('../figs/posteiror_brms_fg_only.pdf', width = 6, height = 6)
```


## is the nutritional drink signal simply because they are not eating so the stool diversity is low

```{r}
# do the patients have EN and regular eating at the same time?
 # no , when they had EN, they don't eat regular food
overlap <- en %>% 
  select(mrn, fdrt) %>% 
  inner_join(cur8)

# what's the correlation between daily total calories and daily consumption of the nutritional drink(not the EN ones)
# the patients and the days that they had nutritional drink
ndrink_pt <- dtbnv %>% 
  filter(str_detect(Food_code, '^N')) %>% 
  distinct(mrn, fdrt)

# the total daily calories of the patients on those days

all_daily_ndrink_pt <- dtbnv %>% 
  group_by(mrn, fdrt) %>% 
  summarise(total_daily = sum(Calories_kcal)) %>% 
  inner_join(ndrink_pt)


# per day sum up the amount of nutritional drink
ndrink_pt_sum <- dtbnv %>% 
  filter(str_detect(Food_code, '^N')) %>% 
  group_by(mrn, fdrt) %>% 
  summarise(ndrink_sum = sum(dehydrated_weight))

combined <- all_daily_ndrink_pt %>% 
  full_join(ndrink_pt_sum)

combined %>% 
  ggpubr::ggscatter(x = 'total_daily', y = 'ndrink_sum', 
                    xlab = 'Total daily calories',ylab = 'Total daily nutritional drink (dehydrated wt)',
            add = "reg.line",  # Add regressin line
           add.params = list(color = "blue", fill = "lightgray"), # Customize line
           conf.int = TRUE, # Add confidence interval
           cor.coef = TRUE, # Add correlation coefficient.
           cor.coeff.args = list(method = "pearson", label.x = 3, label.sep = "\n")) 

# in terms of the nutritional drink (excluding the EN) it's not highly correlated with total calories, cor = 0.49
# so the patients are still eating
# so the nutritional drink could be lowering the diversity
```

```{r}
fn <- read_tsv('../../ppp_Nilu/data/lefse_res/lefse_input_table.tsv', col_names = F)
```

