---
title: "Tease apart nutritional drink"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

```{r}
library(tidyverse)  
library(lubridate)
```

```{r}
dtb <- read_csv('../data/cleaned_diet_data/FINAL_97_with_code_all_foods_drt_with_EN_finalized.csv') %>% 
  mutate(Food_code = as.character(Food_code))
```

## updating the food code to split and create the new group : ndrink and EN

**The original 95 is split from the sweets group, and it divides into sub group EN and non-EN**

```{r}
# updating the food code of the nutritional drinks so that I can tease them apart
ndrink <- read_csv('../data/cleaned_diet_data/foods_sweets_group_95_tsoni_annotated.csv') %>% 
  mutate(Food_code = as.character(Food_code)) %>% 
  mutate(grp = if_else(str_detect(Food_code, '^951070'), 'ENfood', 'ndrink')) %>% 
  split(.$grp)

new_ndrink_food_nsc <- ndrink %>% pluck('ndrink') %>% pull(Food_NSC)

en_food_nsc <- ndrink %>% pluck('ENfood') %>% pull(Food_NSC)
 
dtbn <- dtb %>% 
  mutate(Food_code = if_else(Food_NSC %in% new_ndrink_food_nsc, 
                             str_replace(Food_code, '^9','N'),
                             if_else(Food_NSC %in% en_food_nsc, str_replace(Food_code, '^9','E'), Food_code)))

dtbn %>% 
  write_csv('../data/cleaned_diet_data/FINAL_97_with_code_all_foods_drt_with_EN_finalized_UPDATED11.csv')
```
 

## summarise on the dtbn to per day level

```{r}
nodes <- read_tsv('../data/source/NodeLabels.txt', col_types = 'cc') %>% 
  filter(nchar(Level.code) == 1) %>% 
  rename(fgrp1 = Level.code, 
         fdesc = Main.food.description)

fgrps_df <- dtbn %>% 
  select(mrn, fdrt, dehydrated_weight, Food_code) %>% 
  mutate(fgrp1 = str_sub(Food_code, 1, 1))


# the total dehydated weight for a patient for one day for a single food group
total_per_group <- fgrps_df %>% 
  group_by(mrn, fdrt, fgrp1) %>% 
  summarise(grp_tol = sum(dehydrated_weight)) %>% 
  mutate(fg1_name = case_when(
    fgrp1 == '1' ~ 'fg_milk',
    fgrp1 == '2' ~ 'fg_meat',
    fgrp1 == '3' ~ 'fg_egg',
    fgrp1 == '4' ~ 'fg_legume',
    fgrp1 == '5' ~ 'fg_grain',
    fgrp1 == '6' ~ 'fg_fruit',
    fgrp1 == '7' ~ 'fg_veggie',
    fgrp1 == '8' ~ 'fg_oils', 
    fgrp1 == '9' ~ 'fg_sweets',
    fgrp1 == 'N' ~ 'fg_ndrink',
    fgrp1 == 'E' ~ 'fg_EN'
  ))
```

## average on the previous two days diet food groups

```{r}
stb <- read_csv('../data/cleaned_stool/selected_stool_samples_type_abx.csv') 


stb_pair <- stb %>%  
  select(mrn, sdrt) %>% 
  transmute(mrn = mrn,
            p1d = sdrt-1,
            p2d = sdrt-2) 

mean_p2d_diet <-  function(mrn_, p1d_, p2d_){
  df = total_per_group %>% 
    filter(mrn == mrn_) %>% 
    filter(fdrt %in% c(p1d_, p2d_  )) %>% 
    group_by(fg1_name) %>% 
    summarise(ave_fg = sum(grp_tol)/2)
  return(df)
}

mean_p2d_df <- pmap(stb_pair, function(mrn, p1d, p2d){
    mean_p2d_diet(mrn, p1d, p2d)
  }) %>% 
  set_names(stb %>% pull(sampleid)) %>% 
  bind_rows(.id = 'sampleid') %>% 
  spread(key = 'fg1_name', value = 'ave_fg', fill = 0) %>% 
  inner_join(stb, by = "sampleid") %>% 
  select(-age, -sex, -source)
```

```{r}
# a new column to mark the TPN
# if a patient had TPN in the previous two days before a stool sample collected,
# it will be marked TRUE or otherwise FALSE
tpn <- read_csv('../data/!question/NutritionSupport_DietPatients_peter.csv') %>% 
  filter(`Nutrition Support` == 'TPN') 

ptb <- read_csv('../data/cleaned_patients/diet_patients_97.csv')

tpn_df <- tpn %>% 
  select(mrn = MRN,
         date_support = `Date Nutrition Support`) %>% 
  left_join(ptb %>% select(mrn, hct), by = 'mrn') %>% 
  mutate(date_support = mdy(date_support)) %>% 
  mutate(tdrt = date_support - hct) %>% 
  filter(tdrt <= (max(stb$sdrt - 2))) %>% 
  mutate(tdrt = as.numeric(tdrt))

p2d_TPN_T_or_F <- function(mrn_, p1d_, p2d_){
  num = tpn_df %>% 
    filter(mrn == mrn_) %>% 
    filter(tdrt %in% c(p1d_, p2d_  )) %>% 
    nrow 
  
  if(num > 0) { return(T)}
  else {return(F)}
}

p2d_TPN_T_or_F_df <-  pmap(stb_pair, function(mrn, p1d, p2d){
    p2d_TPN_T_or_F(mrn, p1d, p2d)
  }) %>% 
  set_names(stb %>% pull(sampleid)) %>% 
  bind_rows(.id = 'sampleid') %>% 
  gather(key = 'sampleid', value = 'TPN')
```


```{r}
mean_p2d_df_final <- mean_p2d_df %>% 
  full_join(p2d_TPN_T_or_F_df)

mean_p2d_df_final %>% 
  write_csv('../data/cleaned_stool/all_samples_meta_p2d_fg11.csv')
```

## linear mixed model again

```{r}
library(lmerTest)     

dat <- mean_p2d_df_final %>% 
  mutate(timebin = cut_width(sdrt, 7, boundary=0, closed = 'left')) %>% 
  mutate(intensity = factor(intensity, levels = c('nonablative','reduced','ablative'))) %>% 
  mutate(mrn = as.factor(mrn))

dat %>% 
  select(-mrn) %>%  
  summary
```


```{r}
mod_fg11 <-  lmer(log(simpson_reciprocal) ~ 
              fg_fruit+
               fg_meat+
               fg_milk+
               fg_oils+
                fg_egg+ 
                fg_grain+
                fg_sweets+  
                fg_legume+
                fg_veggie+
                fg_ndrink+
                fg_EN+
               empirical +  
               intensity + 
                TPN +
               (1 | mrn) +
                (1 | timebin), REML = F, data = dat )
summary(mod_fg11) 
```
 
## brms model

```{r} 
library(brms)  
library(ggmcmc)
model_brm <- brm( log(simpson_reciprocal)~ 
                fg_fruit+
                fg_meat+
                fg_milk+
                fg_oils+
                fg_egg+ 
                fg_grain+
                fg_sweets+
                fg_legume+
                fg_veggie+
                fg_ndrink+
                fg_EN+
                TPN +
               empirical + 
               intensity + 
               (1 | mrn) +
                (1 | timebin),  
              data = dat, 
              warmup = 1000, iter = 3000, 
              cores = 8, chains = 2, 
              seed = 123) 

summary(model_brm)
```

```{r}
# trying to make the centered at 0 posterior distribution plot 
posterior_model <- posterior_samples(model_brm) %>% 
  select(starts_with('b')) %>% 
  select(-b_Intercept) %>% 
  gather('item', 'value') %>% 
  group_by(item) %>% 
  mutate(meanperitem = mean(value))%>%
  ungroup()

library(ggthemes)
library(ggridges)
ggplot()+
  ggridges::geom_density_ridges(data  = posterior_model, 
                                aes(x      = value,
                                    y      = reorder(as.factor(item), meanperitem),
                                    height = ..density..),
                                scale = 0.8, size = 0.25,
                                #fill = 'forestgreen',
                                alpha = 0.6) +
  geom_vline(xintercept = 0, col = "red") +
  theme_ridges() +
  scale_x_continuous(limits = c(-1.38,0.25)) +
  ggsave('../figs/posteiror_brms.pdf', width = 10, height = 10)
```


```{r}
# make a version that only has food groups
posterior_model_only_food_group <- posterior_model %>% 
  filter(str_detect(item, '_fg_'))

ggplot()+
  ggridges::geom_density_ridges(data  = posterior_model_only_food_group, 
                                aes(x      = value,
                                    y      = reorder(as.factor(item), meanperitem),
                                    height = ..density..),
                                scale = 1.5, size = 0.25,
                                alpha = 0.6) +
  geom_vline(xintercept = 0, col = "red") +
  #scale_x_continuous(limits = c(-0.02,0.03)) +
  labs(y = 'Food groups',
       title = 'Posterior distribution of regression coefficient for food groups') +
  theme(axis.text.y = element_text(size=20)) +
  theme_tufte() +
  ggsave('../figs/posteiror_brms_fg_only.pdf', width = 6, height = 6) 
```

```{r}
# make modifications for the abstract
# calculate the sugar percentage in each food group using dtb
sugar_perc_all <- dtbn %>% 
  select(Food_code, ends_with('_g')) %>% 
  mutate(fgrp1 = str_sub(Food_code, 1, 1)) %>% 
  select(-Food_code) %>% 
  gather('type', 'gram', Protein_g:Sodium_g) 

sugar_perc_total <- sugar_perc_all  %>% 
  group_by(fgrp1) %>% 
  summarise(total = sum(gram))

sugar_perc_sugar <- sugar_perc_all  %>% 
  filter(type == 'Sugars_g') %>% 
  group_by(fgrp1) %>% 
  summarise(total_sugar = sum(gram))

sugar_df <- sugar_perc_total %>% 
  full_join(sugar_perc_sugar) %>% 
  mutate(sugar_frac = total_sugar/total) %>% 
  mutate(fg1_name = case_when(
    fgrp1 == '1' ~ 'fg_milk',
    fgrp1 == '2' ~ 'fg_meat',
    fgrp1 == '3' ~ 'fg_egg',
    fgrp1 == '4' ~ 'fg_legume',
    fgrp1 == '5' ~ 'fg_grain',
    fgrp1 == '6' ~ 'fg_fruit',
    fgrp1 == '7' ~ 'fg_veggie',
    fgrp1 == '8' ~ 'fg_oils', 
    fgrp1 == '9' ~ 'fg_sweets',
    fgrp1 == 'N' ~ 'fg_ndrink',
    fgrp1 == 'E' ~ 'fg_EN'
  )) %>% 
  mutate(sugar_perc = round(sugar_frac*100, 1)) %>% 
  mutate(sugar_perc = str_glue('{sugar_perc}%')) %>% 
  mutate(fg1_name = str_replace(fg1_name, 'fg_','')) 


# what % of patients out of the 97 had any amount of that food group?
# you might find that its 100% for all of them and 2% for EN, which might lead us to exclude that
pt_perc <- dtbn %>% 
  select(mrn, Food_code) %>% 
  mutate(fgrp1 = str_sub(Food_code, 1, 1)) %>% 
  distinct(mrn, fgrp1) %>% 
  count(fgrp1) %>% 
  mutate(fg1_name = case_when(
    fgrp1 == '1' ~ 'fg_milk',
    fgrp1 == '2' ~ 'fg_meat',
    fgrp1 == '3' ~ 'fg_egg',
    fgrp1 == '4' ~ 'fg_legume',
    fgrp1 == '5' ~ 'fg_grain',
    fgrp1 == '6' ~ 'fg_fruit',
    fgrp1 == '7' ~ 'fg_veggie',
    fgrp1 == '8' ~ 'fg_oils', 
    fgrp1 == '9' ~ 'fg_sweets',
    fgrp1 == 'N' ~ 'fg_ndrink',
    fgrp1 == 'E' ~ 'fg_EN'
  )) %>% 
  mutate(fg1_name = str_replace(fg1_name, 'fg_','')) %>% 
  mutate(pt_perc = round(n/97*100, 1)) %>% 
  mutate(pt_perc = str_glue('{pt_perc}%')) 
```


```{r}
# list the 2.5% and 97.5% quantile and mean for each food group
posterior_fg_df <- posterior_model_only_food_group %>% 
  group_by(item) %>% 
  summarise(q2.5 = quantile(value, probs = 0.025),
            q97.5 = quantile(value, probs = 0.975),
            q50 = quantile(value, probs = 0.5)) %>% 
  mutate(item = str_replace(item, 'b_fg_',''))  %>% 
  full_join(sugar_df %>% select(item = fg1_name, sugar_perc)) %>% 
  full_join(pt_perc %>% select(item = fg1_name, pt_perc)) %>% 
  # regain the full name of the food groups
  mutate(fgrp1 = case_when(
    item ==  'milk' ~ '1',
    item == 'meat' ~ '2',
    item ==  'egg' ~ '3',
    item ==  'legume' ~ '4',
    item == 'grain' ~ '5',
    item == 'fruit' ~ '6',
    item == 'veggie' ~ '7',
    item ==  'oils' ~ '8', 
    item ==  'sweets' ~ '9',
    item == 'ndrink'~  'N',
    item == 'EN'~'E'
  )) %>% 
  left_join(nodes) %>% 
  mutate(fdesc = case_when(
    item ==  'EN' ~ 'Enteral nutrition',
    item == 'ndrink' ~ 'Nutritional drink',
    TRUE  ~ fdesc
  )) %>% 
  mutate(fdesc = str_replace_all(fdesc, '_',' '))
```


```{r}
pos_interval <- posterior_fg_df %>%  
  arrange(q50) %>% 
  ggscatter(x = 'fdesc', y = 'q50', color = 'red',
            ylab = 'Association with fecal microbiome',
            title = '95% interval of regression coefficient in the posterior distribution\nfrom Bayesian multilevel model',
            xlab = 'High-level food categories') +
  geom_errorbar(aes( ymin  = q2.5,
                   ymax  = q97.5,
                   width = 0.25)) +
  geom_hline(yintercept = 0, col = "firebrick", linetype = 'dashed') +
  geom_text(aes(x = fdesc, y = -0.015, label = sugar_perc)) +
  geom_text(aes(x = fdesc[11], y = -0.015, label = 'Sugar\ncontent'), nudge_x = +0.45) +
  geom_text(aes(x = fdesc, y = -0.011, label = pt_perc)) +
  geom_text(aes(x = fdesc[11], y = -0.011, label = 'Patients\nexposed'), nudge_x = +0.45) +
  geom_segment(aes(x=-1, y=-0.005, xend=-1, yend=-0.008), 
               arrow=arrow(length=unit(0.3, "cm")) , 
               nudge_x = -2) +
  coord_flip() + 
  ylim(-0.015, 0.018)
pos_interval
```


## is the nutritional drink signal simply because they are not eating so the stool diversity is low

```{r}
# do the patients have EN and regular eating at the same time?
 # no , when they had EN, they don't eat regular food
# overlap <- en %>% 
#   select(mrn, fdrt) %>% 
#   inner_join(cur8)
# 
# # what's the correlation between daily total calories and daily consumption of the nutritional drink(not the EN ones)
# # the patients and the days that they had nutritional drink
# ndrink_pt <- dtbnv %>% 
#   filter(str_detect(Food_code, '^N')) %>% 
#   distinct(mrn, fdrt)
# 
# # the total daily calories of the patients on those days
# 
# all_daily_ndrink_pt <- dtbnv %>% 
#   group_by(mrn, fdrt) %>% 
#   summarise(total_daily = sum(Calories_kcal)) %>% 
#   inner_join(ndrink_pt)
# 
# 
# # per day sum up the amount of nutritional drink
# ndrink_pt_sum <- dtbnv %>% 
#   filter(str_detect(Food_code, '^N')) %>% 
#   group_by(mrn, fdrt) %>% 
#   summarise(ndrink_sum = sum(dehydrated_weight))
# 
# combined <- all_daily_ndrink_pt %>% 
#   full_join(ndrink_pt_sum)
# 
# combined %>% 
#   ggpubr::ggscatter(x = 'total_daily', y = 'ndrink_sum', 
#                     xlab = 'Total daily calories',ylab = 'Total daily nutritional drink (dehydrated wt)',
#             add = "reg.line",  # Add regressin line
#            add.params = list(color = "blue", fill = "lightgray"), # Customize line
#            conf.int = TRUE, # Add confidence interval
#            cor.coef = TRUE, # Add correlation coefficient.
#            cor.coeff.args = list(method = "pearson", label.x = 3, label.sep = "\n")) 

# in terms of the nutritional drink (excluding the EN) it's not highly correlated with total calories, cor = 0.49
# so the patients are still eating
# so the nutritional drink could be lowering the diversity
```
