---
title: "Start from 97 corrected dtb"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
--- 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

```{r}
library(tidyverse)
library(readxl)
library(lubridate)
```

Comments from Peter: The only concern is how much is "each" that is being served in some foods. Have to ask the kitchen how big is the portion.

(About a quarter in the small_cor has the calories corrected. )

# Diet data

## Integrate the Peter curated smallest portion into the dataset

```{r}
small_cor <- read_csv('../data/!question/double_check_for_calories_also_nutrients_peter.csv') %>% 
  unite(col = 'food_full', Food_NSC:Unit, sep = '__', remove = T) 

cur <- read_csv('../data/!question/all_97_cor.csv') %>% 
  rename(Protein_g = Proteing_g) %>% 
  mutate(Sodium_g = round(Sodium_mg/1000, 3)) %>% 
  select(-Sodium_mg) %>% 
  unite(col = 'food_full', Food_NSC:Unit, sep = '__', remove = T) 


all_small_cor <- cur %>% 
  filter(food_full %in% small_cor$food_full) %>% 
  distinct(food_full, Por_eaten, .keep_all = T) %>% 
  select(MRN:Por_eaten) %>% 
  left_join(small_cor) %>% 
  arrange(food_full, Calories_kcal)

# use the ratio to correct other portions
all_small_cor_res <- all_small_cor %>% 
  group_by(food_full) %>% 
  mutate(Calories_kcal = Calories_kcal[1],
         Protein_g = Protein_g[1],
         Fat_g = Fat_g[1],
         Carbohydrates_g = Carbohydrates_g[1],
         Fibers_g = Fibers_g[1],
         Sugars_g = Sugars_g[1],
         Sodium_g = Sodium_g[1]) %>% 
  split(.$food_full) %>% 
  map_dfr(~ mutate(.data = ., ratio = Por_eaten/Por_eaten[1]) %>% 
        split(.$Por_eaten) %>% 
        map_dfr(~ mutate(.data = .,   
                   Calories_kcal = Calories_kcal*ratio,
                   Protein_g = Protein_g*ratio,
                   Fat_g = Fat_g*ratio,
                   Carbohydrates_g = Carbohydrates_g*ratio,
                   Fibers_g = Fibers_g*ratio,
                   Sugars_g = Sugars_g*ratio,
                   Sodium_g = Sodium_g*ratio)) %>% 
        select(-ratio))

all_small_cor_res %>% 
  summary 
####### 3 of 2017 Dressing, Creamy Italian (.5)__pkt 274g fat?? (I probably should focus on the calories for now...)

# then scale the different food_full and portion to repeated occurences 
cur2 <- cur %>%  
  select(MRN: Por_eaten) %>% 
  left_join(all_small_cor_res %>% select(food_full:Sodium_g) %>% distinct(), by = c('food_full','Por_eaten'))
```


## Load other dataset

```{r}

all_code <- read_csv('../data/cleaned_diet_data/all_food_nsc_with_code_FINAL.csv') %>% 
  distinct(Food_NSC, .keep_all = T)

```

## Foods with zero calorie and their gram weight
 
```{r}
cal0_peter <- read_excel('../data/cal_0_food_nsc_all_with_unit_Peter.xls')

# the things that should be removed from combined
na_rm <- cal0_peter %>% 
  filter(`gram weight` == 'NA') %>% 
  distinct(Food_NSC) %>% 
  pull(Food_NSC)

# remove the items that are not valid food like salt and pepper and ice
cur3 <- cur2 %>% 
  separate(food_full, into = c('Food_NSC','Unit'), sep = '__') %>% 
  filter(! Food_NSC %in% na_rm)

# the items that are watery and now I have a total weight
cal0_peter_items <- cal0_peter %>%   
  filter(`gram weight` != 'NA')  %>% 
  rename(total_weight = `gram weight`) %>% 
  mutate(total_weight = as.numeric(total_weight)) %>% 
  distinct(Food_NSC) %>% 
  pull(Food_NSC) 

cal0_peter_clean <- cal0_peter %>% 
  filter(`gram weight` != 'NA')  %>% 
  rename(total_weight = `gram weight`) %>% 
  mutate(total_weight = as.numeric(total_weight)) %>% 
  distinct()

cal0_split <- cur3 %>% 
  split(.$Food_NSC %in% cal0_peter_items)

cal0_join_with_other <- cal0_split %>% 
  pluck('TRUE') %>%  
  left_join(cal0_peter_clean) 

# there are several NA because they are of other portions
# correct this 
other_portion_NA_items <- cal0_join_with_other %>% 
  filter(is.na(total_weight)) %>% 
  distinct(Food_NSC) %>% 
  pull(Food_NSC)

other_portion_NA_foods <- cal0_join_with_other %>% 
  filter(Food_NSC %in% other_portion_NA_items) %>% 
  arrange(Food_NSC, Unit, total_weight) %>% 
  unite(col = 'food_full', Food_NSC:Unit, sep = '__', remove = T) %>% 
  group_by(food_full) %>% 
  mutate(total_weight = total_weight[1]) %>% 
  split(.$food_full) %>% 
  map_dfr(~ mutate(.data = ., ratio = Por_eaten/Por_eaten[1]) %>% 
        split(.$Por_eaten) %>% 
        map_dfr(~ mutate(.data = .,   
                   total_weight = total_weight*ratio)) %>% 
        select(-ratio)) %>% 
  separate(food_full, into = c('Food_NSC','Unit'), sep = '__') 

cur4 <- bind_rows(
  cal0_split %>% pluck('FALSE'),
  cal0_join_with_other %>% filter(!Food_NSC %in% other_portion_NA_items),
  other_portion_NA_foods
)

```


## Get the food code for the items

```{r}
current <- cur4 %>% 
  distinct(Food_NSC) %>% pull(Food_NSC)

all_c <- all_code %>% 
  distinct(Food_NSC) %>% pull(Food_NSC)

length(intersect(cur4 %>% 
  distinct(Food_NSC) %>% pull(Food_NSC), all_code %>% 
  distinct(Food_NSC) %>% pull(Food_NSC)))

ignore <- setdiff(current, all_c)

# this one the patient only had one ounce, could be ignored

cur5 <- cur4 %>% 
  filter(Food_NSC != ignore)


# join the food code to the table 

cur6 <- cur5 %>% 
  left_join(all_code, by = 'Food_NSC') %>% 
  rename(Food_code = Food.Code,
         description = FNDDS.Main.Food.Description)
summary(cur6)

# high fat items still makes me feel .....but i guess fat is not that important ?? 
```

## Get the dehydrated weight for the non zero calorie foods

```{r}
# load the fndds table and do the joining
engy <- read_excel('../data/source/2015-2016 FNDDS At A Glance - FNDDS Nutrient Values.xlsx', skip = 1, col_types  = 'text') %>% 
  rename_all(funs(str_replace_all(., ' ','_'))) %>% 
  rename_all(funs(str_replace_all(., '\\(|\\)',''))) %>% 
  select(Food_code, Energy_kcal, Water_g) %>% 
  mutate_all(as.numeric)

# there are three records that are 0 cal but should not be, manually correct it.
cur6 %>% 
  filter(is.na(total_weight))  %>% 
  arrange(Calories_kcal) %>% 
  write_csv('../data/!question/cur6.csv')
 
# load back the manually corrected table
cur6_NA_cor <- read_csv('../data/!question/cur6_corrected.csv')

# get the dehydated weight for cur6_NA_cor
cur6_non0_cal <- cur6_NA_cor %>% 
  select(-total_weight) %>% 
  left_join(engy, by  = 'Food_code')  %>% 
  mutate(total_weight = if_else(Energy_kcal == 0, 99999, Calories_kcal/Energy_kcal*100)) %>% 
  mutate(dehydrated_weight = if_else(total_weight == 99999, 0.0001, total_weight*(1 - Water_g/100))) %>% 
  select(-Energy_kcal, -total_weight, -Water_g) %>% 
  mutate(Date = mdy(Date))

# calculate the dehydrated weight for the 0 calories items that I have the total weight 
cal0_weight_conversion <- cur6 %>% 
  filter(!is.na(total_weight)) %>% 
  left_join(engy %>% select(Food_code, Water_g), by  = 'Food_code')  %>% 
  mutate(dehydrated_weight = total_weight*(1 - Water_g/100)) %>% 
  select(-total_weight, -Water_g)

cur7 <- bind_rows(cal0_weight_conversion, cur6_non0_cal)
```


## Convert the date to fdrt

```{r}
ptb <- read_csv('../data/cleaned_patients/diet_patients_97.csv')

cur8 <- cur7 %>% 
  rename(mrn = MRN) %>% 
  left_join(ptb %>% select(mrn, hct), by = "mrn") %>% 
  mutate(fdrt = Date - hct) %>% 
  select(-hct, -Date)  
```


## Add the EN and TPN data

Note the range of the TPN and EN's fdrt days 

```{r}
## combine the EN data (that has a dedicated food code for them)
en <- read_csv('../data/!question/NutritionSupport_DietPatients_peter.csv') %>% 
  filter(!is.na(Food.Code)) %>% 
  mutate(meal = 'EN',
         Por_eaten = 1) %>% 
  select(mrn = MRN, 
         Date = `Date Nutrition Support`,
         Meal = meal,
         Food_NSC = `EN formula`,
         Por_eaten,
         Calories_kcal = `EN calories`,
         Protein_g = `EN PRO`,
         Fat_g = `EN fat`,
         Carbohydrates_g = `EN CHO`,
         Fibers_g = `EN Fiber`,
         Sugars_g = `EN Sugar`,
         Sodium_g = `EN Sodium`,
         Food_code = Food.Code,
         dehydrated_weight = `EN volume_weight_g`,
         description = FNDDS.Main.Food.Description
         ) %>% 
  mutate(Sodium_g = Sodium_g/1000) %>% 
  mutate(Date = mdy(Date),
         Unit = "EN unit") %>% 
  left_join(ptb %>% select(mrn, hct), by = 'mrn') %>% 
  mutate(fdrt = Date - hct) %>% 
  select(colnames(cur8)) %>% 
  filter(fdrt <= max(cur8$fdrt)) # keep only the records of the days that not exceed the max of the normal diet days

# the 5 EN patients are in the diet cohort, confirmed 
length(intersect(cur$MRN, en$mrn))


# how many patients had EN
en %>% 
  distinct(mrn) %>% 
  nrow
```

```{r}
# how many patients had TPN
# the most updated TPN EN table is below:
tpn <- read_csv('../data/!question/NutritionSupport_DietPatients_peter.csv') %>% 
  filter(`Nutrition Support` == 'TPN')


tpn %>% 
  distinct(MRN)
```

## Correct some records that have abnormally high fat_g that got missed      

```{r}
fat_problematic <- cur8 %>% 
  filter(Food_NSC %in% c('2017 Milkshake, Oreo (3) 11 oz','2017 Dressing, Creamy Italian (.5)','2017 Dressing, Creamy Italian')) %>% 
  arrange(Food_NSC, Unit, Por_eaten) %>% 
  group_by(Food_NSC, Unit) %>% 
  slice(1) %>% 
  select(Food_NSC:description)

fat_problematic %>% 
  write_csv('../data/!question/fat_problematic.csv')
```


```{r}
# correct the above mentioned few -- the fat_g column 
fat_cor <- read_csv('../data/!question/fat_problematic_peter.csv') %>% 
  unite(col = 'food_full', Food_NSC:Unit, sep = '__', remove = T) 

all_fat <- cur8 %>% 
  unite(col = 'food_full', Food_NSC:Unit, sep = '__', remove = T)  %>% 
  filter(food_full %in% fat_cor$food_full) %>% 
  select(mrn:Por_eaten, fdrt) %>% 
  left_join(fat_cor) %>% 
  arrange(food_full, Calories_kcal)

all_fat_res <- all_fat %>% 
  group_by(food_full) %>% 
  mutate(Calories_kcal = Calories_kcal[1],
         Protein_g = Protein_g[1],
         Fat_g = Fat_g[1],
         Carbohydrates_g = Carbohydrates_g[1],
         Fibers_g = Fibers_g[1],
         Sugars_g = Sugars_g[1],
         Sodium_g = Sodium_g[1],
         Food_code = Food_code[1],
         description = description[1]) %>% 
  split(.$food_full) %>% 
  map_dfr(~ mutate(.data = ., ratio = Por_eaten/Por_eaten[1]) %>% 
        split(.$Por_eaten) %>% 
        map_dfr(~ mutate(.data = .,   
                   Calories_kcal = Calories_kcal*ratio,
                   Protein_g = Protein_g*ratio,
                   Fat_g = Fat_g*ratio,
                   Carbohydrates_g = Carbohydrates_g*ratio,
                   Fibers_g = Fibers_g*ratio,
                   Sugars_g = Sugars_g*ratio,
                   Sodium_g = Sodium_g*ratio)) %>% 
        select(-ratio)) %>% 
   separate(food_full, into = c('Food_NSC','Unit'), sep = '__') 

# add the dehydrated weight and fdrt to the above table
all_fat_res_full <- all_fat_res %>% 
  left_join(engy, by  = 'Food_code')  %>% 
  mutate(total_weight = if_else(Energy_kcal == 0, 99999, Calories_kcal/Energy_kcal*100)) %>% 
  mutate(dehydrated_weight = if_else(total_weight == 99999, 0.0001, total_weight*(1 - Water_g/100))) %>% 
  select(-Energy_kcal, -total_weight, -Water_g) %>% 
  select(colnames(cur8))

# now merge this corrected portion back to the whole dataset 
fat_cor_split <- cur8 %>% 
  unite(col = 'food_full', Food_NSC:Unit, sep = '__', remove = T)  %>% 
  split(.$food_full %in% fat_cor$food_full)

cur9 <- bind_rows(
  all_fat_res_full,
  fat_cor_split %>% pluck('FALSE') %>% separate(food_full, into = c('Food_NSC','Unit'), sep = '__') 
)
  
summary(cur9)

```

## regrouping fruit juice (citrus and non-citrus ) into sweet beverage group

```{r}
all_level_label <- read_tsv('../data/source/NodeLabelsMCT.txt', col_types = 'cc')

fruit2 <- read_tsv('../data/source/NodeLabelsMCT.txt', col_types = 'cc')  %>% 
  filter(str_detect(Level.code, '^6')) %>% 
  filter(str_length(Level.code) == 2)

sweets2 <- read_tsv('../data/source/NodeLabelsMCT.txt', col_types = 'cc')  %>% 
  filter(str_detect(Level.code, '^9')) %>% 
  filter(str_length(Level.code) == 2)

all_fruit <- cur9 %>% 
  mutate(Food_code = as.character(Food_code)) %>% 
  filter(str_detect(Food_code, '^6')) %>% 
  distinct(Food_NSC, Food_code) %>% 
  mutate(L2_grp = str_sub(Food_code, start = 1, end = 2))  %>% 
  left_join(fruit2 %>% rename(L2_grp = Level.code), by  = 'L2_grp') %>% 
  arrange(L2_grp) %>% 
  # to revise the food code for the fruit juice citrus and non-citrus
  mutate(is_juice = if_else(str_detect(Food_NSC, fixed('juice', ignore_case=TRUE)), T, F)) %>% 
  # change the food code for the citrus juice 
  mutate(Food_code = if_else(L2_grp == '61' & is_juice == 'TRUE', '92530610', Food_code)) %>% 
  # change the food code for the non-citrus juice
  mutate(Food_code = if_else(L2_grp == '64', '92510610', Food_code))  %>% 
  # also updating the description 
  mutate(L2_grp = str_sub(Food_code, start = 1, end = 2)) %>% 
  mutate(Main.food.description = if_else(Food_code == '92510610', 'Fruit juice drink',
                                                    if_else(Food_code == '92530610', 'Fruit juice drink, with high vitamin C', Main.food.description)))

juice_changed <- all_fruit %>% 
  filter(str_detect(Food_code, '^9')) %>% 
  select(-is_juice, -L2_grp) %>% 
  rename(description = Main.food.description)

# no need to change the food code for the raw orange, it is in its right form, the citrus fruit is separated from "other fruit". 

# so what really happened is I split the fruit juice (citrus and non-citrus ) into sweet beverage group
```

```{r}
# updating the above info in the dtb 
cur10_split  <- cur9 %>% 
  mutate(Food_code = as.character(Food_code)) %>% 
  split(.$Food_NSC %in% juice_changed$Food_NSC)
  
# replace the food code and description for the things that changed 
cur10_split_change <-  cur10_split %>% 
  pluck('TRUE') %>% 
  select(-Food_code, -description) %>% 
  left_join(juice_changed, by  = 'Food_NSC') %>% 
  select(colnames(dtb))

cur10 <- bind_rows( 
  cur10_split_change,
  cur10_split %>% pluck('FALSE')
) %>% 
  mutate(Food_code = as.numeric(Food_code))
```

## correcting few citrus entries that are mistakenly 0 calories

```{r}
# double check the calo ==0 stuff
cal0 <- cur10 %>% 
  filter(Calories_kcal == 0) %>% 
  distinct(Food_NSC, Food_code, .keep_all = T) %>% 
  arrange(Food_code) 

# the "2017 Juice,  Orange (1)" and "2017 Condiments, Lemon, Wedge" has some problems
# it couldn't have a cal of 0, and the current dehydrated wt calculated from total weight is even smaller than a component like sugar.

question_citrus <- cur10 %>% 
  filter(Food_NSC %in% c('2017 Juice,  Orange (1)','2017 Condiments, Lemon, Wedge')) %>% 
  arrange(Food_NSC, Unit, Por_eaten) %>% 
  group_by(Food_NSC, Unit) %>% 
  slice(1) %>% 
  select(Food_NSC:Sodium_g)

question_citrus %>% 
  write_csv('../data/!question/citrus_cal_0_problem.csv')
```


```{r}
# corrcting the few citrus  --- orange juice and lemon
citrus_cor <- read_csv('../data/!question/citrus_cal_0_problem_peter.csv') %>% 
  unite(col = 'food_full', Food_NSC:Unit, sep = '__', remove = T)

all_citrus <- cur10 %>% 
  unite(col = 'food_full', Food_NSC:Unit, sep = '__', remove = T)  %>% 
  filter(food_full %in% citrus_cor$food_full) %>% 
  select(mrn:Por_eaten, fdrt, Food_code, description) %>% 
  left_join(citrus_cor) %>% 
  arrange(food_full, Calories_kcal)

all_citrus_res <- all_citrus %>% 
  group_by(food_full) %>% 
  mutate(Calories_kcal = Calories_kcal[1],
         Protein_g = Protein_g[1],
         Fat_g = Fat_g[1],
         Carbohydrates_g = Carbohydrates_g[1],
         Fibers_g = Fibers_g[1],
         Sugars_g = Sugars_g[1],
         Sodium_g = Sodium_g[1],
         Food_code = Food_code[1],
         description = description[1]) %>% 
  split(.$food_full) %>% 
  map_dfr(~ mutate(.data = ., ratio = Por_eaten/Por_eaten[1]) %>% 
        split(.$Por_eaten) %>% 
        map_dfr(~ mutate(.data = .,   
                   Calories_kcal = Calories_kcal*ratio,
                   Protein_g = Protein_g*ratio,
                   Fat_g = Fat_g*ratio,
                   Carbohydrates_g = Carbohydrates_g*ratio,
                   Fibers_g = Fibers_g*ratio,
                   Sugars_g = Sugars_g*ratio,
                   Sodium_g = Sodium_g*ratio)) %>% 
        select(-ratio)) %>% 
   separate(food_full, into = c('Food_NSC','Unit'), sep = '__') 


# add the dehydrated weight 
all_citrus_res_full <- all_citrus_res %>% 
  left_join(engy, by  = 'Food_code')  %>% 
  mutate(total_weight = if_else(Energy_kcal == 0, 99999, Calories_kcal/Energy_kcal*100)) %>% 
  mutate(dehydrated_weight = if_else(total_weight == 99999, 0.0001, total_weight*(1 - Water_g/100))) %>% 
  select(-Energy_kcal, -total_weight, -Water_g) %>% 
  select(colnames(cur10))


# now merge this corrected portion back to the whole dataset 
citrus_cor_split <- cur10 %>% 
  unite(col = 'food_full', Food_NSC:Unit, sep = '__', remove = T)  %>% 
  split(.$food_full %in% citrus_cor$food_full)

cur11 <- bind_rows(
  all_citrus_res_full,
  citrus_cor_split %>% pluck('FALSE') %>% separate(food_full, into = c('Food_NSC','Unit'), sep = '__') 
)
```

## Finalize the data

```{r}
# actual food records (excluding fake TPN)
actual <- bind_rows(cur10, en)

actual %>% nrow 

actual %>% 
  write_csv('../data/cleaned_diet_data/FINAL_97_with_code_all_foods_drt_with_EN_finalized.csv')

```

```{r}
# a histgram to see the distribution of the cal and nutrients?
# cur9 %>% 
#   select(ends_with('_g')) %>% 
#   gather('item', 'value') %>% 
#   mutate(logvalue = log(value +0.001)) %>% 
#   ggpubr::gghistogram(x = 'logvalue', bins = 60, fill = 'forestgreen') +
#   facet_wrap(~item, scales = 'free') +
#   ggsave('../figs/cur9_nutrients.pdf', width = 12, height = 7)
# 
# cur9 %>% 
#   select(Calories_kcal, ends_with('_g')) %>% 
#   gather('item', 'value') %>% 
#   ggpubr::ggboxplot(y = 'value') +
#   facet_wrap(~item, scales = 'free')
```

 
# Stool data

go to script 011 for selecting qualified stool samples (Note, the dtb used is **actual**)

# The Microbiome and p2d diet association using lmer and brm

go to script 035 (Also split out the nutritional drinks)

# Also rerun the dominant species script and show the results of that  


