---
title: "Start from 97 corrected dtb"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
--- 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

```{r}
library(tidyverse)
library(readxl)
library(lubridate)
```

Comments from Peter: The only concern is how much is "each" that is being served in some foods. Have to ask the kitchen how big is the portion.

(About a quarter in the small_cor has the calories corrected. )

# Diet data

## Integrate the Peter curated smallest portion into the dataset

```{r}
small_cor <- read_csv('../data/!question/double_check_for_calories_also_nutrients_peter.csv') %>% 
  unite(col = 'food_full', Food_NSC:Unit, sep = '__', remove = T) 

cur <- read_csv('../data/!question/all_97_cor.csv') %>% 
  rename(Protein_g = Proteing_g) %>% 
  mutate(Sodium_g = round(Sodium_mg/1000, 3)) %>% 
  select(-Sodium_mg) %>% 
  unite(col = 'food_full', Food_NSC:Unit, sep = '__', remove = T) 


all_small_cor <- cur %>% 
  filter(food_full %in% small_cor$food_full) %>% 
  distinct(food_full, Por_eaten, .keep_all = T) %>% 
  select(MRN:Por_eaten) %>% 
  left_join(small_cor) %>% 
  arrange(food_full, Calories_kcal)

# use the ratio to correct other portions
all_small_cor_res <- all_small_cor %>% 
  group_by(food_full) %>% 
  mutate(Calories_kcal = Calories_kcal[1],
         Protein_g = Protein_g[1],
         Fat_g = Fat_g[1],
         Carbohydrates_g = Carbohydrates_g[1],
         Fibers_g = Fibers_g[1],
         Sugars_g = Sugars_g[1],
         Sodium_g = Sodium_g[1]) %>% 
  split(.$food_full) %>% 
  map_dfr(~ mutate(.data = ., ratio = Por_eaten/Por_eaten[1]) %>% 
        split(.$Por_eaten) %>% 
        map_dfr(~ mutate(.data = .,   
                   Calories_kcal = Calories_kcal*ratio,
                   Protein_g = Protein_g*ratio,
                   Fat_g = Fat_g*ratio,
                   Carbohydrates_g = Carbohydrates_g*ratio,
                   Fibers_g = Fibers_g*ratio,
                   Sugars_g = Sugars_g*ratio,
                   Sodium_g = Sodium_g*ratio)) %>% 
        select(-ratio))

all_small_cor_res %>% 
  summary 
####### 3 of 2017 Dressing, Creamy Italian (.5)__pkt 274g fat?? (I probably should focus on the calories for now...)

# then scale the different food_full and portion to repeated occurences 
cur2 <- cur %>%  
  select(MRN: Por_eaten) %>% 
  left_join(all_small_cor_res %>% select(food_full:Sodium_g) %>% distinct(), by = c('food_full','Por_eaten'))
```


## Load other dataset

```{r}

all_code <- read_csv('../data/cleaned_diet_data/all_food_nsc_with_code_FINAL.csv') %>% 
  distinct(Food_NSC, .keep_all = T)

```

## Foods with zero calorie and their gram weight
 
```{r}
cal0_peter <- read_excel('../data/cal_0_food_nsc_all_with_unit_Peter.xls')

# the things that should be removed from combined
na_rm <- cal0_peter %>% 
  filter(`gram weight` == 'NA') %>% 
  distinct(Food_NSC) %>% 
  pull(Food_NSC)

# remove the items that are not valid food like salt and pepper and ice
cur3 <- cur2 %>% 
  separate(food_full, into = c('Food_NSC','Unit'), sep = '__') %>% 
  filter(! Food_NSC %in% na_rm)

# the items that are watery and now I have a total weight
cal0_peter_items <- cal0_peter %>%   
  filter(`gram weight` != 'NA')  %>% 
  rename(total_weight = `gram weight`) %>% 
  mutate(total_weight = as.numeric(total_weight)) %>% 
  distinct(Food_NSC) %>% 
  pull(Food_NSC) 

cal0_peter_clean <- cal0_peter %>% 
  filter(`gram weight` != 'NA')  %>% 
  rename(total_weight = `gram weight`) %>% 
  mutate(total_weight = as.numeric(total_weight)) %>% 
  distinct()

cal0_split <- cur3 %>% 
  split(.$Food_NSC %in% cal0_peter_items)

cal0_join_with_other <- cal0_split %>% 
  pluck('TRUE') %>%  
  left_join(cal0_peter_clean) 

# there are several NA because they are of other portions
# correct this 
other_portion_NA_items <- cal0_join_with_other %>% 
  filter(is.na(total_weight)) %>% 
  distinct(Food_NSC) %>% 
  pull(Food_NSC)

other_portion_NA_foods <- cal0_join_with_other %>% 
  filter(Food_NSC %in% other_portion_NA_items) %>% 
  arrange(Food_NSC, Unit, total_weight) %>% 
  unite(col = 'food_full', Food_NSC:Unit, sep = '__', remove = T) %>% 
  group_by(food_full) %>% 
  mutate(total_weight = total_weight[1]) %>% 
  split(.$food_full) %>% 
  map_dfr(~ mutate(.data = ., ratio = Por_eaten/Por_eaten[1]) %>% 
        split(.$Por_eaten) %>% 
        map_dfr(~ mutate(.data = .,   
                   total_weight = total_weight*ratio)) %>% 
        select(-ratio)) %>% 
  separate(food_full, into = c('Food_NSC','Unit'), sep = '__') 

cur4 <- bind_rows(
  cal0_split %>% pluck('FALSE'),
  cal0_join_with_other %>% filter(!Food_NSC %in% other_portion_NA_items),
  other_portion_NA_foods
)

```


## Get the food code for the items

```{r}
current <- cur4 %>% 
  distinct(Food_NSC) %>% pull(Food_NSC)

all_c <- all_code %>% 
  distinct(Food_NSC) %>% pull(Food_NSC)

length(intersect(cur4 %>% 
  distinct(Food_NSC) %>% pull(Food_NSC), all_code %>% 
  distinct(Food_NSC) %>% pull(Food_NSC)))

ignore <- setdiff(current, all_c)

# this one the patient only had one ounce, could be ignored

cur5 <- cur4 %>% 
  filter(Food_NSC != ignore)


# join the food code to the table 

cur6 <- cur5 %>% 
  left_join(all_code, by = 'Food_NSC') %>% 
  rename(Food_code = Food.Code,
         description = FNDDS.Main.Food.Description)
summary(cur6)

# high fat items still makes me feel .....but i guess fat is not that important ?? 
```

## Get the dehydrated weight for the non zero calorie foods

```{r}
# load the fndds table and do the joining
engy <- read_excel('../data/source/2015-2016 FNDDS At A Glance - FNDDS Nutrient Values.xlsx', skip = 1, col_types  = 'text') %>% 
  rename_all(funs(str_replace_all(., ' ','_'))) %>% 
  rename_all(funs(str_replace_all(., '\\(|\\)',''))) %>% 
  select(Food_code, Energy_kcal, Water_g) %>% 
  mutate_all(as.numeric)

# there are three records that are 0 cal but should not be, manually correct it.
cur6 %>% 
  filter(is.na(total_weight))  %>% 
  arrange(Calories_kcal) %>% 
  write_csv('../data/!question/cur6.csv')
 
# load back the manually corrected table
cur6_NA_cor <- read_csv('../data/!question/cur6_corrected.csv')

# get the dehydated weight for cur6_NA_cor
cur6_non0_cal <- cur6_NA_cor %>% 
  select(-total_weight) %>% 
  left_join(engy, by  = 'Food_code')  %>% 
  mutate(total_weight = if_else(Energy_kcal == 0, 99999, Calories_kcal/Energy_kcal*100)) %>% 
  mutate(dehydrated_weight = if_else(total_weight == 99999, 0.0001, total_weight*(1 - Water_g/100))) %>% 
  select(-Energy_kcal, -total_weight, -Water_g) %>% 
  mutate(Date = mdy(Date))

# calculate the dehydrated weight for the 0 calories items that I have the total weight 
cal0_weight_conversion <- cur6 %>% 
  filter(!is.na(total_weight)) %>% 
  left_join(engy %>% select(Food_code, Water_g), by  = 'Food_code')  %>% 
  mutate(dehydrated_weight = total_weight*(1 - Water_g/100)) %>% 
  select(-total_weight, -Water_g)

cur7 <- bind_rows(cal0_weight_conversion, cur6_non0_cal)
```


## Convert the date to fdrt

```{r}
ptb <- read_csv('../data/cleaned_patients/diet_patients_97.csv')

cur8 <- cur7 %>% 
  rename(mrn = MRN) %>% 
  left_join(ptb %>% select(mrn, hct), by = "mrn") %>% 
  mutate(fdrt = Date - hct) %>% 
  select(-hct, -Date)  
```


## Add the EN and TPN data

Note the range of the TPN and EN's fdrt days 

```{r}
## combine the EN data (that has a dedicated food code for them)
en <- read_csv('../data/cleaned_diet_data/NutritionSupport_DietPatients.csv') %>% 
  filter(!is.na(Food.Code)) %>% 
  mutate(meal = 'EN',
         Por_eaten = 1) %>% 
  select(mrn = MRN, 
         Date = `Date Nutrition Support`,
         Meal = meal,
         Food_NSC = `EN formula`,
         Por_eaten,
         Calories_kcal = `EN calories`,
         Protein_g = `EN PRO`,
         Fat_g = `EN fat`,
         Carbohydrates_g = `EN CHO`,
         Fibers_g = `EN Fiber`,
         Sugars_g = `EN Sugar`,
         Sodium_g = `EN Sodium`,
         Food_code = Food.Code,
         description = FNDDS.Main.Food.Description
         ) %>% 
  mutate(Sodium_g = Sodium_g/1000) %>% 
  mutate(dehydrated_weight = Protein_g + Fat_g + Carbohydrates_g + Fibers_g + Sugars_g + Sodium_g) %>% 
  mutate(Date = mdy(Date),
         Unit = "EN unit") %>% 
  left_join(ptb %>% select(mrn, hct), by = 'mrn') %>% 
  mutate(fdrt = Date - hct) %>% 
  select(colnames(cur8)) %>% 
  filter(fdrt <= max(cur8$fdrt)) # keep only the records of the days that not exceed the max of the normal diet days

```

```{r}
# create *0* counts for the foods on TPN days since TPN doesn't go thru colon but they are actual foods  
TPN <- read_csv('../data/cleaned_diet_data/TPN_EN patients.csv') %>% 
  filter(mrn %in% cur8$mrn) %>% 
  rename(fdrt = Days.Post.HSCT) %>% 
  mutate(nut_type = if_else(TPN.1..EN.2 == 1, "TPN", "EN")) %>% 
  select(mrn, fdrt, nut_type) %>% 
  filter(nut_type == 'TPN') 

# to see if there is overlap between the normal diet and the above one
overlap <- TPN %>% 
  inner_join(cur8 %>% mutate(fdrt = as.numeric(fdrt))) %>% 
  distinct(mrn, fdrt, nut_type)
 
overlap %>% 
  distinct(nut_type)

# the days the patients only have TPN
TPN_clean <- TPN %>% 
  left_join(overlap, by = c("mrn", "fdrt")) %>% 
  filter(is.na(nut_type.y))  %>% 
  rename(nut_type = nut_type.x) %>% 
  select(-nut_type.y) %>% 
  select(- nut_type)

ftb_colnames <- colnames(cur8)

# use all of the foods this patients has ever eaten as the food_nsc
create_fake_food_cts_for_TPN_patient <- function(mrn_, fdrt_){
  this_pt = cur8 %>% 
    filter(mrn == mrn_) %>% 
    distinct(mrn, Food_NSC, Food_code, description) %>% 
    mutate(fdrt = fdrt_, 
           Meal = "fake meal",
           Por_eaten = 0,
           Unit = "fake unit",
           Calories_kcal = 0, 
           Protein_g = 0, 
           Fat_g = 0,
           Carbohydrates_g = 0,
           Fibers_g = 0,
           Sugars_g =0,
           Sodium_g =0,
           dehydrated_weight = 0) %>% 
    select(all_of(ftb_colnames))
}
         
fake_TPN <- pmap_dfr(TPN_clean, function(mrn, fdrt){
  create_fake_food_cts_for_TPN_patient(mrn, fdrt)
} )
```

```{r}
# some metric
en %>% 
  distinct(mrn) %>% 
  nrow

TPN %>% 
  distinct(mrn) %>% 
  nrow
```

## Finalize the data

```{r}
# actual food records (excluding fake TPN)
actual <- bind_rows(cur8, en)

actual %>% nrow

actual %>% 
  write_csv('../data/cleaned_diet_data/FINAL_97_with_code_all_foods_drt_with_EN_Oct8_actual.csv')

# including the fake TPN
ALL <- bind_rows(actual %>%  mutate(fdrt = as.numeric(fdrt)), fake_TPN)

ALL %>% nrow

ALL %>% 
  write_csv('../data/cleaned_diet_data/FINAL_97_with_code_all_foods_drt_with_EN_and_TPN_Oct8.csv')
```


```{r}
ALL %>% 
  select(-mrn) %>% 
  summary
```

# Stool data

go to script 011 for selecting qualified stool samples (Note, the dtb used is **actual**)

# The Microbiome and p2d diet association using lmer and brm

go to script 035 (Also split out the nutritional drinks)

