---
title: "For Marcel's question"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

```{r}
library(tidyverse)
library(ggpubr)
```

- Does calorie intake (either total intake or % of needs met, as Peter calculates it) correlate with Blautia genus abundance, either in a simple correlation or in your fancier adjusted model?
- Does calorie intake correlate with stool alpha diversity, either in a simple correlation or in your fancier adjusted model?
- Does fiber intake correlate with Blautia abundance, either in a simple correlation or in your fancier adjusted model

## load the data

```{r}
dtb <- read_csv('../data/cleaned_diet_data/FINAL_97_with_code_all_foods_drt_with_EN_finalized.csv', col_types = cols(.default = col_guess(), Food_code = col_character())) 

# genus level counts 
cts <- read_csv('../data/cleaned_stool/ALL_stool_samples_genus_counts.csv')

meta <- read_csv('../data/cleaned_stool/all_samples_meta_p2d_fg9_updated.csv') %>% 
  mutate(timebin = cut_width(sdrt, 7, boundary=0, closed = 'left')) %>% 
  mutate(intensity = factor(intensity, levels = c('nonablative','reduced','ablative'))) 
```


## total caloric intake VS Blautia

```{r}
# the daily caloric intake :
all_daily_pt <- dtb %>%
  group_by(mrn, fdrt) %>%
  summarise(total_daily = sum(Calories_kcal))

# the previous two days average for each stool sample
stb <- read_csv('../data/cleaned_stool/selected_stool_samples_type_abx.csv') 

stb_pair <- stb %>%  
  select(mrn, sdrt) %>% 
  transmute(mrn = mrn,
            p1d = sdrt-1,
            p2d = sdrt-2) 

mean_p2d_cal <-  function(mrn_, p1d_, p2d_){
  df = all_daily_pt %>% 
    filter(mrn == mrn_) %>% 
    filter(fdrt %in% c(p1d_, p2d_  )) %>% 
    group_by(mrn) %>% 
    summarise(ave_cal = sum(total_daily)/2)
  return(df)
}

mean_p2d_df_cal <- pmap(stb_pair, function(mrn, p1d, p2d){
    mean_p2d_cal(mrn, p1d, p2d)
  }) %>% 
  set_names(stb %>% pull(sampleid)) %>% 
  bind_rows(.id = 'sampleid') 

# find the relab of the blautia genus
blautia <- cts %>% 
  filter(str_detect(genus, 'Blautia')) %>% 
  filter(sampleid %in% meta$sampleid) %>% 
  mutate(relablog = log10(relab + 2*10^-6)) %>% 
  select(sampleid,
         blautia_relab_log = relablog)

meta_blautia <- meta %>% 
  full_join(mean_p2d_df_cal) %>% 
  full_join(blautia)

# find the relab of the Enterococcus genus
Enterococcus <- cts %>% 
  filter(str_detect(genus, 'Enterococcus')) %>% 
  filter(sampleid %in% meta$sampleid) %>% 
  mutate(relablog = log10(relab + 2*10^-6)) %>% 
  dplyr::select(sampleid,
         Enterococcus_relab_log = relablog)
```

```{r}
# the modeling about total caloric intake VS Blautia
library(lmerTest) 
mod_cal <-  lmer(blautia_relab_log ~ 
               ave_cal +
               empirical +  
               intensity + 
                TPN +
               (1 | mrn) +
                (1 | timebin), REML = F, data = meta_blautia )
summary(mod_cal) 
```
```{r}
library(brms)  
library(ggmcmc)
model_brm <- brm(blautia_relab_log ~ 
               ave_cal +
                TPN +
               empirical + 
               intensity + 
               (1 | mrn) +
                (1 | timebin),  
              data = meta_blautia, 
              warmup = 1000, iter = 3000, 
              cores = 8, chains = 2, 
              seed = 123) 
plot(model_brm)

# looks very very weak
```

```{r}
# the version that has zero
meta_blautia %>% 
    ggscatter(x = 'ave_cal', y = 'blautia_relab_log', alpha = 0.7,
             xlab = 'Average of total daily caloric intake during previous two days',
             ylab = 'Blautia genus relative abundance log10 transformed',
               add = "reg.line",  # Add regressin line
           add.params = list(color = "blue", fill = "lightgray"), # Customize line
           conf.int = TRUE, # Add confidence interval
           cor.coef = TRUE, # Add correlation coefficient.
           cor.coeff.args = list(method = "spearman",  label.sep = "\n", label.y = -5)) 
  
meta_blautia %>% 
  distinct(mrn)
```

## caloric intake VS stool alpha diversity

```{r}
mod_cal_alpha <-  lmer(log(simpson_reciprocal) ~ 
               ave_cal +
               empirical +  
               intensity + 
                TPN +
               (1 | mrn) +
                (1 | timebin), REML = F, data = meta_blautia )
summary(mod_cal_alpha) 
```
```{r}
model_brm <- brm(log(simpson_reciprocal) ~  
               ave_cal +
                TPN +
               empirical + 
               intensity + 
               (1 | mrn) +
                (1 | timebin),  
              data = meta_blautia, 
              warmup = 1000, iter = 3000, 
              cores = 8, chains = 2, 
              seed = 123) 
plot(model_brm)
```

No it doesn't look like there is an association

```{r}
# meta_blautia %>%   
#   filter(empirical == 'FALSE') %>% 
#   ggscatter(x = 'ave_cal', y = 'simpson_reciprocal',
#             alpha = 0.7,
#             xlab = 'Average of total daily caloric intake during previous two days',
#             ylab = 'Simpson reciprocal on log10 scale',
#             #facet.by = 'empirical',
#              add = "reg.line",  # Add regressin line
#            add.params = list(color = "blue", fill = "lightgray"), # Customize line
#            conf.int = TRUE, # Add confidence interval
#            cor.coef = TRUE, # Add correlation coefficient.
#            cor.coeff.args = list(method = "pearson",  label.sep = "\n",  label.x = 2500) ) +
#   scale_y_log10()  
# 
# # how many stool samples and how many patients 
# no_empirical <- meta_blautia %>% 
#   filter(empirical == 'FALSE')
# 
# no_empirical %>% 
#   distinct(mrn)
```
oh wait, the scatter plot is doing a simple correlation and it looks like the correlation holds up but the model doesn't say so after adjusting other factors.

## fiber VS blautia

```{r}
# I need to calculate the fiber value per day per patient 
daily_fiber <- dtb %>% 
  group_by(mrn, fdrt) %>%
  summarise(total_daily_fiber = sum(Fibers_g))

# the previous two days average for each stool sample

mean_p2d_fiber <-  function(mrn_, p1d_, p2d_){
  df = daily_fiber %>% 
    filter(mrn == mrn_) %>% 
    filter(fdrt %in% c(p1d_, p2d_  )) %>% 
    group_by(mrn) %>% 
    summarise(ave_fiber = sum(total_daily_fiber)/2)
  return(df)
}

mean_p2d_df_fiber <- pmap(stb_pair, function(mrn, p1d, p2d){
    mean_p2d_fiber(mrn, p1d, p2d)
  }) %>% 
  set_names(stb %>% pull(sampleid)) %>% 
  bind_rows(.id = 'sampleid') 

meta_blautia_fiber <- meta_blautia %>% 
  full_join(mean_p2d_df_fiber)
```

```{r}
mod_blautia_fiber <-  lmer(blautia_relab_log ~ 
               ave_fiber +
               empirical +  
               intensity + 
                TPN +
               (1 | mrn) +
                (1 | timebin), REML = F, data = meta_blautia_fiber )
summary(mod_blautia_fiber) 
```
```{r}
# model_brm <- brm(blautia_relab_log ~  
#                ave_fiber +
#                 TPN +
#                empirical + 
#                intensity + 
#                (1 | mrn) +
#                 (1 | timebin),  
#               data = meta_blautia_fiber, 
#               warmup = 1000, iter = 3000, 
#               cores = 8, chains = 2, 
#               seed = 123) 
# plot(model_brm)
```

```{r}
meta_blautia_fiber %>% 
  mutate(ave_fiber_log = log10(ave_fiber + 0.01)) %>% 
    ggscatter(x = 'ave_fiber_log', y = 'blautia_relab_log', alpha = 0.7,
              xlab = 'Average of fiber intake during previous two days log10 transformed',
              ylab = 'Blautia genus relative abundance log10 transformed',
               add = "reg.line",  # Add regressin line
           add.params = list(color = "blue", fill = "lightgray"), # Customize line
           conf.int = TRUE, # Add confidence interval
           cor.coef = TRUE, # Add correlation coefficient.
           cor.coeff.args = list(method = "spearman",  label.y = -5, label.sep = "\n")) 

meta_blautia_fiber %>% 
  distinct(mrn)
```

## integrating the log of the relab of the Enterococcus

```{r}
combined <- meta_blautia_fiber %>% 
  full_join(Enterococcus)%>% 
  mutate(kcal = ave_cal/1000) %>% 
  mutate(ave_fiber_log = log10(ave_fiber + 0.01),
         simpson_reciprocal_log = log10(simpson_reciprocal)) %>% 
  mutate(blautia_relab_log = if_else(is.na(blautia_relab_log), 0, blautia_relab_log))

combined %>% 
  filter(ave_fiber < 35) %>% 
  ggscatter( x= 'ave_fiber', y = 'Enterococcus_relab_log',
               add = "reg.line",  # Add regressin line
           add.params = list(color = "blue", fill = "lightgray"), # Customize line
           conf.int = TRUE, # Add confidence interval
           cor.coef = TRUE, # Add correlation coefficient.
           cor.coeff.args = list(method = "spearman", 
                                 label.sep = "\n", 
                                 label.x.npc = "middle", 
                                 label.y.npc = "bottom",
                                 size = 5,
                                 color = 'black'))
```

## plotting 

```{r}
combined  %>% 
  dplyr::select(simpson_reciprocal_log, blautia_relab_log, Enterococcus_relab_log,  ave_fiber_log, kcal) %>% 
  gather('yaxis', 'value', simpson_reciprocal_log:Enterococcus_relab_log) %>% 
  mutate(yaxis = factor(yaxis, levels = c('simpson_reciprocal_log','blautia_relab_log','Enterococcus_relab_log'))) %>% 
  gather('xaxis', 'number', kcal:ave_fiber_log) %>% 
  mutate(xaxis = factor(xaxis, levels = c('kcal','ave_fiber'))) %>% 
  ggscatter(x = 'number', y = 'value', alpha = 0.2,xlab = '', ylab = '',size = 1.2, 
             add = "reg.line",  # Add regressin line
           add.params = list(color = "blue", fill = "lightgray"), # Customize line
           conf.int = TRUE, # Add confidence interval
           cor.coef = TRUE, # Add correlation coefficient.
           cor.coeff.args = list(method = "spearman", 
                                 label.sep = "\n", 
                                 label.x.npc = "left", 
                                 label.y.npc = "bottom",
                                 size = 2.5,
                                 color = 'black')) +
  facet_grid(yaxis ~ xaxis, scales = 'free') + 
  theme_bw(base_size = 11) +
  theme(aspect.ratio = 1,
        panel.grid.major = element_blank(), 
        strip.background = element_blank(),
        strip.text = element_blank(),
        panel.grid.minor = element_blank()
       ) 

ggsave('../figs/marcel_figures.pdf', width = 3, height = 4)  
```

```{r}
combined %>% 
  distinct(mrn)
```

