---
title: "Swap diet/fruit between extreme eaters"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
dat <- read_csv('../data/cleaned_stool/all_samples_meta_p2d_fg11.csv') %>% 
  mutate(timebin = cut_width(sdrt, 7, boundary=0, closed = 'left')) %>% 
  mutate(intensity = factor(intensity, levels = c('nonablative','reduced','ablative'))) 

dat %>% 
  select(-mrn) %>% 
  summary
```
```{r}
library(lmerTest)     
mod102 <-  lmer(log(simpson_reciprocal) ~ 
              fg_fruit+
               fg_meat+
               fg_milk+
               fg_oils+
                fg_egg+ 
                fg_grain+
                fg_sweets+  
                fg_legume+
                fg_veggie+
                fg_ndrink+
                fg_EN+
               empirical +  
               intensity + 
                TPN +
               (1 | mrn) +
                (1 | timebin), REML = F, data = dat )
```


## how this model predicts diversity for all patients  

```{r}
# the top 10 and bottom 10 fruit eaters
mean_fruit <- dat %>% 
  group_by(mrn) %>% 
  summarise(fruit_mean = mean(fg_fruit)) %>% 
  arrange(-fruit_mean) %>% 
  ungroup() %>% 
  pull(mrn)

dat %>% 
  ggboxplot(x = 'mrn', y = 'fg_fruit', add = 'jitter', order = mean_fruit) +
  geom_boxplot(outlier.shape = NA)  +
  ggsave('../figs/fruit_boxplot_mrn_mean.pdf', width = 10, height = 6)

N <- 10
top10 <- as.character(head(mean_fruit, N))
bottom10 <- as.character(tail(mean_fruit, N))

prediction_subset <- dat %>% 
  filter(mrn %in% c(top10, bottom10))
```


```{r}
# used all of the samples here so its for all patients 
predicted_ <- predict(mod102, newdata=dat, allow.new.levels=TRUE)

tibble(
  observed = dat$simpson_reciprocal,
  predicted = predicted_,
  mrn = dat$mrn
) %>% 
  ggscatter(x = 'observed', y = 'predicted', 
            #color = 'mrn',
            add = "reg.line",  # Add regressin line
           add.params = list(color = "blue", fill = "lightgray"), # Customize line
           conf.int = TRUE, # Add confidence interval
           cor.coef = TRUE, # Add correlation coefficient.
           cor.coeff.args = list(method = "pearson", label.x = 3, label.sep = "\n"),
           xlab = 'Observed simpson reciprocal',
           ylab = 'Predicted simpson reciprocal',
           title = 'Observed VS predicted diversity from LMEM (Pearson correlation)') +
  theme(legend.position = 'none')

```


For scoring this is enough


## swap fruit consumptions for high and low eaters 

what if low fruit eaters had eaten a lot of fruit or what if high fruit eaters had eaten less fruit, then compare the predictions with the real vs the swabbed diet. so the point is to see how the coeffs of fruit is in effects when you consider a patients amounts of fruit consumption

```{r}
tope <- dat %>% 
  filter(mrn %in% top10)

bote <- dat %>% 
  filter(mrn %in% bottom10)

# make sure they have same number of rows
bote__ <- bote[1:nrow(tope),]

high_eater_swab <- tope %>% 
  mutate(fg_fruit = bote__$fg_fruit)

low_eater_swab <- bote__ %>% 
  mutate(fg_fruit = tope$fg_fruit)

high_eater_swab_pred <- predict(mod102,newdata=high_eater_swab,allow.new.levels=TRUE)
high_eater_real_pred <- predict(mod102,newdata=tope,allow.new.levels=TRUE)

high_eater_swab_both <- tibble(
  true = tope$simpson_reciprocal,
  swab_predict = high_eater_swab_pred,
  real_predict = high_eater_real_pred
) %>% 
  gather('type', 'value', swab_predict:real_predict) %>% 
  mutate(value =  exp(value))

high_eater_swab_both %>% 
  ggscatter(x = 'true', y = 'value', color = 'type', palette = 'lancet',
            xlab = 'Observed diversity', ylab = 'Predicted diversity',
            title = 'Top eaters')  +
  geom_line(aes(group = true), color = "grey", alpha = 0.8) +
  theme_cleveland()
  
```

```{r}
low_eater_swab_pred <- predict(mod102,newdata=low_eater_swab,allow.new.levels=TRUE)
low_eater_real_pred <- predict(mod102,newdata=bote__,allow.new.levels=TRUE)

low_eater_swab_both <- tibble(
  true = bote__$simpson_reciprocal,
  swab_predict = low_eater_swab_pred,
  real_predict = low_eater_real_pred
) %>% 
  gather('type', 'value', swab_predict:real_predict) %>% 
  mutate(value =  exp(value))
 

low_eater_swab_both %>% 
  ggscatter(x = 'true', y = 'value', color = 'type', palette = 'lancet',
            xlab = 'Observed diversity', ylab = 'Predicted diversity',
            title = 'Bottom eaters')  +
  geom_line(aes(group = true), color = "grey", alpha = 0.8)
```

## not only swapping the fruit but also swapping the other food group comsumptions  

```{r}
high_eater_swab_all <- tope %>% 
  mutate(fg_egg = bote__$fg_egg,
         fg_EN = bote__$fg_EN,
         fg_fruit = bote__$fg_fruit,
         fg_grain = bote__$fg_grain,
         fg_legume = bote__$fg_legume,
         fg_meat = bote__$fg_meat,
         fg_milk = bote__$fg_milk,
         fg_ndrink = bote__$fg_ndrink,
         fg_oils = bote__$fg_oils,
         fg_sweets = bote__$fg_sweets,
         fg_veggie = bote__$fg_veggie)

low_eater_swab_all <- bote__ %>% 
  mutate(fg_egg = tope$fg_egg,
         fg_EN = tope$fg_EN,
         fg_fruit = tope$fg_fruit,
         fg_grain = tope$fg_grain,
         fg_legume = tope$fg_legume,
         fg_meat = tope$fg_meat,
         fg_milk = tope$fg_milk,
         fg_ndrink = tope$fg_ndrink,
         fg_oils = tope$fg_oils,
         fg_sweets = tope$fg_sweets,
         fg_veggie = tope$fg_veggie) 
```

```{r}
high_eater_swab_pred <- predict(mod102,newdata=high_eater_swab_all,allow.new.levels=TRUE)
high_eater_real_pred <- predict(mod102,newdata=tope,allow.new.levels=TRUE)

high_eater_swab_both <- tibble(
  true = tope$simpson_reciprocal,
  swab_predict = high_eater_swab_pred,
  real_predict = high_eater_real_pred
) %>% 
  gather('type', 'value', swab_predict:real_predict) %>% 
  mutate(value =  exp(value))

high_eater_swab_both %>% 
  ggscatter(x = 'true', y = 'value', color = 'type', palette = 'lancet',
            xlab = 'Observed diversity', ylab = 'Predicted diversity',
            title = 'Top eaters')  +
  geom_line(aes(group = true), color = "grey", alpha = 0.8) +
  theme_cleveland()
```

```{r}
low_eater_swab_pred <- predict(mod102,newdata=low_eater_swab_all,allow.new.levels=TRUE)
low_eater_real_pred <- predict(mod102,newdata=bote__,allow.new.levels=TRUE)

low_eater_swab_both <- tibble(
  true = bote__$simpson_reciprocal,
  swab_predict = low_eater_swab_pred,
  real_predict = low_eater_real_pred
) %>% 
  gather('type', 'value', swab_predict:real_predict) %>% 
  mutate(value =  exp(value))
 

low_eater_swab_both %>% 
  ggscatter(x = 'true', y = 'value', color = 'type', palette = 'lancet',
            xlab = 'Observed diversity', ylab = 'Predicted diversity',
            title = 'Bottom eaters')  +
  geom_line(aes(group = true), color = "grey", alpha = 0.8)+
  theme_cleveland()
```

