---
title: "Domination in patient time course"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr) 
```

# Selecting samples for shotgun sequencing (the ones that have the domination happening )

1.	Peruse the 92 patients and look at which dominations they have
2.	Identify patients have any domination in the diet-analysis sample set (even >=10% abundance) from any family Enterococcaceae or family Enterobacteriaceae)
3.	Tsoni Print the Ying shiny timecourses for those patients
4.	Identify which ones have qPCRs available (can request more))

```{r}
meta <- read_csv('../data/cleaned_stool/all_samples_meta_p2d_fg9_updated.csv')

# the current 92 mrns being use
fam <- read_csv('../data/cleaned_stool/ALL_stool_samples_family_counts.csv') %>% 
  filter(sampleid %in% meta$sampleid) %>% 
  mutate(family = str_extract(taxa_family, 'f__.+$')) %>% 
  mutate(family = str_replace(family, 'f__',''))

# find the two families relab for every stool sample
fam2 <- fam %>% 
  filter(family %in% c('Enterococcaceae','Enterobacteriaceae')) %>% 
  select(sampleid, family, relab) %>% 
  spread('family','relab')

meta <- read_csv('../data/cleaned_stool/all_samples_meta_p2d_fg9_updated.csv')  %>% 
  full_join(fam2)
```

```{r}
# the patients that have the domination samples occur
target_fam <- c('Enterococcaceae','Enterobacteriaceae')

target_fam %>% 
  map(function(var){

    mrns <- fam %>% 
      filter(family == var) %>% 
      filter(relab > 0.1) %>% 
      left_join(meta %>% select(mrn, sampleid, sdrt)) %>% 
      distinct(mrn) %>% 
      pull(mrn)
    
    # visualize all the samples for these patients
    mrn_samp <- meta %>% 
      filter(mrn %in% mrns)
    
    mrn_samp %>% 
      ggdotchart(x = 'sdrt', y = var,  facet.by = 'mrn',  add = "segments",
                       sorting = 'none', font.label = list(size = 5, face = "plain"),
                 title = str_glue('{var}')) +
            theme(axis.text.x = element_text(size=5, angle=90, hjust=1)) +
            geom_hline(yintercept = 0.1, linetype = 'dashed', color = 'red') +
            ggsave(str_glue('../figs/family_2_dom_{var}.jpg'), width = 15, height = 7)
  })
```

```{r}
# to see the overlap between these two groups of patients
target_fam <- c('Enterococcaceae','b')

fam_patients <- target_fam %>% 
  map(function(var){

    mrns <- fam %>% 
      filter(family == var) %>% 
      filter(relab > 0.1) %>% 
      left_join(meta %>% select(mrn, sampleid, sdrt)) %>% 
      distinct(mrn) %>% 
      pull(mrn)
  }) 

length(intersect(fam_patients[[1]], fam_patients[[2]]))
length(fam_patients[[1]])
length(fam_patients[[2]])
# the 2 patients that were not in the overlap  
setdiff(fam_patients[[2]], fam_patients[[1]]) # these two can be ignored doesn't reach the requirement 

# how many patients I selected
length(fam_patients[[1]])
```
```{r}
# the current stool sample table 
samps <- meta %>% 
      filter(mrn %in% fam_patients[[1]])

stb_ <- read_csv('../data/cleaned_stool/selected_stool_samples_type_abx.csv') %>% 
  filter(sampleid %in% samps$sampleid)
```

```{r}
# to get whether those stool samples have at least one diet data in the previous three days 
qual_stool_df <- stb_ %>% 
  mutate(p1d = sdrt-1,
         p2d = sdrt-2,
         p3d = sdrt-3) %>% 
  select(mrn, p1d, p2d, p3d)

# use the no fake diet counts table and get the stool samples that should be removed -- not even one day
dtb <- read_csv('../data/cleaned_diet_data/FINAL_97_with_code_all_foods_drt_with_EN_finalized.csv')

no_diet_d_to_be_rm_stool_s <- function(mrn_, p1d_, p2d_, p3d_){
  df = dtb %>% 
    filter(mrn == mrn_) %>% 
    filter(fdrt == p1d_ | fdrt == p2d_  | fdrt == p3d_) %>% 
    nrow()
  return(df)
}

qual_stool<- pmap(qual_stool_df, function(mrn, p1d, p2d, p3d){
    no_diet_d_to_be_rm_stool_s(mrn, p1d, p2d, p3d)
  }) %>% 
  set_names(stb_ %>% pull(sampleid)) %>% 
  bind_rows() %>% 
  gather(key = 'sampleid', value = 'num_d_records') %>% 
  arrange(num_d_records) 

nrow(qual_stool[qual_stool$num_d_records> 0,]) == nrow(qual_stool)
# so all of these stool samples have at least one diet record in the previous three days 
# aka qualified 
```
  
**Need to remove the 11 that have already been sequenced **
**Need also remove the ones that already been sequenced by Oriana batch**

**Remember that very new diet patient that wasn't included before**
I can now be assure that the 8 samples from that one patient didn't have any dominance happening so good!!!

# selecting samples for shotgun sequencing 

```{r}
# The 16 patients that will all be sequenced 
all16 <- intersect(fam_patients[[1]], fam_patients[[2]])

all16_samps <- meta %>%    
  filter(mrn %in% all16)  
```


```{r}
# the remaining patients that have enterococcaeae domination occuring 
enter_only <- setdiff(fam_patients[[1]], fam_patients[[2]])

# find the patients that have no non-dominating samples before a dominating one
# if the earliest sample has Enterococcaceae relab > 0.1 ....
enter_rm <- meta %>%  
  filter(mrn %in% enter_only) %>% 
  select(sampleid, mrn, sdrt, Enterococcaceae) %>% 
  arrange(mrn, sdrt) %>% 
  group_by(mrn) %>% 
  slice(1) %>% 
  filter(Enterococcaceae > 0.1) %>% 
  pull(mrn)

# remove the above patients  
enter_only_updated <- setdiff(enter_only, enter_rm)
```

```{r}
# find all the non-dominating samples until the first dominating one for the above patients 
enter_dom_samples <- meta %>%  
  filter(mrn %in% enter_only_updated)

# sequence all the non-dominating ones up till the first dominating one
# first visualize 
var <- 'Enterococcaceae'
enter_dom_samples %>% 
      ggdotchart(x = 'sdrt', y = var,  facet.by = 'mrn',  add = "segments",
                       sorting = 'none', font.label = list(size = 5, face = "plain"),
                 title = str_glue('{var}')) +
            theme(axis.text.x = element_text(size=5, angle=90, hjust=1)) +
            geom_hline(yintercept = 0.1, linetype = 'dashed', color = 'red') +
            ggsave(str_glue('../figs/family_dom_{var}.jpg'), width = 15, height = 7)
```
```{r}
# find for each patient the day of the first dominance
enter_dom_samples_day <- enter_dom_samples %>% 
  filter(Enterococcaceae > 0.1) %>% 
  arrange(mrn, sdrt) %>% 
  group_by(mrn) %>% 
  slice(1) %>% 
  select(mrn, sdrt)

# use some loop to find for each patient stool sample up till that day 
select_up_until <- function(mrn_, sdrt_){
  enter_dom_samples %>% 
    filter(mrn == mrn_) %>% 
    filter(sdrt <= sdrt_)
}

select_up_until_res <- pmap(enter_dom_samples_day, function(mrn, sdrt){
    select_up_until(mrn, sdrt)
  }) %>% 
  set_names(enter_dom_samples_day %>%  pull(mrn)) %>% 
  bind_rows(.id = 'mrn') %>% 
  mutate(mrn = as.numeric(mrn))
 
```

```{r}
# the current all (16 pt that have all to be sequneced and then some that have non-dom samples before a dom one)
current <- bind_rows(
  all16_samps, select_up_until_res
)
```

```{r}
# remove the ones that have already been sequenced
library(vdbR)
connect_database('~/dbConfig.txt')
get_table_from_database('shotgun_lookup_ad')
length(intersect(shotgun_lookup_ad$sampleid, current$sampleid))
# any one included in the Oriana list?
# the current diet samples 

O500 <- readxl::read_excel('../data/ShotgunSeqSubmissionandQC.xlsx') %>% 
  rename_all(~ gsub(" |:", "_", .))
O500_seq <- O500$Samples_Submitted_

fail <- O500 %>% 
  filter(Samples_that_did_not_pass_iGO_QC_!='NA') %>% 
  pull(Samples_that_did_not_pass_iGO_QC_)

O500_ <- setdiff(O500_seq, O500_no)
length(intersect(O500_, meta$sampleid))

length(intersect(O500_, current$sampleid))
O <- intersect(O500_, current$sampleid)

low <-  O500 %>% 
  filter(Samples_that_iGO_will_attempted_!='NA') %>% 
  pull(Samples_that_iGO_will_attempted_)

intersect(O, low)

already <- union(shotgun_lookup_ad$sampleid, O500_seq)
  
length(intersect(already, current$sampleid))
length(unique(intersect(already, current$sampleid)))
overlap <- unique(intersect(already, current$sampleid))
no <- union(low, fail)
length(intersect(overlap, no))
# ok i made sure only 20 will be removed from the current list (3 already in the db and 17 going to be returned. and none of the 17 is low quality or failed qc)
```

```{r}
# the final list
final <- current %>% 
  filter(!sampleid %in% overlap) 

final %>% 
  select(sampleid) %>% 
  write_csv('../data/growth/shotgun_request_diet_sampleid_Angel.csv')
```

```{r}
# 329 is a bit too much might select 1/3 or 1/2 of those
# look at the 16 patients that have both domination happening
both <- all16_samps %>% 
  distinct(mrn) %>% 
  pull(mrn)

mrn_samp_df <- meta %>% 
      filter(mrn %in% both) %>% 
  mutate(mrn = str_glue('P{as.numeric(as.factor(mrn))}'))
    
target_fam %>% 
  map(function(var){

    mrn_samp_df %>% 
      ggdotchart(x = 'sdrt', y = var,  facet.by = 'mrn',  add = "segments",
                       sorting = 'none', font.label = list(size = 5, face = "plain"),
                 ylab = str_glue('{var} family relative abundance'),
                 xlab = 'Day relative to transplant',
                 title = str_glue('Patients who had {var} dominance incidents (relab > 10%)')) +
            theme(axis.text.x = element_text(size=5, angle=90, hjust=1)) +
            geom_hline(yintercept = 0.1, linetype = 'dashed', color = 'red') +
            ggsave(str_glue('../figs/bothpatients_family_2_dom_{var}.pdf'), width = 15, height = 7)
  })
```


```{r}
# the final final list of 197 samples
last <- all16_samps 

last %>% 
  write_csv('../data/growth/shotgun_request_meta_full.csv')

last %>% 
  select(sampleid) %>% 
  write_csv('../data/growth/shotgun_request_nutrition_sampleid_Angel_final.csv')
```


```{r}
# mrns for the blood stream infections
meta %>% 
  distinct(mrn)  %>% 
  write_csv('../data/mrns_in_nutrition_for_BSI.csv')
```

