---
title: "The model re-done with Bayesian"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

```{r}
library(tidyverse)
library(brms)  
library(vdbR)
library(ggpubr)
dtb <- read_csv('../data/cleaned_diet_data/FINAL_97_with_code_all_foods_drt_with_EN_finalized.csv')
```

## Loading meta data table

```{r load_meta}
meta <- read_csv('../data/cleaned_stool/all_samples_meta_p2d_fg9_updated.csv') %>% 
  mutate(timebin = cut_width(sdrt, 7, boundary=0, closed = 'left')) %>% 
  mutate(intensity = factor(intensity, levels = c('nonablative','reduced','ablative'))) %>% 
  mutate(mrn = factor(mrn)) %>% 
  mutate(fg_egg = fg_egg/100,
         fg_fruit = fg_fruit/100,
         fg_grain = fg_grain/100,
         fg_legume = fg_legume/100,
         fg_meat = fg_meat/100,
         fg_milk = fg_milk/100,
         fg_oils = fg_oils/100,
         fg_sweets = fg_sweets/100,
         fg_veggie = fg_veggie/100) %>% 
  mutate(empiT = if_else(empirical == 'TRUE', 1, 0),
         empiF = if_else(empirical == 'FALSE', 1, 0),
         tpnT = if_else(TPN == 'TRUE', 1, 0),
         tpnF = if_else(TPN == 'FALSE', 1, 0),
         enT = if_else(TPN == 'TRUE', 1, 0),
         enF = if_else(TPN == 'FALSE', 1, 0),
         inten_non = if_else(intensity == 'nonablative', 1, 0),
         inten_ab = if_else(intensity == 'ablative', 1, 0),
         inten_re = if_else(intensity == 'reduced', 1, 0))
```

## Priors 

```{r get_prior}
# what priors I cann set 
model_formular <- log(simpson_reciprocal)~ 1 +
                fg_fruit+
               fg_meat+
               fg_milk+
               fg_oils+
                fg_egg+ 
                fg_grain+
                fg_sweets+  
                fg_legume+
                fg_veggie+
               inten_non + inten_ab + inten_re +
               empiT +  empiF+
                tpnT +tpnF+
                enT +enF+
               (1 | mrn) +
                (1 | timebin)
get_prior( model_formular,  
              data = meta)
```

### Setting priors

```{r set_prior}
priors <- c(# for the food group variables
            prior(normal(0, 1), class = 'b', coef = "fg_egg"),
            prior(normal(0, 1), class = 'b', coef = "fg_fruit"),
            prior(normal(0, 1), class = 'b', coef = "fg_grain"),
            prior(normal(0, 1), class = 'b', coef = "fg_legume"),
            prior(normal(0, 1), class = 'b', coef = "fg_meat"),
            prior(normal(0, 1), class = 'b', coef = "fg_milk"),
            prior(normal(0, 1), class = 'b', coef = "fg_oils"),
            prior(normal(0, 1), class = 'b', coef = "fg_sweets"),
            prior(normal(0, 1), class = 'b', coef = "fg_veggie"),
            # for the TPN 
            prior(normal(0, 0.1), class = 'b', coef = "tpnF"),
            prior(normal(0, 0.1), class = 'b', coef = "tpnT"),
            # for the EN
            prior(normal(0, 0.1), class = 'b', coef = "enF"),
            prior(normal(0, 0.1), class = 'b', coef = "enT"),
            # for the empirical 
            prior(normal(0, 0.5), class = 'b', coef = "empiF"),
            prior(normal(0, 0.5), class = 'b', coef = "empiT"),
            # for the intensity 
            prior(normal(0, 0.1), class = 'b', coef = "inten_re"),
            prior(normal(0, 0.1), class = 'b', coef = "inten_ab"),
            prior(normal(0, 0.1), class = 'b', coef = "inten_non"),
            prior(normal(2, 0.1), class = 'Intercept'))

# samples are drawn solely from the priors ignoring the likelihood
model_prior <- brm( model_formular,  
              data = meta, 
              warmup = 1000, iter = 3000, 
              prior = priors,
              cores = 16, 
              chains = 2, 
              seed = 123, sample_prior = T) 

# look at the priors set for the model
prior_summary(model_prior)

# the model in stan code 
#stancode(model_prior)

# actually draw samples from the priors 
samples_prior <- prior_samples(model_prior)

```

```{r set_prior}
# the mean food group intake from our data 
ave_fg <- meta %>% 
  select(starts_with('fg')) %>% 
  summarise_all(funs(mean)) %>% 
  gather

mean_fg <-  meta %>% 
  select(starts_with('fg')) %>% 
  summarise_all(funs(mean))

max_fg <- meta %>% 
  select(starts_with('fg')) %>% 
  summarise_all(funs(max))

min_fg <- meta %>% 
  select(starts_with('fg')) %>% 
  summarise_all(funs(min))

# the max fruit intake 
max_fruit <- max(meta$fg_fruit)

# the min of the fruit intake
min_fruit <- min(meta$fg_fruit) 

max_log_div <- max(log(meta$simpson_reciprocal)) 

```


```{r}
# create a df with max and min and ave in 9 groups
# create a df with the corresponding value that the samples prior coeff can multiply
max_ <- ave_fg %>% 
  rename(mean_ = value) %>% 
  mutate(max_fg_egg = if_else(key == 'fg_egg', max_fg$fg_egg, mean_),
         max_fg_fruit = if_else(key == 'fg_fruit', max_fg$fg_fruit, mean_),
         max_fg_grain = if_else(key == 'fg_grain', max_fg$fg_grain, mean_),
         max_fg_legume = if_else(key == 'fg_legume', max_fg$fg_legume, mean_),
         max_fg_meat = if_else(key == 'fg_meat', max_fg$fg_meat, mean_),
         max_fg_milk = if_else(key == 'fg_milk', max_fg$fg_milk, mean_),
         max_fg_oils = if_else(key == 'fg_oils', max_fg$fg_oils, mean_),
         max_fg_sweets = if_else(key == 'fg_sweets', max_fg$fg_sweets, mean_),
         max_fg_veggie = if_else(key == 'fg_veggie', max_fg$fg_veggie, mean_)) %>% 
  select(-mean_) %>% 
  mutate(key = if_else(str_detect(key, '^fg'), str_glue('b_{key}'), key))  %>% 
  arrange(key)

prior_coeffs <- samples_prior %>% 
  select(Intercept, b_inten_re, b_empiF, b_tpnF, b_enF, starts_with('b_fg')) %>% 
  gather('term','prior_coeff') %>% 
  split(.$term)
  
values <- max_ %>% 
  select(-key) %>% 
  as.list()

values <- values %>% 
  map(~ set_names(x = ., max_$key))

# calculate the resulted prior log div 
vals <- names(values)

max_each_fg <- vals  %>% 
  set_names(vals) %>% 
  map(function(val) {
    res = prior_coeffs$b_empiF$prior_coeff  +
   prior_coeffs$b_enF$prior_coeff  +
   prior_coeffs$b_fg_egg$prior_coeff * values %>% pluck(val) %>% pluck('b_fg_egg') +
   prior_coeffs$b_fg_fruit$prior_coeff * values %>% pluck(val) %>% pluck('b_fg_fruit') +
   prior_coeffs$b_fg_grain$prior_coeff * values %>% pluck(val) %>% pluck('b_fg_grain') +
   prior_coeffs$b_fg_legume$prior_coeff * values %>% pluck(val) %>% pluck('b_fg_legume') +
   prior_coeffs$b_fg_meat$prior_coeff * values %>% pluck(val) %>% pluck('b_fg_meat') +
   prior_coeffs$b_fg_milk$prior_coeff * values %>% pluck(val) %>% pluck('b_fg_milk') +
   prior_coeffs$b_fg_oils$prior_coeff * values %>% pluck(val) %>% pluck('b_fg_oils') +
   prior_coeffs$b_fg_sweets$prior_coeff * values %>% pluck(val) %>% pluck('b_fg_sweets') +
   prior_coeffs$b_fg_veggie$prior_coeff * values %>% pluck(val) %>% pluck('b_fg_veggie') +
   prior_coeffs$b_inten_re$prior_coeff  +
   prior_coeffs$b_tpnF$prior_coeff +
   prior_coeffs$Intercept$prior_coeff 
  }) %>% 
  bind_rows(.id = 'grp') %>% 
  gather('grp', 'prior_res')

```


```{r plot_prior}
# combinne the max and min and ave together and plot
ave_each_fg <- samples_prior %>% 
  mutate(mean_fg = Intercept + 
           b_fg_fruit*mean_fg$fg_fruit +
           b_fg_meat*mean_fg$fg_meat + 
           b_fg_milk*mean_fg$fg_milk + 
           b_fg_oils*mean_fg$fg_oils + 
           b_fg_egg*mean_fg$fg_egg + 
           b_fg_grain*mean_fg$fg_grain +
           b_fg_sweets*mean_fg$fg_sweets +
           b_fg_legume*mean_fg$fg_legume +
           b_fg_veggie*mean_fg$fg_veggie +
           b_inten_re +
           b_empiF +
           b_tpnF +
           b_enF) %>% 
  select(mean_fg) %>% 
  mutate(grp = 'ave_each') %>% 
  rename(prior_res = mean_fg) %>% 
  select(grp, prior_res)

prior_all <- bind_rows(ave_each_fg, max_each_fg)

prior_all %>%  write_csv('../figs/paper/data/prior_ave_max_min_res_68.csv')

grps <- prior_all %>% 
  distinct(grp)

fg_cha <- meta %>% 
  select(starts_with('fg')) %>% 
  colnames()

prior_check_list <- fg_cha %>% 
  set_names(fg_cha) %>% 
  map(function(char){
    prior_all %>% 
      filter(str_detect(grp, str_glue('ave_each|{char}'))) %>% 
      gghistogram(x = 'prior_res', fill = 'grp', color = 'white', palette = 'lancet', xlab = 'log(diversity)',
                  add_density = TRUE, bins = 30,
                  title = str_glue('Prior check {char}'))
    })

g <- cowplot::plot_grid(prior_check_list[[1]],prior_check_list[[2]],prior_check_list[[3]],prior_check_list[[4]],prior_check_list[[5]],prior_check_list[[6]],prior_check_list[[7]],prior_check_list[[8]],prior_check_list[[9]],
    ncol = 3, nrow= 3, 
    o = 'hv',
    #labels = 'AUTO',
    axis = 'bltr') 

ggsave('../figs/paper/prior_check_9fg.pdf', width = 10, height = 10, dpi = 300, plot = g)

```





## Posterior

```{r}
model_brm <- brm( log(simpson_reciprocal) ~ 
                fg_fruit+
               fg_meat+
               fg_milk+
               fg_oils+
                fg_egg+ 
                fg_grain+
                fg_sweets+  
                fg_legume+
                fg_veggie+
               empirical +  
               intensity + 
                TPN +
                EN +
               (1 | mrn) +
                (1 | timebin),  
              data = meta, 
              warmup = 1000, iter = 3000, 
              prior = priors,
              cores = 16, chains = 2, 
              seed = 123) 

summary(model_brm)


# extract the coefficients table
posterior_model <- posterior_samples(model_brm) %>% 
  select(starts_with('b')) %>% 
  #select(-b_Intercept) %>% 
  gather('item', 'value') %>% 
  group_by(item) %>% 
  mutate(meanperitem = mean(value))%>%
  ungroup() 

posterior_model %>% 
  write_csv('../data/brms_model_var_res_updated.csv')
# 
# get_prior(log(simpson_reciprocal)~ 
#                 fg_fruit+
#                fg_meat+
#                fg_milk+
#                fg_oils+
#                 fg_egg+ 
#                 fg_grain+
#                 fg_sweets+  
#                 fg_legume+
#                 fg_veggie+
#                empirical +  
#                intensity + 
#                 TPN +
#                 EN +
#                (1 | mrn) +
#                 (1 | timebin),  
#               data = meta)
```

### Extract the posterior samples 

```{r}
# extract posterior samples for the coeff
post_samples  <- posterior_samples(model_brm, "^b")

# sample posterior coefficients and multiply with the mean diet
# sample posterior coefficients and multiply with the mean diet changed so that fruits are the max fruit values

post_2 <- post_samples %>% 
  mutate(mean_post = b_Intercept + 
           b_fg_fruit*mean_fg$fg_fruit +
           b_fg_meat*mean_fg$fg_meat + 
           b_fg_milk*mean_fg$fg_milk + 
           b_fg_oils*mean_fg$fg_oils + 
           b_fg_egg*mean_fg$fg_egg + 
           b_fg_grain*mean_fg$fg_grain +
           b_fg_sweets*mean_fg$fg_sweets +
           b_fg_legume*mean_fg$fg_legume +
           b_fg_veggie*mean_fg$fg_veggie,
         
         max_fruit_post = b_Intercept + 
           b_fg_fruit*max_fruit +
           b_fg_meat*mean_fg$fg_meat + 
           b_fg_milk*mean_fg$fg_milk + 
           b_fg_oils*mean_fg$fg_oils + 
           b_fg_egg*mean_fg$fg_egg + 
           b_fg_grain*mean_fg$fg_grain +
           b_fg_sweets*mean_fg$fg_sweets +
           b_fg_legume*mean_fg$fg_legume +
           b_fg_veggie*mean_fg$fg_veggie,
         
        min_fruit_post = b_Intercept + 
           b_fg_fruit*min_fruit +
           b_fg_meat*mean_fg$fg_meat + 
           b_fg_milk*mean_fg$fg_milk + 
           b_fg_oils*mean_fg$fg_oils + 
           b_fg_egg*mean_fg$fg_egg + 
           b_fg_grain*mean_fg$fg_grain +
           b_fg_sweets*mean_fg$fg_sweets +
           b_fg_legume*mean_fg$fg_legume +
           b_fg_veggie*mean_fg$fg_veggie ) %>% 
  select(mean_post, max_fruit_post, min_fruit_post) %>% 
  gather()

post_2 %>% 
  gghistogram(x = 'value', fill = 'key', color = 'white', palette = 'lancet', add_density = T,
              xlab = 'log(diversity)',
              #add_density = TRUE,  
              rug = TRUE,
              title = 'Post check\nmax_fruit(87) with others average;mean_post:all groups average(11)\nmin_fruit:(0)')
```

another thing to do is for both prior and posterior checks: use the ACTUAL diet from the 5 least fruit rich, and 5 most fruit rich diets.

something like :
for i in 5:
    top = topdiets[i]
    least = leastfruitdiets[i]
    coeffs = sample_posterior(n=1000) // sample_prior
    topdietdiversity_posterior = predict(top, coeffs)
    leastdietdiversity_posterior = predict(least,coeffs)
plot(topdietsdiversity_posterior) // prior 

```{r}
# the top and bottom fruit eaters diet 
set.seed(123)
N <- 5
N_sample <- 1000
topdiet <- meta %>% 
  top_n(n = N, wt = fg_fruit) %>% 
  mutate(TPNTRUE = if_else(TPN == TRUE, 1, 0),
         ENTRUE = if_else(EN == TRUE, 1, 0),
         empiricalTRUE = if_else(empirical == TRUE, 1, 0),
         intensityreduced = if_else(intensity == 'reduced', 1, 0),
         intensityablative = if_else(intensity == 'ablative', 1, 0)) %>% 
  select(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie)

botdiet <- meta %>% 
  filter(fg_fruit == 0) %>% 
  slice_sample(n = 5) %>% 
  mutate(TPNTRUE = if_else(TPN == TRUE, 1, 0),
         ENTRUE = if_else(EN == TRUE, 1, 0),
         empiricalTRUE = if_else(empirical == TRUE, 1, 0),
         intensityreduced = if_else(intensity == 'reduced', 1, 0),
         intensityablative = if_else(intensity == 'ablative', 1, 0))%>% 
  select(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie)


per_diet_post_pred <- function(fg_egg_, fg_fruit_, fg_grain_, fg_legume_, fg_meat_, fg_milk_, fg_oils_, 
                               fg_sweets_, fg_veggie_) {
  post_samp <- post_samples %>% 
    sample_n(size = N_sample, replace = F)
  
  ret = pmap(post_samp, function(b_Intercept, b_fg_fruit, b_fg_meat, b_fg_milk, b_fg_oils, b_fg_egg,
                                 b_fg_grain, b_fg_sweets, b_fg_legume, b_fg_veggie) {
    b_Intercept + b_fg_fruit*fg_fruit_ + b_fg_meat*fg_meat_ + b_fg_milk*fg_milk_ + b_fg_oils*fg_oils_ +
      b_fg_egg*fg_egg_ + b_fg_grain*fg_grain_ + b_fg_sweets*fg_sweets_ + b_fg_legume*fg_legume_ + 
      b_fg_veggie*fg_veggie_ 
  }) %>% 
    set_names(seq(1, N_sample)) %>% 
    bind_rows() %>% 
    gather()
  return(ret)
                               }


res_top <- pmap(topdiet, function(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie){
  per_diet_post_pred(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie)
}) %>% 
  set_names(paste('D', seq(1, nrow(topdiet)), sep = '')) %>% 
  bind_rows(.id = 'diet')


res_bot <- pmap(botdiet, function(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie){
  per_diet_post_pred(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie)
}) %>% 
  set_names(paste('D', seq(1, nrow(topdiet)), sep = '')) %>% 
  bind_rows(.id = 'diet')


# plot them together in one 
bind_rows(
  res_top %>% mutate(grp = 'top'),
  res_bot %>% mutate(grp = 'bot')
) %>% 
  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white', add = 'mean',
                  xlab = 'Predicted log diversity', ylab = 'Probability density') +
  ggsave('../figs/predicted_logdiversi_top_tybottom_5.pdf', width = 7, height = 5)
```

