---
title: "The model re-done with Bayesian"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

```{r}
library(tidyverse)
library(brms)  
library(vdbR)
library(ggpubr)
```

## Loading meta data table

```{r}
meta <- read_csv('../data/cleaned_stool/all_samples_meta_p2d_fg9_updated.csv') %>% 
  mutate(timebin = cut_width(sdrt, 7, boundary=0, closed = 'left')) %>% 
  mutate(intensity = factor(intensity, levels = c('nonablative','reduced','ablative'))) %>% 
  mutate(mrn = factor(mrn)) %>% 
  mutate(fg_egg = fg_egg/100,
         fg_fruit = fg_fruit/100,
         fg_grain = fg_grain/100,
         fg_legume = fg_legume/100,
         fg_meat = fg_meat/100,
         fg_milk = fg_milk/100,
         fg_oils = fg_oils/100,
         fg_sweets = fg_sweets/100,
         fg_veggie = fg_veggie/100)

```

## Priors 

### Setting priors

```{r}
priors <- c(prior(normal(0, 0.5), class = 'b'),   
            prior(normal(0, 1), class = 'b', coef = "fg_egg"),
            prior(normal(0, 1), class = 'b', coef = "fg_fruit"),
            prior(normal(0, 1), class = 'b', coef = "fg_grain"),
            prior(normal(0, 1), class = 'b', coef = "fg_legume"),
            prior(normal(0, 1), class = 'b', coef = "fg_meat"),
            prior(normal(0, 1), class = 'b', coef = "fg_milk"),
            prior(normal(0, 1), class = 'b', coef = "fg_oils"),
            prior(normal(0, 1), class = 'b', coef = "fg_sweets"),
            prior(normal(0, 1), class = 'b', coef = "fg_veggie"),
            prior(normal(0, 0.1), class = 'b', coef = "TPNTRUE"),
            prior(normal(0, 0.1), class = 'b', coef = "ENTRUE"),
            prior(normal(0, 0.1), class = 'b', coef = "empiricalTRUE"),
            prior(normal(0, 0.1), class = 'b', coef = "intensityreduced"),
            prior(normal(0, 0.1), class = 'b', coef = "intensityablative"),
            prior(normal(2, 0.5), class = 'Intercept'))
```

### Draw samples only from the priors

```{r}
# samples are drawn solely from the priors ignoring the likelihood
model_prior <- brm( log(simpson_reciprocal)~ 
                fg_fruit+
               fg_meat+
               fg_milk+
               fg_oils+
                fg_egg+ 
                fg_grain+
                fg_sweets+  
                fg_legume+
                fg_veggie+
               empirical +  
               intensity + 
                TPN +
                EN +
               (1 | mrn) +
                (1 | timebin),  
              data = meta, 
              warmup = 1000, iter = 3000, 
              prior = priors,
              cores = 16, 
              chains = 2, 
              seed = 123, sample_prior = T) 

# look at the priors set for the model
prior_summary(model_prior)

# the model in stan code 
#stancode(model_prior)

# actually draw samples from the priors 
samples_prior <- prior_samples(model_prior)

#plot(hypothesis(model_prior, "fg_fruit < 0"))
#plot(hypothesis(model_prior, "empiricalTRUE < 0"))

# the mean food group intake from our data 
mean_fg <- meta %>% 
  select(starts_with('fg')) %>% 
  summarise_all(funs(mean))

# the max fruit intake 
max_fruit <- max(meta$fg_fruit)

# the min of the fruit intake
min_fruit <- min(meta$fg_fruit) 

# get the following two 
# sample prior coefficients and multiply with the mean diet
# sample prior coefficients and multiply with the mean diet changed so that fruits are the max fruit values
prior_mean <- samples_prior %>% 
  mutate(mean_prior = Intercept + 
           b_fg_fruit*mean_fg$fg_fruit +
           b_fg_meat*mean_fg$fg_meat + 
           b_fg_milk*mean_fg$fg_milk + 
           b_fg_oils*mean_fg$fg_oils + 
           b_fg_egg*mean_fg$fg_egg + 
           b_fg_grain*mean_fg$fg_grain +
           b_fg_sweets*mean_fg$fg_sweets +
           b_fg_legume*mean_fg$fg_legume +
           b_fg_veggie*mean_fg$fg_veggie,
         
         max_fruit_prior = Intercept + 
           b_fg_fruit*max_fruit +
           b_fg_meat*mean_fg$fg_meat + 
           b_fg_milk*mean_fg$fg_milk + 
           b_fg_oils*mean_fg$fg_oils + 
           b_fg_egg*mean_fg$fg_egg + 
           b_fg_grain*mean_fg$fg_grain +
           b_fg_sweets*mean_fg$fg_sweets +
           b_fg_legume*mean_fg$fg_legume +
           b_fg_veggie*mean_fg$fg_veggie) %>% 
  select(mean_prior, max_fruit_prior) %>% 
  gather()

# plot 
prior_mean %>% 
  gghistogram(x = 'value', fill = 'key', color = 'white', palette = 'nejm', xlab = 'log(diversity)',
              #add_density = TRUE,  
              rug = TRUE,
              title = 'Prior check\nred:max fruit(87) with others average;blue:all groups average(11)')
```


## The full model incorporating the data 

```{r}
model_brm <- brm( log(simpson_reciprocal) ~ 
                fg_fruit+
               fg_meat+
               fg_milk+
               fg_oils+
                fg_egg+ 
                fg_grain+
                fg_sweets+  
                fg_legume+
                fg_veggie+
               empirical +  
               intensity + 
                TPN +
                EN +
               (1 | mrn) +
                (1 | timebin),  
              data = meta, 
              warmup = 1000, iter = 3000, 
              prior = priors,
              cores = 16, chains = 2, 
              seed = 123) 

summary(model_brm)
# 
# get_prior(log(simpson_reciprocal)~ 
#                 fg_fruit+
#                fg_meat+
#                fg_milk+
#                fg_oils+
#                 fg_egg+ 
#                 fg_grain+
#                 fg_sweets+  
#                 fg_legume+
#                 fg_veggie+
#                empirical +  
#                intensity + 
#                 TPN +
#                 EN +
#                (1 | mrn) +
#                 (1 | timebin),  
#               data = meta)
```

### Extract the posterior samples 

```{r}
# extract posterior samples for the coeff
post_samples  <- posterior_samples(model_brm, "^b")

# sample posterior coefficients and multiply with the mean diet
# sample posterior coefficients and multiply with the mean diet changed so that fruits are the max fruit values

post_2 <- post_samples %>% 
  mutate(mean_post = b_Intercept + 
           b_fg_fruit*mean_fg$fg_fruit +
           b_fg_meat*mean_fg$fg_meat + 
           b_fg_milk*mean_fg$fg_milk + 
           b_fg_oils*mean_fg$fg_oils + 
           b_fg_egg*mean_fg$fg_egg + 
           b_fg_grain*mean_fg$fg_grain +
           b_fg_sweets*mean_fg$fg_sweets +
           b_fg_legume*mean_fg$fg_legume +
           b_fg_veggie*mean_fg$fg_veggie,
         
         max_fruit_post = b_Intercept + 
           b_fg_fruit*max_fruit +
           b_fg_meat*mean_fg$fg_meat + 
           b_fg_milk*mean_fg$fg_milk + 
           b_fg_oils*mean_fg$fg_oils + 
           b_fg_egg*mean_fg$fg_egg + 
           b_fg_grain*mean_fg$fg_grain +
           b_fg_sweets*mean_fg$fg_sweets +
           b_fg_legume*mean_fg$fg_legume +
           b_fg_veggie*mean_fg$fg_veggie,
         
        min_fruit_post = b_Intercept + 
           b_fg_fruit*min_fruit +
           b_fg_meat*mean_fg$fg_meat + 
           b_fg_milk*mean_fg$fg_milk + 
           b_fg_oils*mean_fg$fg_oils + 
           b_fg_egg*mean_fg$fg_egg + 
           b_fg_grain*mean_fg$fg_grain +
           b_fg_sweets*mean_fg$fg_sweets +
           b_fg_legume*mean_fg$fg_legume +
           b_fg_veggie*mean_fg$fg_veggie ) %>% 
  select(mean_post, max_fruit_post, min_fruit_post) %>% 
  gather()

post_2 %>% 
  gghistogram(x = 'value', fill = 'key', color = 'white', palette = 'lancet', add_density = T,
              xlab = 'log(diversity)',
              #add_density = TRUE,  
              rug = TRUE,
              title = 'Post check\nmax_fruit(87) with others average;mean_post:all groups average(11)\nmin_fruit:(0)')
```

another thing to do is for both prior and posterior checks: use the ACTUAL diet from the 5 least fruit rich, and 5 most fruit rich diets.

something like :
for i in 5:
    top = topdiets[i]
    least = leastfruitdiets[i]
    coeffs = sample_posterior(n=1000) // sample_prior
    topdietdiversity_posterior = predict(top, coeffs)
    leastdietdiversity_posterior = predict(least,coeffs)
plot(topdietsdiversity_posterior) // prior 

```{r}
# the top and bottom fruit eaters diet 
set.seed(123)
N <- 5
N_sample <- 1000
topdiet <- meta %>% 
  top_n(n = N, wt = fg_fruit) %>% 
  mutate(TPNTRUE = if_else(TPN == TRUE, 1, 0),
         ENTRUE = if_else(EN == TRUE, 1, 0),
         empiricalTRUE = if_else(empirical == TRUE, 1, 0),
         intensityreduced = if_else(intensity == 'reduced', 1, 0),
         intensityablative = if_else(intensity == 'ablative', 1, 0)) %>% 
  select(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie)

botdiet <- meta %>% 
  filter(fg_fruit == 0) %>% 
  slice_sample(n = 5) %>% 
  mutate(TPNTRUE = if_else(TPN == TRUE, 1, 0),
         ENTRUE = if_else(EN == TRUE, 1, 0),
         empiricalTRUE = if_else(empirical == TRUE, 1, 0),
         intensityreduced = if_else(intensity == 'reduced', 1, 0),
         intensityablative = if_else(intensity == 'ablative', 1, 0))%>% 
  select(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie)


per_diet_post_pred <- function(fg_egg_, fg_fruit_, fg_grain_, fg_legume_, fg_meat_, fg_milk_, fg_oils_, 
                               fg_sweets_, fg_veggie_) {
  post_samp <- post_samples %>% 
    sample_n(size = N_sample, replace = F)
  
  ret = pmap(post_samp, function(b_Intercept, b_fg_fruit, b_fg_meat, b_fg_milk, b_fg_oils, b_fg_egg,
                                 b_fg_grain, b_fg_sweets, b_fg_legume, b_fg_veggie) {
    b_Intercept + b_fg_fruit*fg_fruit_ + b_fg_meat*fg_meat_ + b_fg_milk*fg_milk_ + b_fg_oils*fg_oils_ +
      b_fg_egg*fg_egg_ + b_fg_grain*fg_grain_ + b_fg_sweets*fg_sweets_ + b_fg_legume*fg_legume_ + 
      b_fg_veggie*fg_veggie_ 
  }) %>% 
    set_names(seq(1, N_sample)) %>% 
    bind_rows() %>% 
    gather()
  return(ret)
                               }


res_top <- pmap(topdiet, function(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie){
  per_diet_post_pred(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie)
}) %>% 
  set_names(paste('D', seq(1, nrow(topdiet)), sep = '')) %>% 
  bind_rows(.id = 'diet')


res_bot <- pmap(botdiet, function(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie){
  per_diet_post_pred(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie)
}) %>% 
  set_names(paste('D', seq(1, nrow(topdiet)), sep = '')) %>% 
  bind_rows(.id = 'diet')


# plot them together in one 
bind_rows(
  res_top %>% mutate(grp = 'top'),
  res_bot %>% mutate(grp = 'bot')
) %>% 
  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white', add = 'mean',
                  xlab = 'Predicted log diversity', ylab = 'Probability density') +
  ggsave('../figs/predicted_logdiversi_top_tybottom_5.pdf', width = 7, height = 5)
```

