---
title: "The model re-done with Bayesian"  
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```


```{r}
library(tidyverse)
library(brms)   
library(ggpubr)
library(tidybayes)
library(cowplot)
dtb <- read_csv('../data/cleaned_diet_data/FINAL_97_with_code_all_foods_drt_with_EN_finalized.csv')
theme_set(theme_tidybayes() + panel_border())
ncores <- parallel::detectCores()
key <- read_csv('../data/cleaned_diet_data/food_group_color_key_final.csv', col_types = 'ccccc')
axis_text_size <- 10
axis_title_size <- 10

dtb %>% 
  distinct(mrn) %>% 
  write_csv('../data/nutrition_patients_97_mrns.csv', col_names = T)
```

```{r}
# the top 10 most consumed in fruits , sweets and veggie item 
# total number of meals 
n_meal <- dtb %>% 
  count(mrn, fdrt, Meal) %>% 
  nrow()
  
top3grp <- dtb %>% 
  mutate(fgrp1 = str_sub(Food_code, 1,1)) %>% 
  filter(fgrp1 %in% c('9','6','7')) %>% 
  group_by(Food_code, description, fgrp1) %>% 
  summarise(total_per_code = sum(dehydrated_weight)) %>% 
  group_by(fgrp1) %>% 
  arrange(desc(total_per_code), .by_group = T)  %>% 
  top_n(10) %>% 
  left_join(key) %>% 
  ungroup() %>% 
  mutate(gram_per_meal = total_per_code/n_meal) %>% 
  mutate(fg1_full_name = case_when(
    fgrp1 == '6' ~ 'Fruits',
    fgrp1 == '7' ~ 'Vegetables',
    fgrp1 == '9' ~ 'Sugars, Sweets & Beverages'
  )) %>% 
  mutate(fg1_full_name = factor(fg1_full_name, levels = c('Fruits','Sugars, Sweets & Beverages', 'Vegetables')))

ggdotchart(top3grp, x = "description", y = "gram_per_meal",
           xlab = 'Food item names', 
           ylab = 'Per meal average consumption (gram)',
   group = "fg1_full_name", color = "fg1_full_name",
   palette = c('#7D3C98','#db2589', '#229954'),
   rotate = TRUE,
   sorting = "descending", 
   ggtheme = theme_bw(),
   y.text.col = TRUE ) +
  theme(legend.position = 'bottom',
        legend.title = element_blank())

ggsave('../figs/paper/S7_top10_eaten_foods_of_3_068_Jan24.pdf', width = 7.5, height = 6)
``` 


## Loading meta data table

```{r load_meta}
meta <- read_csv('../data/cleaned_stool/all_samples_meta_p2d_fg9_updated.csv') %>% 
  mutate(timebin = cut_width(sdrt, 7, boundary=0, closed = 'left')) %>% 
  mutate(intensity = factor(intensity, levels = c('nonablative','reduced','ablative'))) %>% 
  mutate(mrn = factor(mrn)) %>% 
  mutate(fg_egg = fg_egg/100,
         fg_fruit = fg_fruit/100,
         fg_grain = fg_grain/100,
         fg_legume = fg_legume/100,
         fg_meat = fg_meat/100,
         fg_milk = fg_milk/100,
         fg_oils = fg_oils/100,
         fg_sweets = fg_sweets/100,
         fg_veggie = fg_veggie/100) %>% 
  mutate(inten_non = if_else(intensity == 'nonablative', 1, 0),
         inten_ab = if_else(intensity == 'ablative', 1, 0),
         inten_re = if_else(intensity == 'reduced', 1, 0)) 

colnames(meta)


```

## Priors 

```{r get_prior}
# what priors I cann set 
model_formular <- log(simpson_reciprocal)~ 1 +
                fg_fruit+
               fg_meat+
               fg_milk+
               fg_oils+
                fg_egg+ 
                fg_grain+
                fg_sweets+  
                fg_legume+
                fg_veggie+
               inten_non + inten_ab + inten_re +
               empirical+
                TPN+
                EN+
               (1 | mrn) +
                (1 | timebin)
get_prior( model_formular,  
              data = meta)
```

### Setting priors

```{r set_prior}

priors <- c(# for the food group variables
            prior(normal(0, 1), class = 'b', coef = "fg_egg"),
            prior(normal(0, 1), class = 'b', coef = "fg_fruit"),
            prior(normal(0, 1), class = 'b', coef = "fg_grain"),
            prior(normal(0, 1), class = 'b', coef = "fg_legume"),
            prior(normal(0, 1), class = 'b', coef = "fg_meat"),
            prior(normal(0, 1), class = 'b', coef = "fg_milk"),
            prior(normal(0, 1), class = 'b', coef = "fg_oils"),
            prior(normal(0, 1), class = 'b', coef = "fg_sweets"),
            prior(normal(0, 1), class = 'b', coef = "fg_veggie"),
            # for the TPN 
            prior(normal(0, 0.1), class = 'b', coef = "TPNTRUE"),
            # for the EN
            prior(normal(0, 0.1), class = 'b', coef = "ENTRUE"),
            # for the empirical 
            prior(normal(0, 0.5), class = 'b', coef = "empiricalTRUE"),
            # for the intensity 
            prior(normal(0, 0.1), class = 'b', coef = "inten_re"),
            prior(normal(0, 0.1), class = 'b', coef = "inten_ab"),
            prior(normal(0, 0.1), class = 'b', coef = "inten_non"),
            prior(normal(2, 0.1), class = 'Intercept'))

# samples are drawn solely from the priors ignoring the likelihood
# model_wo_EN  means the usual model that the EN is still a binary variable and not ann inndivicual group
model_div <- brm( model_formular,  
              data = meta, 
              warmup = 1000, iter = 3000, 
              prior = priors,
              cores = ncores, 
              chains = 2, 
              seed = 123, sample_prior = T) 

# actually draw samples from the priors 
samples_prior <- prior_draws(model_div)

```

```{r }
# the mean food group intake from our data 
ave_fg <- meta %>% 
  select(starts_with('fg')) %>% 
  summarise_all(funs(mean)) %>% 
  gather

mean_fg <-  meta %>% 
  select(starts_with('fg')) %>% 
  summarise_all(funs(mean))

max_fg <- meta %>% 
  select(starts_with('fg')) %>% 
  summarise_all(funs(max)) 


```


```{r}
# create a df with max and min and ave in 9 groups
# create a df with the corresponding value that the samples prior coeff can multiply
max_ <- ave_fg %>%
  rename(mean_ = value) %>%
  mutate(max_fg_egg = if_else(key == 'fg_egg', max_fg$fg_egg, mean_),
         max_fg_fruit = if_else(key == 'fg_fruit', max_fg$fg_fruit, mean_),
         max_fg_grain = if_else(key == 'fg_grain', max_fg$fg_grain, mean_),
         max_fg_legume = if_else(key == 'fg_legume', max_fg$fg_legume, mean_),
         max_fg_meat = if_else(key == 'fg_meat', max_fg$fg_meat, mean_),
         max_fg_milk = if_else(key == 'fg_milk', max_fg$fg_milk, mean_),
         max_fg_oils = if_else(key == 'fg_oils', max_fg$fg_oils, mean_),
         max_fg_sweets = if_else(key == 'fg_sweets', max_fg$fg_sweets, mean_),
         max_fg_veggie = if_else(key == 'fg_veggie', max_fg$fg_veggie, mean_)) %>%
  select(-mean_) %>%
  mutate(key = if_else(str_detect(key, '^fg'), str_glue('b_{key}'), key))  %>%
  arrange(key)



values <- max_ %>%
  select(-key) %>%
  as.list()

values <- values %>%
  map(~ set_names(x = ., max_$key))

# calculate the resulted prior log div
vals <- names(values)
# 
# max_each_fg <- vals  %>% 
#   set_names(vals) %>% 
#   map(function(val) {
#     res = prior_coeffs$b_empiF$prior_coeff  +
#    prior_coeffs$b_enF$prior_coeff  +
#    prior_coeffs$b_fg_egg$prior_coeff  * values %>% pluck(val) %>% pluck('b_fg_egg') +
#    prior_coeffs$b_fg_fruit$prior_coeff  * values %>% pluck(val) %>% pluck('b_fg_fruit') +
#    prior_coeffs$b_fg_grain$prior_coeff  * values %>% pluck(val) %>% pluck('b_fg_grain') +
#    prior_coeffs$b_fg_legume$prior_coeff  * values %>% pluck(val) %>% pluck('b_fg_legume') +
#    prior_coeffs$b_fg_meat$prior_coeff * values %>% pluck(val) %>% pluck('b_fg_meat') +
#    prior_coeffs$b_fg_milk$prior_coeff  * values %>% pluck(val) %>% pluck('b_fg_milk') +
#    prior_coeffs$b_fg_oils$prior_coeff * values %>% pluck(val) %>% pluck('b_fg_oils') +
#    prior_coeffs$b_fg_sweets$prior_coeff  * values %>% pluck(val) %>% pluck('b_fg_sweets') +
#    prior_coeffs$b_fg_veggie$prior_coeff  * values %>% pluck(val) %>% pluck('b_fg_veggie') +
#    prior_coeffs$b_inten_re$prior_coeff  +
#    prior_coeffs$b_tpnF$prior_coeff +
#    prior_coeffs$Intercept$prior_coeff 
#   }) %>% 
#   bind_rows(.id = 'grp') %>% 
#   gather('grp', 'prior_res')
```


```{r plot_prior}
# combine the max and min and ave together and plot
# ave_each_fg <- samples_prior %>% 
#   mutate(mean_fg = Intercept + 
#           b_fg_fruit *mean_fg$fg_fruit+
#            b_fg_meat *mean_fg$fg_meat+ 
#            b_fg_milk *mean_fg$fg_milk+ 
#            b_fg_oils * mean_fg$fg_oils+ 
#            b_fg_egg * mean_fg$fg_egg+ 
#            b_fg_grain* mean_fg$fg_grain +
#            b_fg_sweets * mean_fg$fg_sweets+
#            b_fg_legume * mean_fg$fg_legume+
#            b_fg_veggie* mean_fg$fg_veggie +
#            b_inten_re +
#            b_empiF +
#            b_tpnF +
#            b_enF) %>% 
#   select(mean_fg) %>% 
#   mutate(grp = 'ave_each') %>% 
#   rename(prior_res = mean_fg) %>% 
#   select(grp, prior_res)
# 
# prior_all <- bind_rows(ave_each_fg, max_each_fg) 
# 
# prior_all %>%  write_csv('../figs/paper/data/prior_ave_max_min_res_68.csv')
# 
# grps <- prior_all %>% 
#   distinct(grp)
# 
# fg_cha <- meta %>% 
#   select(starts_with('fg')) %>% 
#   colnames()
```

## Posterior

### Extract the posterior samples 

```{r post_sample}
# extract posterior samples for the coeff
post_samples  <- posterior_samples(model_div, '^b_')

# save it out for the forest plot of the coeff
post_fg_coeff <- post_samples %>% 
  select(starts_with('b_fg')) %>% 
  gather('item', 'value')

post_fg_coeff_mean <- post_fg_coeff  %>% 
  group_by(item) %>% 
  summarise(meanperitem = mean(value))

out <- post_fg_coeff %>% 
  left_join(post_fg_coeff_mean, by = "item")

out %>% write_csv('../data/brms_model_var_res_updated.csv') 
```

```{r post_coeff_forest}
# make the forest plot of the post coeff 
# like the one on the ash abstract

post_coeff <- post_samples %>% 
  select(starts_with('b_fg')) %>% 
  gather('item', 'coeff') %>% 
  mutate(item = str_replace(item, 'b_fg_','')) %>% 
   mutate(fgrp1 = case_when(
    item ==  'milk' ~ '1',
    item == 'meat' ~ '2',
    item ==  'egg' ~ '3',
    item ==  'legume' ~ '4',
    item == 'grain' ~ '5',
    item == 'fruit' ~ '6',
    item == 'veggie' ~ '7',
    item ==  'oils' ~ '8', 
    item ==  'sweets' ~ '9'
  ))  %>% 
  left_join(key %>% select(fgrp1, color, shortname))  %>% 
  mutate(shortname = fct_reorder(shortname, coeff, .fun=median, .desc = F)) 

fg_labels <- levels(post_coeff$shortname)
```
```{r}
fg_colors <- post_coeff %>% 
  distinct(shortname, color) %>% 
  select(shortname, color) %>% 
  deframe()

title1 <- bquote(Outcome: log[e]~(Simpsons~reciprocal))

div_post <- post_coeff %>% 
  ggplot(aes(x = coeff, y = shortname, col = shortname)) +
  stat_pointinterval(.width = c(.66, .95)) +
  #scale_color_manual(values = c('#EC0000','#00468B')) +
  geom_vline(xintercept = 0, col = 'black', linetype = 'dashed') +
  labs(x = 'Regression coefficients',
       y = '', 
       title = 'Diversity') +
  theme_classic() +
  theme(legend.position = 'none') +
  scale_color_manual(values = fg_colors) +
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8),
        aspect.ratio=1)

div_post %>% 
  write_rds('../data/068_div_post.rds')

ggsave('../figs/paper/068_fig2_model_coeff_forest.eps',  
       width = 80,
       height = 80,
         #height = 60,
         units = c("mm"),
         dpi = 400, device = 'eps')
```


```{r post_coeff_binary}
coeff_bi <- post_samples %>% 
  select(!starts_with('b_fg')) %>% 
  select(-b_Intercept) %>% 
  gather('item', 'coeff') %>% 
  mutate(item_name = case_when(
    item ==  'b_inten_non' ~ 'Intensity: nonablative',
    item == 'b_inten_ab' ~ 'Intensity: ablative',
    item ==  'b_inten_re' ~ 'Intensity: reduced',
    item ==  'b_empiricalTRUE' ~ 'Empirical abx exposure',
    item == 'b_TPNTRUE' ~ 'TPN exposure',
    item ==  'b_ENTRUE' ~ 'EN exposure'
  )) %>% 
  mutate(item = factor(item, levels = c('b_empiF','b_empiT','b_inten_non','b_inten_re','b_inten_ab','b_enF','b_enT','b_tpnF','b_tpnT'))) %>% 
  mutate(item_name = fct_reorder(item_name, coeff, .fun=median, .desc = F)) 

diversity_factor <- coeff_bi %>% 
  ggplot(aes(x = coeff, y = item_name)) +
  stat_pointinterval(.width = c(.66, .95)) +
  #scale_color_manual(values = c('#EC0000','#00468B')) +
  geom_vline(xintercept = 0, col = 'black', linetype = 'dashed') +
  labs(x = 'Regression coefficients',
       y = 'Factor variables',
       title = 'Diversity') +
  theme_classic() +
  theme(legend.position = 'none',
        aspect.ratio=1)

diversity_factor %>% 
  write_rds('../data/068_diversity_factor_forest.rds')

ggsave('../figs/paper/068_fig2_model_coeff_forest_binary.eps',  
       width = 100,
       height = 100,
         #height = 60,
         units = c("mm"),
         dpi = 400, device = 'eps')
```

 
```{r post_sample_food_ave}
# when the food intake is average mean

###THE RESULTS IS THE LOG TRANSFORMED DIVERSITY !!!!##
ave_each_fg_post <- post_samples %>% 
  mutate(mean_fg = b_Intercept + 
           b_fg_fruit*mean_fg$fg_fruit +
           b_fg_meat*mean_fg$fg_meat + 
           b_fg_milk*mean_fg$fg_milk + 
           b_fg_oils*mean_fg$fg_oils + 
           b_fg_egg*mean_fg$fg_egg + 
           b_fg_grain*mean_fg$fg_grain +
           b_fg_sweets*mean_fg$fg_sweets +
           b_fg_legume*mean_fg$fg_legume +
           b_fg_veggie*mean_fg$fg_veggie +
           b_inten_re ) %>% 
  select(mean_fg) %>% 
  mutate(grp = 'ave_each') %>% 
  rename(post_res = mean_fg) %>% 
  select(grp, post_res)

# when the intake is max
post_coeffs <- post_samples %>% 
  select(b_Intercept, b_inten_re,  starts_with('b_fg')) %>% 
  gather('term','post_coeff') %>% 
  split(.$term) 

max_each_fg_post <- vals  %>% 
  set_names(vals) %>% 
  map(function(val) {
    res =
   post_coeffs$b_fg_egg$post_coeff * values %>% pluck(val) %>% pluck('b_fg_egg') +
   post_coeffs$b_fg_fruit$post_coeff * values %>% pluck(val) %>% pluck('b_fg_fruit') +
   post_coeffs$b_fg_grain$post_coeff * values %>% pluck(val) %>% pluck('b_fg_grain') +
   post_coeffs$b_fg_legume$post_coeff * values %>% pluck(val) %>% pluck('b_fg_legume') +
   post_coeffs$b_fg_meat$post_coeff * values %>% pluck(val) %>% pluck('b_fg_meat') +
   post_coeffs$b_fg_milk$post_coeff * values %>% pluck(val) %>% pluck('b_fg_milk') +
   post_coeffs$b_fg_oils$post_coeff * values %>% pluck(val) %>% pluck('b_fg_oils') +
   post_coeffs$b_fg_sweets$post_coeff * values %>% pluck(val) %>% pluck('b_fg_sweets') +
   post_coeffs$b_fg_veggie$post_coeff * values %>% pluck(val) %>% pluck('b_fg_veggie') +
   post_coeffs$b_inten_re$post_coeff  +
   post_coeffs$b_Intercept$post_coeff 
  }) %>% 
  bind_rows(.id = 'grp')
```

```{r post_diff}
# calculate the absolute difference between the max in each fg and the mean 
# note the mean is from EN-free records.
maxs <- max_each_fg_post %>% 
  as.list()

diff_value_lt0 <-  maxs %>% 
  map(function(vec) {
    tibble(
      max_=vec,
      ave=ave_each_fg_post$post_res
    ) %>% 
      mutate(post_diff_ = max_ - ave)
  }) %>% 
  map( ~ count(x = ., post_diff_ < 0) %>% 
          mutate(perc = n/sum(n) * 100) %>% 
          slice(2) %>% 
          select(perc)) %>% 
  bind_rows(.id = 'term')


theme_set(theme_tidybayes() + cowplot::panel_border())
post_diff_df <- maxs %>% 
  map(function(vec) {
    tibble(
      max_=vec,
      ave=ave_each_fg_post$post_res
    ) %>% 
      mutate(post_diff_original_scale = exp(max_) - exp(ave))
  }) %>% 
  bind_rows(.id = 'term') %>% 
  mutate(term = str_replace(term, 'max_','')) %>% 
  left_join(key %>% 
              select(term = fg1_name, color, shortname)) %>% 
  mutate(shortname = fct_reorder(shortname, post_diff_original_scale, .fun=median, .desc = F))
```
```{r post_diff_3}
# only plot the post diff forest in fruits and sweets
only3 <- post_diff_df %>% 
  filter(term %in% c('fg_fruit','fg_sweets', 'fg_veggie')) 

only3_sim <- only3 %>% 
  group_by(term, shortname) %>% 
  median_qi(post_diff_original_scale, .width = c(.95, .8, .5)) %>%
  ggplot(aes(y = shortname, x = post_diff_original_scale)) +
  geom_interval(aes(xmin = .lower, xmax = .upper, fill = term)) + 
  geom_vline(xintercept = 0, linetype = "dashed", col = 'black') +
  labs(x = expression(Change~`in`~alpha~diversity),
       y = '', 
       title = 'Diversity predictions') +
  theme_classic() + 
    scale_fill_manual(values = key %>% 
                       select(fg1_name, color) %>% deframe()) +
  scale_x_continuous(breaks = seq(-10, 20, 2)) +
 #scale_color_brewer()+
  xlim(-6, 4.5) +
  theme(axis.text=element_text(size=8),
        legend.position = 'right',
        axis.title=element_text(size=8),
        aspect.ratio=1)

only3_sim 
```


```{r post_diff_3}
only3_sim %>% 
  write_rds('../data/068_only3_sim.rds')

ggsave('../figs/paper/068_fig2_fg3_post_diff.eps',  width = 80,
       height = 80,
         #height = 60, 
         units = c("mm"),
         dpi = 400, device = 'eps')  
```



```{r post_diff}
post_diff_df %>% 
  ggplot(aes(y = term, x = post_diff_original_scale, col = term)) +
  stat_halfeye() +
  #scale_fill_manual(values = c("gray80", "skyblue")) +
  geom_vline(xintercept = 0, linetype = "dashed", col = 'red') +
  labs(x = "Difference in Simpson's reciprocal diversity",
       y = '',
       title = 'The diff in simpson reciprocal diversity if one food group is consumed\nat max amount than the average') +
  theme(legend.position = 'none') +
  scale_x_continuous(breaks = seq(-10, 20, 2)) +
  scale_color_manual(values = key %>% 
                       select(fg1_name, color) %>% deframe()) +
  xlim(-4.1, 12)


```

## the forest plot with prior and post 

```{r both}
both <- bind_rows(
  samples_prior %>% 
    select(starts_with('b_fg')) %>% mutate(grp = 'prior'),
  post_samples %>% 
    select(starts_with('b_fg')) %>% mutate(grp = 'post')
)  %>% 
  gather('item', 'coeff', b_fg_fruit:b_fg_veggie) 

summ <- both %>% 
  group_by(grp, item) %>% 
  summarise(q025 = quantile(coeff, 0.025),
            q975 = quantile(coeff, 0.975),
            q50 = quantile(coeff, 0.5)) %>% 
  arrange(grp, q50) 

fg_order <- summ %>%
  filter(grp == 'post') %>%
  arrange(-q50) %>%
  pull(item)
```


```{r pinterval}
pinterval <- both %>%  
  #mutate(item = str_replace(item, 'b_','')) %>% 
  left_join(key %>% 
              select(item = fg1_name)) %>% 
  mutate(item = factor(item, levels = fg_order)) %>% 
  ggplot(aes(x = coeff, y = grp, color = grp)) +
  stat_pointinterval(.width = c(.66, .95)) +
  scale_color_manual(values = c('#EC0000','#00468B')) +
  geom_vline(xintercept = 0, col = 'gray', linetype = 'dashed') +
  facet_wrap(item ~ .) +
  labs(x = 'Regression coefficients',
       y = '') +
  theme(legend.position = 'none')
 
pinterval
ggsave('../figs/paper/prior_post_pinterval_coeff.pdf', width = 10, height = 4)
```

## others

```{r post_mean_max}
# post_all <- bind_rows(ave_each_fg_post, max_each_fg_post) %>% 
#   mutate(div = exp(post_res))
# post_all %>%  write_csv('../figs/paper/data/post_ave_max_min_res_68.csv')
# 
# post_check_list <- fg_cha %>% 
#   set_names(fg_cha) %>% 
#   map(function(char){
#     post_all %>% 
#       filter(str_detect(grp, str_glue('ave_each|{char}'))) %>% 
#       gghistogram(x = 'div', fill = 'grp', color = 'white', palette = 'lancet', xlab = 'logdiversity',
#                   #add_density = TRUE, 
#                   bins = 30,
#                   title = str_glue('Post check {char}'))
#     })
# 
# g_post <- cowplot::plot_grid(post_check_list[[1]],post_check_list[[2]],post_check_list[[3]],post_check_list[[4]],post_check_list[[5]],post_check_list[[6]],post_check_list[[7]],post_check_list[[8]],post_check_list[[9]],
#     ncol = 3, nrow= 3, 
#     o = 'hv',
#     #labels = 'AUTO',
#     axis = 'bltr') 
# g_post
# 
# ggsave('../figs/paper/post_check_9fg.pdf', width = 10, height = 10, dpi = 300, plot = g_post)
```

## Post_top_patients

another thing to do is for both prior and posterior checks: use the ACTUAL diet from the 5 least fruit rich, and 5 most fruit rich diets.

something like :
for i in 5:
    top = topdiets[i]
    least = leastfruitdiets[i]
    coeffs = sample_posterior(n=1000) // sample_prior
    topdietdiversity_posterior = predict(top, coeffs)
    leastdietdiversity_posterior = predict(least,coeffs)
plot(topdietsdiversity_posterior) // prior 

```{r}
# the top and bottom fruit eaters diet 
# set.seed(123)
# N <- 5
# N_sample <- 1000
# topdiet <- meta %>% 
#   top_n(n = N, wt = fg_fruit) %>% 
#   mutate(TPNTRUE = if_else(TPN == TRUE, 1, 0),
#          ENTRUE = if_else(EN == TRUE, 1, 0),
#          empiricalTRUE = if_else(empirical == TRUE, 1, 0),
#          intensityreduced = if_else(intensity == 'reduced', 1, 0),
#          intensityablative = if_else(intensity == 'ablative', 1, 0)) %>% 
#   select(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie)
# 
# botdiet <- meta %>% 
#   filter(fg_fruit == 0) %>% 
#   slice_sample(n = 5) %>% 
#   mutate(TPNTRUE = if_else(TPN == TRUE, 1, 0),
#          ENTRUE = if_else(EN == TRUE, 1, 0),
#          empiricalTRUE = if_else(empirical == TRUE, 1, 0),
#          intensityreduced = if_else(intensity == 'reduced', 1, 0),
#          intensityablative = if_else(intensity == 'ablative', 1, 0))%>% 
#   select(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie)
# 
# 
# per_diet_post_pred <- function(fg_egg_, fg_fruit_, fg_grain_, fg_legume_, fg_meat_, fg_milk_, fg_oils_, 
#                                fg_sweets_, fg_veggie_) {
#   post_samp <- post_samples %>% 
#     sample_n(size = N_sample, replace = F)
#   
#   ret = pmap(post_samp, function(b_Intercept, b_fg_fruit, b_fg_meat, b_fg_milk, b_fg_oils, b_fg_egg,
#                                  b_fg_grain, b_fg_sweets, b_fg_legume, b_fg_veggie) {
#     b_Intercept + b_fg_fruit*fg_fruit_ + b_fg_meat*fg_meat_ + b_fg_milk*fg_milk_ + b_fg_oils*fg_oils_ +
#       b_fg_egg*fg_egg_ + b_fg_grain*fg_grain_ + b_fg_sweets*fg_sweets_ + b_fg_legume*fg_legume_ + 
#       b_fg_veggie*fg_veggie_ 
#   }) %>% 
#     set_names(seq(1, N_sample)) %>% 
#     bind_rows() %>% 
#     gather()
#   return(ret)
#                                }
# 
# 
# res_top <- pmap(topdiet, function(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie){
#   per_diet_post_pred(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie)
# }) %>% 
#   set_names(paste('D', seq(1, nrow(topdiet)), sep = '')) %>% 
#   bind_rows(.id = 'diet')
# 
# 
# res_bot <- pmap(botdiet, function(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie){
#   per_diet_post_pred(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie)
# }) %>% 
#   set_names(paste('D', seq(1, nrow(topdiet)), sep = '')) %>% 
#   bind_rows(.id = 'diet')
# 
# 
# # plot them together in one 
# bind_rows(
#   res_top %>% mutate(grp = 'top'),
#   res_bot %>% mutate(grp = 'bot')
# ) %>% 
#   gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white', add = 'mean',
#                   xlab = 'Predicted log diversity', ylab = 'Probability density') +
#   ggsave('../figs/predicted_logdiversi_top_tybottom_5.pdf', width = 7, height = 5)
```
