---
title: "The model re-done with Bayesian"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

```{r}
library(tidyverse)
library(brms)  
library(ggpubr)
library(tidybayes)
dtb <- read_csv('../data/cleaned_diet_data/FINAL_97_with_code_all_foods_drt_with_EN_finalized.csv')
```

## Loading meta data table

```{r load_meta}
meta <- read_csv('../data/cleaned_stool/all_samples_meta_p2d_fg9_updated.csv') %>% 
  mutate(timebin = cut_width(sdrt, 7, boundary=0, closed = 'left')) %>% 
  mutate(intensity = factor(intensity, levels = c('nonablative','reduced','ablative'))) %>% 
  mutate(mrn = factor(mrn)) %>% 
  mutate(fg_egg = fg_egg/100,
         fg_fruit = fg_fruit/100,
         fg_grain = fg_grain/100,
         fg_legume = fg_legume/100,
         fg_meat = fg_meat/100,
         fg_milk = fg_milk/100,
         fg_oils = fg_oils/100,
         fg_sweets = fg_sweets/100,
         fg_veggie = fg_veggie/100) %>% 
  mutate(empiT = if_else(empirical == 'TRUE', 1, 0),
         empiF = if_else(empirical == 'FALSE', 1, 0),
         tpnT = if_else(TPN == 'TRUE', 1, 0),
         tpnF = if_else(TPN == 'FALSE', 1, 0),
         enT = if_else(TPN == 'TRUE', 1, 0),
         enF = if_else(TPN == 'FALSE', 1, 0),
         inten_non = if_else(intensity == 'nonablative', 1, 0),
         inten_ab = if_else(intensity == 'ablative', 1, 0),
         inten_re = if_else(intensity == 'reduced', 1, 0)) %>% 
  mutate(fg_EN = if_else(EN == 'TRUE', fg_sweets, 0),
         fg_sweets = if_else(EN == 'TRUE', 0, fg_sweets))


# I should very strictly split the sweets annd the EN group
# recalculate the sweets and EN for that one patient with age 59.1
person <- meta %>% 
  filter(str_detect(sampleid, '^1498')) %>% distinct(mrn) %>% pull(mrn)

change <- dtb %>% 
  filter(mrn == person) %>% 
  filter(fdrt %in% c(14,15)) %>% 
  filter(str_detect(Food_code, '^9')) %>% 
  mutate(grp = if_else(dehydrated_weight > 100, 'fg_EN', 'fg_sweets')) %>% 
  group_by(grp) %>% 
  summarise(value = sum(dehydrated_weight)/2/100)

that_value <- meta %>% 
  filter(mrn == person) %>% 
  filter(fg_EN > 0) %>% 
  pull(fg_EN)

meta <- meta %>% 
  mutate(fg_EN = if_else(fg_EN == that_value, 2.1525, fg_EN)) %>% 
  mutate(fg_sweets = if_else(fg_EN == 2.1525, 0.03932517, fg_sweets))

meta %>% write_csv('../figs/paper/data/meta_fg10_bayes.csv')
```
```{r}

# the EN patients
en <- meta %>% 
  filter(fg_EN >0)


en %>% 
  select(mrn, sdrt, enteral_nutrition_intake = fg_EN, simpson_reciprocal) %>% 
  mutate(mrn = str_glue('Patient{as.numeric(factor(mrn))}')) %>% 
  mutate(enteral_nutrition_intake = enteral_nutrition_intake*100) %>% 
  gather('type', 'value' ,enteral_nutrition_intake: simpson_reciprocal) %>% 
  ggscatter(x = 'sdrt', y = 'value',  color = 'type', xlab = 'day relative to transplant') +
  facet_grid(type ~ mrn, scales = 'free_y') +
  theme_light()
  
```

## Priors 

```{r get_prior}
# what priors I cann set 
model_formular <- log(simpson_reciprocal)~ 1 +
                fg_fruit+
               fg_meat+
               fg_milk+
               fg_oils+
                fg_egg+ 
                fg_grain+
                fg_sweets+  
                fg_legume+
                fg_veggie+
                fg_EN+
               inten_non + inten_ab + inten_re +
               empiT +  empiF+
                tpnT +tpnF+
               (1 | mrn) +
                (1 | timebin)
get_prior( model_formular,  
              data = meta)
```

### Setting priors

```{r set_prior}

priors <- c(# for the food group variables
            prior(normal(0, 1), class = 'b', coef = "fg_egg"),
            prior(normal(0, 1), class = 'b', coef = "fg_fruit"),
            prior(normal(0, 1), class = 'b', coef = "fg_grain"),
            prior(normal(0, 1), class = 'b', coef = "fg_legume"),
            prior(normal(0, 1), class = 'b', coef = "fg_meat"),
            prior(normal(0, 1), class = 'b', coef = "fg_milk"),
            prior(normal(0, 1), class = 'b', coef = "fg_oils"),
            prior(normal(0, 1), class = 'b', coef = "fg_sweets"),
            prior(normal(0, 1), class = 'b', coef = "fg_veggie"),
            prior(normal(0, 1), class = 'b', coef = "fg_EN"),
            # for the TPN 
            prior(normal(0, 0.1), class = 'b', coef = "tpnF"),
            prior(normal(0, 0.1), class = 'b', coef = "tpnT"),
            # for the EN
            # prior(normal(0, 0.1), class = 'b', coef = "enF"),
            # prior(normal(0, 0.1), class = 'b', coef = "enT"),
            # for the empirical 
            prior(normal(0, 0.5), class = 'b', coef = "empiF"),
            prior(normal(0, 0.5), class = 'b', coef = "empiT"),
            # for the intensity 
            prior(normal(0, 0.1), class = 'b', coef = "inten_re"),
            prior(normal(0, 0.1), class = 'b', coef = "inten_ab"),
            prior(normal(0, 0.1), class = 'b', coef = "inten_non"),
            prior(normal(2, 0.1), class = 'Intercept'))

# samples are drawn solely from the priors ignoring the likelihood
model_w_EN <- brm( model_formular,  
              data = meta, 
              warmup = 1000, iter = 3000, 
              prior = priors,
              cores = 16, 
              chains = 2, 
              seed = 123, sample_prior = T) 

# look at the priors set for the model
summary(model_w_EN)

# actually draw samples from the priors 
samples_prior <- prior_samples(model_w_EN)

```

```{r }
# the mean food group intake from our data 
ave_fg <- meta %>% 
  select(starts_with('fg')) %>% 
  summarise_all(funs(mean)) %>% 
  gather

mean_fg <-  meta %>% 
  select(starts_with('fg')) %>% 
  summarise_all(funs(mean))

max_fg <- meta %>% 
  select(starts_with('fg')) %>% 
  summarise_all(funs(max))

min_fg <- meta %>% 
  select(starts_with('fg')) %>% 
  summarise_all(funs(min))

max_log_div <- max(log(meta$simpson_reciprocal)) 
range(log(meta$simpson_reciprocal))
```

```{r}
# create a df with max  and ave in 10 groups
# create a df with the corresponding value that the samples prior coeff can multiply
max_ <- ave_fg %>% 
  rename(mean_ = value) %>% 
  mutate(max_fg_egg = if_else(key == 'fg_egg', max_fg$fg_egg, mean_),
         max_fg_fruit = if_else(key == 'fg_fruit', max_fg$fg_fruit, mean_),
         max_fg_grain = if_else(key == 'fg_grain', max_fg$fg_grain, mean_),
         max_fg_legume = if_else(key == 'fg_legume', max_fg$fg_legume, mean_),
         max_fg_meat = if_else(key == 'fg_meat', max_fg$fg_meat, mean_),
         max_fg_milk = if_else(key == 'fg_milk', max_fg$fg_milk, mean_),
         max_fg_oils = if_else(key == 'fg_oils', max_fg$fg_oils, mean_),
         max_fg_sweets = if_else(key == 'fg_sweets', max_fg$fg_sweets, mean_),
         max_fg_veggie = if_else(key == 'fg_veggie', max_fg$fg_veggie, mean_),
         max_fg_EN = if_else(key == 'fg_EN', max_fg$fg_EN, mean_)) %>% 
  select(-mean_) %>% 
  mutate(key = if_else(str_detect(key, '^fg'), str_glue('b_{key}'), key))  %>% 
  arrange(key)

prior_coeffs <- samples_prior %>% 
  select(Intercept, b_inten_re, b_empiF, b_tpnF, starts_with('b_fg')) %>% 
  gather('term','prior_coeff') %>% 
  split(.$term)
  
values <- max_ %>% 
  select(-key) %>% 
  as.list()

values <- values %>% 
  map(~ set_names(x = ., max_$key))

# calculate the resulted prior log div 
vals <- names(values)

max_each_fg <- vals  %>% 
  set_names(vals) %>% 
  map(function(val) {
    res = prior_coeffs$b_empiF$prior_coeff  +
   #prior_coeffs$b_enF$prior_coeff  +
   prior_coeffs$b_fg_egg$prior_coeff  * values %>% pluck(val) %>% pluck('b_fg_egg') +
   prior_coeffs$b_fg_fruit$prior_coeff  * values %>% pluck(val) %>% pluck('b_fg_fruit') +
   prior_coeffs$b_fg_grain$prior_coeff  * values %>% pluck(val) %>% pluck('b_fg_grain') +
   prior_coeffs$b_fg_legume$prior_coeff  * values %>% pluck(val) %>% pluck('b_fg_legume') +
   prior_coeffs$b_fg_meat$prior_coeff * values %>% pluck(val) %>% pluck('b_fg_meat') +
   prior_coeffs$b_fg_milk$prior_coeff  * values %>% pluck(val) %>% pluck('b_fg_milk') +
   prior_coeffs$b_fg_oils$prior_coeff * values %>% pluck(val) %>% pluck('b_fg_oils') +
   prior_coeffs$b_fg_sweets$prior_coeff  * values %>% pluck(val) %>% pluck('b_fg_sweets') +
   prior_coeffs$b_fg_veggie$prior_coeff  * values %>% pluck(val) %>% pluck('b_fg_veggie') +
   prior_coeffs$b_fg_EN$prior_coeff  * values %>% pluck(val) %>% pluck('b_fg_EN') +
   prior_coeffs$b_inten_re$prior_coeff   +
   prior_coeffs$b_tpnF$prior_coeff +
   prior_coeffs$Intercept$prior_coeff 
  }) %>% 
  bind_rows(.id = 'grp') %>% 
  gather('grp', 'prior_res')

```


```{r plot_prior}
# combinne the max and min and ave together and plot
ave_each_fg <- samples_prior %>% 
  mutate(mean_fg = Intercept + 
           b_fg_fruit *mean_fg$fg_fruit+
           b_fg_meat *mean_fg$fg_meat+ 
           b_fg_milk *mean_fg$fg_milk+ 
           b_fg_oils * mean_fg$fg_oils+ 
           b_fg_egg * mean_fg$fg_egg+ 
           b_fg_grain* mean_fg$fg_grain +
           b_fg_sweets * mean_fg$fg_sweets+
           b_fg_legume * mean_fg$fg_legume+
           b_fg_veggie* mean_fg$fg_veggie +
           b_fg_EN* mean_fg$fg_EN +
           b_inten_re +
           b_empiF +
           b_tpnF) %>% 
  select(mean_fg) %>% 
  mutate(grp = 'ave_each') %>% 
  rename(prior_res = mean_fg) %>% 
  select(grp, prior_res)

prior_all <- bind_rows(ave_each_fg, max_each_fg) 


grps <- prior_all %>% 
  distinct(grp)

fg_cha <- meta %>% 
  select(starts_with('fg')) %>% 
  colnames()


```

## Posterior

### Extract the posterior samples 

```{r post_sample}
# extract posterior samples for the coeff
post_samples  <- posterior_samples(model_w_EN, '^b_')


# when the food intake is average mean

ave_each_fg_post <- post_samples %>% 
  mutate(mean_fg = b_Intercept + 
           b_fg_fruit*mean_fg$fg_fruit +
           b_fg_meat*mean_fg$fg_meat + 
           b_fg_milk*mean_fg$fg_milk + 
           b_fg_oils*mean_fg$fg_oils + 
           b_fg_egg*mean_fg$fg_egg + 
           b_fg_grain*mean_fg$fg_grain +
           b_fg_sweets*mean_fg$fg_sweets +
           b_fg_legume*mean_fg$fg_legume +
           b_fg_veggie*mean_fg$fg_veggie +
           b_fg_EN*mean_fg$fg_EN +
           b_inten_re +
           b_empiF +
           b_tpnF) %>% 
  select(mean_fg) %>% 
  mutate(grp = 'ave_each') %>% 
  rename(post_res = mean_fg) %>% 
  select(grp, post_res)

# when the intake is max
post_coeffs <- post_samples %>% 
  select(b_Intercept, b_inten_re, b_empiF, b_tpnF, starts_with('b_fg')) %>% 
  gather('term','post_coeff') %>% 
  split(.$term) 

max_each_fg_post <- vals  %>% 
  set_names(vals) %>% 
  map(function(val) {
    res = post_coeffs$b_empiF$post_coeff  +
   post_coeffs$b_fg_egg$post_coeff * values %>% pluck(val) %>% pluck('b_fg_egg') +
   post_coeffs$b_fg_fruit$post_coeff * values %>% pluck(val) %>% pluck('b_fg_fruit') +
   post_coeffs$b_fg_grain$post_coeff * values %>% pluck(val) %>% pluck('b_fg_grain') +
   post_coeffs$b_fg_legume$post_coeff * values %>% pluck(val) %>% pluck('b_fg_legume') +
   post_coeffs$b_fg_meat$post_coeff * values %>% pluck(val) %>% pluck('b_fg_meat') +
   post_coeffs$b_fg_milk$post_coeff * values %>% pluck(val) %>% pluck('b_fg_milk') +
   post_coeffs$b_fg_oils$post_coeff * values %>% pluck(val) %>% pluck('b_fg_oils') +
   post_coeffs$b_fg_sweets$post_coeff * values %>% pluck(val) %>% pluck('b_fg_sweets') +
   post_coeffs$b_fg_veggie$post_coeff * values %>% pluck(val) %>% pluck('b_fg_veggie') +
   post_coeffs$b_fg_EN$post_coeff * values %>% pluck(val) %>% pluck('b_fg_EN') +
   post_coeffs$b_inten_re$post_coeff  +
   post_coeffs$b_tpnF$post_coeff +
   post_coeffs$b_Intercept$post_coeff 
  }) %>% 
  bind_rows(.id = 'grp') 

post_all <- bind_rows(ave_each_fg_post, max_each_fg_post)

```

```{r post_diff}
# calculate the absolute difference between the max in each fg and the mean 

maxs <- max_each_fg_post %>% 
  as.list()

diff_value_lt0 <-  maxs %>% 
  map(function(vec) {
    tibble(
      max_=vec,
      ave=ave_each_fg_post$post_res
    ) %>% 
      mutate(post_diff_ = max_ - ave)
  }) %>% 
  map( ~ count(x = ., post_diff_ < 0) %>% 
          mutate(perc = n/sum(n) * 100) %>% 
          slice(2) %>% 
          select(perc)) %>% 
  bind_rows(.id = 'term')


theme_set(theme_tidybayes() + cowplot::panel_border())
post_diff_df <- maxs %>% 
  map(function(vec) {
    tibble(
      max_=vec,
      ave=ave_each_fg_post$post_res
    ) %>% 
      mutate(post_diff_original_scale = exp(max_) - exp(ave))
  }) %>% 
  bind_rows(.id = 'term') %>% 
  mutate(term = fct_reorder(term, post_diff_original_scale, .fun=median, .desc = F)) %>% 
  ggplot(aes(y = term, x = post_diff_original_scale, fill = stat(x < 0))) +
  stat_halfeye() +
  scale_fill_manual(values = c("gray80", "skyblue")) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(x = 'The diff in simpson reciprocal diversity if one food group is consumed\nat max amount than the average',
       y = '',
       title = '') +
  theme(legend.position = 'top') +
  scale_x_continuous(breaks = seq(-10, 20, 2))

ggsave('../figs/paper/fg10_post_diff.png', width = 7, height = 6)


post_diff_df  
```

## Post_top_patients

another thing to do is for both prior and posterior checks: use the ACTUAL diet from the 5 least fruit rich, and 5 most fruit rich diets.

something like :
for i in 5:
    top = topdiets[i]
    least = leastfruitdiets[i]
    coeffs = sample_posterior(n=1000) // sample_prior
    topdietdiversity_posterior = predict(top, coeffs)
    leastdietdiversity_posterior = predict(least,coeffs)
plot(topdietsdiversity_posterior) // prior 

```{r}
# the top and bottom fruit eaters diet 
# set.seed(123)
# N <- 5
# N_sample <- 1000
# topdiet <- meta %>% 
#   top_n(n = N, wt = fg_fruit) %>% 
#   mutate(TPNTRUE = if_else(TPN == TRUE, 1, 0),
#          ENTRUE = if_else(EN == TRUE, 1, 0),
#          empiricalTRUE = if_else(empirical == TRUE, 1, 0),
#          intensityreduced = if_else(intensity == 'reduced', 1, 0),
#          intensityablative = if_else(intensity == 'ablative', 1, 0)) %>% 
#   select(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie)
# 
# botdiet <- meta %>% 
#   filter(fg_fruit == 0) %>% 
#   slice_sample(n = 5) %>% 
#   mutate(TPNTRUE = if_else(TPN == TRUE, 1, 0),
#          ENTRUE = if_else(EN == TRUE, 1, 0),
#          empiricalTRUE = if_else(empirical == TRUE, 1, 0),
#          intensityreduced = if_else(intensity == 'reduced', 1, 0),
#          intensityablative = if_else(intensity == 'ablative', 1, 0))%>% 
#   select(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie)
# 
# 
# per_diet_post_pred <- function(fg_egg_, fg_fruit_, fg_grain_, fg_legume_, fg_meat_, fg_milk_, fg_oils_, 
#                                fg_sweets_, fg_veggie_) {
#   post_samp <- post_samples %>% 
#     sample_n(size = N_sample, replace = F)
#   
#   ret = pmap(post_samp, function(b_Intercept, b_fg_fruit, b_fg_meat, b_fg_milk, b_fg_oils, b_fg_egg,
#                                  b_fg_grain, b_fg_sweets, b_fg_legume, b_fg_veggie) {
#     b_Intercept + b_fg_fruit*fg_fruit_ + b_fg_meat*fg_meat_ + b_fg_milk*fg_milk_ + b_fg_oils*fg_oils_ +
#       b_fg_egg*fg_egg_ + b_fg_grain*fg_grain_ + b_fg_sweets*fg_sweets_ + b_fg_legume*fg_legume_ + 
#       b_fg_veggie*fg_veggie_ 
#   }) %>% 
#     set_names(seq(1, N_sample)) %>% 
#     bind_rows() %>% 
#     gather()
#   return(ret)
#                                }
# 
# 
# res_top <- pmap(topdiet, function(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie){
#   per_diet_post_pred(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie)
# }) %>% 
#   set_names(paste('D', seq(1, nrow(topdiet)), sep = '')) %>% 
#   bind_rows(.id = 'diet')
# 
# 
# res_bot <- pmap(botdiet, function(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie){
#   per_diet_post_pred(fg_egg, fg_fruit, fg_grain, fg_legume, fg_meat, fg_milk, fg_oils, fg_sweets, fg_veggie)
# }) %>% 
#   set_names(paste('D', seq(1, nrow(topdiet)), sep = '')) %>% 
#   bind_rows(.id = 'diet')
# 
# 
# # plot them together in one 
# bind_rows(
#   res_top %>% mutate(grp = 'top'),
#   res_bot %>% mutate(grp = 'bot')
# ) %>% 
#   gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white', add = 'mean',
#                   xlab = 'Predicted log diversity', ylab = 'Probability density') +
#   ggsave('../figs/predicted_logdiversi_top_tybottom_5.pdf', width = 7, height = 5)
```


