---
title: "macronutrients_analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
library(vdbR)
library(tidybayes)
connect_database('~/dbConfig.txt')
library(brms)
library(rstan)
options(mc.cores = parallel::detectCores())
ncores = parallel::detectCores()
rstan_options(auto_write = TRUE)
axis_text_size <- 10
axis_title_size <- 10
```

```{r}
# the % needs met data
```

```{r}
dtb <- read_csv('../data/cleaned_diet_data/FINAL_97_with_code_all_foods_drt_with_EN_finalized.csv')
meta <- read_csv('../data/cleaned_stool/all_samples_meta_p2d_fg9_updated.csv')
```

## microbiome 

```{r genus}
cts <- read_csv('../data/cleaned_stool/ALL_stool_samples_genus_counts.csv') %>% 
  filter(sampleid %in% meta$sampleid) 

thre <- seq(0.0001, 0.002, 0.0001)
thre %>% 
  set_names(thre) %>% 
  map_dfr(function(num){
    cts %>% 
    group_by(genus) %>% 
    count(relab > num) %>% 
    rename(criteria = names(.)[2]) %>% 
    filter(criteria == 'TRUE') %>% 
    arrange(-n) %>% 
    filter(genus != 'NA') %>% 
    mutate(perc = round(n/nrow(meta)*100, 0)) %>% 
    filter(perc > 10) %>% 
      nrow
  }) %>% 
  gather('thre', 'num')

target_genera <-  cts %>% 
  group_by(genus) %>% 
  count(relab > 0.002) %>% 
  rename(criteria = names(.)[2]) %>% 
  filter(criteria == 'TRUE') %>% 
  arrange(-n) %>% 
  filter(genus != 'NA') %>% 
  mutate(perc = round(n/nrow(meta)*100, 0)) %>% 
  filter(perc > 10) %>% 
  pull(genus) 

gcts <- cts %>% 
  filter(genus %in% target_genera) %>% 
  mutate(relablog = log10(relab + 2*10^-6)) %>% 
  dplyr::select(-relab) %>% 
  spread(key = 'genus', value = 'relablog', fill = 0)
```

## food and p2d average in cal and nutrients

```{r p2d_cal}
# the daily caloric intake :
all_daily_pt <- dtb %>%
  group_by(mrn, fdrt) %>%
  summarise(total_daily = sum(Calories_kcal))

# the previous two days average for each stool sample

stb_pair <- meta %>%  
  select(mrn, sdrt) %>% 
  transmute(mrn = mrn,
            p1d = sdrt-1,
            p2d = sdrt-2) 

mean_p2d_cal <-  function(mrn_, p1d_, p2d_){
  df = all_daily_pt %>% 
    filter(mrn == mrn_) %>% 
    filter(fdrt %in% c(p1d_, p2d_  )) %>% 
    group_by(mrn) %>% 
    summarise(ave_cal = sum(total_daily)/2)
  return(df)
}

mean_p2d_df_cal <- pmap(stb_pair, function(mrn, p1d, p2d){
    mean_p2d_cal(mrn, p1d, p2d)
  }) %>% 
  set_names(meta %>% pull(sampleid)) %>% 
  bind_rows(.id = 'sampleid') 
```

```{r p2d_nutrients}
# the daily total of each nutrients 
nutrients <- dtb %>%
  select(mrn, fdrt, Protein_g:Sugars_g) %>% 
  gather('type','gram', Protein_g:Sugars_g) %>% 
  group_by(mrn, fdrt, type) %>%
  summarise(total_daily = sum(gram)) %>% 
  split(.$type)

mean_p2d_fiber <-  function(mrn_, p1d_, p2d_){
  df = nutrients$Fibers_g %>% 
    filter(mrn == mrn_) %>% 
    filter(fdrt %in% c(p1d_, p2d_  )) %>% 
    group_by(mrn) %>% 
    summarise(ave_fiber = sum(total_daily)/2)
  return(df)
}
mean_p2d_carb <-  function(mrn_, p1d_, p2d_){
  df = nutrients$Carbohydrates_g %>% 
    filter(mrn == mrn_) %>% 
    filter(fdrt %in% c(p1d_, p2d_  )) %>% 
    group_by(mrn) %>% 
    summarise(ave_carb = sum(total_daily)/2)
  return(df)
}

mean_p2d_fat <-  function(mrn_, p1d_, p2d_){
  df = nutrients$Fat_g %>% 
    filter(mrn == mrn_) %>% 
    filter(fdrt %in% c(p1d_, p2d_  )) %>% 
    group_by(mrn) %>% 
    summarise(ave_fat = sum(total_daily)/2)
  return(df)
}
mean_p2d_Protein <-  function(mrn_, p1d_, p2d_){
  df = nutrients$Protein_g %>% 
    filter(mrn == mrn_) %>% 
    filter(fdrt %in% c(p1d_, p2d_  )) %>% 
    group_by(mrn) %>% 
    summarise(ave_Protein = sum(total_daily)/2)
  return(df)
}
mean_p2d_Sugars <-  function(mrn_, p1d_, p2d_){
  df = nutrients$Sugars_g %>% 
    filter(mrn == mrn_) %>% 
    filter(fdrt %in% c(p1d_, p2d_  )) %>% 
    group_by(mrn) %>% 
    summarise(ave_Sugars = sum(total_daily)/2)
  return(df)
}

mean_p2d_fiber <- pmap(stb_pair, function(mrn, p1d, p2d){
    mean_p2d_fiber(mrn, p1d, p2d)
  }) %>% 
  set_names(meta %>% pull(sampleid)) %>% 
  bind_rows(.id = 'sampleid') 

mean_p2d_fat <- pmap(stb_pair, function(mrn, p1d, p2d){
    mean_p2d_fat(mrn, p1d, p2d)
  }) %>% 
  set_names(meta %>% pull(sampleid)) %>% 
  bind_rows(.id = 'sampleid') 

mean_p2d_Protein <- pmap(stb_pair, function(mrn, p1d, p2d){
    mean_p2d_Protein(mrn, p1d, p2d)
  }) %>% 
  set_names(meta %>% pull(sampleid)) %>% 
  bind_rows(.id = 'sampleid') 

mean_p2d_Sugars <- pmap(stb_pair, function(mrn, p1d, p2d){
    mean_p2d_Sugars(mrn, p1d, p2d)
  }) %>% 
  set_names(meta %>% pull(sampleid)) %>% 
  bind_rows(.id = 'sampleid') 

mean_p2d_carb <- pmap(stb_pair, function(mrn, p1d, p2d){
    mean_p2d_carb(mrn, p1d, p2d)
  }) %>% 
  set_names(meta %>% pull(sampleid)) %>% 
  bind_rows(.id = 'sampleid') 

p2d_diet <- mean_p2d_df_cal %>% 
  full_join(mean_p2d_fiber, by = c("sampleid", "mrn")) %>% 
  full_join(mean_p2d_fat, by = c("sampleid", "mrn")) %>% 
  full_join(mean_p2d_Protein, by = c("sampleid", "mrn")) %>% 
  full_join(mean_p2d_Sugars, by = c("sampleid", "mrn")) %>% 
  full_join(mean_p2d_carb, by = c("sampleid", "mrn")) 

full <- meta %>% 
  full_join(p2d_diet, by = c("sampleid", "mrn")) %>% 
  full_join(gcts, by = "sampleid") %>% 
  mutate(timebin = cut_width(sdrt, 7, boundary=0, closed = 'left')) %>% 
  mutate(mrn = factor(mrn)) %>% 
  mutate(ave_fiber = ave_fiber/100,
         ave_fat = ave_fat/100,
         ave_Protein = ave_Protein/100,
         ave_Sugars = ave_Sugars/100,
         ave_carb = ave_carb/100,
         ave_cal = ave_cal/1000
         ) %>% 
  mutate(inten_non = if_else(intensity == 'nonablative', 1, 0),
         inten_ab = if_else(intensity == 'ablative', 1, 0),
         inten_re = if_else(intensity == 'reduced', 1, 0))

full %>% 
  mutate(ave_fat_n_pro = ave_fat + ave_Protein) %>% 
  select(-fat_n_pro ) %>% 
  write_csv('../data/cleaned_stool/all_samples_meta_p2d_fg9_dietall_genera36.csv')

full <- read_csv('../data/cleaned_stool/all_samples_meta_p2d_fg9_dietall_genera36.csv')
summary(full)
```

```{r}
library(psych)
macro_columns4 <- full %>% 
  select(starts_with('ave')) %>% 
  select(-ave_carb, -ave_cal)

pairs.panels(macro_columns4, 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
             )
ggsave('../figs/078_macro_4_pair_cor.pdf')
```

```{r}
# to solve the high correlation between the fat and protein, combine them
macro_columns <- full %>% 
  select(starts_with('ave')) %>% 
  select(-ave_carb, -ave_cal) %>% 
  mutate(fat_n_pro = ave_fat + ave_Protein) %>% 
  select(-ave_fat, -ave_Protein)


pairs.panels(macro_columns, 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
             )
```

## modeling part 

### simpson reciprocal ~ p2d_cal + empirical + intensity + TPN + EN + ...

```{r}
alpha_cal <- log(simpson_reciprocal) ~ 1 +
                ave_cal +
               inten_non + inten_ab + inten_re +
               EN +
                TPN +
                empirical +
               (1 | mrn) +
                (1 | timebin)

priors_alpha_cal <- c(
            prior(normal(0, 0.5), class = 'b', coef = "ave_cal"),
            # for the TPN 
            prior(normal(0, 0.1), class = 'b', coef = "TPNTRUE"),
            # for the EN
            prior(normal(0, 0.1), class = 'b', coef = "ENTRUE"),
            # for the empirical 
            prior(normal(0, 0.6), class = 'b', coef = "empiricalTRUE"),
            # for the intensity 
            prior(normal(0, 0.1), class = 'b', coef = "inten_re"),
            prior(normal(0, 0.1), class = 'b', coef = "inten_ab"),
            prior(normal(0, 0.1), class = 'b', coef = "inten_non"),
            # for the intercept
            prior(normal(2, 0.1), class = 'Intercept'))

# vet the prior 
model_alpha_cal  <- brm( alpha_cal,  
              data = full, 
              warmup = 1000, iter = 3000, 
              prior = priors_alpha_cal,
              cores = ncores,
              chains = 2, 
              seed = 123, sample_prior = T)

mean_ave_cal <- mean(full$ave_cal)

mean_ave_cal_prior <- prior_draws(model_alpha_cal) %>% 
  mutate(mean_ave_cal_div = Intercept + 
          b_ave_cal * mean_ave_cal+
           b_inten_re ) %>% 
  select(mean_ave_cal_div) %>% 
  mutate(grp = 'mean_ave_cal') %>% 
  rename(prior_res = mean_ave_cal_div) %>% 
  select(grp, prior_res)

mean_ave_cal_prior %>% 
  count(prior_res >= min(full$simpson_reciprocal) & prior_res <= max(full$simpson_reciprocal)) %>% 
  mutate(fraction = round(n /sum(n), 2))  %>% 
  slice(2) %>% 
  pull(fraction)

# look at the post
post_coeff <- posterior_samples(model_alpha_cal, '^b_') %>% 
  select(starts_with('b_ave')) %>% 
  gather('item', 'coeff') %>% 
  mutate(item = str_replace(item, 'b_ave_','')) %>% 
  mutate(item = fct_reorder(item, coeff, .fun=median, .desc = F)) 

post_coeff %>% 
  ggplot(aes(x = coeff, y = item)) +
  stat_pointinterval(.width = c(.66, .95)) +
  geom_vline(xintercept = 0, col = 'red', linetype = 'dashed') +
  labs(x = 'Regression coefficients',
       y = 'Calories') +
  theme_classic() +
  theme(legend.position = 'none') +
  theme(axis.text=element_text(size=axis_text_size),
        axis.title=element_text(size=axis_title_size),
        aspect.ratio=1)

ggsave('../figs/paper/078_cal_alpha_model_coeff_forest.eps',  
       width = 80,
       height = 80,
         #height = 60,
         units = c("mm"),
         dpi = 400, device = 'eps') 

ggsave('../figs/paper/078_cal_alpha_model_coeff_forest.jpg',  
       width = 80,
       height = 80,
         #height = 60,
         units = c("mm"),
         dpi = 400, device = 'jpg') 

```

### genus ~ p2d_cal + empirical + intensity + TPN + EN + ...

```{r}
Enterococcus_cal <- Enterococcus ~ 1 +
                ave_cal +
               inten_non + inten_ab + inten_re +
               EN +
                TPN +
                empirical +
               (1 | mrn) +
                (1 | timebin)

priors_genus_cal <- c(
            prior(normal(0, 0.5), class = 'b', coef = "ave_cal"),
            # for the TPN 
            prior(normal(0, 0.1), class = 'b', coef = "TPNTRUE"),
            # for the EN
            prior(normal(0, 0.1), class = 'b', coef = "ENTRUE"),
            # for the empirical 
            prior(normal(0, 0.6), class = 'b', coef = "empiricalTRUE"),
            # for the intensity 
            prior(normal(0, 0.1), class = 'b', coef = "inten_re"),
            prior(normal(0, 0.1), class = 'b', coef = "inten_ab"),
            prior(normal(0, 0.1), class = 'b', coef = "inten_non"),
            # for the intercept
            prior(normal(-3, 1), class = 'Intercept'))

# vet the prior 
model_Enterococcus_cal  <- brm( Enterococcus_cal,  
              data = full, 
              warmup = 1000, iter = 3000, 
              prior = priors_genus_cal,
              cores = ncores,
              control = list(adapt_delta = 0.99),
              chains = 2, 
              seed = 123, sample_prior = T)

mean_ave_cal <- mean(full$ave_cal)

mean_Enterococcus_cal_prior <- prior_draws(model_Enterococcus_cal) %>% 
  mutate(mean_ave_Enterococcus_div = Intercept + 
          b_ave_cal * mean_ave_cal+
           b_inten_re ) %>% 
  select(mean_ave_Enterococcus_div) %>% 
  mutate(grp = 'mean_ave_genus') %>% 
  rename(prior_res = mean_ave_Enterococcus_div) %>% 
  select(grp, prior_res)

mean_Enterococcus_cal_prior %>% 
  count(prior_res >= min(full$Enterococcus) & prior_res <= max(full$Enterococcus)) %>% 
  mutate(fraction = round(n /sum(n), 2))  %>% 
  slice(2) %>% 
  pull(fraction)
```

```{r}
# doing it on all the genera
ret_genus_cal <- target_genera %>% 
  set_names(target_genera) %>% 
  purrr::map(function(genus) {
    mod =  brm( as.formula(str_glue('{genus}  ~ 
                1 +
                ave_cal +
                inten_non + inten_ab + inten_re +
                empirical+
                TPN+
                EN+
               (1 | mrn) +
                (1 | timebin)')),  
                data = full, 
              warmup = 1000, iter = 3000, 
              prior = priors_genus_cal,
              cores = ncores, 
              chains = 2, 
              control = list(adapt_delta = 0.99),
              seed = 456, sample_prior = T) 
  })

# save the data
prior_df <- ret_genus_cal %>% 
  imap(~ prior_draws(.x) ) %>% 
  bind_rows(.id = 'genus')
prior_df %>% 
  write_csv('../data/078_genus_cal_model_prior.csv')

post_df <- ret_genus_cal %>% 
  imap(~ (suppressWarnings(posterior_samples(.x)) %>% 
            select(-starts_with('r_'))))  %>% 
  bind_rows(.id = 'genus')  
post_df %>% 
  write_csv('../data/078_genus_cal_model_post.csv')
```

```{r}
# look at the post
post_df <- read_csv('../data/078_genus_cal_model_post.csv') 

post_res_genera <- post_df %>% 
  select(genus, starts_with('b_ave')) %>% 
  gather('item','coeff', names(.)[2]:names(.)[ncol(.)]) %>% 
  group_by(genus, item) %>% 
  summarise(q50 = median(coeff),
            q2.5 = quantile(coeff, probs = 0.025),
            q97.5 = quantile(coeff, probs = 0.975),
            q12.5 = quantile(coeff, probs = 0.125),
            q87.5 = quantile(coeff, probs = 0.875),
            q1.25 = quantile(coeff, probs = 0.0125),
            q98.75 = quantile(coeff, probs = 0.9875),
            q0.5 = quantile(coeff, probs = 0.005),
            q99.5 = quantile(coeff, probs = 0.995)
            ) %>% 
  ungroup() %>% 
  mutate(mark = if_else(q99.5 < 0 | q0.5 > 0, '***', if_else(q98.75 < 0 | q1.25 > 0, '**', if_else(q97.5 < 0 | q2.5 > 0, '*', '')))) %>% 
  mutate(color = if_else(q87.5 < 0, 'steelblue', if_else(q12.5 > 0, 'maroon', 'white'))) %>% 
  mutate(item = str_replace(item, 'b_ave_','')) %>% 
  mutate(item = str_to_title(item))

col_key <- post_res_genera %>% 
  ungroup() %>% 
  distinct(color) %>% 
  pull(color)
names(col_key) <- col_key
 
ggplot(post_res_genera, aes(x = genus, y = item)) +
  geom_tile(aes(fill = color,  x = genus, y =  item), alpha = 0.5, color='white', width=0.95, height=0.95) +
  geom_text(aes(label = mark, x = genus,y =  item),
            nudge_y = -0.1, nudge_x = 0,size = 5) +
  scale_fill_manual(values = col_key, labels = c('Less than 75% CI crosses 0','75% CI < 0 negative', '75% CI > 0 positive')) +
  theme_light() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(angle=50, hjust=1, size=11),
        axis.text.y=element_text(size=11),
        axis.title=element_text(size=axis_title_size),
        #panel.grid.major = element_blank(), 
        #panel.grid.minor = element_blank(),
        legend.position = 'top',
        legend.key = element_rect( colour = "gray50"),
        panel.background = element_blank()) 

ggsave('../figs/paper/078_heatmap_genus_36_cal_ggplot.pdf', width = 8, height = 3)
ggsave('../figs/paper/078_heatmap_genus_36_cal_ggplot.jpg', width = 8, height = 3)
```

### simpson reciprocal ~ ave_fiber + empirical + intensity + TPN + EN + ...

```{r}
alpha_fiber <- log(simpson_reciprocal) ~ 1 +
                ave_fiber +
               inten_non + inten_ab + inten_re +
               EN +
                TPN +
                empirical +
               (1 | mrn) +
                (1 | timebin)

priors_alpha_fiber <- c(
            # for the macro nutrients
            prior(normal(0, 1), class = 'b', coef = "ave_fiber"),
            # for the TPN 
            prior(normal(0, 0.1), class = 'b', coef = "TPNTRUE"),
            # for the EN
            prior(normal(0, 0.1), class = 'b', coef = "ENTRUE"),
            # for the empirical 
            prior(normal(0, 0.5), class = 'b', coef = "empiricalTRUE"),
            # for the intensity 
            prior(normal(0, 0.1), class = 'b', coef = "inten_re"),
            prior(normal(0, 0.1), class = 'b', coef = "inten_ab"),
            prior(normal(0, 0.1), class = 'b', coef = "inten_non"),
            # for the intercept
            prior(normal(2, 0.5), class = 'Intercept'))

# vet the prior 
model_alpha_fiber  <- brm( alpha_fiber,  
              data = full, 
              warmup = 1000, iter = 3000, 
              prior = priors_alpha_fiber,
               control = list(adapt_delta = 0.99),
              cores = ncores,
              chains = 2, 
              seed = 123, sample_prior = T)

mean_macro <-  full %>% 
  select(starts_with('ave_')) %>% 
  summarise_all(funs(mean))

mean_ave_fiber_prior <- prior_draws(model_alpha_fiber) %>% 
  mutate(mean_div = Intercept + 
          b_ave_fiber * mean_macro$ave_fiber+
           b_inten_re ) %>% 
  select(mean_div) %>% 
  mutate(grp = 'mean_macro') %>% 
  rename(prior_res = mean_div) %>% 
  select(grp, prior_res)
```

```{r}
# plot the above model post coeff
post_coeff <- posterior_samples(model_alpha_fiber, '^b_') %>% 
  select(starts_with('b_ave')) %>% 
  gather('item', 'coeff') %>% 
  mutate(item = str_replace(item, 'b_ave_','')) %>% 
   mutate(item = case_when(
    item ==  'fiber' ~ 'Fibers',
    TRUE ~ item
  ))  %>% 
  mutate(item = fct_reorder(item, coeff, .fun=median, .desc = F)) 

post_coeff %>% 
  ggplot(aes(x = coeff, y = item)) +
  stat_pointinterval(.width = c(.66, .95)) +
  geom_vline(xintercept = 0, col = 'red', linetype = 'dashed') +
  labs(x = 'Regression coefficients',
       y = 'Macronutrients') +
  theme_classic() +
  theme(legend.position = 'none') +
  theme(axis.text=element_text(size=axis_text_size),
        axis.title=element_text(size=axis_title_size),
        aspect.ratio=1)

ggsave('../figs/paper/078_fiber_alpha_model_coeff_forest.eps',  
       width = 80,
       height = 80,
         #height = 60,
         units = c("mm"),
         dpi = 400, device = 'eps')

coeff_bi <- post_samples %>% 
  select(!starts_with('b_fg')) %>% 
  select(-b_Intercept) %>% 
  gather('item', 'coeff') %>% 
  mutate(item_name = case_when(
    item ==  'b_inten_non' ~ 'Intensity: nonablative',
    item == 'b_inten_ab' ~ 'Intensity: ablative',
    item ==  'b_inten_re' ~ 'Intensity: reduced',
    item ==  'b_empiricalTRUE' ~ 'Empirical abx exposure',
    item == 'b_TPNTRUE' ~ 'TPN exposure',
    item ==  'b_ENTRUE' ~ 'EN exposure'
  )) %>% 
  mutate(item = factor(item, levels = c('b_empiF','b_empiT','b_inten_non','b_inten_re','b_inten_ab','b_enF','b_enT','b_tpnF','b_tpnT'))) %>% 
  mutate(item_name = fct_reorder(item_name, coeff, .fun=median, .desc = F)) 

coeff_bi %>% 
  ggplot(aes(x = coeff, y = item_name)) +
  stat_pointinterval(.width = c(.66, .95)) +
  #scale_color_manual(values = c('#EC0000','#00468B')) +
  geom_vline(xintercept = 0, col = 'red', linetype = 'dashed') +
  labs(x = 'Regression coefficients',
       y = 'Binary variables') +
  theme_classic() +
  theme(legend.position = 'none',
        aspect.ratio=1)
```


### simpson reciprocal ~ ave_fiber + ave_fat_n_pro + ave_Sugars + empirical + intensity + TPN + EN + ...

```{r}
alpha_macro <- log(simpson_reciprocal) ~ 1 +
                ave_fiber +
                ave_fat_n_pro +
                ave_Sugars +
               inten_non + inten_ab + inten_re +
               EN +
                TPN +
                empirical +
               (1 | mrn) +
                (1 | timebin)

priors_alpha_macro <- c(
            # for the macro nutrients
            prior(normal(0, 0.4), class = 'b', coef = "ave_fiber"),
            prior(normal(0, 0.4), class = 'b', coef = "ave_fat_n_pro"),
            prior(normal(0, 0.4), class = 'b', coef = "ave_Sugars"),
            # for the TPN 
            prior(normal(0, 0.1), class = 'b', coef = "TPNTRUE"),
            # for the EN
            prior(normal(0, 0.1), class = 'b', coef = "ENTRUE"),
            # for the empirical 
            prior(normal(0, 0.5), class = 'b', coef = "empiricalTRUE"),
            # for the intensity 
            prior(normal(0, 0.1), class = 'b', coef = "inten_re"),
            prior(normal(0, 0.1), class = 'b', coef = "inten_ab"),
            prior(normal(0, 0.1), class = 'b', coef = "inten_non"),
            # for the intercept
            prior(normal(2, 0.5), class = 'Intercept'))

# vet the prior 
model_alpha_macro  <- brm( alpha_macro,  
              data = full, 
              warmup = 1000, iter = 3000, 
              prior = priors_alpha_macro,
               control = list(adapt_delta = 0.99),
              cores = ncores,
              chains = 2, 
              seed = 123, sample_prior = T)

mean_macro <-  full %>% 
  select(starts_with('ave_')) %>% 
  summarise_all(funs(mean))

mean_ave_macro_prior <- prior_draws(model_alpha_macro) %>% 
  mutate(mean_div = Intercept + 
          b_ave_fiber * mean_macro$ave_fiber+
           b_ave_fat_n_pro * mean_macro$ave_fat_n_pro+
           b_ave_Sugars * mean_macro$ave_Sugars+
           b_inten_re ) %>% 
  select(mean_div) %>% 
  mutate(grp = 'mean_macro') %>% 
  rename(prior_res = mean_div) %>% 
  select(grp, prior_res)


# the posterior
model_alpha_macro
colnames(full)
```

```{r}
# plot the above model post coeff
post_coeff <- posterior_samples(model_alpha_macro, '^b_') %>% 
  select(starts_with('b_ave')) %>% 
  gather('item', 'coeff') %>% 
  mutate(item = str_replace(item, 'b_ave_','')) %>% 
   mutate(item = case_when(
    item ==  'fiber' ~ 'Fibers',
    item == 'fat' ~ 'Fat',
    TRUE ~ item
  ))  %>% 
  mutate(item = fct_reorder(item, coeff, .fun=median, .desc = F)) 

post_coeff %>% 
  ggplot(aes(x = coeff, y = item)) +
  stat_pointinterval(.width = c(.66, .95)) +
  geom_vline(xintercept = 0, col = 'red', linetype = 'dashed') +
  labs(x = 'Regression coefficients',
       y = 'Macronutrients') +
  theme_classic() +
  theme(legend.position = 'none') +
  theme(axis.text=element_text(size=axis_text_size),
        axis.title=element_text(size=axis_title_size),
        aspect.ratio=1)

ggsave('../figs/paper/078_macronutrients_alpha_model_coeff_forest.eps',  
       width = 80,
       height = 80,
         #height = 60,
         units = c("mm"),
         dpi = 400, device = 'eps')
```

```{r}
# the binary vars
coeff_bi <- posterior_samples(model_alpha_macro, '^b_')  %>% 
  select(!starts_with('b_ave')) %>% 
  select(-b_Intercept) %>% 
  gather('item', 'coeff') %>% 
  mutate(item_name = case_when(
    item ==  'b_inten_non' ~ 'Intensity: nonablative',
    item == 'b_inten_ab' ~ 'Intensity: ablative',
    item ==  'b_inten_re' ~ 'Intensity: reduced',
    item ==  'b_empiricalTRUE' ~ 'Empirical abx exposure',
    item == 'b_TPNTRUE' ~ 'TPN exposure',
    item ==  'b_ENTRUE' ~ 'EN exposure'
  )) %>% 
  mutate(item = factor(item, levels = c('b_empiF','b_empiT','b_inten_non','b_inten_re','b_inten_ab','b_enF','b_enT','b_tpnF','b_tpnT'))) %>% 
  mutate(item_name = fct_reorder(item_name, coeff, .fun=median, .desc = F)) 

coeff_bi %>% 
  ggplot(aes(x = coeff, y = item_name)) +
  stat_pointinterval(.width = c(.66, .95)) +
  #scale_color_manual(values = c('#EC0000','#00468B')) +
  geom_vline(xintercept = 0, col = 'red', linetype = 'dashed') +
  labs(x = 'Regression coefficients',
       y = 'Factor variables') +
  theme_classic() +
  theme(legend.position = 'none',
        aspect.ratio=1)

ggsave('../figs/paper/078_macronutrients_alpha_model_coeff_forest_binary.eps',  
       width = 80,
       height = 80,
         #height = 60,
         units = c("mm"),
         dpi = 400, device = 'eps')
```

### genus ~ ave_fiber + ave_fat_n_pro + ave_Sugars + empirical + intensity + TPN + EN + ...

```{r}
# The outcome being each of the genus
# to test and confirm the prior
Enterococcus_mod <- Enterococcus ~ 1 +
              ave_fiber +
              ave_fat_n_pro +
                ave_Sugars +
               inten_non + inten_ab + inten_re +
                empirical+
                TPN+
                EN+ 
               (1 | mrn) +
                (1 | timebin)

priors_genus_macro <- c(
            # for the macro nutrients
            prior(normal(0, 0.4), class = 'b', coef = "ave_fiber"),
            prior(normal(0, 0.4), class = 'b', coef = "ave_fat_n_pro"),
            prior(normal(0, 0.4), class = 'b', coef = "ave_Sugars"),
            # for the TPN 
            prior(normal(0, 0.1), class = 'b', coef = "TPNTRUE"),
            # for the EN
            prior(normal(0, 0.1), class = 'b', coef = "ENTRUE"),
            # for the empirical 
            prior(normal(0, 0.5), class = 'b', coef = "empiricalTRUE"),
            # for the intensity 
            prior(normal(0, 0.1), class = 'b', coef = "inten_re"),
            prior(normal(0, 0.1), class = 'b', coef = "inten_ab"),
            prior(normal(0, 0.1), class = 'b', coef = "inten_non"),
            # for the intercept
            prior(normal(-3, 1), class = 'Intercept'))


Enterococcus_model <- brm( Enterococcus_mod,  
              data = full, 
              warmup = 1000, iter = 3000, 
              prior = priors_genus_macro,
              cores = ncores, 
              chains = 2, 
              control = list(adapt_delta = 0.99),
              seed = 123, sample_prior = T )



coeff_bi <- post_samples %>% 
  select(!starts_with('b_fg')) %>% 
  select(-b_Intercept) %>% 
  gather('item', 'coeff') %>% 
  mutate(item_name = case_when(
    item ==  'b_inten_non' ~ 'Intensity: nonablative',
    item == 'b_inten_ab' ~ 'Intensity: ablative',
    item ==  'b_inten_re' ~ 'Intensity: reduced',
    item ==  'b_empiricalTRUE' ~ 'Empirical abx exposure',
    item == 'b_TPNTRUE' ~ 'TPN exposure',
    item ==  'b_ENTRUE' ~ 'EN exposure'
  )) %>% 
  mutate(item = factor(item, levels = c('b_empiF','b_empiT','b_inten_non','b_inten_re','b_inten_ab','b_enF','b_enT','b_tpnF','b_tpnT'))) %>% 
  mutate(item_name = fct_reorder(item_name, coeff, .fun=median, .desc = F)) 

coeff_bi %>% 
  ggplot(aes(x = coeff, y = item_name)) +
  stat_pointinterval(.width = c(.66, .95)) +
  #scale_color_manual(values = c('#EC0000','#00468B')) +
  geom_vline(xintercept = 0, col = 'red', linetype = 'dashed') +
  labs(x = 'Regression coefficients',
       y = 'Binary variables') +
  theme_classic() +
  theme(legend.position = 'none',
        aspect.ratio=1)
```


```{r}
# check the priors
mean_enterococcus_prior <- prior_draws(Enterococcus_model) %>% 
  mutate(mean_genus = Intercept + 
          b_ave_fiber * mean_macro$ave_fiber+
           b_ave_fat_n_pro * mean_macro$ave_fat_n_pro+
           b_ave_Sugars * mean_macro$ave_Sugars+
           b_inten_re ) %>% 
  select(mean_genus) %>% 
  mutate(grp = 'mean_enterococcus') %>% 
  rename(prior_res = mean_genus) %>% 
  select(grp, prior_res)

# calculate the percentage that falls within the actual range of the data
mean_enterococcus_prior %>% 
  count(prior_res >= min(full$Enterococcus) & prior_res <= max(full$Enterococcus)) %>% 
  mutate(fraction = round(n /sum(n), 2))  %>% 
  slice(2) %>% 
  pull(fraction)

# look at the post
post_coeff <- posterior_samples(Enterococcus_model, '^b_') %>% 
  select(starts_with('b_ave')) %>% 
  gather('item', 'coeff') %>% 
  mutate(item = str_replace(item, 'b_ave_','')) %>% 
   mutate(item = case_when(
    item ==  'fiber' ~ 'Fibers',
    item == 'fat' ~ 'Fat',
    item ==  'carb' ~ 'Carbohydrates',
    TRUE ~ item
  ))  %>% 
  mutate(item = fct_reorder(item, coeff, .fun=median, .desc = F)) 

post_coeff %>% 
  ggplot(aes(x = coeff, y = item)) +
  stat_pointinterval(.width = c(.66, .95)) +
  geom_vline(xintercept = 0, col = 'red', linetype = 'dashed') +
  labs(x = 'Regression coefficients',
       y = 'Macronutrients') +
  theme_classic() +
  theme(legend.position = 'none') +
  theme(axis.text=element_text(size=axis_text_size),
        axis.title=element_text(size=axis_title_size),
        aspect.ratio=1)
  


```

```{r}
ret_genus_macro <- target_genera %>% 
  set_names(target_genera) %>% 
  purrr::map(function(genus) {
    mod =  brm( as.formula(str_glue('{genus}  ~ 
               1 +
                 ave_fiber +
                ave_fat_n_pro +
                ave_Sugars +
               inten_non + inten_ab + inten_re +
                empirical+
                TPN+
                EN+
               (1 | mrn) +
                (1 | timebin)')),  
                data = full, 
              warmup = 1000, iter = 3000, 
              prior = priors_genus_macro,
              cores = ncores, 
              chains = 2, 
              control = list(adapt_delta = 0.99),
              seed = 456, sample_prior = T) 
  })

# save the data
prior_df <- ret_genus_macro %>% 
  imap(~ prior_draws(.x) ) %>% 
  bind_rows(.id = 'genus')
prior_df %>% 
  write_csv('../data/078_genus_macro_model_prior.csv')

post_df <- ret_genus_macro %>% 
  imap(~ (suppressWarnings(posterior_samples(.x)) %>% 
            select(-starts_with('r_'))))  %>% 
  bind_rows(.id = 'genus')  

post_df %>% 
  write_csv('../data/078_genus_macro_model_post.csv')
```

```{r}
## look at the all genera priors
valid_interval_perc <- ret_genus_macro %>% 
  imap(~ prior_draws(.x) %>% 
            mutate(mean_genera = Intercept + 
                    b_ave_fiber * mean_macro$ave_fiber+
                     b_ave_fat_n_pro * mean_macro$ave_fat_n_pro+
                     b_ave_Sugars * mean_macro$ave_Sugars+
                     b_inten_re ) %>% 
            select(mean_genera) %>% 
            mutate(grp = 'ave_each') %>% 
            rename(prior_res = mean_genera) %>% 
            select(grp, prior_res) %>% 
            count(prior_res > -5 & prior_res < 0) %>% 
            mutate(perc = n/sum(n)*100) %>% 
            select(perc) %>% 
            slice(2) %>% 
            pull(perc) %>% 
            round(1) ) %>% 
  bind_rows(.id = .y) %>% 
  gather('genus','within_interval_perc')
# looks ok to me
```

```{r}
post_df <- read_csv('../data/078_genus_macro_model_post.csv')
  
post_res_genera <- post_df %>% 
  select(genus, starts_with('b_ave')) %>% 
  gather('item','coeff', names(.)[2]:names(.)[ncol(.)]) %>% 
  group_by(genus, item) %>% 
  summarise(q50 = median(coeff),
            q2.5 = quantile(coeff, probs = 0.025),
            q97.5 = quantile(coeff, probs = 0.975),
            q12.5 = quantile(coeff, probs = 0.125),
            q87.5 = quantile(coeff, probs = 0.875),
            q1.25 = quantile(coeff, probs = 0.0125),
            q98.75 = quantile(coeff, probs = 0.9875),
            q0.5 = quantile(coeff, probs = 0.005),
            q99.5 = quantile(coeff, probs = 0.995)
            ) %>% 
  ungroup() %>% 
  mutate(mark = if_else(q99.5 < 0 | q0.5 > 0, '***', if_else(q98.75 < 0 | q1.25 > 0, '**', if_else(q97.5 < 0 | q2.5 > 0, '*', '')))) %>% 
  mutate(color = if_else(q87.5 < 0, 'steelblue', if_else(q12.5 > 0, 'maroon', 'white'))) %>% 
  mutate(item = str_replace(item, 'b_ave_','')) %>% 
  mutate(item = str_to_title(item))
```
```{r}
# trying to order the rows and the columns by similarity in euclidean distances
# make it a numerical matrix and then calculate the distance
dat <- post_res_genera %>% 
  select(genus:q50) %>% 
  spread('item', 'q50') %>% 
  column_to_rownames('genus')

hc <- hclust(dist(dat))
dend <- as.dendrogram(hc)
genus_order <- labels(dend)

# the macro order
dat <- post_res_genera %>% 
  select(genus:q50) %>% 
  spread('genus', 'q50') %>% 
  column_to_rownames('item')

hc <- hclust(dist(dat))
dend <- as.dendrogram(hc)
macro_order <- labels(dend)

post_res_ordered <- post_res_genera %>% 
  mutate(genus = factor(genus, levels = genus_order),
         item = factor(item, levels = macro_order))
```

```{r}
col_key <- post_res_genera %>% 
  ungroup() %>% 
  distinct(color) %>% 
  pull(color)
names(col_key) <- col_key
 
ggplot(post_res_ordered, aes(x = genus, y = item)) +
  geom_tile(aes(fill = color,  x = genus, y =  item), alpha = 0.5, color='white', width=0.95, height=0.95) +
  geom_text(aes(label = mark, x = genus,y =  item),
            nudge_y = -0.1, nudge_x = 0,size = 5) +
  scale_fill_manual(values = col_key, labels = c('Less than 75% CI crosses 0','75% CI > 0 positive', '75% CI < 0 negative')) +
  theme_light() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(angle=50, hjust=1, size=11),
        axis.text.y=element_text(size=11),
        axis.title=element_text(size=axis_title_size),
        #panel.grid.major = element_blank(), 
        #panel.grid.minor = element_blank(),
        legend.position = 'top',
        legend.key = element_rect( colour = "gray50"),
        panel.background = element_blank()) 

ggsave('../figs/paper/078_heatmap_genus_36_macro_ggplot.pdf', width = 8, height = 3)
ggsave('../figs/paper/078_heatmap_genus_36_macro_ggplot.jpg', width = 8, height = 3)
```

```{r}
full %>% 
  mutate(logsugar = log10(ave_Sugars*100) )%>% 
  filter(ave_Sugars < 2) %>% 
  ggscatter(x = 'ave_Sugars', y = 'Enterococcus',
             add = "reg.line",  # Add regressin line
           add.params = list(color = "blue", fill = "lightgray"), # Customize line
           conf.int = TRUE, # Add confidence interval
           cor.coef = TRUE, # Add correlation coefficient.
           cor.coeff.args = list(method = "pearson",  label.sep = "\n")) 
  
```

```{r}
# why: when looking scatter plot of sugar VS enterococcus there is about 0 cor,
# but when including sugar as macronutrients 
```

