---
title: "FOOD diversity visualization: alpha and beta"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
---

```{r setup, include=FALSE, out.width='150%'}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

```{r}
library(tidyverse)
library(ape)
library(vegan)
library(ggpubr)
```

```{r}
# load the pheno data of the patients 
pheno <- read_csv('../data/pheno/final_pheno.csv')
```


# alpha diversity

```{r}
alpha <- read_csv('../data/finalized/alpha_faith_daily_all.csv') %>% 
  left_join(pheno, by = 'mrn')
```

## change over time

```{r}
# see different intensity
alpha %>% 
  ggscatter('DayRT', 'faith_pd', 
            color = 'intensity', palette = 'lancet', alpha = 0.8,
            #ellipse = TRUE, mean.point = TRUE, star.plot = TRUE,
            ylab = 'alpha diversity (Faith)', xlab = 'Day relative to transplant',
            title = 'Food alpha diversity split by intensity',
            add = "loess", conf.int = TRUE)
```

```{r}
# see different source 
alpha %>% 
  ggscatter('DayRT', 'faith_pd', 
            color = 'source', palette = 'jco', alpha = 0.8,
            #ellipse = TRUE, mean.point = TRUE, star.plot = TRUE,
            ylab = 'alpha diversity (Faith)', xlab = 'Day relative to transplant',
            title = 'Food alpha diversity split by source',
            add = "loess", conf.int = TRUE)
```

##### **Food** alpha diversity drops probably because of the chemo, so the patients didn't want to eat. Specifically, the ablative/reduced and cord/TCD have the biggest drop. 

# beta diversity (unweighted unifrac)

```{r}
#Check the beta distance matrix
beta_un_uni <- read.delim('../data/finalized/beta_phylo/distance-matrix.tsv', sep = '\t', row.names = 1)
food_dist <- as.dist(beta_un_uni)

#pcoa
pcoa_f <- as.data.frame(pcoa(food_dist)$vectors) %>% 
  select(Axis.1, Axis.2) %>% 
  rownames_to_column('mrnDay') %>% 
  separate(mrnDay, into = c('mrn','DayRT'), sep = 'd') %>% 
  mutate(DayRT = str_replace(DayRT,'_','-'),
         DayRT = as.numeric(DayRT),
         mrn = as.numeric(mrn))

# the percent variance
eigen <- pcoa(food_dist)$values$Eigenvalues

percent_var <- signif(eigen/sum(eigen), 3)*100
```

```{r}
# join the coords with the pheno
beta_un_uni_df <- pcoa_f %>% 
  left_join(pheno, by  = 'mrn')
```

## PCoA plot for all patients in all time points

### Colored by day relative to transplant

```{r}
# don't split and color all of them by DRT
beta_un_uni_df %>% 
  mutate(mrn = factor(as.character(mrn))) %>% 
ggplot( aes(x = Axis.1, y = Axis.2, col = DayRT)) +
  geom_point(alpha=0.75, size = 2) +
  #stat_ellipse(color = "dark grey", type = "norm", linetype = 2, level = 0.95, alpha = 0.5) +
  xlab(paste0("PC 1 [",percent_var[1],"%]")) +
  ylab(paste0("PC 2 [",percent_var[2],"%]")) +
  theme_classic() +
  theme(legend.position = 'top') +
  labs(title = 'PCoA unifrac distance from all patients and all time points colored by transplant day') +
  scale_color_gradient(low = "yellow", high = "darkblue")
```

##### Not able to see a clear separation

### Color/split by intensity and source 

```{r}
# plot mark the source and the intensity
beta_un_uni_df %>% 
  mutate(mrn = factor(as.character(mrn))) %>% 
ggplot( aes(x = Axis.1, y = Axis.2, shape = source , col = intensity)) +
  geom_point(alpha=0.75, size = 2) +
  #stat_ellipse(color = "dark grey", type = "norm", linetype = 2, level = 0.95, alpha = 0.5) +
  theme_classic() +
  scale_color_brewer(palette = 'Set1') +
  labs(title = 'PCoA unifrac distance from all patients and all time points colored by intensity') +
  facet_wrap(~source) +
  theme(legend.position = 'top') +
  ggsave('../figs/beta_un_unifrac_all_pts_source_intensity.jpg', width = 12, height = 8)

```

##### No separation either

```{r}
# beta_un_uni_df %>% 
#   mutate(mrn = factor(as.character(mrn))) %>% 
# ggplot( aes(x = Axis.1, y = Axis.2,  col = DayRT)) +
#   geom_point(alpha=0.9, size = 2) +
#   #stat_ellipse(color = "dark grey", type = "norm", linetype = 2, level = 0.95, alpha = 0.5) +
#   theme_classic() +
#   scale_color_gradient2(low = "yellow", mid = "red", high = "blue", midpoint = 10) +
#   facet_wrap(intensity~source) +
#   ggsave('../figs/beta_un_unifrac_all_pts_DayRT.jpg', width = 12, height = 8)
```

### Thinking about zooming in on earlier days

```{r}
# look at the distribution of the DayRT cuz have fewer large values
beta_un_uni_df %>% 
  gghistogram('DayRT', bins = 120, fill = 'forestgreen', color = 'white',
              title = 'Distribution of day relative to transplant of the food records',
              xlab = 'Day relative to transplant')
```

##### day -10 to day 15 seem to have most food records.

```{r}
# include the earlier time points records only
beta_un_uni_df %>% 
  filter(DayRT %in% -10:15) %>% 
  mutate(mrn = factor(as.character(mrn))) %>% 
ggplot( aes(x = Axis.1, y = Axis.2,  col = DayRT)) +
  geom_point(alpha=0.9, size = 2) +
  #stat_ellipse(color = "dark grey", type = "norm", linetype = 2, level = 0.95, alpha = 0.5) +
  theme_classic() +
  theme(legend.position = 'top') +
  scale_color_gradient(low = "yellow", high = "darkblue")  +
  labs(title = 'PCoA of unifrac distance between day -10 to day 15') +
  facet_wrap(intensity~source) +
  ggsave('../figs/beta_un_unifrac_all_pts_DayRT_early.jpg', width = 12, height = 8)
```

##### The separation in transplant day is more clear in ablative + TCD

### Statistical testing on the Axis 2 before and after day -2 of all the data

```{r}
beta_un_uni_df %>% 
  mutate(grp = if_else(DayRT <= -2, 'before', 'after')) %>% 
  mutate(grp = factor(grp, levels = c('before','after'))) %>% 
  ggboxplot(x = 'grp', y = 'Axis.2', add = 'jitter', color = 'grp', palette = 'nejm',
            title = 'Axis 2 unifrac distance PCOA between before and after day -2') +
  stat_compare_means(comparisons = list(c('before', 'after')),
                      label = "p.signif",
                      method = 'wilcox.test',
                      correct=FALSE)
```

##### Using day -2 since that's the earliest abx time points. Although it's not visually obvious from the PCoA plot, the statistics show that the food composition is significantly different between the two time period.  

```{r}
# # or log transform the time and see if it will be clearer
# beta_un_uni_df %>% 
#   mutate(DayRT = DayRT + 50,
#          log2DayRT = log2(DayRT)) %>% 
#   mutate(mrn = factor(as.character(mrn))) %>% 
# ggplot( aes(x = Axis.1, y = Axis.2,  col = log2DayRT)) +
#   geom_point(alpha=0.9, size = 2) +
#   #stat_ellipse(color = "dark grey", type = "norm", linetype = 2, level = 0.95, alpha = 0.5) +
#   theme_classic() +
#   scale_color_gradient(low = "yellow", high = "darkblue")  +
#   facet_wrap(intensity~source) +
#   ggsave('../figs/beta_un_unifrac_all_pts_DayRT_log_time.jpg', width = 12, height = 8)
```


```{r out.width="100%"}
# split by each patient and color by day 

# beta_un_uni_df %>% 
#   ggplot(aes(x = Axis.1, y = Axis.2, color = DayRT, shape = intensity)) +
#   geom_point(size = 2) +
#   scale_color_continuous(type = 'viridis') +
#   theme_classic() +
#   facet_wrap(source ~ mrn)  +
#   ggsave('../figs/beta_un_unifrac_all_pts.jpg', width = 20, height = 15)
```

```{r}
# one figure for each patient
# beta_un_uni_df %>% 
#   mutate(source = factor(source)) %>% 
#   split(.$mrn) %>% 
#   map(~ ggplot(data = ., aes(x = Axis.1, y = Axis.2, color = DayRT, shape = intensity, size = source)) +
#   geom_point(size = 2) +
#   scale_color_continuous(type = 'viridis') +
#   theme_classic())
  
```


