---
title: "Check the BSI"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
# see how does the 16 patients overlap here with the 95
combined <- read_csv('../data/growth/069_irep_combined_res.csv')  %>% 
  distinct(mrn)

irep_res <- read_csv('../data/growth/069_irep_combined_res.csv')
```

Note that there are multiple rows of results for each ‘episode’ of a bloodstream infection, because every time someone gets their blood tested, we test many spots around their body (specimen.description) for example the various tubes/lumens they have, often described as colors (red, blue) and peripheral blood, meaning an actual fresh needle stuck into a vein, not a catheter that’s sitting in them for weeks)
also, once someone has bacteria detected, then we keep checking daily until it goes away, so need to pay attention to test_date .  Once it is positive, that only counts as one bloodstream infection event for the next many days (1 week or so).
also, if the result is Coagulase Negative Staphylococcus then we ignore that almost always as a false positive / contaminant.
and some of the results require some manual review because they use various free-text vocabulary terms

```{r}
bsi <- read_csv('/Volumes/vandenBrinkLab/Angel_Dai/Nutrition_project/data/bsi_nutrition_cohort-2022-02-24.csv') %>% 
  filter(result != 'Coagulase Negative Staphylococcus') %>% 
  mutate(mrn = as.numeric(MRN)) %>% 
  select(mrn, test_name, test_date, test_datetime, specimen.description, result, pos)

ptb <- read_csv('../data/cleaned_patients/diet_patients_97.csv')

# how many positive entries
bsi %>% 
  count(pos)

colnames(bsi)

pos_res <- bsi %>% 
  filter(pos == "TRUE") %>% 
  left_join(ptb %>% 
              select(mrn, hct)) %>% 
  mutate(test_day = test_date - hct) %>% 
  mutate(test_day = as.numeric(test_day)) %>% 
  filter(test_day %in% -8:53)  %>% # the time relavant to the nutrition project 
  relocate(test_day, .after = mrn) %>% 
  mutate(pid = str_glue('P{as.numeric(as.factor(mrn))}')) %>% 
  relocate(pid, .after = mrn)  %>% 
  mutate(short_res = str_extract(result, "[^\\s]+")) %>% 
  relocate(short_res, .after = pid)  %>% 
  arrange(pid, test_day) %>% 
  filter(short_res != 'Candida') %>% 
  mutate(irep = if_else(mrn %in% combined$mrn, T, F))
  
pos_res %>% 
  count(mrn, sort = T)


# scrutinize each patient carefully
# look at the test_day distribution
each <- pos_res %>% 
  gghistogram(x = 'test_day', facet.by = 'pid', label = 'short_res', repel = T, fill = 'irep') +
  scale_fill_manual(values = c('white','red'))

ggsave('../data/098_bsi_pos.pdf', width = 10)

# unfortunately P11 that is in f3c does have klebsielle at d40 
# nothing about P klebsiella




```

