---
title: "Filtering batch 2 stool samples"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
library(lubridate)
library(vdbR)
connect_database()
get_table_from_database('samples_castori_ag')
```
```{r}
ptb <- read_csv('../data/129_new_76_full_patient_info.csv')
stb_ <- read_csv('../data/130_batch2_pt_stool_samples.csv')
dtb <- read_rds('../data/136_updated_dtb2.rds')
```

# selecting stool samples for the patients and also close to diet data

```{r}
stb <- stb_ %>% 
  left_join(ptb %>% select(mrn, hct)) %>% 
  mutate(sdrt = datecollection - hct) %>% 
  mutate(sdrt = as.numeric(sdrt))


# remove all the stool samples collected >= 2 days after the latest diet sample for every patient
# get the latest diet sample for every patient
dtb_latest_dsample <- dtb %>% 
  arrange(mrn, desc(fdrt)) %>% 
  distinct(mrn, .keep_all = T) %>% 
  select(mrn, fdrt)  

# join the lastest dsample day to the stb 
stb_dsample_diff_tobe_removed <- stb %>% 
  left_join(dtb_latest_dsample, by = 'mrn')  %>% 
  mutate(diff = sdrt - fdrt) %>% 
  filter(diff >= 2) %>% 
  pull(sampleid)

stb1 <- stb %>% 
  filter(!sampleid %in% stb_dsample_diff_tobe_removed)

# also remove the stool samples that are earlier than the earliest diet sample for each patient
earliest_dsample <- dtb %>% 
  arrange(mrn, fdrt) %>% 
  distinct(mrn, .keep_all = T) %>% 
  select(mrn, fdrt)

early_2b_rm <- stb1 %>% 
  left_join(earliest_dsample, by = 'mrn')  %>% 
  mutate(diff = sdrt - fdrt) %>% 
  filter(diff <= 0) %>% 
  pull(sampleid)

stb2 <- stb1 %>% 
  filter(!sampleid %in% early_2b_rm) %>% 
  select(-hct, -datecollection)
```

```{r}
# now find the samples that  have a satisfying total count
stb3 <- stb2 %>% 
  filter(count_total >= 400)
```

# remove the duplicated samples of the same patient on the same day 

select one of the samples and ignore the other. take the one with the earlier sampleid by alphanumeric sorting

```{r}
stb4 <- stb3 %>% 
  arrange(mrn, sdrt, sampleid) %>% 
  distinct(mrn, sdrt, .keep_all = T)
```

# has to have at least one real diet data in the previous 2 days (for the later pairing)

```{r}
# list the previous 2 days drt for every stool sample 
qual_stool_df <- stb4 %>% 
  mutate(p1d = sdrt-1,
         p2d = sdrt-2) %>% 
  select(mrn, p1d, p2d)



no_real_diet_d_to_be_rm_stool_s <- function(mrn_, p1d_, p2d_){
  df = dtb %>% 
    filter(mrn == mrn_) %>% 
    filter(fdrt == p1d_ | fdrt == p2d_  ) %>% 
    nrow()
  return(df)
}

qual_stool2 <- pmap(qual_stool_df, function(mrn, p1d, p2d){
    no_real_diet_d_to_be_rm_stool_s(mrn, p1d, p2d)
  }) %>% 
  set_names(stb4 %>% pull(sampleid)) %>% 
  bind_rows(.id = 'sampleid') %>% 
  gather(key = 'sampleid', value = 'num_d_records') %>% 
  arrange(num_d_records) 

no_diet_to_be_removed_stool_samples <- qual_stool2 %>% 
  filter(num_d_records == 0) %>% 
  pull(sampleid)
  
stb5 <- stb4 %>% 
  filter(!sampleid %in% no_diet_to_be_removed_stool_samples) 
```

# no missing data in the previous two days before a stool sample collection

```{r}
missingdays <- readxl::read_excel('../data/MissingDataCheck_DietStudy_New80Patients.xlsx') %>% 
  filter(diet_data_status == 'missing') %>% 
  mutate(mrn = as.numeric(mrn),
         date = ymd(date)) %>% 
  left_join(ptb %>% select(mrn, hct))  %>% 
  mutate(missing_int = interval(date, date))


# the previous two days interval for every stool sample
p2d_df <- stb5 %>% 
  select(sampleid, mrn) %>% 
  inner_join(samples_castori_ag %>% select(sampleid, datecollection)) %>% 
  mutate(p1d = datecollection -1,
         p2d = datecollection -2) %>% 
  mutate(p2d_int = interval(p2d, p1d)) %>% 
  select(sampleid, datecollection, mrn, p1d, p2d, p2d_int) %>% 
  inner_join(missingdays, by = "mrn") %>% 
  mutate(had_missing = int_overlaps(p2d_int, missing_int)) %>% 
  filter(had_missing == 'TRUE')

have_missing_diet_need_to_be_rm <- p2d_df %>% 
  pull(sampleid)

# final qualified stool samples
stb6 <- stb5 %>% 
  filter(!sampleid %in% have_missing_diet_need_to_be_rm) 
```

# Clean the antibiotics data from this cohort

```{r}


# censor_abx <- c('active_atb_vanco_po',	'active_atb_imipenem',	'active_atb_meropenem',	'active_atb_ertapenem',		'active_atb_cefepime',		'active_atb_linezolid',	'active_atb_metro',	'active_atb_piptazo')
picked <-  c('vancomycin', 'imipenem_cilastatin','meropenem','ertapenem', 'cefepime', 'linezolid','metronidazole','piperacillin_tazobactam')


med <- read_rds('../data/batch2_medication_exposures.rds') %>% 
  select(MRN:stop_date, drug_name, route) %>% 
  mutate(drug_name_clean = str_replace_all(drug_name, '-', '_')) %>% 
  mutate(drug_name_clean = str_to_lower(drug_name_clean)) %>% 
  mutate(drug_name_clean = case_when(
    drug_name_clean == 'levofloxacin (levaquin)' ~ 'levofloxacin',
    drug_name_clean == 'vancomycin invest' ~ 'vancomycin',
    TRUE ~ drug_name_clean
  )) %>% 
  mutate(mrn = as.numeric(MRN))

now <- med %>% count(drug_name_clean) %>% arrange(desc(n)) 
allroute <- med %>% count(route)

has <- med %>% 
  filter(drug_name_clean %in% picked) %>% 
  distinct(drug_name_clean)

# find all the ones that may have extra part of the strings
allpickes <- tibble(picked = picked) %>% 
  summarise(text = str_c(picked, collapse = "|"))
allhas <- now %>% 
  filter(str_detect(drug_name_clean, allpickes$text))

# no "metronidazole topical"
# no "ciprofloxacin otic" it's for ear
# no "ciprofloxacin_hydrocortisone otic" also for ear

# empirical med exposure
empirical_med_oral <- med %>% 
  filter(drug_name_clean %in% picked) %>% 
  select(-drug_name) %>% 
  filter(route == 'PO') %>% 
  filter(mrn %in% ptb$mrn)



empirical_med_oral %>% write_csv('../data/137_batch2_empircal_oral.csv')
```

```{r}
# find if exposure to oral empirical abx in the two days prior to a stool sample collection
p2d_df <- stb6 %>% 
  select(sampleid, mrn) %>% 
  inner_join(samples_castori_ag %>% select(sampleid, datecollection)) %>% 
  mutate(p1d = datecollection -1,
         p2d = datecollection -2) %>% 
  mutate(p2d_int = interval(p2d, p1d)) %>% 
  select(sampleid, datecollection, mrn, p1d, p2d, p2d_int)

all_empirical <- empirical_med_oral %>% 
  distinct(drug_name_clean) %>% pull(drug_name_clean)

find_exposure <- all_empirical %>% 
  set_names(all_empirical) %>% 
  map_dfr(function(d){
    each_drug = empirical_med_oral %>% 
        filter(drug_name_clean == d) %>% 
        select(mrn, drug_name_clean, start_date, stop_date) %>% 
        mutate(drug_int =  interval(start_date, stop_date)) %>% 
        rename(abx_start = start_date,
               abx_stop = stop_date)
    
    had_drug_res = p2d_df %>% 
        inner_join(each_drug, by = "mrn") %>% 
        mutate(had_drug = int_overlaps(p2d_int, drug_int)) %>% 
        filter(had_drug == 'TRUE')
    
    return(had_drug_res) 
  }) %>%  
  left_join(ptb %>% select(mrn, hct)) %>% 
  mutate(abx_start_day = abx_start - hct,
         abx_stop_day = abx_stop - hct) 

exposed_empirical <- empirical_med_oral %>% 
  mutate(drug_int =  interval(start_date, stop_date)) %>% 
  inner_join(p2d_df, by = "mrn") %>% 
   mutate(had_drug = int_overlaps(p2d_int, drug_int)) %>% 
        filter(had_drug == 'TRUE')
exposed_empirical %>% distinct(sampleid)

# the above samples are flagged for exposed to empirical
stb7 <- stb6 %>% 
  mutate(empirical = if_else(sampleid %in% find_exposure$sampleid, T, F))

stb7 %>% count(empirical) 


# so few samples have oral empirical abx???
```
# find the EN and TPN info of this cohort 

```{r}
# only few patients had TPN
# no patients had EN
support <- read_csv('../data/NutritionSupport_DietPatients.csv') %>% 
  filter(MRN %in% ptb$mrn) %>% 
  mutate(date = mdy(`Date Nutrition Support`)) %>% 
  filter(`Nutrition Support` == 'TPN') %>% 
  select(mrn = MRN,
         date,
         nutrition_support = `Nutrition Support`) %>% 
  mutate(TPN_int = interval(date, date))

# to find if any of those stool samples had TPN in the prior two days
has_tpn <- p2d_df %>% 
  inner_join(support, by = "mrn") %>% 
  mutate(had_tpn = int_overlaps(p2d_int, TPN_int)) %>% 
  filter(had_tpn == 'TRUE')

stb8 <- stb7 %>% 
  mutate(TPN = if_else(sampleid %in% has_tpn$sampleid, T, F)) %>% 
  mutate(EN = F)
 
# join with the intensity level of the patients 
stb9 <- stb8 %>% 
  left_join(ptb %>% select(mrn, intensity), by = "mrn") %>% 
   mutate(intensity = case_when(
    intensity == 'Ablative' ~ 'ablative',
    intensity == 'Reduced Intensity' ~ 'reduced',
    intensity == 'Nonablative' ~ 'nonablative'
  ))

stb9 %>% write_rds('../data/137_stb.rds')
```


