---
title: "Two batches comparison"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

```{r}
library(tidyverse)
library(ggpubr)
```

# please summarize how many patients and samples were exposed to our list of key antibiotics in batch 1 vs batch2 and list the antibiotics.

The list of antibiotics: 'vancomycin', 'imipenem_cilastatin','meropenem','ertapenem', 'cefepime', 'linezolid','metronidazole','piperacillin_tazobactam'

```{r}
# Below is the code I did to find out the exposure to the abx we picked orally
library(lubridate)
library(vdbR)
connect_database()
get_table_from_database('samples_castori_ag')
# the table that has the stool samples qualified to be in the model
meta <- read_csv('../data/cleaned_stool/all_samples_meta_p2d_fg9_updated.csv')

# the table that has the patient information
ptb <- read_csv('../data/cleaned_patients/diet_patients_97.csv') %>% 
  mutate(patient_in_model = if_else(mrn %in% meta$mrn, T, F))

# below is the empirical abx that we picked 
picked <-  c('vancomycin', 'imipenem_cilastatin','meropenem','ertapenem', 'cefepime', 'linezolid','metronidazole','piperacillin_tazobactam')

# the below rds is the medication data Tsoni pulled on Nov 29 for batch 1 
# clean the drug names
med1 <- read_rds('../data/tasks_nutrition_batch1_2022-11-29.rds') %>% 
  select(MRN, start_date, stop_date, drug_name, route) %>% 
  mutate(drug_name_clean = str_replace_all(drug_name, '-', '_')) %>% 
  mutate(drug_name_clean = str_to_lower(drug_name_clean),
         mrn = as.numeric(MRN)) %>% 
  filter(mrn %in% ptb$mrn)

# select only the entries that represent the oral exposure to the empirical drugs
empirical_med_oral1 <- med1 %>% 
  filter(drug_name_clean %in% picked) %>% 
  select(-drug_name) %>% 
  filter(route == 'oral') %>% 
  filter(mrn %in% ptb$mrn)

# make a time interval that means the day -2 to day -1 for each qualified stool sample
p2d_df <- meta %>% 
  select(sampleid, mrn) %>% 
  inner_join(samples_castori_ag %>% select(sampleid, datecollection)) %>% 
  mutate(p1d = datecollection -1,
         p2d = datecollection -2) %>% 
  mutate(p2d_int = interval(p2d, p1d)) %>% 
  select(sampleid, datecollection, mrn, p1d, p2d, p2d_int)

# find if there is an overlap between the prior two days of every stool sample and the administering of the oral empirical drugs, and identify the exposed stool samples
exposed_empirical <- empirical_med_oral1 %>% 
  mutate(drug_int =  interval(start_date, stop_date)) %>% 
  inner_join(p2d_df, by = "mrn") %>% 
   mutate(had_drug = int_overlaps(p2d_int, drug_int)) %>% 
        filter(had_drug == 'TRUE')

exposed_empirical %>% distinct(sampleid) %>% nrow
exposed_empirical %>% distinct(MRN) %>% nrow
```

```{r}
# for batch 2, the same method is used 
stb2 <- read_rds('../data/137_stb.rds')

stb2_exposed <- stb2 %>% 
  filter(empirical == 'TRUE')

stb2_exposed %>% distinct(sampleid) %>% nrow
stb2_exposed %>% distinct(mrn) %>% nrow
```

**In summary, 98 samples from 15 patients were exposed to empirical drugs in batch 1; 11 samples from 7 patients were exposed in batch 2**

# please plot the diet data from both batches on the same UMAP and color  them according to batch1 vs batch 2 to see if they are superimposable or separate

```{r}
# prepare the tables for taxUMAP


dtb1 <- read_csv('../data/cleaned_diet_data/FINAL_97_with_code_all_foods_drt_with_EN_finalized.csv')
dtb2 <- read_rds('../data/136_updated_dtb2.rds') %>% mutate(Food_code = as.numeric(Food_code))
all.equal(colnames(dtb1), colnames(dtb2))
# the weight need to be relative so needs to normalize them 


fc_table <- bind_rows(dtb1, dtb2) %>% 
  select(mrn, fdrt, Food_code, dehydrated_weight) %>% 
  group_by(mrn, fdrt, Food_code) %>% 
  summarise(total = sum(dehydrated_weight)) %>% 
  ungroup() %>% 
  mutate(index_column = str_glue('P{mrn}_{fdrt}')) %>% 
  select(index_column, Food_code, total)

fc_total <- fc_table %>% 
  group_by(index_column) %>% 
  summarise(daily_total = sum(total))

fc_df <- fc_table %>% 
  left_join(fc_total) %>% 
  mutate(daily_relative = total/daily_total) %>% 
  select(index_column,Food_code, daily_relative ) %>% 
  spread('Food_code', 'daily_relative', fill = 0)

fc_df %>% write_csv('../data/143_food_code_relative.csv')
```
```{r}
actual_foodids <- as.character(fc_table$Food_code)

final.table <- read_csv('../data/source/final_table_for_writing_out_to_newick.csv') %>% 
  mutate(FoodID = as.character(FoodID))

food_taxa <- final.table %>% 
  filter(FoodID %in% actual_foodids) %>% 
  select(ASV = FoodID, Kingdom = L1, Phylum = L2, Class = L3, Order = L4) 

food_taxa %>% write_csv('../data/143_food_taxa.csv')
```

```{r}
 # ~/pipeline/scripts/food_tree/softwares/taxumap/run_taxumap.py -t 143_food_taxa.csv -m 143_food_code_relative.csv --agg_levels Kingdom/Phylum -n 173 --weights 1/0.5 --outdir batch_all_taxumap_kingdom_and_phylum 
```

```{r}
# the resulted coords from taxUMAP
pt1 <- unique(dtb1$mrn)
pt2 <- unique(dtb2$mrn)

out <- read_csv('../data/batch_all_taxumap_kingdom_and_phylum/taxumap_embedding.csv') %>% 
  separate(index_column, into = c('mrn','fdrt'), sep = '_') %>% 
  mutate(mrn = as.numeric(str_replace(mrn, 'P','')),
         fdrt = as.numeric(fdrt)) %>% 
  mutate(batch = if_else(mrn %in% pt1, 'batch1', if_else(mrn %in% pt2, 'batch2', 'NA')))

# to see the distribution of the two batches
out %>% 
  arrange(fdrt ) %>% 
  ggscatter(x = 'taxumap1', y = 'taxumap2', color = 'batch', palette = 'nejm', alpha = 0.1, facet.by = 'batch') +
  theme(aspect.ratio = 1, legend.position = 'right')
```

**It looks the two batches overlap well, and don't have fundamental difference**


# please plot fecal alpha diversity over time from both batch 1 and batch 2 side by side (can be faceted or on the same plot).

```{r}
meta1  <- meta 
meta2 <- read_rds('../data/137_stb.rds')

stb_both <- bind_rows(
  meta1 %>% select(sampleid, sdrt, simpson_reciprocal, intensity, mrn ) %>% mutate(batch = 'batch1'),
  meta2 %>% select(sampleid, sdrt, simpson_reciprocal , intensity, mrn) %>% mutate(batch = 'batch2')
)

stb_both %>% 
  ggscatter(x = 'sdrt', y = 'simpson_reciprocal', alpha = 0.5, size = 1, color = 'batch',palette = 'nejm',
            xlab = 'Day relative to HCT', shape = 'triangle',
            ylab = 'Microbiome alpha diversity',
            title = 'Microbiome alpha diversity',
            add = "loess", conf.int = TRUE, 
            add.params = list(color = 'batch',  size = 1)) +
  theme(aspect.ratio=1, legend.position = 'right') 
```
```{r}
# split then to two panels each to different conditioning intensity
stb_both %>% 
  ggscatter(x = 'sdrt', y = 'simpson_reciprocal', alpha = 0.5, size = 1, color = 'intensity',palette = 'lancet',
            facet.by = 'batch',
            xlab = 'Day relative to HCT', shape = 'triangle',
            ylab = 'Microbiome alpha diversity',
            title = 'Microbiome alpha diversity',
            add = "loess", conf.int = TRUE, 
            add.params = list(color = 'intensity',  size = 1)) +
  #facet_grid(~ batch) +
  theme(aspect.ratio=1, legend.position = 'right') 

stb_both %>% 
  count(batch)

stb_both %>% 
  distinct(mrn, batch) %>% 
  count(batch)
```

