---
title: "Pair the stool with previous N day of diet for every patient"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

# load the tables

```{r}
# load the current dtable(diet) and stb(stool)
dtb <- read_tsv('../data/finalized/meta_data_67.tsv')

# for stool samples, now only looking at samples marked with "keep"
stb <- read_csv('../data/finalized/stool/filter_stool_close_to_diet_meta.csv') %>% 
  filter(keep == 'keep') %>% 
  select(sampleid, mrn, drt) 
```

# Pair the stool with previous 1 day of diet and count how many pairs for each patient

```{r}
# looking at 23575
d1 <- dtb %>% 
  filter(mrn == 23575) %>% 
  select(mrn, foodDayRT) 

s1 <- stb %>% 
  filter(mrn == 23575) %>% 
  select(mrn, drt) %>% 
  mutate(desired_Fdrt = drt - 1) %>% 
  filter(desired_Fdrt %in% d1$foodDayRT)
```

# Pair the stool with previous 2 day of diet and count how many pairs for each patient
 
```{r}
pts <- stb %>% 
  distinct(mrn) %>%   
  pull(mrn)
 
for (pt in pts){ 
    
  stool_samp_for_pt_df =  stb %>% filter(mrn == pt) %>% 
    mutate(p1dF = drt - 1,
           p2dF = drt - 2)
  diet_sapmle_for_pt_df = dtb %>% filter(mrn == pt)
  
  apply(stool_samp_for_pt_df, 1, function(Row){
    if(Row[['p1dF']] %in% diet_sapmle_for_pt_df$foodDayRT & Row[['p2dF']] %in% diet_sapmle_for_pt_df$foodDayRT){

      ret = data_frame(
        p1dF = Row[['p1dF']],
        p2dF = Row[['p2dF']],
        sampleid = Row[['sampleid']],
        mrn = Row[['mrn']],
        drt = Row[['drt']]
      )
      write_csv(ret, str_glue('../data/finalized/paired/p2day/{pt}_{Row[["sampleid"]]}.csv'))
    }
  })
}
```

# Pair the stool with previous 3 day of diet and count how many pairs for each patient
 
```{r}
pts <- stb %>% 
  distinct(mrn) %>% 
  pull(mrn)
 
for (pt in pts){ 
    
  stool_samp_for_pt_df =  stb %>% filter(mrn == pt) %>% 
    mutate(p1dF = drt - 1,
           p2dF = drt - 2,
           p3dF = drt - 3)
  diet_sapmle_for_pt_df = dtb %>% filter(mrn == pt)
  
  apply(stool_samp_for_pt_df, 1, function(Row){
    if(Row[['p1dF']] %in% diet_sapmle_for_pt_df$foodDayRT & 
       Row[['p2dF']] %in% diet_sapmle_for_pt_df$foodDayRT &
       Row[['p3dF']] %in% diet_sapmle_for_pt_df$foodDayRT){

      ret = data_frame(
        p1dF = Row[['p1dF']],
        p2dF = Row[['p2dF']],
        p3dF = Row[['p3dF']],
        sampleid = Row[['sampleid']],
        mrn = Row[['mrn']],
        drt = Row[['drt']]
      )
      write_csv(ret, str_glue('../data/finalized/paired/p3day/{pt}_{Row[["sampleid"]]}.csv'))
    }
  }) 
}
```

# Pair the stool with previous 4 day of diet and count how many pairs for each patient 
 
```{r}
pts <- stb %>% 
  distinct(mrn) %>% 
  pull(mrn)
 
for (pt in pts){ 
    
  stool_samp_for_pt_df =  stb %>% filter(mrn == pt) %>% 
    mutate(p1dF = drt - 1,
           p2dF = drt - 2,
           p3dF = drt - 3,
           p4dF = drt - 4)
  diet_sapmle_for_pt_df = dtb %>% filter(mrn == pt)
  
  apply(stool_samp_for_pt_df, 1, function(Row){
    if(Row[['p1dF']] %in% diet_sapmle_for_pt_df$foodDayRT & 
       Row[['p2dF']] %in% diet_sapmle_for_pt_df$foodDayRT &
       Row[['p3dF']] %in% diet_sapmle_for_pt_df$foodDayRT &
       Row[['p4dF']] %in% diet_sapmle_for_pt_df$foodDayRT){

      ret = data_frame(
        p1dF = Row[['p1dF']],
        p2dF = Row[['p2dF']],
        p3dF = Row[['p3dF']],
        p4dF = Row[['p4dF']],
        sampleid = Row[['sampleid']],
        mrn = Row[['mrn']],
        drt = Row[['drt']]
      )
      write_csv(ret, str_glue('../data/finalized/paired/p4day/{pt}_{Row[["sampleid"]]}.csv'))
    }
  }) 
}
```

# Pair the stool with previous 5 day of diet and count how many pairs for each patient 
 
```{r}
pts <- stb %>% 
  distinct(mrn) %>% 
  pull(mrn)
 
for (pt in pts){ 
    
  stool_samp_for_pt_df =  stb %>% filter(mrn == pt) %>% 
    mutate(p1dF = drt - 1,
           p2dF = drt - 2,
           p3dF = drt - 3,
           p4dF = drt - 4,
           p5dF = drt - 5)
  diet_sapmle_for_pt_df = dtb %>% filter(mrn == pt)
  
  apply(stool_samp_for_pt_df, 1, function(Row){
    if(Row[['p1dF']] %in% diet_sapmle_for_pt_df$foodDayRT & 
       Row[['p2dF']] %in% diet_sapmle_for_pt_df$foodDayRT &
       Row[['p3dF']] %in% diet_sapmle_for_pt_df$foodDayRT &
       Row[['p4dF']] %in% diet_sapmle_for_pt_df$foodDayRT &
       Row[['p5dF']] %in% diet_sapmle_for_pt_df$foodDayRT ){

      ret = data_frame(
        p1dF = Row[['p1dF']],
        p2dF = Row[['p2dF']],
        p3dF = Row[['p3dF']],
        p4dF = Row[['p4dF']],
        p5dF = Row[['p5dF']],
        sampleid = Row[['sampleid']],
        mrn = Row[['mrn']],
        drt = Row[['drt']]
      )
      write_csv(ret, str_glue('../data/finalized/paired/p5day/{pt}_{Row[["sampleid"]]}.csv'))
    }
  }) 
}
```


# check how many qualified pairs for each patient in each group

```{r}
folders <- paste('../data/finalized/paired/', paste(paste('p',seq(2,5), sep = ''),'day', sep = ''), sep = '')

fns <- folders %>% 
  set_names(folders) %>% 
  map(~ data_frame(fn = list.files(path = .))) %>% 
  bind_rows(.id = 'grp') %>% 
  mutate(grp = str_replace(grp, '../data/finalized/paired/','')) %>% 
  separate(fn, into = c('mrn','sampleid'), sep = '_') %>% 
  mutate(sampleid = str_replace(sampleid, '.csv$',''))

 
# n meaning how many samples 
summary_grp <- fns %>% 
  count(grp, mrn) %>% 
  mutate(mrn = as.numeric(mrn))

summary_grp %>% 
  count(grp)

fns %>% 
  count(mrn)
# 
summary_grp %>% 
  ggboxplot(y = 'n', add = 'jitter', 
            color = 'grp', palette = 'aaas',
            xlab = '',
            title = 'Number of paired samples for each patient in each group') +
  facet_wrap(~ grp) +
  theme_cleveland() +
  scale_y_continuous(breaks  = seq(0, 26, 2)) +
  ggsave('../figs/paired_group_summary_samples_n.jpg', width = 8, height = 7)

```

```{r}
summary_grp %>% 
  arrange(n) %>% 
  write_csv('../data/finalized/paired/paired_group_summary_stats.csv')


```

