---
title: "taxUMAP of the combined cohort"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
library(vegan)
library(vdbR)
connect_database()
umap_pt_size <- 0.5
key <- read_csv('../data/cleaned_diet_data/food_group_color_key_final.csv', col_types = 'ccccc')
axis_text_size <- 10
axis_title_size <- 10
```

```{r}
dtb <- read_csv('../data/152_combined_DTB.csv')

# the weight need to be relative so needs to normalize them 


fc_table <- dtb %>% 
  select(mrn, fdrt, Food_code, dehydrated_weight) %>% 
  group_by(mrn, fdrt, Food_code) %>% 
  summarise(total = sum(dehydrated_weight)) %>% 
  ungroup() %>% 
  mutate(index_column = str_glue('P{mrn}_{fdrt}')) %>% 
  select(index_column, Food_code, total)

fc_total <- fc_table %>% 
  group_by(index_column) %>% 
  summarise(daily_total = sum(total))

fc_df <- fc_table %>% 
  left_join(fc_total) %>% 
  mutate(daily_relative = total/daily_total) %>% 
  select(index_column,Food_code, daily_relative ) %>% 
  spread('Food_code', 'daily_relative', fill = 0)

fc_df %>% write_csv('../data/162_food_code_relative.csv')

dtb %>% distinct(mrn)
```

The taxonomy table (e.g., examples/example_data/taxonomy.csv) is expected to resolve higher taxonomic groups for each ASV/OTU in the microbiota table. The index of the taxonomy table should be ASV/OTU labels, while the columns of the taxonomy table should be higher taxonomic categories (e.g., kingdom, phylum, etc.). The columns must be ordered from left to right in decreasing taxonomic hierarchy,

```{r}
actual_foodids <- colnames(fc_df)[2:length(colnames(fc_df))]

final.table <- read_csv('../data/source/final_table_for_writing_out_to_newick.csv') %>% 
  mutate(FoodID = as.character(FoodID))

food_taxa <- final.table %>% 
  filter(FoodID %in% actual_foodids) %>% 
  select(ASV = FoodID, Kingdom = L1, Phylum = L2, Class = L3, Order = L4) 

food_taxa %>% write_csv('../data/162_food_taxa.csv')

top2level <- food_taxa %>% 
  distinct(Kingdom, Phylum)

```
```{bash}
<!-- ~/pipeline/scripts/food_tree/softwares/taxumap/run_taxumap.py -t 162_food_taxa.csv -m 162_food_code_relative.csv --agg_levels Kingdom/Phylum/Class/Order -n 173 --weights 0.01/2/10/10  -->
```

```{r}
# calculate the dominant food group of that day
key <- read_csv('../data/cleaned_diet_data/food_group_color_key_final.csv') %>% mutate(fgrp1 = as.character(fgrp1))
fg_daily <- dtb %>% 
  group_by(mrn, fdrt ) %>% 
  summarise(daily_total = sum(dehydrated_weight))

fg_sum <- dtb %>% 
  mutate(fgrp1 = str_sub(Food_code, 1, 1)) %>% 
   group_by(mrn, fdrt , fgrp1) %>% 
  summarise(daily_fg = sum(dehydrated_weight)) %>% 
  left_join(fg_daily) %>% 
  mutate(fg_relab = daily_fg/daily_total) %>% 
  ungroup() %>% 
  left_join(key) %>% 
  group_by(mrn, fdrt) %>% 
  arrange(desc(fg_relab)) %>% 
  slice(1)
```

```{r}
# find out why the few dots are on that far island
OUT <- read_csv('../data/taxumap_embedding.csv') %>% 
  separate(index_column, into = c('mrn','fdrt'), sep = '_') %>% 
  mutate(mrn = as.numeric(str_replace(mrn, 'P','')),
         fdrt = as.numeric(fdrt)) %>% 
  full_join(fg_sum, by = c("mrn", "fdrt"))

color_key <- OUT %>% 
  distinct( color, shortname) %>% select(shortname, color) %>% 
  deframe()
```
# color by the dominant food group

```{r}
out2_umap <- OUT %>% 
  #arrange(desc(n)) %>% 
  ggscatter(x = 'taxumap1', y = 'taxumap2', color = 'shortname', alpha = 1,size = umap_pt_size , shape = 16,
            xlab = 'TaxUMAP1', ylab = 'TaxUMAP2') +
  scale_color_manual(values = color_key) +
  theme_classic() +
  #lemon::coord_capped_cart(bottom = 'both', left = 'both') +
  theme(legend.position = "none",
        line = element_blank(),
        legend.title = element_blank(),
        axis.title=element_blank(),
        axis.text= element_blank(),
        axis.ticks = element_blank(),
                 panel.background = element_rect(fill='transparent'), #transparent panel bg
    plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
    panel.grid.major = element_blank(), #remove major gridlines
    panel.grid.minor = element_blank(), #remove minor gridlines
    legend.background = element_rect(fill='transparent'), #transparent legend bg
    legend.box.background = element_rect(fill='transparent') ,
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha=1)))
ggsave( '../data/162_umap_fg.png', dpi = 300, width = 2.5, height = 2.5,plot = out2_umap, bg = "transparent")
```

# color by the daily cal intake

```{r}
# calculate daily caloric intake   
cal <- dtb %>% 
  group_by(mrn, fdrt) %>% 
  summarise(total_cal = sum(Calories_kcal)) 
out2_cal <- OUT %>% full_join(cal)
out2_caloric <- out2_cal %>% 
  #arrange(desc(n)) %>% 
  mutate(transform_cal = sqrt(total_cal)) %>% 
  ggscatter(x = 'taxumap1', y = 'taxumap2', color = 'transform_cal', alpha = 1,size = umap_pt_size , shape = 16,
            xlab = '', ylab = '') +
  paletteer::scale_color_paletteer_c("viridis::plasma") +
  theme_classic() +
  #lemon::coord_capped_cart(bottom = 'both', left = 'both') +
  theme(legend.position = "none",
        line = element_blank(),
        legend.title = element_blank(),
        axis.title=element_blank(),
        axis.text= element_blank(),
        axis.ticks = element_blank(),
         panel.background = element_rect(fill='transparent'), #transparent panel bg
    plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
    panel.grid.major = element_blank(), #remove major gridlines
    panel.grid.minor = element_blank(), #remove minor gridlines
    legend.background = element_rect(fill='transparent'), #transparent legend bg
    legend.box.background = element_rect(fill='transparent') ,
        aspect.ratio=1)
ggsave( '../data/162_umap_cal.png', dpi = 300, width = 2.5, height = 2.5,plot = out2_caloric, bg = "transparent")
```

# color by diet alpha diversity

```{r}
faith1 <- read_tsv('../data/156_batch1_faith_pd/alpha-diversity.tsv')  %>% 
  separate(...1, into = c('mrn', 'fdrt'), sep = '_', convert = T)  %>% 
  mutate(mrn = as.numeric(str_replace(mrn, 'P','')))

faith2 <-  read_tsv('../data/139_batch2_faith_pd/alpha-diversity.tsv') %>% 
  separate(...1, into = c('mrn', 'fdrt'), sep = '_', convert = T)  %>% 
  mutate(mrn = as.numeric(str_replace(mrn, 'P','')))

faith <- bind_rows(faith1, faith2)
```
```{r}
out_faith <- OUT %>% 
  full_join(faith)
library(viridis)
umap_faith <- out_faith %>% 
  #arrange(desc(n)) %>% 
  mutate(log10faith = log10(faith_pd + 1)) %>% 
  ggscatter(x = 'taxumap1', y = 'taxumap2', color = 'faith_pd', alpha = 1,size = umap_pt_size, shape = 16,
            xlab = '', ylab = '') +
  scale_color_viridis() +
  theme_classic() +
  theme(line = element_blank(),
        legend.title = element_blank(),
        axis.title=element_blank(),
        axis.text= element_blank(),
        axis.ticks = element_blank(),
        aspect.ratio=1,
                 panel.background = element_rect(fill='transparent'), #transparent panel bg
    plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
    panel.grid.major = element_blank(), #remove major gridlines
    panel.grid.minor = element_blank(), #remove minor gridlines
    legend.background = element_rect(fill='transparent'), #transparent legend bg
    legend.box.background = element_rect(fill='transparent') ,
        legend.position = 'none')
ggsave( '../data/162_umap_faith.png', dpi = 300, width = 2.5,height = 2.5, plot = umap_faith, bg = "transparent")
```


# color by the time relative to transplant

```{r}
# cuztomizing the times bins of the drt
cutoff = 0 # Day 0 is conveneint but this can be varied
n_groups_pre = 3;  # all HCT days <0 will be split into 3 bins; can vary this number to play
n_groups_post = 6; # all HCT days >0 will be split into 6 bins; can vary this number to play
# split the fdrt into groups:
umap_time <- OUT %>% 
  select(taxumap1, taxumap2, fdrt)  
times <- umap_time %>% select( fdrt)  
splits <- times %>% 
  split(.$fdrt <= 0)
pre <- splits %>% 
  pluck('TRUE') %>% 
  mutate(bin = cut_number(fdrt, n_groups_pre))
post <- splits %>% 
  pluck('FALSE') %>% 
  mutate(bin = cut_number(fdrt, n_groups_post)) 
both <- bind_rows(pre, post) %>% 
  distinct(fdrt, .keep_all = T)

umap_time_df <- umap_time %>% 
  left_join(both)

umap_drt_high_to_low_bin <- umap_time_df %>% 
  arrange(desc(fdrt)) %>% 
  #arrange(fdrt) %>% 
   ggscatter(x = 'taxumap1', y = 'taxumap2', color = 'bin', alpha = 1, size = umap_pt_size, shape = 16,
             xlab = '', ylab = '') + 
   scale_color_manual(values = RColorBrewer::brewer.pal(9,"Spectral"))  +
   theme_classic() +
   theme(legend.position = "none",
         line = element_blank(),
         legend.title = element_blank(),
         axis.title=element_blank(),
         axis.text= element_blank(),
         axis.ticks = element_blank(),
              panel.background = element_rect(fill='transparent'), #transparent panel bg
    plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
    panel.grid.major = element_blank(), #remove major gridlines
    panel.grid.minor = element_blank(), #remove minor gridlines
    legend.background = element_rect(fill='transparent'), #transparent legend bg
    legend.box.background = element_rect(fill='transparent') ,
        aspect.ratio=1)
ggsave( '../data/162_umap_drt_high_to_low_bin.png', dpi = 300, width = 2.5, height = 2.5,plot = umap_drt_high_to_low_bin, bg = "transparent")
```


