---
title: "Random Forests for Survival"
output: html_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readxl)
library(randomForestSRC)
library(ggRandomForests)
```

the outcome of interest can be overall survival or Progression-Free-Survival, which Nilu has nicely organized for this cohort



## The patient data including clinical variables

Need to have a more comprehensive tables of the important clinical variables

```{r}
suv_pt <- read_excel('../data/pttable_anon_040920.xlsx', sheet = 1)
dic <- read_csv('../data/mrn_pid_key07-Apr-2020 12.20.csv')

pt_sub <- suv_pt %>% 
  full_join(dic) %>% 
  select(-pid.anon.v1, -X1, -CleanMyeloma)

# also the table that has the median div in peri period
# get the median and do a summarization
samps <- read_csv('../data/sdtable_anon_040720.csv') %>% 
  group_by(pid.anon.v1) %>% 
  summarise(med_SR = median(simpson_reciprocal)) %>% 
  full_join(dic) %>% 
  select(mrn, med_SR)


# this is the more comprehensive table and subset to have only the 240 included in the survival analysis
pt_all <- read_csv('../data/msk duke pttable 03 02 20.csv')
  
  
pt <- pt_all %>% 
  select(mrn, institution, Sex , Age, disease_cat, Pt.Status_cat, Regimen_cat) %>% 
  filter(mrn %in% pt_sub$mrn) %>% 
  full_join(pt_sub) %>% 
  full_join(samps)

# divide the pt to groups low div and high div 
thre_div <- median(pt$med_SR)

pt <- pt %>% 
  mutate(div_grp = if_else(med_SR > thre_div, 'high', 'low'))

```


```{r}
# any NA?
pt %>% 
  summary


# change the column types
pt_final <- pt %>% 
  mutate(institution = if_else(institution == 'MSK Auto', T, F),
         Sex = if_else(Sex == 'M', T, F),
         div_grp = if_else(div_grp == 'high', T, F)) %>% 
  select(-mrn, -Pt.Status_cat,-med_SR)

pt_final %>% 
  summary

colnames(pt_final)

```


## build the model

There can't be any factor strings !!! you need to convert to numeric or logic!!!!!

### For the PFS

```{r}
res <- rfsrc(Surv(OS_landmark_time, DeadorAlive_24) ~ . , data = pt_final %>% select(-PFS_landmrk_time, -POD_24), importance = TRUE)

#plot(gg_error(res))


plot(gg_vimp(res)) +
  theme(legend.position = c(0.8, 0.2)) +
  labs(fill = "VIMP > 0",
       title = 'Overall survival')
```

### For the OS

```{r}
res_os <- rfsrc(Surv(PFS_landmrk_time, POD_24) ~ . , data = pt_final %>% select(-OS_landmark_time, -DeadorAlive_24), importance = TRUE)

plot(gg_vimp(res_os)) +
  theme(legend.position = c(0.8, 0.2)) +
  labs(fill = "VIMP > 0",
       title = 'Progression free survival')
```

## The patient data including only microbiota features

top 20 most abundant genera count matrix and summarize to get the median abundance for each patient

```{r}
cts_genus <- read.csv('../data/cts_genus.csv', check.names = F, row.names = 1)
dim(cts_genus)

N <- 40
topN_genera <- sort(apply(cts_genus, 1, sum), decreasing = T)[1:N]


topN_ma <- cts_genus[rownames(cts_genus) %in% names(topN_genera),]


# get the pt samples corresponding table and do the median summarization
samps <- read_csv('../data/msk duke sdtable 03 02 20.csv') %>% 
  rename(sampleid = sampleid_clean)

topN_df <- topN_ma %>% 
  as.data.frame() %>% 
  rownames_to_column('genus') %>% 
  gather(key = 'sampleid', value = 'relab', names(.)[2]:names(.)[ncol(.)]) %>% 
  left_join(samps %>% select(sampleid, mrn)) %>% 
  group_by(mrn, genus) %>% 
  summarise(med_relab = median(relab)) %>% 
  spread(key = 'mrn', value = 'med_relab', fill = 0) %>% 
  column_to_rownames('genus') %>% 
  as.matrix()
```
