---
title: "Analysis of the fiber types of the patients"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse) 
library(ggpubr)
```

-Calculate the population median of fiber intake (days -10 to +30), from all the patients. Including zeros. 
-Based on the population median, calculate the high and low fiber intake distribution (all patients combined).
-Calculate the high and low fiber distribution in GVHD vs. No-GVHD patients.
-Calculate the absolute fiber intake in grams for each fiber type (days -10 to +30), allocating values of zero for the remaining fiber types, in GVHD and No-GVHD patients.  

```{r}
type <- read_csv('../data/10_patients_foods_fiber_type_all.csv')
DTB <- read_csv('../../food_tree/data/152_combined_DTB.csv')
# join with gvhd info of the patients
allpt <- read_csv('../../food_tree/data/156_transplant_data.csv') %>% 
  mutate(d100_a_gvhd = if_else(d100_a_gvhd == 'NE', 'N/E', d100_a_gvhd))

pts %>% count(d100_a_gvhd)
```


```{r}
# new pts
pts <- read_csv('../data/selected_pts.csv') %>% 
  filter(!is.na(mrn)) %>% inner_join(allpt)
```

```{r}
# to get the patient info of 
# -Median age
# -Male/Female distribution
# -Race and Ethnicity
# -Hematological diagnosis
# get the race from the original table
ptb  <-  read_csv('../data/11_ptb_characteristics.csv')
library(vdbR)
connect_database()
get_table_from_database('patient_allo_ks_20221104')
race <- patient_allo_ks_20221104 %>% 
  filter(mrn %in% pts$mrn) 

ptbr <- ptb %>% 
  left_join(race %>% select(mrn, race_dscrp,ethnicity ))
  
ptbr %>% write_csv('../data/11_ptb_characteristics.csv')

```


```{r}
# calculate daily intake of each food code 
dailycode <- DTB %>%   
  group_by(mrn, fdrt, Food_code) %>% 
  summarise(dcodefiber = sum(Fibers_g)) %>% 
  janitor::clean_names() %>% 
  #filter(dcodefiber > 0) %>% 
  left_join(type, by = "food_code") %>% 
  ungroup() %>% 
  relocate(description, .after = 'soluble') %>% 
  left_join(pts %>% select(mrn, d100_a_gvhd))

dailycode %>% write_csv('../data/patients_16s_data/11_pt_daily_fiber_type.csv')
dailycode %>% write_csv('../data/11_pt_daily_fiber_type.csv')  

key <- read_csv('../../food_tree/data/cleaned_diet_data/food_group_color_key_final.csv', col_types = 'ccccc')
dailyfgfiber <-  DTB %>%   
  mutate(fgrp1 = str_sub(Food_code, 1, 1)) %>% 
  group_by(mrn, fdrt, fgrp1) %>% 
  summarise(dfgfiber = sum(Fibers_g)) %>% 
  left_join(pts %>% select(mrn, d100_a_gvhd)) %>% 
  left_join(key %>% select(fgrp1, shortname))  

dailyfgfiber %>% 
  filter( d100_a_gvhd %in% c('Y','N') & fdrt <= 30) %>% 
  ggscatter(x = 'fdrt', y = 'dfgfiber',  alpha = 0.05, shape = 16,xlab = 'transplant day',
             add = "loess", conf.int = TRUE, ylab = 'Daily fiber intake from each group',
            #title = 'soluble',
            add.params = list(color = 'd100_a_gvhd', size = 1)) +
   scale_color_manual(values = color_key) +
   scale_fill_manual(values = color_key) +
  facet_wrap(~ shortname, scales = 'free')
ggsave('../data/11_daily_fiber_fg.pdf', width = 8, height = 6)

DTB %>% 
  distinct(mrn, fdrt) %>% 
  ungroup() %>% 
  count(mrn) %>% 
  summary(n)
  gghistogram(x = 'n')
  
pts %>% select(starts_with('d100')) %>% 
  inner_join(read_csv('../../'))

dailytotalfiber <- DTB %>% 
  group_by(mrn, fdrt) %>% 
  summarise(dailyfiber = sum(Fibers_g)) %>% 
  ungroup %>% 
  left_join(pts %>% select(mrn, d100_a_gvhd))

dailytotalfiber %>% 
  filter( d100_a_gvhd %in% c('Y','N') & fdrt <= 30) %>% 
  ggscatter(x = 'fdrt', y = 'dailyfiber',  alpha = 0.05, shape = 16,xlab = 'transplant day',
             add = "loess", conf.int = TRUE, ylab = 'Daily fiber intake from each group',
            #title = 'soluble',
            add.params = list(color = 'd100_a_gvhd', size = 1)) +
   scale_color_manual(values = color_key) +
   scale_fill_manual(values = color_key) 


```

```{r}
#   looking at cellulose over time in patients with GVHD vs no GVHD (geom_point and geom_smooth, colored by GVHD day -30 to +30)
# similarly looking at soluble fibers over time in patients with GVHD vs no GVHD (geom_point and geom_smooth, colored by GVHD day -30 to +30)
# similarly looking at insoluble fibers over time in patients with GVHD vs no GVHD (geom_point and geom_smooth, colored by GVHD day -30 to +30)

dailytotaldry <- DTB %>% 
  group_by(mrn, fdrt) %>% 
  summarise(alldry = sum(dehydrated_weight)) 

all_types <- dailycode %>% 
  group_by(mrn, fdrt, fiber_type) %>% 
  summarise(gram = sum(dcodefiber)) %>% 
  ungroup() %>% 
  filter(!is.na(fiber_type)) %>% 
  spread('fiber_type','gram', fill = 0) %>% 
  gather('fiber_type','gram', names(.)[3]:names(.)[ncol(.)])%>% 
  left_join(pts %>% select(mrn, d100_a_gvhd))

all_types %>% write_csv('../data/patients_16s_data/11_patients_fiber_type_intake.csv')
all_types %>% write_csv('../data/11_patients_fiber_type_intake.csv')
```

```{r}
all_types %>% 
  filter(fdrt %in% -10:30) %>% 
  ggscatter(x = 'fdrt', y = 'gram', alpha = 0.05, shape = 16,xlab = 'transplant day',
             add = "loess", conf.int = TRUE, 
            add.params = list(color = 'forestgreen', size = 1)) +
  facet_grid(~ fiber_type) 
ggsave('../data/patients_16s_data/11_fiber_all_type_daily_intake.pdf', width = 15, height = 5)
ggsave('../data/11_fiber_all_type_daily_intake.pdf', width = 15, height = 5)
```

```{r}
# what is with the inulin a horizontal line
inulin <- all_types %>% 
  filter(fiber_type == 'Inulin')

inulin_same_value <- inulin%>% 
  count(gram, sort = T) %>% 
  slice(1) %>% 
  pull(gram)

pt_ate <- inulin %>% 
  filter(gram == inulin_same_value) %>% 
  count(mrn, sort = T)   

what_ate <- pt_ate %>% 
  slice(1) %>% pull(mrn)
# one banana a day ....


# what about the Beta-Glucans
all_types %>% 
  filter(fiber_type == 'Beta-Glucans') %>% 
   count(gram, sort = T) %>% 
  slice(1) %>% 
  pull(gram)
 
# it's because of breakfast cereal 
```

```{r soluble}
all_soluble <- dailycode %>% 
  group_by(mrn, fdrt, soluble) %>% 
  summarise(gram = sum(dcodefiber)) %>% 
  ungroup() %>% 
  filter(!is.na(soluble)) %>% 
  spread('soluble','gram', fill = 0) %>% 
  gather('soluble','gram', names(.)[3]:names(.)[ncol(.)])%>% 
  left_join(pts %>% select(mrn, d100_a_gvhd))

all_soluble %>% write_csv('../data/patients_16s_data/11_patients_fiber_soluble_intake.csv')
all_soluble %>% write_csv('../data/11_patients_fiber_soluble_intake.csv')

all_soluble %>% 
  ggscatter(x = 'fdrt', y = 'gram', alpha = 0.05, shape = 16,xlab = 'transplant day',
             add = "loess", conf.int = TRUE, title = 'soluble',
            add.params = list(color = 'brown', size = 1)) +
  facet_grid(~ soluble) +
  scale_y_sqrt()
ggsave('../data/patients_16s_data/11_fiber_all_soluble_daily_intake.pdf', width = 6, height = 4)
ggsave('../data/11_fiber_all_soluble_daily_intake.pdf', width = 6, height = 4)
```

```{r}
# now join a line of gvhd or not
color_key <- c('#42B440','#EC0000')
names(color_key) <- c('N','Y')
all_types %>% 
  distinct(mrn, d100_a_gvhd) %>% 
  count(d100_a_gvhd)
colnames(pts)
```

```{r}
# cleaning the gvhd information
ptsg <- pts %>% 
  select(hct, d100_a_gvhd:d100_ibmtr_highest_overall_grade) %>% 
  mutate(d100_a_gvhd_onset = ymd(d100_a_gvhd_onset),
         hct = ymd(hct)) %>% 
  
ptsg
```


```{r}
all_soluble %>%  
  filter( fdrt %in% -10:30 & d100_a_gvhd %in% c('Y','N')) %>% 
  ggscatter(x = 'fdrt', y = 'gram', alpha = 0.05, shape = 16,xlab = 'transplant day',
             add = "loess", conf.int = TRUE, title = 'soluble',
            add.params = list(color = 'd100_a_gvhd', size = 1)) +
  facet_grid(~ soluble)  +
  scale_color_manual(values = color_key) +
  scale_fill_manual(values = color_key)+
  scale_y_sqrt() 

all_soluble %>%  
  filter( fdrt %in% -10:30 & d100_a_gvhd %in% c('Y','N')) %>% 
  ggscatter(x = 'fdrt', y = 'gram', alpha = 0.05, shape = 16,xlab = 'transplant day',
             add = "loess", conf.int = TRUE, title = 'soluble',
            add.params = list(color = 'soluble', size = 1)) +
  facet_grid(~ d100_a_gvhd)  +
  scale_color_manual(values = c('lightblue','orange')) +
  scale_fill_manual(values = c('blue','yellow'))+
  scale_y_sqrt() 
ggsave('../data/11_fiber_all_soluble_daily_intake_gvhd_2.pdf', width = 6, height = 4)
ggsave('../data/patients_16s_data/11_fiber_all_soluble_daily_intake_gvhd.pdf', width = 6, height = 4)
ggsave('../data/11_fiber_all_soluble_daily_intake_gvhd.pdf', width = 6, height = 4)
```

```{r}
# maybe break the fdrt down to weeks
soluble_week <- all_soluble %>% 
   mutate(timebin = cut_width(fdrt, 7, boundary=0, closed = 'left'))

soluble_week %>% 
  filter(d100_a_gvhd %in% c('Y','N')) %>% 
  ggboxplot(x = 'timebin', y = 'gram', add = 'jitter', add.params = list(alpha = 0.06),
            xlab = 'transplant day', title = 'Soluble',shape = 16,
            color = 'd100_a_gvhd') +
   facet_grid(~ soluble)  +
  scale_color_manual(values = color_key) +
  scale_fill_manual(values = color_key) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_y_sqrt()

# subsample to select the median value for repeated sampling
soluble_week %>% 
  filter(d100_a_gvhd %in% c('Y','N')) %>% 
  group_by(mrn, d100_a_gvhd, soluble, timebin) %>% 
  summarise(med_week = median(gram)) %>% 
  ggboxplot(x = 'timebin', y = 'med_week', add = 'jitter', add.params = list(alpha = 0.3),
            xlab = 'transplant day', title = 'Soluble',shape = 16,
            color = 'd100_a_gvhd') +
   facet_grid(~ soluble)  +
  scale_color_manual(values = color_key) +
  scale_fill_manual(values = color_key) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_y_sqrt()

ggsave('../data/patients_16s_data/11_fiber_all_soluble_daily_intake_gvhd_week_median.pdf', width = 8, height = 4)
ggsave('../data/11_fiber_all_soluble_daily_intake_gvhd_week_median.pdf', width = 8, height = 4)
```
```{r}
# cellulose
# cel <- all_types %>% 
#   filter(fiber_type == 'Cellulose')
# cel %>% 
#   filter(fdrt %in% -30:30 & d100_a_gvhd %in% c('Y','N')) %>% 
#   ggscatter(x = 'fdrt', y = 'gram', alpha = 0.05, shape = 16,xlab = 'transplant day',
#              add = "loess", conf.int = TRUE, title = 'Cellulose',
#             add.params = list(color = 'd100_a_gvhd', size = 1)) +
#   scale_color_manual(values = color_key) +
#   scale_fill_manual(values = color_key)
# 
# ggsave('../data/patients_16s_data/11_fiber_Cellulose_daily_intake_gvhd.pdf', width = 4, height = 4)
# ggsave('../data/11_fiber_Cellulose_daily_intake_gvhd.pdf', width = 4, height = 4)
```
```{r}
# all of the types of fiber

all_types %>% 
  filter(fdrt %in% -10:30 & d100_a_gvhd %in% c('Y','N')) %>% 
  ggscatter(x = 'fdrt', y = 'gram', alpha = 0.05, shape = 16,xlab = 'transplant day',
             add = "loess", conf.int = TRUE, title = 'All fiber types',
            add.params = list(color = 'd100_a_gvhd', size = 1)) +
  scale_color_manual(values = color_key) +
  scale_fill_manual(values = color_key) +
  facet_grid(~ fiber_type) +
  scale_y_sqrt()

ggsave('../data/patients_16s_data/11_fiber_all_daily_intake_gvhd.pdf', width = 15, height = 4)
ggsave('../data/11_fiber_all_daily_intake_gvhd.pdf', width = 15, height = 4)
```

```{r}
# randomly select one sample per patient 
set.seed(666)

type_unisample <- all_types %>% 
  split(.$fiber_type) %>% 
  map_dfr(function(df){
    df %>%  group_by(mrn) %>% sample_n(size = 1, replace = F)
  })


type_unisample %>% 
  filter( d100_a_gvhd %in% c('Y','N')) %>% 
  ggscatter(x = 'fdrt', y = 'gram', alpha = 0.2, shape = 16,xlab = 'transplant day',
             add = "loess", conf.int = TRUE, title = 'All fiber types one sample per pt',
            add.params = list(color = 'd100_a_gvhd', size = 1)) +
  scale_color_manual(values = color_key) +
  scale_fill_manual(values = color_key) +
  scale_y_sqrt() +
  facet_grid(~ fiber_type)

ggsave('../data/patients_16s_data/11_fiber_all_daily_intake_gvhd_one_sample_per_pt.pdf', width = 15, height = 4)
ggsave('../data/11_fiber_all_daily_intake_gvhd_one_sample_per_pt.pdf', width = 15, height = 4)

```

```{r}

solu_unisample <- all_soluble %>% 
  split(.$soluble) %>% 
  map_dfr(function(df){
    df %>%  group_by(mrn) %>% sample_n(size = 1, replace = F)
  })

solu_unisample %>% 
  filter( d100_a_gvhd %in% c('Y','N')) %>% 
  ggscatter(x = 'fdrt', y = 'gram', alpha = 0.2, shape = 16,xlab = 'transplant day',
             add = "loess", conf.int = TRUE, title = 'soluble',
            add.params = list(color = 'd100_a_gvhd', size = 1)) +
  facet_grid(~ soluble)  +
  scale_color_manual(values = color_key) +
  scale_fill_manual(values = color_key)+
  scale_y_sqrt()

ggsave('../data/patients_16s_data/11_fiber_all_soluble_daily_intake_one_sample_per_pt.pdf', width = 6, height = 4)
ggsave('../data/11_fiber_all_soluble_daily_intake_one_sample_per_pt.pdf', width = 6, height = 4)
```
```{r}
# use the percentage to combat the zero fiber eating thing 
```

```{r}
median_now <- median(meta$mean_fiber)
meta <- read_csv('~/pipeline/scripts/food_tree/data/090_all_samples_meta_p2d_fg9_dietall_genera90.csv') %>% 
  inner_join( pts)%>% 
  filter(sdrt %in% 7:21)%>% 
  group_by(mrn) %>% 
  summarise(ave_div = mean(simpson_reciprocal),
            mean_fiber = mean(ave_fiber)) %>% 
  mutate(mean_fiber = mean_fiber * 100) %>% 
  mutate(grp = if_else(mean_fiber >= median_now, 'high', 'low'))

meta  %>% 
  ggscatter(x = 'mean_fiber', y = 'ave_div',
            xlab = 'Prior 2 day fiber intake averaged by patient', ylab = 'microbiome alpha diversity averaged by patient',
            add = "reg.line",  # Add regressin line
           add.params = list(color = "blue", fill = "lightgray"), # Customize line
           conf.int = TRUE, # Add confidence interval
           cor.coef = TRUE, # Add correlation coefficient.
           cor.coeff.args = list(method = "pearson",  label.sep = "\n")) +
  theme(aspect.ratio=1)
ggsave('../data/11_alpha_and_fiber.pdf')
```

```{r}
meta %>% 
  ggboxplot(x = 'grp', y = 'ave_div', add = 'jitter') +
  stat_compare_means(comparisons= list(c('high', 'low')),
            label= "p.format",
            method= 'wilcox.test',
						correct=FALSE)
ggsave('../data/11_box_low_and_high.pdf')
```

```{r}

```

