---
title: "New user connects to vdb database"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

This document lists the steps how a new user can connect to the vdbr database and also some important notes.

**Please make sure you are compliant with the protocol and be careful of using the information in this database since there is sensitive PHI contained.**

**AAA** refers to the current database manager you should contact. Currently it's Angel (daia1@mskcc.org)

## AAA notify you with a temporary password with your new account

You get information like this: (an example)

- user: mab4025
- temporary password: test123456
- pass expiration: 2020-1-17

**Please act fast and finish the below steps before the expiration date. Thank you.**

## Install the vdbR package

```{r }
remotes::install_git("https://github.mskcc.org/gomesa/vdbR.git",quiet=T)
```

**Choose 1 to install or update all necessary package**

If the above line doen't run:

- Your R version might need to be updated, for example at the time of this document's creation, you need to use R 4.3.0
- You need to have the package `remotes`

```{r}
# to test if this package has been successfully installed
library(vdbR)
```

## Create a config file and put it at your home directory

Create a config file and put it at your home directory: `~`

The config file should be named "dbConfig.txt" and the format should be following the example file in this folder. (See the file dbConfig.txt)

Don't worry you will only need to do this once. After you get this done you don't need to worry about it further.
 

```{r}
# run the below lines 
library(vdbR)
connect_database(config_file = "~/dbConfig.txt")
```
 
Then you are prompted to create a new password. Doesn't matter how complicated it is, as long as it's different from the temp. And your new password will be writen to your config file.

## Notify AAA of your password change so he/she can unlock you

AAA rerun the create_new_postgres_user_vbdlab.R script

## Then you should be good to go, try it with commands like

```{r}
list_table_from_database()
get_table_from_database("asv_alpha_diversity_ag")
```

