---
title: "Creating heatmap using pheatmap pkg"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(tidyverse)
library(pheatmap) # installing with install.packages("pheatmap")
```

## An example of formatting the input table

```{r}
# load the pheno and counts table of your analysis.
pheno <- read_csv('~/melodyP/data/cleaned_pheno/final_pheno.csv')

genus <- read_tsv('~/melodyP/data/humann2_res/genus_level/humann2_pathabundance_genus_level_cpm_stratified.tsv') %>% 
  dplyr::rename(pw = names(.)[1]) %>% 
  rename_all(funs(str_replace(., '_genefamilies_.+$','')))

genus_entero <- genus %>% 
  filter(str_detect(pw, 'g__Enterococcus')) 

# The input counts table HAS TO be a matrix
# with feature name as row names and sample names as column names.
genus_entero_mat <- genus_entero %>% 
  column_to_rownames('pw') %>% 
  as.matrix

# this below table links your sample with different pheno data you are interested in
# and the row names HAVE TO be the sample name or whatever sample ID that is in your counts table.
annot <- data.frame(row.names = pheno$sampleid, 
                    tox = pheno$Toxicity, 
                    CR = pheno$CR)
```

## Plot heatmap with customization

If you don't provide a path to the filename, the plot will show up in the Rstudio directly.

```{r}
# a heatmap to see the counts distribution


pheatmap(
  genus_entero_mat,
  annotation_col = annot,
  show_rownames = F,
  show_colnames = T,
  #filename = '../../figs/Enterococcus_heatmap.jpg',
  height = 18,
  width = 18,
  scale = 'row',
  cluster_rows = T,
  cluster_cols = T
)
```

**Some important parameters, please refer to the help page for more details**

* scale: whether the values should be centered and scaled
* cluster_rows/cluster_cols: whether to cluster the rows or the columns. There are several distance metric that could be used.

## Customize the pheno annotation's color

If you wanna customize the pheno annotation's color, to align with your other plots color palette for example.

```{r}
annotCol <- list(tox = c('#00468B', '#EC0000'),
                 CR = c('#42B540','#925E9F'))
names(annotCol$tox) = c('Y','N')
names(annotCol$CR) = c('Y','N')

annotCol
```
```{r}
# now the color of the pheno label is changed.
 
pheatmap(
  genus_entero_mat,
  annotation_col = annot,
  annotation_colors = annotCol,
  show_rownames = F,
  show_colnames = T,
  #filename = '../figs/Enterococcus_heatmap.jpg',
  height = 18,
  width = 18,
  scale = 'row',
  cluster_rows = T,
  cluster_cols = T
)
```

And of course you can experiment with different color palette for the heatmap too. 

